### 2025-01-23 hw1/submission_notebook.Rmd
Data localization, notebook - Localized geodata data with default paths - Added R markdown notebook

---
title: "Geospatial - Assignment 1 - Moritz Peist, Noemi Lucchi, Simon Vellin"
output:
  html_document:
    toc: true
  pdf_document:
    toc: true
---

```{r setup, include=FALSE} 
# Disable warning messages
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# Methodology & general code

In the following notebook we replicate the charts for all 5 papers. Contentwise, the code follows always the same structure. We first load the data, then we create the chart and finally we save the chart as a .png file.

First we load the needed packages:

```{r}
library(osmdata)
library(sf)
library(spData)
library(tidyverse)
library(geodata)
library(readxl)
library(ggplot2)
library(raster)
library(terra)
```

## [Paper 1](https://www.sciencedirect.com/science/article/pii/S0304387818300166#appsec1)

### Data loading

The authors of the paper provide a link to the data, however, the link is unfortunately outdated. Yet, the data is available in a newer version on the website (authors used 2013 data, we use 2024 data).

Moreover, the river gradient data was not directly available, so we use the elevation data from the geodata package to retrieve and later calculate gradients.

```{r}
# Dams, Source:https://www.dws.gov.za/DSO/publications.aspx
# Data was only available as .KMZ and not readable by st_read so needed to be converted by:
# https://mygeodata.cloud/converter/kml-to-geojson
dams <- st_read(
  "data/paper1/Registered Dams Oct20241.geojson"
)

# Dam metadata (same source but an Excel file)
dam_met <- read_excel(
  "data/paper1/List of Registered Dams Oct2024.xlsx",
  col_types = rep("text", ncol(read_excel("data/paper1/List of Registered Dams Oct2024.xlsx")))
) %>%
  mutate(`Completion date` = na_if(as.numeric(`Completion date`), 0))

# RSA boundaries - downloaded from GDAM and temporarily stored
sa_districts_sf <- st_as_sf(gadm(country = "ZAF", level = 3, path = tempdir()))

# River data - Download elevation data for South Africa
elevation <- elevation_30s(country = "ZAF", path = tempdir())
```

### Data wrangling

We try to apply the same data cleaning as the authors though we deal with more recent data, i.e.:

> "The data was accessed on August 23, 2013, including all dams listed as of March 2013. About 15 percent of dams were dropped because of missing critical information that could not be verified (690/4,830). For most missing dams (658/690), the missing information was completion date. Online internet searches for this information suggest that most were very old (prior to the analysis period) or no longer operational. Of those dropped, 499 reported a purpose of “irrigation.” The resulting dataset included 4,140  dams, which were restricted to the subset of irrigation dams (3,176)."

Furthermore, we calculate the mean gradient for each district and add it to the districts data.

```{r}
# Now filter dams, as done in the paper, i.e, omit missing completion date and filter for irrigation
dams <- dams %>%
  left_join(dam_met,
    by = join_by("COL52A1FC9ABE076254" == "No of dam"),
    # relationship = "many-to-many"
  ) %>%
  filter(!is.na(`Completion date`), str_detect(tolower(Purpose), "irrigation")) %>%
  distinct(across(-`geometry`), .keep_all = TRUE) # Delete the 2 duplicate cases identified

# Convert districts to terra format for raster operations
districts_vect <- vect(sa_districts_sf)

# Calculate slope from elevation
slope <- terrain(elevation, v = "slope", unit = "degrees")

# Calculate zonal statistics (mean slope) for each district
district_gradients <- terra::extract(slope, districts_vect, fun = mean, na.rm = TRUE)

# Add gradient data to districts
sa_districts_sf$gradient <- district_gradients$slope

# Ensure same CRS
dams <- st_transform(dams, st_crs(sa_districts_sf))

# Clip dams to South Africa boundaries
sa_boundary <- st_union(sa_districts_sf)
dams_clipped <- st_intersection(dams, sa_boundary)
```

### Plotting

Finally, we output the plot:

```{r}
# Create plot
ggplot() +
  # Add districts with gradient fill
  geom_sf(
    data = sa_districts_sf,
    aes(fill = gradient),
    color = "white",
    size = 0.1
  ) +
  # Add clipped dams
  geom_sf(
    data = dams_clipped,
    size = 0.2,
    color = "black",
    aes(shape = "Dam Location")
  ) + # Add to legend
  # Style the gradient fill
  scale_fill_gradientn(
    colors = grey.colors(6, start = 0.9, end = 0.2),
    name = "Average District\nRiver Gradient",
    breaks = c(
      0, 2.229103, 3.821713, 6.593167,
      9.879707, 13.130823, 19.829017
    ),
    labels = function(x) format(x, digits = 6),
    guide = guide_colorbar(
      order = 1,
      nbin = 100, # For smoother color transition
      ticks.colour = "black",
      frame.colour = "black",
      frame.linewidth = 0.5
    )
  ) +
  # Add shape scale for dam points in legend
  scale_shape_manual(
    values = c("Dam Location" = 16),
    name = NULL,
    guide = guide_legend(order = 2)
  ) +
  # Customize theme
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 14),
    legend.position = "left",
    legend.box = "vertical",
    legend.margin = margin(0, 0, 0, 0),
    legend.spacing = unit(0.1, "cm"),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    legend.key.size = unit(0.5, "cm")
  ) +
  # Add titles
  labs(
    title = "South African Dams and River Gradients",
    caption = "Source: Based on Mettetal (2019) but with 2024 instead of 2013 data"
  )
```

In comparison the first paper map looked as follows:

![Figure 2](data/paper1/1-s2.0-S0304387818300166-gr2.jpg)

## [Paper 2](https://www.sciencedirect.com/science/article/pii/S0166046220303100)

### Data loading

Unfortunately, again the data was not directly available. Therefore, we leverage some r packages to get the data for highways, powergrids (osmdata) and population data from 2020 (geodata). The ESSR survey data was directly available via the World Bank data repositories as a CSV (2011/12) data used as in the paper. We load all of these in the following:

```{r}
# Get Ethiopia districts
eth_districts_sf <- gadm(country = "ETH", level = 3, path = tempdir())
eth_districts_sf <- st_as_sf(eth_districts_sf)

# Get population data and process it for Ethiopia
ethiopia_pop <- population(2020, res = 5, path = tempdir())
ethiopia_pop <- crop(ethiopia_pop, ext(eth_districts_sf))
ethiopia_pop <- mask(ethiopia_pop, vect(eth_districts_sf))

# Convert districts to same CRS as population data if needed
eth_districts_sf <- st_transform(eth_districts_sf, crs(ethiopia_pop))

# WB data: https://microdata.worldbank.org/index.php/catalog/2053/data-dictionary
# Read in the survey data and get it in the right shape
wb_data <- read_csv("data/paper2/pub_eth_householdgeovariables_y1.csv") %>%
  group_by(ea_id) %>%
  summarise(
    LAT_DD_MOD = mean(LAT_DD_MOD, na.rm = TRUE),
    LON_DD_MOD = mean(LON_DD_MOD, na.rm = TRUE)
  )

# Transform it to GEO data and clip to Ethiopia
wb_data_geo <- st_as_sf(wb_data, coords = c("LON_DD_MOD", "LAT_DD_MOD"), crs = 4326)
wb_data_geo <- st_transform(wb_data_geo, crs(ethiopia_pop))

# Get high voltage power lines with comprehensive voltage range
ethiopia_hv_power <- opq("Ethiopia") %>%
  add_osm_feature(key = "power", value = c("line", "cable")) %>%
  add_osm_feature(key = "voltage", value = paste0(seq(66, 500, by = 1), "000")) %>%
  osmdata_sf()

# Clean and clip power lines
hv_lines_clean <- ethiopia_hv_power$osm_lines %>%
  mutate(voltage = str_extract(voltage, "\\d+")) %>%
  mutate(voltage_kv = as.numeric(voltage) / 1000) %>%
  filter(!is.na(voltage_kv)) %>%
  st_intersection(eth_districts_sf)

# Get major roads and clip to Ethiopia
ethiopia_roads <- opq("Ethiopia") %>%
  add_osm_feature(
    key = "highway",
    value = c("motorway", "trunk")
  ) %>%
  osmdata_sf()
```

### Data wrangling

We calculate the population density for each district and clip the roads to the Ethiopia boundary.

```{r}
# Clip roads to Ethiopia boundary
roads_clipped <- ethiopia_roads$osm_lines %>%
  st_intersection(eth_districts_sf)

# Calculate population density by district
districts_pop <- terra::extract(ethiopia_pop, vect(eth_districts_sf), fun = sum, na.rm = TRUE)

# Add population to districts dataset
eth_districts_sf$total_pop <- districts_pop[, 2] # The second column contains the values

# Calculate area in square kilometers
eth_districts_sf$area_km2 <- as.numeric(st_area(eth_districts_sf)) / 1000000

# Calculate density (people per km2)
eth_districts_sf$pop_density <- eth_districts_sf$total_pop / eth_districts_sf$area_km2
```

### Plotting

Finally, we output the plot:

```{r}
# Create the final plot
ggplot() +
  # Add Ethiopia districts with population density fill
  geom_sf(
    data = eth_districts_sf,
    aes(fill = pop_density),
    color = "gray50",
    linewidth = 0.1
  ) +
  # Add power lines in red
  geom_sf(
    data = hv_lines_clean,
    color = "red",
    linewidth = 0.35,
    size = 0.35
  ) +
  # Add major roads in black
  geom_sf(
    data = roads_clipped,
    color = "black",
    linewidth = 0.5,
    size = 0.5
  ) +
  # Add villages as points
  geom_sf(
    data = wb_data_geo,
    color = "black",
    size = 1,
    alpha = 0.6,
    shape = 16 # solid circle
  ) +
  # Customize the fill scale to match paper's blue shading
  scale_fill_gradientn(
    colors = c("white", "lightblue", "royalblue", "navy"),
    trans = "log",
    name = "Population\nDensity\n(people/km²)",
    labels = scales::comma
  ) +
  theme_minimal() +
  labs(
    title = "High Voltage Power Grid Network and Major Roads in Ethiopia",
    subtitle = "Showing transmission lines ≥66kV, population density by district, and villages",
    caption = "Data sources: OpenStreetMap, WorldPop 2020, World Bank\n
      Source: Based on Fried and Lagakos (2021)"
  ) +
  theme(
    legend.position = "right",
    plot.title = element_text(size = 10),
    plot.subtitle = element_text(size = 8),
    plot.caption = element_text(size = 6)
  )
```

In comparison the second paper map looked as follows: ![Figure 4](data/paper2/1-s2.0-S0166046220303100-gr4.jpg)

The different color labelling of the surveyed villages is unfortunately not explained in the paper. Therefore, we leave the points black.

## Paper 3

### Data loading

### Data wrangling

### Plotting

In comparison the third paper map looked as follows: ![Figure x]()

## Paper 4

### Data loading

### Data wrangling

### Plotting

In comparison the fourth paper map looked as follows: ![Figure x]()

## Paper 5

### Data loading

### Data wrangling

### Plotting

In comparison the fifth paper map looked as follows: ![Figure x]()

### 2025-01-23 hw1/submission_notebook.tex
Data localization, notebook - Localized geodata data with default paths - Added R markdown notebook

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Geospatial - Assignment 1 - Moritz Peist, Noemi Lucchi, Simon Vellin},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Geospatial - Assignment 1 - Moritz Peist, Noemi Lucchi, Simon
Vellin}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\section{Methodology}\label{methodology}

In the following notebook we replicate the charts for all 5 papers.
Contentwise, the code follows always the same structure. We first load
the data, then we create the chart and finally we save the chart as a
.png file.

\subsection{\texorpdfstring{\href{https://www.sciencedirect.com/science/article/pii/S0304387818300166\#appsec1}{Paper
1}}{Paper 1}}\label{paper-1}

\subsubsection{Data loading}\label{data-loading}

The first paper map looked as follows:
\includegraphics{hw1/data/paper1/1-s2.0-S0304387818300166-gr2.jpg}

\subsection{Paper 2}\label{paper-2}

The second paper map looked as follows:
\includegraphics{hw1/data/paper2/1-s2.0-S0166046220303100-gr4.jpg}

\subsection{Paper 3}\label{paper-3}

\subsection{Paper 4}\label{paper-4}

\subsection{Paper 5}\label{paper-5}

This is an \href{http://rmarkdown.rstudio.com}{R Markdown} Notebook.
When you execute code within the notebook, the results appear beneath
the code.

Try executing this chunk by clicking the \emph{Run} button within the
chunk or by placing your cursor inside it and pressing
\emph{Ctrl+Shift+Enter}.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot}\NormalTok{(cars)}
\end{Highlighting}
\end{Shaded}

\includegraphics{submission_notebook_files/figure-latex/unnamed-chunk-2-1.pdf}

Add a new chunk by clicking the \emph{Insert Chunk} button on the
toolbar or by pressing \emph{Ctrl+Alt+I}.

When you save the notebook, an HTML file containing the code and output
will be saved alongside it (click the \emph{Preview} button or press
\emph{Ctrl+Shift+K} to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the
editor. Consequently, unlike \emph{Knit}, \emph{Preview} does not run
any R code chunks. Instead, the output of the chunk when it was last run
in the editor is displayed.

\end{document}

### 2025-01-23 hw1/paper1.R

@@ -7,6 +7,9 @@ library(readxl)
 
 # Load data (paper: https://www.sciencedirect.com/science/article/pii/S0304387818300166#appsec1)
 
+# Set geodata default path
+geodata_path("hw1/data/paper1")
+
 # Dams, Source:https://www.dws.gov.za/DSO/publications.aspx
 # Data was only available as .KMZ and not readable by st_read so needed to be converted by:
 # https://mygeodata.cloud/converter/kml-to-geojson
@@ -14,12 +17,18 @@ dams <- st_read(
   "hw1/data/paper1/Registered Dams Oct20241.geojson"
 )
 
-# Dam metadata
+# Dam metadata (same source but an Excel file)
 dam_met <- read_excel(
   "hw1/data/paper1/List of Registered Dams Oct2024.xlsx",
   col_types = rep("text", ncol(read_excel("hw1/data/paper1/List of Registered Dams Oct2024.xlsx")))
 ) %>%
   mutate(`Completion date` = na_if(as.numeric(`Completion date`), 0))
+
+# RSA boundaries - downloaded from GDAM and temporarily stored
+sa_districts_sf <- st_as_sf(gadm(country = "ZAF", level = 3))
+
+# River data - Download elevation data for South Africa
+elevation <- elevation_30s(country = "ZAF")
 
 
 # Now filter dams, as done in the paper, i.e, omit missing completion date and filter for irrigation
@@ -32,12 +41,6 @@ dams <- dams %>%
   distinct(across(-`geometry`), .keep_all = TRUE) # Delete the 2 duplicate cases identified
 
 
-# RSA boundaries - downloaded from GDAM and temporarily stored
-sa_districts_sf <- st_as_sf(gadm(country = "ZAF", level = 3, path = tempdir()))
-
-# River data - Download elevation data for South Africa
-elevation <- elevation_30s(country = "ZAF", path = tempdir())
-
 # Convert districts to terra format for raster operations
 districts_vect <- vect(sa_districts_sf)
 
@@ -49,8 +52,6 @@ district_gradients <- terra::extract(slope, districts_vect, fun = mean, na.rm = 
 
 # Add gradient data to districts
 sa_districts_sf$gradient <- district_gradients$slope
-
-# Create plot
 
 # Ensure same CRS
 dams <- st_transform(dams, st_crs(sa_districts_sf))

### 2025-01-23 hw1/paper2.R

@@ -9,12 +9,17 @@ library(ggplot2)
 library(raster)
 library(terra)
 
+# Paper: https://www.sciencedirect.com/science/article/pii/S0166046220303100
+
+# Set geodata default path
+geodata_path("hw1/data/paper2")
+
 # Get Ethiopia districts
-eth_districts_sf <- gadm(country = "ETH", level = 3, path = tempdir())
+eth_districts_sf <- gadm(country = "ETH", level = 3)
 eth_districts_sf <- st_as_sf(eth_districts_sf)
 
 # Get population data and process it for Ethiopia
-ethiopia_pop <- population(2020, res = 5, path = tempdir())
+ethiopia_pop <- population(2020, res = 5)
 ethiopia_pop <- crop(ethiopia_pop, ext(eth_districts_sf))
 ethiopia_pop <- mask(ethiopia_pop, vect(eth_districts_sf))
 
@@ -80,28 +85,6 @@ ggplot() +
     color = "gray50",
     linewidth = 0.1
   ) +
-  # Add power lines in red
-  geom_sf(
-    data = hv_lines_clean,
-    color = "red",
-    linewidth = 0.35,
-    size = 0.35
-  ) +
-  # Add major roads in black
-  geom_sf(
-    data = roads_clipped,
-    color = "black",
-    linewidth = 0.5,
-    size = 0.5
-  ) +
-  # Add villages as points
-  geom_sf(
-    data = wb_data_geo,
-    color = "black",
-    size = 1,
-    alpha = 0.6,
-    shape = 16 # solid circle
-  ) +
   # Customize the fill scale to match paper's blue shading
   scale_fill_gradientn(
     colors = c("white", "lightblue", "royalblue", "navy"),
@@ -113,11 +96,41 @@ ggplot() +
   labs(
     title = "High Voltage Power Grid Network and Major Roads in Ethiopia",
     subtitle = "Showing transmission lines ≥66kV, population density by district, and villages",
-    caption = "Data sources: OpenStreetMap, WorldPop 2020, World Bank"
+    caption = "Data sources: OpenStreetMap, WorldPop 2020, World Bank\n
+      Source: Based on Fried and Lagakos (2021)"
   ) +
   theme(
-    legend.position = "right",
+    legend.position = "left",
     plot.title = element_text(size = 10),
     plot.subtitle = element_text(size = 8),
     plot.caption = element_text(size = 6)
+  ) +
[…]

### 2025-01-23 hw1/submission_notebook.Rmd
Paper 1 & 2 done with caching

@@
 ```{r}
+# Set geodata default path
+geodata_path("data/paper1")
+
 # Dams, Source:https://www.dws.gov.za/DSO/publications.aspx
@@
 # RSA boundaries - downloaded from GDAM and temporarily stored
-sa_districts_sf <- st_as_sf(gadm(country = "ZAF", level = 3, path = tempdir()))
+sa_districts_sf <- st_as_sf(gadm(country = "ZAF", level = 3))
 
 # River data - Download elevation data for South Africa
-elevation <- elevation_30s(country = "ZAF", path = tempdir())
+elevation <- elevation_30s(country = "ZAF")
 ```
@@
 ```{r}
+# Set geodata default path
+geodata_path("data/paper2")
+
 # Get Ethiopia districts
-eth_districts_sf <- gadm(country = "ETH", level = 3, path = tempdir())
+eth_districts_sf <- gadm(country = "ETH", level = 3)
 eth_districts_sf <- st_as_sf(eth_districts_sf)
@@
 # Get population data and process it for Ethiopia
-ethiopia_pop <- population(2020, res = 5, path = tempdir())
+ethiopia_pop <- population(2020, res = 5)
 ethiopia_pop <- crop(ethiopia_pop, ext(eth_districts_sf))
@@
 ggplot() +
-  # Add Ethiopia districts with population density fill
   geom_sf(
@@
   ) +
-  # Add power lines in red
-  geom_sf(
-    data = hv_lines_clean,
-    color = "red",
-    linewidth = 0.35,
-    size = 0.35
-  ) +
-  # Add major roads in black
-  geom_sf(
-    data = roads_clipped,
-    color = "black",
-    linewidth = 0.5,
-    size = 0.5
-  ) +
-  # Add villages as points
-  geom_sf(
-    data = wb_data_geo,
-    color = "black",
-    size = 1,
-    alpha = 0.6,
-    shape = 16 # solid circle
-  ) +
-  # Customize the fill scale to match paper's blue shading
   scale_fill_gradientn(
@@
     name = "Population\nDensity\n(people/km²)",
-    labels = scales::comma
+    labels = scales::comma,
+    guide = guide_colorbar(order = 1)
+  ) +
+  geom_sf(
+    data = roads_clipped,
+    aes(color = "Highway"),
+    linewidth = 0.5
+  ) +
+  geom_sf(
+    data = hv_lines_clean,
+    aes(color = "HV grid"),
+    linewidth = 0.35
+  ) +
+  geom_sf(
+    data = wb_data_geo,
+    aes(shape = "Survey village"),
+    color = "black",
+    size = 1,
+    alpha = 0.6
+  ) +
+  scale_color_manual(
+    name = NULL,
+    values = c("Highway" = "black", "HV grid" = "red"),
+    guide = guide_legend(order = 2)
+  ) +
+  scale_shape_manual(
+    name = NULL,
+    values = c("Survey village" = 16),
+    guide = guide_legend(order = 3)
+  ) +
+  guides(
+    fill = guide_colorbar(order = 1),
+    color = guide_legend(
+      order = 2,
+      override.aes = list(
+        linewidth = 1,
+        shape = NA, # Removes the box
+        linetype = 1
+      )
+    ),
+    shape = guide_legend(order = 3)
   ) +
@@
     subtitle = "Showing transmission lines ≥66kV, population density by district, and villages",
-    caption = "Data sources: OpenStreetMap, WorldPop 2020, World Bank\n
-      Source: Based on Fried and Lagakos (2021)"
+    caption = "Data sources: OpenStreetMap, WorldPop 2020, World Bank (2012)\n
+      Based on Fried and Lagakos (2021)"
   ) +
   theme(
-    legend.position = "right",
+    legend.position = "left",
     plot.title = element_text(size = 10),
     plot.subtitle = element_text(size = 8),
-    plot.caption = element_text(size = 6)
+    plot.caption = element_text(size = 6),
+    axis.text = element_blank(),
+    axis.title = element_blank()
   )
@@
 
-The different color labelling of the surveyed villages is unfortunately not explained in the paper. Therefore, we leave the points black.
+The different color labeling of the surveyed villages is unfortunately not explained in the paper. Therefore, we leave the points black.
 

### 2025-01-23 hw1/paper2.R

@@ -78,19 +78,57 @@ eth_districts_sf$pop_density <- eth_districts_sf$total_pop / eth_districts_sf$ar
 
 # Create the final plot
 ggplot() +
-  # Add Ethiopia districts with population density fill
   geom_sf(
     data = eth_districts_sf,
     aes(fill = pop_density),
     color = "gray50",
     linewidth = 0.1
   ) +
-  # Customize the fill scale to match paper's blue shading
   scale_fill_gradientn(
     colors = c("white", "lightblue", "royalblue", "navy"),
     trans = "log",
     name = "Population\nDensity\n(people/km²)",
-    labels = scales::comma
+    labels = scales::comma,
+    guide = guide_colorbar(order = 1)
+  ) +
+  geom_sf(
+    data = roads_clipped,
+    aes(color = "Highway"),
+    linewidth = 0.5
+  ) +
+  geom_sf(
+    data = hv_lines_clean,
+    aes(color = "HV grid"),
+    linewidth = 0.35
+  ) +
+  geom_sf(
+    data = wb_data_geo,
+    aes(shape = "Survey village"),
+    color = "black",
+    size = 1,
+    alpha = 0.6
+  ) +
+  scale_color_manual(
+    name = NULL,
+    values = c("Highway" = "black", "HV grid" = "red"),
+    guide = guide_legend(order = 2)
+  ) +
+  scale_shape_manual(
+    name = NULL,
+    values = c("Survey village" = 16),
+    guide = guide_legend(order = 3)
+  ) +
+  guides(
+    fill = guide_colorbar(order = 1),
+    color = guide_legend(
+      order = 2,
+      override.aes = list(
+        linewidth = 1,
+        shape = NA, # Removes the box
+        linetype = 1
+      )
+    ),
+    shape = guide_legend(order = 3)
   ) +
   theme_minimal() +
   labs(
@@ -103,34 +141,7 @@ ggplot() +
     legend.position = "left",
     plot.title = element_text(size = 10),
     plot.subtitle = element_text(size = 8),
-    plot.caption = element_text(size = 6)
-  ) +
-  # Add shape scale for villlage points in legend
-  scale_color_manual(
-    values = c("HV grid" = "red", "Highway" = "black"),
-    name = NULL,
-    guide = guide_legend(order = 2)
-  ) +
-  scale_size_manual(
-    values = c("Survey village" = 1),
-    name = NULL,
[…]

### 2025-01-24 hw1/paper3.R
Initial paper 3 commit

@@ -1,0 +1,69 @@
+# Clean working directory and import necessary packages 
+rm(list = ls())
+library(sf)
+library(geobr)
+library(dplyr)
+library(ggplot2)
+library(readxl)
+library(tidyr)
+
+# Read the .xls file with the Brasilian population for 2009  
+df <- read_excel("data/paper3/UF_Municipio.xls", skip = 4, col_names = TRUE)
+df
+
+# Get the geographical data from the package geobr
+state <- read_state(year=2010)
+meso <- read_meso_region(year=2010)
+state
+meso
+
+# Convert ESTIMADA (=population) into a numerical variable and handle missing values 
+# by setting them to 0
+df <- df %>%
+  mutate(ESTIMADA = as.numeric(ESTIMADA)) %>%
+  mutate(ESTIMADA = replace_na(ESTIMADA, 0))
+
+# Group by the country code 
+df_grouped <- df %>% 
+  group_by(COD...2) %>%
+  summarise(
+    Total_Estimada = sum(ESTIMADA, na.rm = TRUE),  
+    Count = n()  
+)
+
+# Merge the 'meso' data with the population data 
+df_merged <- meso %>%
+  left_join(df_grouped, by = c("code_state" = "COD...2"))
+
+# Merge the resulting df with the 'state' data to get the orientation 
+df_merged <- df_merged %>%
+  st_join(state)
+
+# Create the column pop_share
+df_merged <- df_merged %>%
+mutate(pop_share = Total_Estimada / sum(Total_Estimada)) 
+
+# Filter for North and Central-West regions
+regions_to_highlight <- df_merged %>%
+  filter(name_region %in% c("Centro Oeste", "Norte"))
+
+# Join the geometries of these regions in a polygon - for better 
+# visualization of the red line we will plot
+regions_union <- regions_to_highlight %>%
+  st_union() %>%  
+  st_as_sf()      
+
+# Final plot  
+ggplot(data = df_merged) +
+  geom_sf(aes(fill = pop_share), color = "white") +  
+  scale_fill_distiller(palette = "Blues", name = "Pop. Share") +  
+  geom_sf(data = regions_union, fill = NA, color = "red", size = 1.5)
+
+# Adjusting the scale of colors 
+ggplot(data = df_merged) +
+  geom_sf(aes(fill = pop_share), color = "white") +
[…]

### 2025-01-24 hw1/paper3.R
fixed path for paper 3

@@ -8,7 +8,7 @@ library(readxl)
 library(tidyr)
 
 # Read the .xls file with the Brasilian population for 2009  
-df <- read_excel("data/paper3/UF_Municipio.xls", skip = 4, col_names = TRUE)
+df <- read_excel("hw1/data/paper3/UF_Municipio.xls", skip = 4, col_names = TRUE)
 df
 
 # Get the geographical data from the package geobr

### 2025-01-25 hw1/paper3.R
updated code

@@ -6,16 +6,13 @@ library(dplyr)
 library(ggplot2)
 library(readxl)
 library(tidyr)
+library(lwgeom)
 
 # Read the .xls file with the Brasilian population for 2009  
 df <- read_excel("hw1/data/paper3/UF_Municipio.xls", skip = 4, col_names = TRUE)
-df
 
 # Get the geographical data from the package geobr
 state <- read_state(year=2010)
-meso <- read_meso_region(year=2010)
-state
-meso
 
 # Convert ESTIMADA (=population) into a numerical variable and handle missing values 
 # by setting them to 0
@@ -29,41 +26,35 @@ df_grouped <- df %>%
   summarise(
     Total_Estimada = sum(ESTIMADA, na.rm = TRUE),  
     Count = n()  
-)
+  )
 
-# Merge the 'meso' data with the population data 
-df_merged <- meso %>%
+# Remove the last row of df_merged - in df the last row was the total
+# and it would have been considered another meso region
+df_grouped <- df_grouped %>% 
+  slice(-n())
+
+# Merge the geographical data with the population data 
+df_merged <- state %>%
   left_join(df_grouped, by = c("code_state" = "COD...2"))
-
-# Merge the resulting df with the 'state' data to get the orientation 
-df_merged <- df_merged %>%
-  st_join(state)
 
 # Create the column pop_share
 df_merged <- df_merged %>%
-mutate(pop_share = Total_Estimada / sum(Total_Estimada)) 
+  mutate(pop_share = Total_Estimada / sum(Total_Estimada)*100) 
 
 # Filter for North and Central-West regions
 regions_to_highlight <- df_merged %>%
   filter(name_region %in% c("Centro Oeste", "Norte"))
 
-# Join the geometries of these regions in a polygon - for better 
-# visualization of the red line we will plot
+# Join the geometries of these regions in a polygon 
 regions_union <- regions_to_highlight %>%
-  st_union() %>%  
-  st_as_sf()      
+  st_union() %>% 
+  st_as_sf()
 
-# Final plot  
-ggplot(data = df_merged) +
-  geom_sf(aes(fill = pop_share), color = "white") +  
-  scale_fill_distiller(palette = "Blues", name = "Pop. Share") +  
-  geom_sf(data = regions_union, fill = NA, color = "red", size = 1.5)
-
-# Adjusting the scale of colors 
+# Final plot 
[…]

### 2025-01-25 hw1/submission_notebook.Rmd
fixed path for paper 3
updating HTML with the third paper

@@
 
+The authors of the paper do not provide a direct link to the data, but they refer to the Brazilian Statistical Bureau (IBGE) when introducing the "meso-regions," which are the political boundaries used in their plots.
+
+We retrieved population data from the aforementioned IBGE website and used geographical data available in the R package geobr.
+While the authors reproduced the same plot for three different years — 1950, 1980, and 2010 — due to a lack of data, we were only able to reproduce the plot for the most recent year.
+Specifically, on the IBGE website, we found data for 2009.
+However, we assume no significant variations in population occurred within a single year, allowing us to proceed with our analysis.
+For consistency, also the geographical data are referred to 2009.
+
+```{r}
+# Read the .xls file with the Brasilian population for 2009  
+df <- read_excel("hw1/data/paper3/UF_Municipio.xls", skip = 4, col_names = TRUE)
+
+# Get the geographical data from the package geobr
+state <- read_state(year=2009)
+meso <- read_meso_region(year=2009)
+```
+
 ### Data wrangling
 
+From the package geobr, data at the meso-regions level are available, but they have a different level of granularity compared to the population data.
+This discrepancy can be easily verified by comparing the number of observations in the two datasets.
+Due to this inconsistency between the geographical data and the population data, we could not exactly replicate the authors' plots, which show the Brazilian population shares for meso-regions.
+Instead, we had to perform the analysis at the state level.
+
+First, we grouped the population data by state to ensure that both datasets referred to the same political boundaries.
+Next, we joined the two datasets based on the state code.
+After this, we created a column indicating the population share, multiplying the ratio by 100 to make the results more consistent with the paper.
+Finally, we filtered the data for the North and Central-West regions to highlight the borders in the plot.
+
+```{r}
+# Convert ESTIMADA (=population) into a numerical variable and # handle missing values by setting them to 0
+df <- df %>%
+  mutate(ESTIMADA = as.numeric(ESTIMADA)) %>%
+  mutate(ESTIMADA = replace_na(ESTIMADA, 0))
+
+# Group by the state code 
+df_grouped <- df %>% 
+  group_by(COD...2) %>%
+  summarise(
+    Total_Estimada = sum(ESTIMADA, na.rm = TRUE),  
+    Count = n()  
+  )
+
+# Remove the last row of df_merged - in df the last row was the # total and it would have been considered another meso region
+df_grouped <- df_grouped %>% 
+  slice(-n())
+
+# Merge the geographical data with the population data 
+df_merged <- state %>%
+  left_join(df_grouped, by = c("code_state" = "COD...2"))
+
+# Create the column pop_share
+df_merged <- df_merged %>%
+  mutate(pop_share = Total_Estimada / sum(Total_Estimada)*100)
+
+# Filter for North and Central-West regions
+regions_to_highlight <- df_merged %>%
+  filter(name_region %in% c("Centro Oeste", "Norte"))
+
+# Join the geometries of these regions in a polygon 
+regions_union <- regions_to_highlight %>%
+  st_union() %>% 
+  st_as_sf()
+```
+
 ### Plotting
+
+```{r}
+ggplot(data = df_merged) +
+  geom_sf(aes(fill = pop_share), color = "white") +
+  scale_fill_gradientn(
+    colors = c("lightblue", "blue", "darkblue"),
+    name = "Pop. Share",
+    limits = c(0, 25)) +
+  geom_sf(data = regions_union, fill = NA, color = "red", linewidth = 1)
+```
 

### 2025-01-25 hw1/paper3.R

@@ -12,7 +12,8 @@ library(lwgeom)
 df <- read_excel("hw1/data/paper3/UF_Municipio.xls", skip = 4, col_names = TRUE)
 
 # Get the geographical data from the package geobr
-state <- read_state(year=2010)
+state <- read_state(year=2009)
+meso <- read_meso_region(year=2009)
 
 # Convert ESTIMADA (=population) into a numerical variable and handle missing values 
 # by setting them to 0

### 2025-01-27 hw1/submission_notebook.Rmd
minor updates

@@
 However, we assume no significant variations in population occurred within a single year, allowing us to proceed with our analysis.
-For consistency, also the geographical data are referred to 2009.
 
@@
 # Get the geographical data from the package geobr
-state <- read_state(year=2009)
-meso <- read_meso_region(year=2009)
+state <- read_state(year=2010)
+meso <- read_meso_region(year=2010)
 ```

### 2025-01-27 hw1/paper3.R

@@ -12,8 +12,8 @@ library(lwgeom)
 df <- read_excel("hw1/data/paper3/UF_Municipio.xls", skip = 4, col_names = TRUE)
 
 # Get the geographical data from the package geobr
-state <- read_state(year=2009)
-meso <- read_meso_region(year=2009)
+state <- read_state(year=2010)
+meso <- read_meso_region(year=2010)
 
 # Convert ESTIMADA (=population) into a numerical variable and handle missing values 
 # by setting them to 0

### 2025-01-27 hw1/submission_notebook.Rmd
adding plot from authors
adding comment about the censobr package

@@
 
+Additionally, we explored the R package censobr, which is expected to contain Brazil's census data for earlier years.
+Unfortunately, despite using the function data_dictionary and experimenting with various arguments (e.g., "population," "households," etc.), we were unable to retrieve population data.
+Instead, the package provided other types of data that were not relevant to our analysis.
+
 ```{r}
@@
 
-In comparison the third paper map looked as follows: ![Figure x]()
+In comparison the third paper map looked as follows: ![Figure 4](data/paper3/original_plot.png)
 

### 2025-01-27 hw1/submission_notebook.Rmd
Added correct 2010 boundaries for paper 3

--- 2025-01-27 hw1/submission_notebook.Rmd
@@
 library(terra)
+library(lwgeom)
+library(geobr)
 ```
@@
 
-## Paper 3
+## [Paper 3](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3776990)
 
@@
 # Read the .xls file with the Brasilian population for 2009
-df <- read_excel("hw1/data/paper3/UF_Municipio.xls", skip = 4, col_names = TRUE)
+df <- read_excel("data/paper3/UF_Municipio.xls", skip = 4, col_names = TRUE)
 
 # Get the geographical data from the package geobr
-state <- read_state(year=2010)
-meso <- read_meso_region(year=2010)
+state <- read_state(year = 2010)
+meso <- read_meso_region(year = 2010)
+region <- read_region(year = 2010)
 ```
@@
 df_merged <- df_merged %>%
-  mutate(pop_share = Total_Estimada / sum(Total_Estimada)*100)
+  mutate(pop_share = Total_Estimada / sum(Total_Estimada) * 100)
 
-# Filter for North and Central-West regions
-regions_to_highlight <- df_merged %>%
-  filter(name_region %in% c("Centro Oeste", "Norte"))
-
-# Join the geometries of these regions in a polygon 
-regions_union <- regions_to_highlight %>%
-  st_union() %>% 
-  st_as_sf()
+# Aggregate regions as the author's did to get the correct boundaries
+region <- region %>%
+  mutate(
+    name_region = case_when(
+      name_region %in% c("Centro Oeste", "Norte") ~ 2,
+      TRUE ~ 1
+    )
+  ) %>%
+  dplyr::select(name_region, geom) %>%
+  group_by(name_region) %>%
+  summarise(geom = st_union(geom)) %>%
+  sfheaders::sf_remove_holes()
 ```
@@
 ```{r}
+# Final plot
 ggplot(data = df_merged) +
+  ggtitle("(c) 2010") +
+  theme(plot.title = element_text(hjust = 0.5)) +
   geom_sf(aes(fill = pop_share), color = "white") +
@@
     name = "Pop. Share",
-    limits = c(0, 25)) +
-  geom_sf(data = regions_union, fill = NA, color = "red", linewidth = 1)
+    limits = c(0, 25)
+  ) +
+  geom_sf(data = region, aes(color = factor(name_region)), fill = NA, linewidth = 1) +
+  scale_color_manual(values = c("black", "red")) +
+  guides(color = "none")
 ```
 
-In comparison the third paper map looked as follows: ![Figure x]()
+In comparison the third paper map looked as follows: ![Figure 4](data/paper3/original_plot.png)
+
+Again, as outlined above, we only replicate the 2010 map, due to data non-availability in the prior periods.
 

### 2025-01-27 hw1/paper3.R

@@ -1,5 +1,3 @@
-# Clean working directory and import necessary packages 
-rm(list = ls())
 library(sf)
 library(geobr)
 library(dplyr)
@@ -8,54 +6,62 @@ library(readxl)
 library(tidyr)
 library(lwgeom)
 
-# Read the .xls file with the Brasilian population for 2009  
+# Read the .xls file with the Brasilian population for 2009
 df <- read_excel("hw1/data/paper3/UF_Municipio.xls", skip = 4, col_names = TRUE)
 
 # Get the geographical data from the package geobr
-state <- read_state(year=2010)
-meso <- read_meso_region(year=2010)
+state <- read_state(year = 2010)
+meso <- read_meso_region(year = 2010)
+region <- read_region(year = 2010)
 
-# Convert ESTIMADA (=population) into a numerical variable and handle missing values 
+# Convert ESTIMADA (=population) into a numerical variable and handle missing values
 # by setting them to 0
 df <- df %>%
   mutate(ESTIMADA = as.numeric(ESTIMADA)) %>%
   mutate(ESTIMADA = replace_na(ESTIMADA, 0))
 
-# Group by the country code 
-df_grouped <- df %>% 
+# Group by the country code
+df_grouped <- df %>%
   group_by(COD...2) %>%
   summarise(
-    Total_Estimada = sum(ESTIMADA, na.rm = TRUE),  
-    Count = n()  
+    Total_Estimada = sum(ESTIMADA, na.rm = TRUE),
+    Count = n()
   )
 
 # Remove the last row of df_merged - in df the last row was the total
 # and it would have been considered another meso region
-df_grouped <- df_grouped %>% 
+df_grouped <- df_grouped %>%
   slice(-n())
 
-# Merge the geographical data with the population data 
+# Merge the geographical data with the population data
 df_merged <- state %>%
   left_join(df_grouped, by = c("code_state" = "COD...2"))
 
 # Create the column pop_share
 df_merged <- df_merged %>%
-  mutate(pop_share = Total_Estimada / sum(Total_Estimada)*100) 
+  mutate(pop_share = Total_Estimada / sum(Total_Estimada) * 100)
 
-# Filter for North and Central-West regions
-regions_to_highlight <- df_merged %>%
-  filter(name_region %in% c("Centro Oeste", "Norte"))
+# Aggregate regions as the author's did to get the correct boundaries
[…]

### 2025-01-27 hw1/submission_notebook.Rmd
Knitted HTML document

@@
 ---
-title: "Geospatial - Assignment 1 - Moritz Peist, Noemi Lucchi, Simon Vellin"
+title: "Geospatial - Assignment 1"
+author: 
+- Moritz Peist
+- Noemi Lucchi
+- Simon Vellin
 output:
@@
 
-In the following notebook we replicate the charts for all 5 papers.
-Contentwise, the code follows always the same structure.
-We first load the data, then we create the chart and finally we save the chart as a .png file.
+In the following notebook we replicate the charts for all 5 papers. Contentwise, the code follows always the same structure. We first load the data, then we create the chart and finally display it.
 
@@
 
-# RSA boundaries - downloaded from GDAM and temporarily stored
+# RSA boundaries - downloaded from GDAM
 sa_districts_sf <- st_as_sf(gadm(country = "ZAF", level = 3))
@@
 
-In comparison the second paper map looked as follows: ![Figure 4](data/paper2/1-s2.0-S0166046220303100-gr4.jpg)
+In comparison the second paper map looked as follows: 
+![Figure 4](data/paper2/1-s2.0-S0166046220303100-gr4.jpg)
 
@@
 
-In comparison the third paper map looked as follows: ![Figure 4](data/paper3/original_plot.png)
+In comparison the third paper map looked as follows:
+
+![Figure 4](data/paper3/original_plot.png)
 
@@
 
-In comparison the fourth paper map looked as follows: ![Figure x]()
+In comparison the fourth paper map looked as follows:
+
+![Figure x]()
 
@@
 
-In comparison the fifth paper map looked as follows: ![Figure x]()
+In comparison the fifth paper map looked as follows:
+
+![Figure x]()

### 2025-01-27 hw1/paper4.R
Initial commit paper 4+5

@@ -1,0 +1,113 @@
+## [Paper 4](https://economics.mit.edu/sites/default/files/publications/Catastrophe_Risk_and_Settlement_Location.pdf)
+
+library(osmdata)
+library(sf)
+library(spData)
+library(tidyverse)
+library(geodata)
+library(readxl)
+library(ggplot2)
+library(raster)
+library(terra)
+library(rmapshaper)
+library(ggplot2)
+library(ggrepel)
+
+path <- ("/Users/simonvellin/Documents/BSE/TERM 2/GEOSPATIAL/0. ASSIGNMENTS/HW1/PAPER 4")
+outpath <- ("/Users/simonvellin/Documents/BSE/TERM 2/GEOSPATIAL/0. ASSIGNMENTS/HW1/PAPER 4")
+
+----------------------------------------
+            # DATA MINING
+----------------------------------------
+
+# We are looking for a shapefile or dataset specifically containing road network data for Vietnam.
+
+# We inspect data from GADM:
+st_layers("/Users/simonvellin/Documents/BSE/TERM 2/GEOSPATIAL/0. ASSIGNMENTS/HW1/PAPER 4/GADM-Vietnam")
+
+# None of the layers in the file appear to contain road network data. All layers are polygons representing administrative boundaries. 
+
+# We inspect data from Humdata:
+st_layers("/Users/simonvellin/Documents/BSE/TERM 2/GEOSPATIAL/0. ASSIGNMENTS/HW1/PAPER 4/ROADS/vnm_rdsl_2015_OSM.shp")
+vietnam_roads <- st_read("/Users/simonvellin/Documents/BSE/TERM 2/GEOSPATIAL/0. ASSIGNMENTS/HW1/PAPER 4/ROADS/vnm_rdsl_2015_OSM.shp")
+view(vietnam_roads)
+
+# Get all the unique values from the column 'type'
+unique_values <- unique(vietnam_roads$type)
+print(unique_values)
+
+# Although the data is from 2015, whereas the paper displays roads from 2010, we decide to move forward with this data as road network.
+
+# We inspect data from SimpleMaps:
+st_layers("/Users/simonvellin/Documents/BSE/TERM 2/GEOSPATIAL/0. ASSIGNMENTS/HW1/PAPER 4/VIETNAM")
+
+# We decide to move forward with this data as Vietnam boundary.
+----------------------------------------
+            # DATA LOADING
+----------------------------------------
+
+# Vietnam's road network shapefile already loaded as vietnam_roads following inspection
+
[…]

### 2025-01-27 hw1/paper5.R

@@ -1,0 +1,89 @@
+## [Paper 5](https://web.stanford.edu/~memorten/ewExternalFiles/Morten_Oliveira_Brasilia.pdf)
+
+library(osmdata)
+library(sf)
+library(spData)
+library(tidyverse)
+library(geodata)
+library(readxl)
+library(ggplot2)
+library(raster)
+library(terra)
+library(rmapshaper)
+library(ggplot2)
+library(ggrepel)
+
+path <- ("hw1/data/paper5/")
+outpath <- ("/Users/simonvellin/Documents/BSE/TERM 2/GEOSPATIAL/0. ASSIGNMENTS/HW1/PAPER 5/OUTPUT")
+
+
+### Data mining
+
+# The author's replication package is accessible, so we move forward with it's code/data (https://www.openicpsr.org/openicpsr/project/183316/version/V1/view).
+
+### Data loading
+
+# Load the shapefile containing state boundaries for Brazil (1940 projection)
+states <- st_read(file.path(path, "GIS_data/uf1940/uf1940_prj.shp"))
+# Simplify the state boundaries to reduce file size while maintaining shape integrity
+states_simple  <- rmapshaper::ms_simplify(states,keep=0.01,keep_shapes=TRUE)
+
+### Data wrangling
+
+# Define the year for the highways data
+year  <- 2000
+
+# Construct the file path for the highways shapefile dynamically based on the year
+file_name = paste0("/",year,"/highways_",year,"_prj.shp")
+
+# Load the highways shapefile for the specified year
+all_highways <- st_read(file.path(path, "GIS_data/roads", file_name))
+# Simplify the highways data for efficient plotting
+all_highways_simple <- rmapshaper::ms_simplify(all_highways,keep=0.01,keep_shapes=TRUE)
+
+# Load and simplify the minimum spanning tree (MST) shapefile for pie distribution
+mst_pie <- st_read(file.path(path, "GIS_data/mst/mst_pie_prj.shp"))
+mst_pie_simple  <- rmapshaper::ms_simplify(mst_pie,keep=0.01,keep_shapes=TRUE)
+
+# Load and simplify MST data excluding pies for Rio region
+mst_all_rio <- st_read(file.path(path, "GIS_data/mst/mst_all_rio_prj.shp"))
+mst_all_rio_simple  <- rmapshaper::ms_simplify(mst_all_rio,keep=0.01,keep_shapes=TRUE)
+
+# Load shapefile for capital cities and extract coordinates for labeling
[…]

### 2025-01-27 hw1/submission_notebook.Rmd
Paper 3, markdown opticals - Refined paper 3 - Text and optical changes to .rmd

@@
 
+<hr>
+
+```{css, echo=FALSE}
+h1.title {
+  text-align: center;
+}
+
+h4.author, h4.date {
+  text-align: center;
+}
+
+.author {
+  text-align: center;
+}
+
+.date {
+  text-align: center;
+}
+```
+
 ```{r setup, include=FALSE}
@@
 ```
+
+<div style="text-align: center;">
+  <img src="Barcelona_School_of_Economics_logo.svg" alt="BSE logo">
+</div>
+
+<hr>
 
@@
 ```{r}
-library(osmdata)
+# Main tidyverse (includes ggplot2)
+library(tidyverse)
+
+# Spatial packages
 library(sf)
-library(spData)
-library(tidyverse)
-library(geodata)
-library(readxl)
-library(ggplot2)
-library(raster)
-library(terra)
-library(lwgeom)
-library(geobr)
+library(terra) # Newer replacement for raster
+library(osmdata) # OpenStreetMap data
+library(spData) # Spatial data
+library(geobr) # Brazilian boundaries
+library(censobr) # Brazilian population data
+library(lwgeom) # Needed for st_make_valid
+
+# Data import
+library(geodata) # For elevation data
+library(readxl) # For Excel files
 ```
 
-## [Paper 1](https://www.sciencedirect.com/science/article/pii/S0304387818300166#appsec1)
+<hr>
 
-### Data loading
+# [Paper 1](https://www.sciencedirect.com/science/article/pii/S0304387818300166#appsec1)
+
+## Data loading
 
 The authors of the paper provide a link to the data, however, the link is unfortunately outdated.
-Yet, the data is available in a newer version on the website (authors used 2013 data, we use 2024 data).
+Yet, the data is available in a newer version on the [website](https://www.dws.gov.za/DSO/publications.aspx) (authors used 2013 data, we use 2024 data) but had to be converted to geojson format before usage.
 
@@
 
-### Data wrangling
+## Data wrangling
 
@@
 
-### Plotting
+## Plotting
 
@@
 
-## [Paper 2](https://www.sciencedirect.com/science/article/pii/S0166046220303100)
+<hr>
 
-### Data loading
+# [Paper 2](https://www.sciencedirect.com/science/article/pii/S0166046220303100)
+
+## Data loading
 
@@
 Therefore, we leverage some r packages to get the data for highways, powergrids (osmdata) and population data from 2020 (geodata).
-The ESSR survey data was directly available via the World Bank data repositories as a CSV (2011/12) data used as in the paper.
+The [ERSS survey](https://microdata.worldbank.org/index.php/catalog/2053/data-dictionary) data was directly available via the World Bank data repositories as a CSV (2011/12) data used as in the paper.
 We load all of these in the following:
@@
 
-### Data wrangling
+## Data wrangling
 
 We calculate the population density for each district and clip the roads to the Ethiopia boundary.
+Furthermore, t looks like the authors used slightly different road and powergrid networks.
+However, overall the grids look broadly similar when utilizing OpenStreetMap data.
 
@@
 
-## [Paper 3](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3776990)
+<hr>
 
-### Data loading
+# [Paper 3](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3776990)
+
+## Data loading
 
@@
 We retrieved population data from the aforementioned IBGE website and used geographical data available in the R package geobr.
-While the authors reproduced the same plot for three different years — 1950, 1980, and 2010 — due to a lack of data, we were only able to reproduce the plot for the most recent year.
-Specifically, on the IBGE website, we found data for 2009.
-However, we assume no significant variations in population occurred within a single year, allowing us to proceed with our analysis.
+While the authors reproduced the same plot for three different years — 1950, 1980, and 2010 — due to a lack of data, we were only able to reproduce the plot for the most recent year, i.e., 2010.
+Specifically, on the IBGE website, we found data for 2009, and the censobr package provides the 2010 data.
+It also provides census data for 1980 and 1960, but in 1980, it seems the population weights are missing.
+Therefore, we decided to focus on the 2010 data.
 
-Additionally, we explored the R package censobr, which is expected to contain Brazil's census data for earlier years.
-Unfortunately, despite using the function data_dictionary and experimenting with various arguments (e.g., "population," "households," etc.), we were unable to retrieve population data.
-Instead, the package provided other types of data that were not relevant to our analysis.
+For geographical boundaries of Brazil for administrative areas, we use the geobr package.
 
 ```{r}
-# Read the .xls file with the Brasilian population for 2009
-df <- read_excel("data/paper3/UF_Municipio.xls", skip = 4, col_names = TRUE)
+# Read in the Brasilian population for 2010 using the censobr package
+# This also contains 1980, and 1960 data, but 1980 without weights, and 1960 only partially
+pop_2010 <- censobr::read_population(
+  year = 2010,
+  columns = c("code_state", "code_region", "code_muni", "V0010"),
+  showProgress = FALSE,
+  as_data_frame = TRUE,
+  # cache = FALSE
+) %>% # Group by municipality code
+  group_by(code_muni) %>%
+  summarise(
+    total_population = sum(V0010, rm.na = TRUE)
+  )
 
-# Get the geographical data from the package geobr
-state <- read_state(year = 2010)
-meso <- read_meso_region(year = 2010)
+# Get the geographical data from the package geobr and validate if necessary
+meso <- read_meso_region(year = 2010) %>%
+  st_make_valid()
 region <- read_region(year = 2010)
+muni <- read_municipality(year = 2010) %>%
+  st_make_valid()
 ```
 
-### Data wrangling
-
-From the package geobr, data at the meso-regions level are available, but they have a different level of granularity compared to the population data.
-This discrepancy can be easily verified by comparing the number of observations in the two datasets.
-Due to this inconsistency between the geographical data and the population data, we could not exactly replicate the authors' plots, which show the Brazilian population shares for meso-regions.
-Instead, we had to perform the analysis at the state level.
+## Data wrangling
 
 First, we grouped the population data by state to ensure that both datasets referred to the same political boundaries.
-Next, we joined the two datasets based on the state code.
+Next, we joined the two datasets based on the meso code.
+We had to take a rather coarse-granular approach by using the `st_centroid()` function to match meso-regions with municipalities.
+this leads us to loosing out roughly on 50'000 inhabitants.
 After this, we created a column indicating the population share, multiplying the ratio by 100 to make the results more consistent with the paper.
 Finally, we filtered the data for the North and Central-West regions to highlight the borders in the plot.
+Needing `sf_remove_holes()` to avoid inner borders, we loose a bit of granularity in the outer borders of the broader regions.
 
 ```{r}
-# Convert ESTIMADA (=population) into a numerical variable and # handle missing values by setting them to 0
-df <- df %>%
-  mutate(ESTIMADA = as.numeric(ESTIMADA)) %>%
-  mutate(ESTIMADA = replace_na(ESTIMADA, 0))
+# First, join the population data with municipal boundaries
+pop_spatial <- muni %>%
+  left_join(pop_2010, by = "code_muni")
 
-# Group by the state code
-df_grouped <- df %>%
-  group_by(COD...2) %>%
+# Use st_centroid to ensure each municipality is counted only once
+pop_meso <- pop_spatial %>%
+  st_centroid() %>% # Convert polygons to points
+  st_join(meso) %>% # Join with meso regions
+  group_by(code_meso) %>%
   summarise(
-    Total_Estimada = sum(ESTIMADA, na.rm = TRUE),
-    Count = n()
-  )
+    total_population = sum(total_population, na.rm = TRUE)
+  ) %>%
+  st_drop_geometry() # Remove geometry before joining back
 
-# Remove the last row of df_merged - in df the last row was the # total and it would have been considered another meso region
-df_grouped <- df_grouped %>%
-  slice(-n())
+# Now join back with meso geometries
+pop_meso <- meso %>%
+  left_join(pop_meso, by = "code_meso")
 
-# Merge the geographical data with the population data
-df_merged <- state %>%
-  left_join(df_grouped, by = c("code_state" = "COD...2"))
-
-# Create the column pop_share
-df_merged <- df_merged %>%
-  mutate(pop_share = Total_Estimada / sum(Total_Estimada) * 100)
+# Create pop_share
+df_merged <- pop_meso %>%
+  mutate(pop_share = total_population / sum(total_population) * 100)
 
@@
 
-### Plotting
+## Plotting
 
 ```{r}
-# Final plot
 ggplot(data = df_merged) +
   ggtitle("(c) 2010") +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  geom_sf(aes(fill = pop_share), color = "white") +
-  scale_fill_gradientn(
-    colors = c("lightblue", "blue", "darkblue"),
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5), # Center title while keeping minimal theme
+    panel.grid = element_line(color = "gray90"), # Make grid lines lighter
+    axis.text = element_blank(),
+    axis.ticks = element_blank()
+  ) +
+  geom_sf(aes(fill = pop_share), color = "white", size = 0.1) +
+  scale_fill_stepsn(
     name = "Pop. Share",
-    limits = c(0, 25)
+    breaks = c(0, 0.216, 0.382, 0.557, 1.02, 11.12),
+    labels = scales::number_format(accuracy = 0.001),
+    colors = c("#eff3ff", "#bdd7e7", "#6baed6", "#3182bd", "#08519c"), # Made lower values darker
+    values = scales::rescale(c(0, 0.216, 0.382, 0.557, 1.02, 11.12))
   ) +
@@
   scale_color_manual(values = c("black", "red")) +
-  guides(color = "none")
+  guides(color = "none") +
+  # Add titles
+  labs(
+    caption = "Source: Based on Pellegrina and Sotelo (2021)"
+  )
 ```
@@
 
-Again, as outlined above, we only replicate the 2010 map, due to data non-availability in the prior periods.
+Again, as outlined above, we only replicate the 2010 map, due to data issues in the prior periods.
 
-## Paper 4
+<hr>
 
-### Data loading
+# Paper 4
 
-### Data wrangling
+## Data loading
 
-### Plotting
+## Data wrangling
+
+## Plotting
 
@@
 
-## Paper 5
+<hr>
 
-### Data loading
+# Paper 5
 
-### Data wrangling
+## Data loading
 
-### Plotting
+## Data wrangling
+
+## Plotting
 

### 2025-01-27 test.Rmd
Paper 3, markdown opticals - Refined paper 3 - Text and optical changes to .rmd

---
title: "Geospatial"
author: 
  - Moritz Peist
  - Noemi Lucchi
  - Simon Vellin
date: "2025-01-27"
output: html_document
---

<hr>

```{css, echo=FALSE}
h1.title {
  text-align: center;
}

h4.author, h4.date {
  text-align: center;
}

.author {
  text-align: center;
}

.date {
  text-align: center;
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<div style="text-align: center;">
  <img src="hw1/Barcelona_School_of_Economics_logo.svg" alt="BSE logo">
</div>

<hr>

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2025-01-27 hw1/paper3.R

@@ -1,45 +1,54 @@
 library(sf)
 library(geobr)
-library(dplyr)
+library(tidyverse)
 library(ggplot2)
 library(readxl)
-library(tidyr)
 library(lwgeom)
+library(censobr)
 
-# Read the .xls file with the Brasilian population for 2009
-df <- read_excel("hw1/data/paper3/UF_Municipio.xls", skip = 4, col_names = TRUE)
-
-# Get the geographical data from the package geobr
-state <- read_state(year = 2010)
-meso <- read_meso_region(year = 2010)
-region <- read_region(year = 2010)
-
-# Convert ESTIMADA (=population) into a numerical variable and handle missing values
-# by setting them to 0
-df <- df %>%
-  mutate(ESTIMADA = as.numeric(ESTIMADA)) %>%
-  mutate(ESTIMADA = replace_na(ESTIMADA, 0))
-
-# Group by the country code
-df_grouped <- df %>%
-  group_by(COD...2) %>%
+# Read in the Brasilian population for 2010 using the censobr package
+# This also contains 1980, and 1960 data, but 1980 without weights, and 1960 only partially
+pop_2010 <- read_population(
+  year = 2010,
+  columns = c("code_state", "code_region", "code_muni", "V0010"),
+  showProgress = FALSE,
+  as_data_frame = TRUE,
+  # cache = FALSE
+) %>% # Group by municipality code
+  group_by(code_muni) %>%
   summarise(
-    Total_Estimada = sum(ESTIMADA, na.rm = TRUE),
-    Count = n()
+    total_population = sum(V0010, rm.na = TRUE)
   )
 
-# Remove the last row of df_merged - in df the last row was the total
-# and it would have been considered another meso region
-df_grouped <- df_grouped %>%
-  slice(-n())
+# Get the geographical data from the package geobr and validate if necessary
+meso <- read_meso_region(year = 2010) %>%
+  st_make_valid()
+region <- read_region(year = 2010)
+muni <- read_municipality(year = 2010) %>%
+  st_make_valid()
 
-# Merge the geographical data with the population data
-df_merged <- state %>%
-  left_join(df_grouped, by = c("code_state" = "COD...2"))
+# First, join the population data with municipal boundaries
+pop_spatial <- muni %>%
+  left_join(pop_2010, by = "code_muni")
 
-# Create the column pop_share
-df_merged <- df_merged %>%
[…]

### 2025-01-27 hw1/test3.R

@@ -1,0 +1,41 @@
+library(censobr)
+library(tidyverse)
+
+
+# return data as arrow Dataset
+pop_2010 <- read_population(
+  year = 2010,
+  # columns = c("code_state", "code_region", "code_muni", "V0010"),
+  showProgress = FALSE,
+  as_data_frame = TRUE,
+  # cache = FALSE
+)
+
+pop_1980 <- read_population(
+  year = 1980,
+  # columns = c("code_meso", "code_region", "code_state", "code_muni", "V198"),
+  showProgress = FALSE,
+  as_data_frame = TRUE,
+  cache = FALSE
+)
+
+pop_1960 <- read_population(
+  year = 1960,
+  # columns = c("code_state", "code_region", "code_muni", "V5060", "V0010"),
+  showProgress = FALSE,
+  as_data_frame = TRUE
+)
+
+df %>%
+  summarise(total_population = sum(V0010, rm.na = TRUE))
+
+pop_1980 %>%
+  mutate(V198 = as.numeric(V198)) %>%
+  summarise(total_population = sum(V198, rm.na = TRUE))
+# group_by(code_region) %>%
+# summarise(total_population = sum(V0010, rm.na = TRUE))
+
+
+pop_1980 %>%
+  select(V601, V602, V603, V604, V605) %>%
+  head()

### 2025-01-28 hw1/submission_notebook.Rmd
Added 1960 for paper 3 and refined markdown

@@
 We retrieved population data from the aforementioned IBGE website and used geographical data available in the R package geobr.
-While the authors reproduced the same plot for three different years — 1950, 1980, and 2010 — due to a lack of data, we were only able to reproduce the plot for the most recent year, i.e., 2010.
+While the authors produced the same plot for three different years — 1950, 1980, and 2010 — due to a lack of data, we were only able to reproduce the plot for the most recent year, i.e., 2010 and with flaws for 1960 (i.e., we're missing around 13 mln people of the overall population in 1960 due to data compilation issues of the earlier census approaches of the IBGE - [see here](https://ipeagit.github.io/censobr/reference/read_population.html#-census) for more information).
 Specifically, on the IBGE website, we found data for 2009, and the censobr package provides the 2010 data.
 It also provides census data for 1980 and 1960, but in 1980, it seems the population weights are missing.
-Therefore, we decided to focus on the 2010 data.
+Therefore, we decided to focus on the 2010 and 1960 data.
 
@@
 # This also contains 1980, and 1960 data, but 1980 without weights, and 1960 only partially
-pop_2010 <- censobr::read_population(
+pop_2010 <- read_population(
   year = 2010,
@@
   summarise(
-    total_population = sum(V0010, rm.na = TRUE)
+    total_population_2010 = sum(V0010, rm.na = TRUE)
+  )
+
+# Get 1960 data (which is missing approx 13 mln people though -
+# it's only a rough estimate and combined from 2 datasets, see censobr description)
+pop_1960 <- read_population(
+  year = 1960,
+  columns = c("code_muni_1960", "censobr_weight"),
+  showProgress = FALSE,
+  as_data_frame = TRUE,
+  # cache = FALSE
+) %>% # Group by municipality code
+  group_by(code_muni_1960) %>%
+  summarise(
+    total_population_1960 = sum(censobr_weight, rm.na = TRUE)
   )
@@
 
-First, we grouped the population data by state to ensure that both datasets referred to the same political boundaries.
-Next, we joined the two datasets based on the meso code.
+First, we join the 1960 and 2010 population data with munipalicity level data  to ensure that both datasets refer to the same political boundaries.
+Next, we joined the two datasets based with the meso code and group on meso level to get the wanted granularity.
 We had to take a rather coarse-granular approach by using the `st_centroid()` function to match meso-regions with municipalities.
-this leads us to loosing out roughly on 50'000 inhabitants.
+This leads us to loosing out roughly on 50'000 inhabitants in 2010.
 After this, we created a column indicating the population share, multiplying the ratio by 100 to make the results more consistent with the paper.
@@
 pop_spatial <- muni %>%
-  left_join(pop_2010, by = "code_muni")
+  left_join(pop_2010, by = "code_muni") %>%
+  left_join(pop_1960, by = join_by("code_muni" == "code_muni_1960"))
 
@@
   summarise(
-    total_population = sum(total_population, na.rm = TRUE)
+    total_population_2010 = sum(total_population_2010, na.rm = TRUE),
+    total_population_1960 = sum(total_population_1960, na.rm = TRUE)
   ) %>%
@@
 df_merged <- pop_meso %>%
-  mutate(pop_share = total_population / sum(total_population) * 100)
+  mutate(
+    pop_share_2010 = total_population_2010 / sum(total_population_2010) * 100,
+    pop_share_1960 = total_population_1960 / sum(total_population_1960) * 100,
+  )
 
@@
 
-```{r}
+We reproduce the same chart for 2010 as the authors did.
+Due to data issues, we rely on a further chart for 1960, instead of 1950.
+The 1960 chart resembles the 1950 chart of the authors surpisingly good, though the aforementioned data issues.
+Lastly, we omit 1980 data, since the population weights were missing in the censobr dataset.
+
+```{r, figures-side, fig.show="hold", out.width="50%"}
+# Final plot - 1960
+ggplot(data = df_merged) +
+  ggtitle("(c) 1960") +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5), # Center title while keeping minimal theme
+    panel.grid = element_line(color = "gray90"), # Make grid lines lighter
+    axis.text = element_blank(),
+    axis.ticks = element_blank()
+  ) +
+  geom_sf(aes(fill = pop_share_1960), color = "white", size = 0.1) +
+  scale_fill_stepsn(
+    name = "Pop. Share",
+    breaks = c(0, 0.216, 0.382, 0.557, 1.02, 11.12),
+    labels = scales::number_format(accuracy = 0.001),
+    colors = c("#eff3ff", "#bdd7e7", "#6baed6", "#3182bd", "#08519c"), # Made lower values darker
+    values = scales::rescale(c(0, 0.216, 0.382, 0.557, 1.02, 11.12))
+  ) +
+  geom_sf(data = region, aes(color = factor(name_region)), fill = NA, linewidth = 1) +
+  scale_color_manual(values = c("black", "red")) +
+  guides(color = "none") +
+  # Add titles
+  labs(
+    caption = "Source: Based on Pellegrina and Sotelo (2021)"
+  )
+
+# Final plot - 2010
 ggplot(data = df_merged) +
@@
   ) +
-  geom_sf(aes(fill = pop_share), color = "white", size = 0.1) +
+  geom_sf(aes(fill = pop_share_2010), color = "white", size = 0.1) +
   scale_fill_stepsn(
@@
 
-In comparison the third paper map looked as follows:
+In comparison the third paper map looked as follows (left 1950, right 2010):
 
-![Figure 4](data/paper3/original_plot.png)
+<table>
+<tr>
+<td><img src="data/paper3/c1960.jpg" width="400"/></td>
+<td><img src="data/paper3/c2010.jpg" width="400"/></td>
+</tr>
+</table>
 

### 2025-01-28 test.Rmd
Added 1960 for paper 3 and refined markdown

@@
 
-<hr>
-
-```{css, echo=FALSE}
-h1.title {
-  text-align: center;
-}
-
-h4.author, h4.date {
-  text-align: center;
-}
-
-.author {
-  text-align: center;
-}
-
-.date {
-  text-align: center;
-}
-```
-
 ```{r setup, include=FALSE}
@@
 
-<div style="text-align: center;">
-  <img src="hw1/Barcelona_School_of_Economics_logo.svg" alt="BSE logo">
-</div>
-
-<hr>
-
-## R Markdown
-
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
-
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
-
-```{r cars}
-summary(cars)
-```
-
-## Including Plots
-
-You can also embed plots, for example:
-
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+<table>
+<tr>
+<td><img src="hw1/data/paper3/c1960.jpg" width="400"/></td>
+<td><img src="hw1/data/paper3/c2010.png" width="400"/></td>
+</tr>
+</table>

### 2025-01-28 hw1/paper3.R

@@ -17,7 +17,21 @@ pop_2010 <- read_population(
 ) %>% # Group by municipality code
   group_by(code_muni) %>%
   summarise(
-    total_population = sum(V0010, rm.na = TRUE)
+    total_population_2010 = sum(V0010, rm.na = TRUE)
+  )
+
+# Get 1960 data (which is missing approx 13 mln people though -
+# it's only a rough estimate and combined from 2 datasets, see censobr description)
+pop_1960 <- read_population(
+  year = 1960,
+  columns = c("code_muni_1960", "censobr_weight"),
+  showProgress = FALSE,
+  as_data_frame = TRUE,
+  # cache = FALSE
+) %>% # Group by municipality code
+  group_by(code_muni_1960) %>%
+  summarise(
+    total_population_1960 = sum(censobr_weight, rm.na = TRUE)
   )
 
 # Get the geographical data from the package geobr and validate if necessary
@@ -29,7 +43,8 @@ muni <- read_municipality(year = 2010) %>%
 
 # First, join the population data with municipal boundaries
 pop_spatial <- muni %>%
-  left_join(pop_2010, by = "code_muni")
+  left_join(pop_2010, by = "code_muni") %>%
+  left_join(pop_1960, by = join_by("code_muni" == "code_muni_1960"))
 
 # Use st_centroid to ensure each municipality is counted only once
 pop_meso <- pop_spatial %>%
@@ -37,9 +52,11 @@ pop_meso <- pop_spatial %>%
   st_join(meso) %>% # Join with meso regions
   group_by(code_meso) %>%
   summarise(
-    total_population = sum(total_population, na.rm = TRUE)
+    total_population_2010 = sum(total_population_2010, na.rm = TRUE),
+    total_population_1960 = sum(total_population_1960, na.rm = TRUE)
   ) %>%
   st_drop_geometry() # Remove geometry before joining back
+
 
 # Now join back with meso geometries
 pop_meso <- meso %>%
@@ -47,8 +64,10 @@ pop_meso <- meso %>%
 
 # Create pop_share
 df_merged <- pop_meso %>%
-  mutate(pop_share = total_population / sum(total_population) * 100)
-
+  mutate(
+    pop_share_2010 = total_population_2010 / sum(total_population_2010) * 100,
+    pop_share_1960 = total_population_1960 / sum(total_population_1960) * 100,
[…]

### 2025-01-28 hw1/test3.R

@@ -21,9 +21,10 @@ pop_1980 <- read_population(
 
 pop_1960 <- read_population(
   year = 1960,
-  # columns = c("code_state", "code_region", "code_muni", "V5060", "V0010"),
+  columns = c("code_muni_1960", "censobr_weight"),
   showProgress = FALSE,
-  as_data_frame = TRUE
+  as_data_frame = TRUE,
+  # cache = FALSE
 )
 
 df %>%

### 2025-01-28 hw1/submission_notebook.Rmd
Beauty fixes
Corrected typos im markdown

@@
 The authors of the paper provide a link to the data, however, the link is unfortunately outdated.
-Yet, the data is available in a newer version on the [website](https://www.dws.gov.za/DSO/publications.aspx) (authors used 2013 data, we use 2024 data) but had to be converted to geojson format before usage.
+Yet, the data is available in a newer version on [this website](https://www.dws.gov.za/DSO/publications.aspx).
+The authors used 2013 data, we use 2024 data but had to convert the downloaded data to geojson format before usage.
 
-Moreover, the river gradient data was not directly available, so we use the elevation data from the geodata package to retrieve and later calculate gradients.
+Moreover, the river gradient data was not directly available, so we use the elevation data from the geodata package to retrieve and later calculate river gradients.
 
@@
 Unfortunately, again the data was not directly available.
-Therefore, we leverage some r packages to get the data for highways, powergrids (osmdata) and population data from 2020 (geodata).
-The [ERSS survey](https://microdata.worldbank.org/index.php/catalog/2053/data-dictionary) data was directly available via the World Bank data repositories as a CSV (2011/12) data used as in the paper.
+Therefore, we leverage some R packages, i.e., to get the data for highways, powergrids we used the osmdata package and for population data from 2020 we used the geodata package.
+The [ERSS survey](https://microdata.worldbank.org/index.php/catalog/2053/data-dictionary) data was directly available via the World Bank data repositories as a CSV (2011/12) and is identical to the data used in the paper.
 We load all of these in the following:
@@
 We calculate the population density for each district and clip the roads to the Ethiopia boundary.
-Furthermore, t looks like the authors used slightly different road and powergrid networks.
-However, overall the grids look broadly similar when utilizing OpenStreetMap data.
+Furthermore, it looks like the authors used slightly different road and powergrid networks.
+However, overall the grids look broadly similar when utilizing OpenStreetMap data via the osmdata package.
 
@@
 ggplot() +
+  # Administrative boundaries
   geom_sf(
@@
   ) +
+  # Population density
   scale_fill_gradientn(
@@
   ) +
+  # Adding roads
   geom_sf(
@@
   ) +
+  # Adding HV powerlines
   geom_sf(
@@
   ) +
+  # Adding survey villages
   geom_sf(
@@
   ) +
+  # Adding legends
   scale_color_manual(
@@
   ) +
+  # Styling
   guides(
@@
 
-The authors of the paper do not provide a direct link to the data, but they refer to the Brazilian Statistical Bureau (IBGE) when introducing the "meso-regions," which are the political boundaries used in their plots.
+The authors of the paper do not provide a direct link to the data, but they refer to the Brazilian Statistical Bureau (IBGE) when talkig about population data and introducing the "meso-regions", which are the political boundaries used in their plots.
 
-We retrieved population data from the aforementioned IBGE website and used geographical data available in the R package geobr.
-While the authors produced the same plot for three different years — 1950, 1980, and 2010 — due to a lack of data, we were only able to reproduce the plot for the most recent year, i.e., 2010 and with flaws for 1960 (i.e., we're missing around 13 mln people of the overall population in 1960 due to data compilation issues of the earlier census approaches of the IBGE - [see here](https://ipeagit.github.io/censobr/reference/read_population.html#-census) for more information).
-Specifically, on the IBGE website, we found data for 2009, and the censobr package provides the 2010 data.
-It also provides census data for 1980 and 1960, but in 1980, it seems the population weights are missing.
+We retrieve population data via the censobr package, which compiled census and population data based on IBGE statistics. Administrative boundary data for is available in the R package geobr.
+
+While the authors produced the same plot for three different years — 1950, 1980, and 2010 — due to a lack of data, we were only able to reproduce the plot for the most recent year, i.e., 2010 and with flaws for 1960 (i.e., we're missing around 13 mln people of the overall population in 1960 due to data compilation issues of the earlier census approaches of the IBGE such as non-digitalized records - [see here](https://ipeagit.github.io/censobr/reference/read_population.html#-census) for more information).
+Specifically, on the IBGE website, we found data for 2009, while the censobr package provides the 2010 data.
+It also provides census data for 1980 and 1960.
+However, for 1980, it seems that the population weights are missing.
 Therefore, we decided to focus on the 2010 and 1960 data.
-
-For geographical boundaries of Brazil for administrative areas, we use the geobr package.
 
@@
 This leads us to loosing out roughly on 50'000 inhabitants in 2010.
+
 After this, we created a column indicating the population share, multiplying the ratio by 100 to make the results more consistent with the paper.
-Finally, we filtered the data for the North and Central-West regions to highlight the borders in the plot.
+Finally, we added the boundary data for the North and Central-West regions to highlight the borders in the plot as done by the authors.
 Needing `sf_remove_holes()` to avoid inner borders, we loose a bit of granularity in the outer borders of the broader regions.
@@
 
-We reproduce the same chart for 2010 as the authors did.
-Due to data issues, we rely on a further chart for 1960, instead of 1950.
-The 1960 chart resembles the 1950 chart of the authors surpisingly good, though the aforementioned data issues.
-Lastly, we omit 1980 data, since the population weights were missing in the censobr dataset.
+We reproduce the same chart for 2010, as the authors did.
+Due to data non-availability, we rely on a further chart for 1960, instead of 1950.
+The 1960 chart resembles the 1950 chart of the authors surprisingly well, despite the aforementioned data issues.
+Lastly, we omit 1980 data, since the population weights were missing in the censobr dataset and we were not able to infer them accurately.
 
@@
   ) +
+  # Add population share fill
   geom_sf(aes(fill = pop_share_1960), color = "white", size = 0.1) +
+  # Color scales
   scale_fill_stepsn(
@@
   ) +
+  # Add region boundaries
   geom_sf(data = region, aes(color = factor(name_region)), fill = NA, linewidth = 1) +
@@
   ) +
+  # Add population share fill
   geom_sf(aes(fill = pop_share_2010), color = "white", size = 0.1) +
+  # Color scales
   scale_fill_stepsn(
@@
 </table>
-
-Again, as outlined above, we only replicate the 2010 map, due to data issues in the prior periods.
 

### 2025-01-28 hw1/submission_notebook.Rmd
Final commit for paper 4+5 and updated main notebook

@@
 library(lwgeom) # Needed for st_make_valid
+library(rmapshaper) # Simplify and edit spatial features
+library(ggrepel) # Repel overlapping text labels in ggplot
 
@@
 library(geodata) # For elevation data
-library(readxl) # For Excel files
+library(readxl) # Import and read Excel files
 ```
@@
 
-# Paper 4
+# [Paper 4](https://economics.mit.edu/sites/default/files/publications/Catastrophe_Risk_and_Settlement_Location.pdf)
+
+We are looking for a shapefile or dataset specifically containing road network data for Vietnam.
+We inspected data from GADM (https://gadm.org/data.html#google_vignette) but none of the layers in the file appear to contain road network data.
+All layers are polygons representing administrative boundaries.
 
@@
 
+```{r}
+# Define the base path
+path <- "data/paper4"
+
+# We inspect data from Humdata:
+st_layers(file.path(path, "ROADS/vnm_rdsl_2015_OSM.shp"))
+# Read the shapefile into an sf object
+vietnam_roads <- st_read(file.path(path, "ROADS/vnm_rdsl_2015_OSM.shp"))
+view(vietnam_roads)
+
+# Get all the unique values from the column 'type'
+unique_values <- unique(vietnam_roads$type)
+print(unique_values)
+```
+
+Although the data is from 2015, whereas the paper displays roads from 2010, we decide to use this data as road networks.
+
+```{r}
+# We inspect data from SimpleMaps:
+st_layers(file.path(path,"VIETNAM/"))
+
+vietnam_map <- st_read(file.path(path,"VIETNAM", layer = "vn.shp"))
+```
+
+We decide to move forward with this data as Vietnam boundary.
+
 ## Data wrangling
+
+We use the author's documentation (https://www.openicpsr.org/openicpsr/project/207641/version/V1/view;jsessionid=3EBB67FC7920B2090C0D74C349E99F93?path=/openicpsr/207641/fcr:versions/V1/Public-replication-files/Replication_README.pdf&type=file) to understand how he categorised the types of roads.
+
+```{r}
+# Categorise roads into group by trying to replicate the author's logic
+vietnam_roads <- vietnam_roads %>%
+  mutate(
+    road_category = case_when(
+      type %in% c("motorway", "motorway_link") ~ "Freeway",
+      type %in% c("trunk", "trunk_link") ~ "Dual Carriageway",
+      type %in% c("secondary") ~ "Major Roads",
+      type %in% c("primary", "primary_link", "secondary_link", "tertiary", "tertiary_link", "residential", "living_street") ~ "Minor Roads",
+      TRUE ~ "Other Roads"
+  )
+)
+```
 
@@
 
+```{r}
+ggplot() +
+  # Adding the map of Vietnam as the base layer, using a light grey fill like the paper
+  geom_sf(data = vietnam_map, fill = "lightgrey", color = "black", linetype = "solid", size = 0.5) +
+
+  # Overlaying the road network on the map.
+  geom_sf(data = vietnam_roads, aes(color = road_category, linewidth = road_category), alpha = 0.8) +
+
+  # The 'color' aesthetic differentiates road categories
+  scale_color_manual(
+    values = c(
+      "Freeway" = "blue",               
+      "Dual Carriageway" = "darkgreen", 
+      "Major Roads" = "red",            
+      "Minor Roads" = "orange",       
+      "Other Roads" = "yellow"        
+    )
+  ) +
+  # The 'linewidth' aesthetic adjusts line thickness based on road types for visual distinction
+  scale_linewidth_manual(
+    values = c(
+      "Freeway" = 1.1,                # Thick line
+      "Dual Carriageway" = 1.1,       # Thick line
+      "Major Roads" = 0.5,            # Thin line
+      "Minor Roads" = 0.5,            # Thin line
+      "Other Roads" = 0.3             # Extra thin line
+    )
+  ) +
+  labs(
+    title = "Road maps of Vietnam",
+    color = "Road Category",
+    linewidth = "Road Category"    # Ensures the size legend matches the color legend
+  ) +
+  theme_minimal() +
+  theme(
+    panel.grid = element_blank(),  # Removes grid lines
+    axis.title = element_blank(),  # Removes axis titles
+    axis.text = element_blank(),   # Removes axis text
+    axis.ticks = element_blank()   # Removes axis ticks
+  )
+```
+
 In comparison the fourth paper map looked as follows:
 
-![Figure x]()
+![Figure 3](data/paper4/original_plot.png)
 
@@
 
-# Paper 5
+# [Paper 5](https://web.stanford.edu/~memorten/ewExternalFiles/Morten_Oliveira_Brasilia.pdf)
+
+The author's replication package is accessible, so we move forward with it's code and data (https://www.openicpsr.org/openicpsr/project/183316/version/V1/view).
 
@@
 
+```{r}
+path <- ("data/paper5/")
+
+# Load the shapefile containing state boundaries for Brazil (1940 projection)
+states <- st_read(file.path(path, "GIS_data/uf1940/uf1940_prj.shp"))
+# Simplify the state boundaries to reduce file size while maintaining shape integrity
+states_simple  <- rmapshaper::ms_simplify(states,keep=0.01,keep_shapes=TRUE)
+```
+
 ## Data wrangling
+
+```{r}
+# Define the year for the highways data
+year  <- 2000
+
+# Construct the file path for the highways shapefile dynamically based on the year
+file_name = paste0("/",year,"/highways_",year,"_prj.shp")
+
+# Load the highways shapefile for the specified year
+all_highways <- st_read(file.path(path, "GIS_data/roads", file_name))
+# Simplify the highways data
+all_highways_simple <- rmapshaper::ms_simplify(all_highways,keep=0.01,keep_shapes=TRUE)
+
+# Load and simplify the minimum spanning tree (MST) shapefile for pie distribution
+mst_pie <- st_read(file.path(path, "GIS_data/mst/mst_pie_prj.shp"))
+mst_pie_simple  <- rmapshaper::ms_simplify(mst_pie,keep=0.01,keep_shapes=TRUE)
+
+# Load and simplify MST data excluding pies for Rio region
+mst_all_rio <- st_read(file.path(path, "GIS_data/mst/mst_all_rio_prj.shp"))
+mst_all_rio_simple  <- rmapshaper::ms_simplify(mst_all_rio,keep=0.01,keep_shapes=TRUE)
+
+# Load shapefile for capital cities and extract coordinates for labeling
+capital_cities <- st_read(file.path(path, "GIS_data/cities/brazil_capital_cities_prj.shp"))
+cities_xy <- cbind(capital_cities,st_coordinates(capital_cities))
+```
 
@@
 
+```{r}
+# Create a plot showing MST and highway networks in Brazil
+fig_5  <- ggplot() +
+  # Add state boundaries as the base layer with light grey outlines
+  geom_sf(data=states_simple, fill="white", color="grey90") +
+  # Add the simplified MST data with a unique linetype and color
+  geom_sf(data=mst_pie_simple, size=.6, linetype = "11", aes(color = "Minimum spanning tree"), show.legend = "line")   +
+  # Add non-radial highways with a dashed line style
+  geom_sf(data=all_highways_simple %>% filter(dm_anlys_p==1 & dm_radial==0), size=.3, linetype = "dashed", aes(color = "Non-radial highways (2000)"), show.legend = "line") +   
+  # Add radial highways with a solid line style
+  geom_sf(data=all_highways_simple %>% filter(dm_anlys_p==1 & dm_radial==1), size=0.6, aes(color = "Radial highways (2000)"), show.legend = "line") +
+  # Minimal theme for simplicity
+  theme_minimal() +
+  # Overlay capital cities and label them with their names
+  geom_point(data=cities_xy, aes(x=X, y=Y)) +
+  geom_text_repel(data=cities_xy, aes(x=X, y=Y, label=CITY_NAME)) +
+  # Add a custom legend for line types and colors
+  labs(color = " ") +
+  scale_color_manual(values=c("#777676","#868686","#565555"),
+                     guide = guide_legend(override.aes = list(linetype = c("11", "dashed", "solid")))) +
+  # Remove axis lines, ticks, and text for a clean map appearance
+  theme(
+    axis.line = element_blank(),
+    axis.text = element_blank(),
+    axis.ticks = element_blank(),
+    axis.title = element_blank()
+  )
+
+print(fig_5)
+```
 In comparison the fifth paper map looked as follows:
 
-![Figure x]()
+![Figure 1](data/paper5/original_plot.png)

### 2025-01-28 hw1/paper4.R

@@ -9,27 +9,24 @@ library(readxl)
 library(ggplot2)
 library(raster)
 library(terra)
+library(geobr)
 library(rmapshaper)
-library(ggplot2)
 library(ggrepel)
 
-path <- ("/Users/simonvellin/Documents/BSE/TERM 2/GEOSPATIAL/0. ASSIGNMENTS/HW1/PAPER 4")
-outpath <- ("/Users/simonvellin/Documents/BSE/TERM 2/GEOSPATIAL/0. ASSIGNMENTS/HW1/PAPER 4")
-
-----------------------------------------
-            # DATA MINING
-----------------------------------------
-
 # We are looking for a shapefile or dataset specifically containing road network data for Vietnam.
+# We inspected data from GADM (https://gadm.org/data.html#google_vignette) but none of the layers in the file appear to contain road network data. All layers are polygons representing administrative boundaries.
 
-# We inspect data from GADM:
-st_layers("/Users/simonvellin/Documents/BSE/TERM 2/GEOSPATIAL/0. ASSIGNMENTS/HW1/PAPER 4/GADM-Vietnam")
+----------------------------------------
+            # DATA LOADING
+----------------------------------------
 
-# None of the layers in the file appear to contain road network data. All layers are polygons representing administrative boundaries. 
+# Define the base path
+path <- "hw1/data/paper4/"
 
 # We inspect data from Humdata:
-st_layers("/Users/simonvellin/Documents/BSE/TERM 2/GEOSPATIAL/0. ASSIGNMENTS/HW1/PAPER 4/ROADS/vnm_rdsl_2015_OSM.shp")
-vietnam_roads <- st_read("/Users/simonvellin/Documents/BSE/TERM 2/GEOSPATIAL/0. ASSIGNMENTS/HW1/PAPER 4/ROADS/vnm_rdsl_2015_OSM.shp")
+st_layers(file.path(path, "ROADS/vnm_rdsl_2015_OSM.shp"))
+# Read the shapefile into an sf object
+vietnam_roads <- st_read(file.path(path, "ROADS/vnm_rdsl_2015_OSM.shp"))
 view(vietnam_roads)
 
 # Get all the unique values from the column 'type'
@@ -39,16 +36,10 @@ print(unique_values)
 # Although the data is from 2015, whereas the paper displays roads from 2010, we decide to move forward with this data as road network.
 
 # We inspect data from SimpleMaps:
[…]

### 2025-01-28 hw1/paper5.R

@@ -9,26 +9,25 @@ library(readxl)
 library(ggplot2)
 library(raster)
 library(terra)
+library(geobr)
 library(rmapshaper)
-library(ggplot2)
 library(ggrepel)
-
-path <- ("hw1/data/paper5/")
-outpath <- ("/Users/simonvellin/Documents/BSE/TERM 2/GEOSPATIAL/0. ASSIGNMENTS/HW1/PAPER 5/OUTPUT")
-
-
-### Data mining
 
 # The author's replication package is accessible, so we move forward with it's code/data (https://www.openicpsr.org/openicpsr/project/183316/version/V1/view).
 
-### Data loading
+----------------------------------------
+          # DATA LOADING
+----------------------------------------
+path <- ("hw1/data/paper5/")
 
 # Load the shapefile containing state boundaries for Brazil (1940 projection)
 states <- st_read(file.path(path, "GIS_data/uf1940/uf1940_prj.shp"))
 # Simplify the state boundaries to reduce file size while maintaining shape integrity
 states_simple  <- rmapshaper::ms_simplify(states,keep=0.01,keep_shapes=TRUE)
 
-### Data wrangling
+----------------------------------------
+          # DATA WRANGLING
+----------------------------------------
 
 # Define the year for the highways data
 year  <- 2000
@@ -38,7 +37,7 @@ file_name = paste0("/",year,"/highways_",year,"_prj.shp")
 
 # Load the highways shapefile for the specified year
 all_highways <- st_read(file.path(path, "GIS_data/roads", file_name))
-# Simplify the highways data for efficient plotting
+# Simplify the highways data
 all_highways_simple <- rmapshaper::ms_simplify(all_highways,keep=0.01,keep_shapes=TRUE)
 
 # Load and simplify the minimum spanning tree (MST) shapefile for pie distribution
@@ -53,10 +52,12 @@ mst_all_rio_simple  <- rmapshaper::ms_simplify(mst_all_rio,keep=0.01,keep_shapes
 capital_cities <- st_read(file.path(path, "GIS_data/cities/brazil_capital_cities_prj.shp"))
 cities_xy <- cbind(capital_cities,st_coordinates(capital_cities))
 
-### Plotting
+----------------------------------------
+          # PLOTTING
+----------------------------------------
 
 # Create a plot showing MST and highway networks in Brazil
-fig_1  <- ggplot() +
+fig_5  <- ggplot() +
[…]

### 2025-01-28 hw1/submission_notebook.Rmd
Final optimizations on markdown file

--- 2025-01-28 hw1/submission_notebook.Rmd
@@
 
-In the following notebook we replicate the charts for all 5 papers. Contentwise, the code follows always the same structure. We first load the data, then we create the chart and finally display it.
+In the following notebook we replicate the charts for all 5 papers.
+Contentwise, the code follows always the same structure.
+We first load the data, secondly manipulate it, then we create the chart and finally display it.
 
@@
 While the authors produced the same plot for three different years — 1950, 1980, and 2010 — due to a lack of data, we were only able to reproduce the plot for the most recent year, i.e., 2010 and with flaws for 1960 (i.e., we're missing around 13 mln people of the overall population in 1960 due to data compilation issues of the earlier census approaches of the IBGE such as non-digitalized records - [see here](https://ipeagit.github.io/censobr/reference/read_population.html#-census) for more information).
-Specifically, on the IBGE website, we found data for 2009, while the censobr package provides the 2010 data.
+
+On the IBGE website, we found data for 2009, while the censobr package provides the 2010 data.
 It also provides census data for 1980 and 1960.
@@
 The 1960 chart resembles the 1950 chart of the authors surprisingly well, despite the aforementioned data issues.
+The Northern regions being significantly different might be explained by the censobr authors who inter alia state that for 1960:
+>"Information is missing for the states of the former Northern Region, Maranhão, Piaui, Guanabara, Santa Catarina, and Espírito Santo".
+
 Lastly, we omit 1980 data, since the population weights were missing in the censobr dataset and we were not able to infer them accurately.
@@
 
-# Paper 4
+# [Paper 4](https://economics.mit.edu/sites/default/files/publications/Catastrophe_Risk_and_Settlement_Location.pdf)
+
+We are looking for a shapefile or dataset specifically containing road network data for Vietnam.
+We inspected data from [GADM](https://gadm.org/data.html#google_vignette) but none of the layers in the file appear to contain road network data.
+All layers are polygons representing administrative boundaries.
 
@@
 
+```{r}
+# Define the base path
+path <- "data/paper4"
+
+# We inspect data from Humdata:
+st_layers(file.path(path, "ROADS/vnm_rdsl_2015_OSM.shp"))
+# Read the shapefile into an sf object
+vietnam_roads <- st_read(file.path(path, "ROADS/vnm_rdsl_2015_OSM.shp"))
+
+# Get all the unique values from the column 'type'
+unique_values <- unique(vietnam_roads$type)
+print(unique_values)
+```
+
+Although the data is from 2015, whereas the paper displays roads from 2010, we decide to use this data as road networks.
+
+```{r}
+# We inspect data from SimpleMaps:
+st_layers(file.path(path, "VIETNAM/"))
+
+vietnam_map <- st_read(file.path(path, "VIETNAM", layer = "vn.shp"))
+```
+
+We decide to move forward with this data as Vietnam boundary.
+
 ## Data wrangling
+
+We use the [author's documentation](https://www.openicpsr.org/openicpsr/project/207641/version/V1/view;jsessionid=3EBB67FC7920B2090C0D74C349E99F93?path=/openicpsr/207641/fcr:versions/V1/Public-replication-files/Replication_README.pdf&type=file) to understand how he categorised the types of roads.
+
+```{r}
+# Categorise roads into group by trying to replicate the author's logic
+vietnam_roads <- vietnam_roads %>%
+  mutate(
+    road_category = case_when(
+      type %in% c("motorway", "motorway_link") ~ "Freeway",
+      type %in% c("trunk", "trunk_link") ~ "Dual Carriageway",
+      type %in% c("secondary") ~ "Major Roads",
+      type %in% c("primary", "primary_link", "secondary_link", "tertiary", "tertiary_link", "residential", "living_street") ~ "Minor Roads",
+      TRUE ~ "Other Roads"
+    )
+  )
+```
 
@@
 
+```{r}
+ggplot() +
+  # Adding the map of Vietnam as the base layer, using a light grey fill like the paper
+  geom_sf(data = vietnam_map, fill = "lightgrey", color = "black", linetype = "solid", size = 0.5) +
+
+  # Overlaying the road network on the map.
+  geom_sf(data = vietnam_roads, aes(color = road_category, linewidth = road_category), alpha = 0.8) +
+
+  # The 'color' aesthetic differentiates road categories
+  scale_color_manual(
+    values = c(
+      "Freeway" = "blue",
+      "Dual Carriageway" = "darkgreen",
+      "Major Roads" = "red",
+      "Minor Roads" = "orange",
+      "Other Roads" = "yellow"
+    )
+  ) +
+  # The 'linewidth' aesthetic adjusts line thickness based on road types for visual distinction
+  scale_linewidth_manual(
+    values = c(
+      "Freeway" = 1.1, # Thick line
+      "Dual Carriageway" = 1.1, # Thick line
+      "Major Roads" = 0.5, # Thin line
+      "Minor Roads" = 0.5, # Thin line
+      "Other Roads" = 0.3 # Extra thin line
+    )
+  ) +
+  labs(
+    title = "Road maps of Vietnam",
+    color = "Road Category",
+    linewidth = "Road Category" # Ensures the size legend matches the color legend
+  ) +
+  theme_minimal() +
+  theme(
+    panel.grid = element_blank(), # Removes grid lines
+    axis.title = element_blank(), # Removes axis titles
+    axis.text = element_blank(), # Removes axis text
+    axis.ticks = element_blank() # Removes axis ticks
+  )
+```
+
 In comparison the fourth paper map looked as follows:
 
-![Figure x]()
+![Figure 3](data/paper4/original_plot.png)
 
@@
 
-# Paper 5
+# [Paper 5](https://web.stanford.edu/~memorten/ewExternalFiles/Morten_Oliveira_Brasilia.pdf)
+
+The author's replication package is accessible, so we move forward with it's [code and data](https://www.openicpsr.org/openicpsr/project/183316/version/V1/view).
 
@@
 
+```{r}
+path <- ("data/paper5/")
+
+# Load the shapefile containing state boundaries for Brazil (1940 projection)
+states <- st_read(file.path(path, "GIS_data/uf1940/uf1940_prj.shp"))
+# Simplify the state boundaries to reduce file size while maintaining shape integrity
+states_simple <- rmapshaper::ms_simplify(states, keep = 0.01, keep_shapes = TRUE)
+```
+
 ## Data wrangling
+
+Since the authors' description is very comprehensive, we proceed also with their data manipulation approach.
+
+```{r}
+# Define the year for the highways data
+year <- 2000
+
+# Construct the file path for the highways shapefile dynamically based on the year
+file_name <- paste0("/", year, "/highways_", year, "_prj.shp")
+
+# Load the highways shapefile for the specified year
+all_highways <- st_read(file.path(path, "GIS_data/roads", file_name))
+# Simplify the highways data
+all_highways_simple <- rmapshaper::ms_simplify(all_highways, keep = 0.01, keep_shapes = TRUE)
+
+# Load and simplify the minimum spanning tree (MST) shapefile for pie distribution
+mst_pie <- st_read(file.path(path, "GIS_data/mst/mst_pie_prj.shp"))
+mst_pie_simple <- rmapshaper::ms_simplify(mst_pie, keep = 0.01, keep_shapes = TRUE)
+
+# Load and simplify MST data excluding pies for Rio region
+mst_all_rio <- st_read(file.path(path, "GIS_data/mst/mst_all_rio_prj.shp"))
+mst_all_rio_simple <- rmapshaper::ms_simplify(mst_all_rio, keep = 0.01, keep_shapes = TRUE)
+
+# Load shapefile for capital cities and extract coordinates for labeling
+capital_cities <- st_read(file.path(path, "GIS_data/cities/brazil_capital_cities_prj.shp"))
+cities_xy <- cbind(capital_cities, st_coordinates(capital_cities))
+```
 
@@
 
+Lastly, we also use their plotting approach:
+
+```{r}
+# Create a plot showing MST and highway networks in Brazil
+ggplot() +
+  # Add state boundaries as the base layer with light grey outlines
+  geom_sf(data = states_simple, fill = "white", color = "grey90") +
+  # Add the simplified MST data with a unique linetype and color
+  geom_sf(data = mst_pie_simple, size = .6, linetype = "11", aes(color = "Minimum spanning tree"), show.legend = "line") +
+  # Add non-radial highways with a dashed line style
+  geom_sf(data = all_highways_simple %>% filter(dm_anlys_p == 1 & dm_radial == 0), size = .3, linetype = "dashed", aes(color = "Non-radial highways (2000)"), show.legend = "line") +
+  # Add radial highways with a solid line style
+  geom_sf(data = all_highways_simple %>% filter(dm_anlys_p == 1 & dm_radial == 1), size = 0.6, aes(color = "Radial highways (2000)"), show.legend = "line") +
+  # Minimal theme for simplicity
+  theme_minimal() +
+  # Overlay capital cities and label them with their names
+  geom_point(data = cities_xy, aes(x = X, y = Y)) +
+  geom_text_repel(data = cities_xy, aes(x = X, y = Y, label = CITY_NAME)) +
+  # Add a custom legend for line types and colors
+  labs(color = " ") +
+  scale_color_manual(
+    values = c("#777676", "#868686", "#565555"),
+    guide = guide_legend(override.aes = list(linetype = c("11", "dashed", "solid")))
+  ) +
+  # Remove axis lines, ticks, and text for a clean map appearance
+  theme(
+    axis.line = element_blank(),
+    axis.text = element_blank(),
+    axis.ticks = element_blank(),
+    axis.title = element_blank()
+  )
+```
+
 In comparison the fifth paper map looked as follows:
 
-![Figure x]()
+![Figure 1](data/paper5/original_plot.png)

### 2025-01-28 hw1/test3.R

@@ -1,42 +1,0 @@
-library(censobr)
-library(tidyverse)
-
-
-# return data as arrow Dataset
-pop_2010 <- read_population(
-  year = 2010,
-  # columns = c("code_state", "code_region", "code_muni", "V0010"),
-  showProgress = FALSE,
-  as_data_frame = TRUE,
-  # cache = FALSE
-)
-
-pop_1980 <- read_population(
-  year = 1980,
-  # columns = c("code_meso", "code_region", "code_state", "code_muni", "V198"),
-  showProgress = FALSE,
-  as_data_frame = TRUE,
-  cache = FALSE
-)
-
-pop_1960 <- read_population(
-  year = 1960,
-  columns = c("code_muni_1960", "censobr_weight"),
-  showProgress = FALSE,
-  as_data_frame = TRUE,
-  # cache = FALSE
-)
-
-df %>%
-  summarise(total_population = sum(V0010, rm.na = TRUE))
-
-pop_1980 %>%
-  mutate(V198 = as.numeric(V198)) %>%
-  summarise(total_population = sum(V198, rm.na = TRUE))
-# group_by(code_region) %>%
-# summarise(total_population = sum(V0010, rm.na = TRUE))
-
-
-pop_1980 %>%
-  select(V601, V602, V603, V604, V605) %>%
-  head()

### 2025-01-29 hw2/hw2_MoritzPeistNoemiLucchiSimonVellin.Rmd
Initial submission for assignment 2 - TODO: Upload world data

--- 2025-01-28 hw1/submission_notebook.Rmd
@@
 ---
-title: "Geospatial - Assignment 1"
+title: "Geospatial - Assignment 2"
 author: 
@@
 <div style="text-align: center;">
-  <img src="Barcelona_School_of_Economics_logo.svg" alt="BSE logo">
+  <img src="../Barcelona_School_of_Economics_logo.svg" alt="BSE logo">
 </div>
@@
 library(sf)
-library(terra) # Newer replacement for raster
-library(osmdata) # OpenStreetMap data
-library(spData) # Spatial data
-library(geobr) # Brazilian boundaries
-library(censobr) # Brazilian population data
-library(lwgeom) # Needed for st_make_valid
-
-# Data import
-library(geodata) # For elevation data
-library(readxl) # For Excel files
 ```
@@
 <hr>
-
-# [Paper 1](https://www.sciencedirect.com/science/article/pii/S0304387818300166#appsec1)
-
-## Data loading
-
-The authors of the paper provide a link to the data, however, the link is unfortunately outdated.
-Yet, the data is available in a newer version on [this website](https://www.dws.gov.za/DSO/publications.aspx).
-The authors used 2013 data, we use 2024 data but had to convert the downloaded data to geojson format before usage.
-
-Moreover, the river gradient data was not directly available, so we use the elevation data from the geodata package to retrieve and later calculate river gradients.
-
-```{r}
-# Set geodata default path
-geodata_path("data/paper1")
-
-# Dams, Source:https://www.dws.gov.za/DSO/publications.aspx
-# Data was only available as .KMZ and not readable by st_read so needed to be converted by:
-# https://mygeodata.cloud/converter/kml-to-geojson
-dams <- st_read(
-  "data/paper1/Registered Dams Oct20241.geojson"
-)
-
-# Dam metadata (same source but an Excel file)
-dam_met <- read_excel(
-  "data/paper1/List of Registered Dams Oct2024.xlsx",
-  col_types = rep("text", ncol(read_excel("data/paper1/List of Registered Dams Oct2024.xlsx")))
-) %>%
-  mutate(`Completion date` = na_if(as.numeric(`Completion date`), 0))
-
-# RSA boundaries - downloaded from GDAM
-sa_districts_sf <- st_as_sf(gadm(country = "ZAF", level = 3))
-
-# River data - Download elevation data for South Africa
-elevation <- elevation_30s(country = "ZAF")
-```
-
-## Data wrangling
-
-We try to apply the same data cleaning as the authors though we deal with more recent data, i.e.:
-
-> "The data was accessed on August 23, 2013, including all dams listed as of March 2013.
-About 15 percent of dams were dropped because of missing critical information that could not be verified (690/4,830).
-For most missing dams (658/690), the missing information was completion date.
-Online internet searches for this information suggest that most were very old (prior to the analysis period) or no longer operational.
-Of those dropped, 499 reported a purpose of “irrigation.” The resulting dataset included 4,140  dams, which were restricted to the subset of irrigation dams (3,176)."
-
-Furthermore, we calculate the mean gradient for each district and add it to the districts data.
-
-```{r}
-# Now filter dams, as done in the paper, i.e, omit missing completion date and filter for irrigation
-dams <- dams %>%
-  left_join(dam_met,
-    by = join_by("COL52A1FC9ABE076254" == "No of dam"),
-    # relationship = "many-to-many"
-  ) %>%
-  filter(!is.na(`Completion date`), str_detect(tolower(Purpose), "irrigation")) %>%
-  distinct(across(-`geometry`), .keep_all = TRUE) # Delete the 2 duplicate cases identified
-
-# Convert districts to terra format for raster operations
-districts_vect <- vect(sa_districts_sf)
-
-# Calculate slope from elevation
-slope <- terrain(elevation, v = "slope", unit = "degrees")
-
-# Calculate zonal statistics (mean slope) for each district
-district_gradients <- terra::extract(slope, districts_vect, fun = mean, na.rm = TRUE)
-
-# Add gradient data to districts
-sa_districts_sf$gradient <- district_gradients$slope
-
-# Ensure same CRS
-dams <- st_transform(dams, st_crs(sa_districts_sf))
-
-# Clip dams to South Africa boundaries
-sa_boundary <- st_union(sa_districts_sf)
-dams_clipped <- st_intersection(dams, sa_boundary)
-```
-
-## Plotting
-
-Finally, we output the plot:
-
-```{r}
-# Create plot
-ggplot() +
-  # Add districts with gradient fill
-  geom_sf(
-    data = sa_districts_sf,
-    aes(fill = gradient),
-    color = "white",
-    size = 0.1
-  ) +
-  # Add clipped dams
-  geom_sf(
-    data = dams_clipped,
-    size = 0.2,
-    color = "black",
-    aes(shape = "Dam Location")
-  ) + # Add to legend
-  # Style the gradient fill
-  scale_fill_gradientn(
-    colors = grey.colors(6, start = 0.9, end = 0.2),
-    name = "Average District\nRiver Gradient",
-    breaks = c(
-      0, 2.229103, 3.821713, 6.593167,
-      9.879707, 13.130823, 19.829017
-    ),
-    labels = function(x) format(x, digits = 6),
-    guide = guide_colorbar(
-      order = 1,
-      nbin = 100, # For smoother color transition
-      ticks.colour = "black",
-      frame.colour = "black",
-      frame.linewidth = 0.5
-    )
-  ) +
-  # Add shape scale for dam points in legend
-  scale_shape_manual(
-    values = c("Dam Location" = 16),
-    name = NULL,
-    guide = guide_legend(order = 2)
-  ) +
-  # Customize theme
-  theme_minimal() +
-  theme(
-    panel.grid = element_blank(),
-    axis.text = element_blank(),
-    axis.title = element_blank(),
-    plot.title = element_text(hjust = 0.5, size = 14),
-    legend.position = "left",
-    legend.box = "vertical",
-    legend.margin = margin(0, 0, 0, 0),
-    legend.spacing = unit(0.1, "cm"),
-    legend.title = element_text(size = 10),
-    legend.text = element_text(size = 8),
-    legend.key.size = unit(0.5, "cm")
-  ) +
-  # Add titles
-  labs(
-    title = "South African Dams and River Gradients",
-    caption = "Source: Based on Mettetal (2019) but with 2024 instead of 2013 data"
-  )
-```
-
-In comparison the first paper map looked as follows:
-
-![Figure 2](data/paper1/1-s2.0-S0304387818300166-gr2.jpg)
-
-<hr>
-
-# [Paper 2](https://www.sciencedirect.com/science/article/pii/S0166046220303100)
-
-## Data loading
-
-Unfortunately, again the data was not directly available.
-Therefore, we leverage some R packages, i.e., to get the data for highways, powergrids we used the osmdata package and for population data from 2020 we used the geodata package.
-The [ERSS survey](https://microdata.worldbank.org/index.php/catalog/2053/data-dictionary) data was directly available via the World Bank data repositories as a CSV (2011/12) and is identical to the data used in the paper.
-We load all of these in the following:
-
-```{r}
-# Set geodata default path
-geodata_path("data/paper2")
-
-# Get Ethiopia districts
-eth_districts_sf <- gadm(country = "ETH", level = 3)
-eth_districts_sf <- st_as_sf(eth_districts_sf)
-
-# Get population data and process it for Ethiopia
-ethiopia_pop <- population(2020, res = 5)
-ethiopia_pop <- crop(ethiopia_pop, ext(eth_districts_sf))
-ethiopia_pop <- mask(ethiopia_pop, vect(eth_districts_sf))
-
-# Convert districts to same CRS as population data if needed
-eth_districts_sf <- st_transform(eth_districts_sf, crs(ethiopia_pop))
-
-# WB data: https://microdata.worldbank.org/index.php/catalog/2053/data-dictionary
-# Read in the survey data and get it in the right shape
-wb_data <- read_csv("data/paper2/pub_eth_householdgeovariables_y1.csv") %>%
-  group_by(ea_id) %>%
-  summarise(
-    LAT_DD_MOD = mean(LAT_DD_MOD, na.rm = TRUE),
-    LON_DD_MOD = mean(LON_DD_MOD, na.rm = TRUE)
-  )
-
-# Transform it to GEO data and clip to Ethiopia
-wb_data_geo <- st_as_sf(wb_data, coords = c("LON_DD_MOD", "LAT_DD_MOD"), crs = 4326)
-wb_data_geo <- st_transform(wb_data_geo, crs(ethiopia_pop))
-
-# Get high voltage power lines with comprehensive voltage range
-ethiopia_hv_power <- opq("Ethiopia") %>%
-  add_osm_feature(key = "power", value = c("line", "cable")) %>%
-  add_osm_feature(key = "voltage", value = paste0(seq(66, 500, by = 1), "000")) %>%
-  osmdata_sf()
-
-# Clean and clip power lines
-hv_lines_clean <- ethiopia_hv_power$osm_lines %>%
-  mutate(voltage = str_extract(voltage, "\\d+")) %>%
-  mutate(voltage_kv = as.numeric(voltage) / 1000) %>%
-  filter(!is.na(voltage_kv)) %>%
-  st_intersection(eth_districts_sf)
-
-# Get major roads and clip to Ethiopia
-ethiopia_roads <- opq("Ethiopia") %>%
-  add_osm_feature(
-    key = "highway",
-    value = c("motorway", "trunk")
-  ) %>%
-  osmdata_sf()
-```
-
-## Data wrangling
-
-We calculate the population density for each district and clip the roads to the Ethiopia boundary.
-Furthermore, it looks like the authors used slightly different road and powergrid networks.
-However, overall the grids look broadly similar when utilizing OpenStreetMap data via the osmdata package.
-
-```{r}
-# Clip roads to Ethiopia boundary
-roads_clipped <- ethiopia_roads$osm_lines %>%
-  st_intersection(eth_districts_sf)
-
-# Calculate population density by district
-districts_pop <- terra::extract(ethiopia_pop, vect(eth_districts_sf), fun = sum, na.rm = TRUE)
-
-# Add population to districts dataset
-eth_districts_sf$total_pop <- districts_pop[, 2] # The second column contains the values
-
-# Calculate area in square kilometers
-eth_districts_sf$area_km2 <- as.numeric(st_area(eth_districts_sf)) / 1000000
-
-# Calculate density (people per km2)
-eth_districts_sf$pop_density <- eth_districts_sf$total_pop / eth_districts_sf$area_km2
-```
-
-## Plotting
-
-Finally, we output the plot:
-
-```{r}
-# Create the final plot
-ggplot() +
-  # Administrative boundaries
-  geom_sf(
-    data = eth_districts_sf,
-    aes(fill = pop_density),
-    color = "gray50",
-    linewidth = 0.1
-  ) +
-  # Population density
-  scale_fill_gradientn(
-    colors = c("white", "lightblue", "royalblue", "navy"),
-    trans = "log",
-    name = "Population\nDensity\n(people/km²)",
-    labels = scales::comma,
-    guide = guide_colorbar(order = 1)
-  ) +
-  # Adding roads
-  geom_sf(
-    data = roads_clipped,
-    aes(color = "Highway"),
-    linewidth = 0.5
-  ) +
-  # Adding HV powerlines
-  geom_sf(
-    data = hv_lines_clean,
-    aes(color = "HV grid"),
-    linewidth = 0.35
-  ) +
-  # Adding survey villages
-  geom_sf(
-    data = wb_data_geo,
-    aes(shape = "Survey village"),
-    color = "black",
-    size = 1,
-    alpha = 0.6
-  ) +
-  # Adding legends
-  scale_color_manual(
-    name = NULL,
-    values = c("Highway" = "black", "HV grid" = "red"),
-    guide = guide_legend(order = 2)
-  ) +
-  scale_shape_manual(
-    name = NULL,
-    values = c("Survey village" = 16),
-    guide = guide_legend(order = 3)
-  ) +
-  # Styling
-  guides(
-    fill = guide_colorbar(order = 1),
-    color = guide_legend(
-      order = 2,
-      override.aes = list(
-        linewidth = 1,
-        shape = NA, # Removes the box
-        linetype = 1
-      )
-    ),
-    shape = guide_legend(order = 3)
-  ) +
-  theme_minimal() +
-  labs(
-    title = "High Voltage Power Grid Network and Major Roads in Ethiopia",
-    subtitle = "Showing transmission lines ≥66kV, population density by district, and villages",
-    caption = "Data sources: OpenStreetMap, WorldPop 2020, World Bank (2012)\n
-      Based on Fried and Lagakos (2021)"
-  ) +
-  theme(
-    legend.position = "left",
-    plot.title = element_text(size = 10),
-    plot.subtitle = element_text(size = 8),
-    plot.caption = element_text(size = 6),
-    axis.text = element_blank(),
-    axis.title = element_blank()
-  )
-```
-
-In comparison the second paper map looked as follows: 
-![Figure 4](data/paper2/1-s2.0-S0166046220303100-gr4.jpg)
-
-The different color labeling of the surveyed villages is unfortunately not explained in the paper. Therefore, we leave the points black.
-
-<hr>
-
-# [Paper 3](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3776990)
-
-## Data loading
-
-The authors of the paper do not provide a direct link to the data, but they refer to the Brazilian Statistical Bureau (IBGE) when talkig about population data and introducing the "meso-regions", which are the political boundaries used in their plots.
-
-We retrieve population data via the censobr package, which compiled census and population data based on IBGE statistics. Administrative boundary data for is available in the R package geobr.
-
-While the authors produced the same plot for three different years — 1950, 1980, and 2010 — due to a lack of data, we were only able to reproduce the plot for the most recent year, i.e., 2010 and with flaws for 1960 (i.e., we're missing around 13 mln people of the overall population in 1960 due to data compilation issues of the earlier census approaches of the IBGE such as non-digitalized records - [see here](https://ipeagit.github.io/censobr/reference/read_population.html#-census) for more information).
-Specifically, on the IBGE website, we found data for 2009, while the censobr package provides the 2010 data.
-It also provides census data for 1980 and 1960.
-However, for 1980, it seems that the population weights are missing.
-Therefore, we decided to focus on the 2010 and 1960 data.
-
-```{r}
-# Read in the Brasilian population for 2010 using the censobr package
-# This also contains 1980, and 1960 data, but 1980 without weights, and 1960 only partially
-pop_2010 <- read_population(
-  year = 2010,
-  columns = c("code_state", "code_region", "code_muni", "V0010"),
-  showProgress = FALSE,
-  as_data_frame = TRUE,
-  # cache = FALSE
-) %>% # Group by municipality code
-  group_by(code_muni) %>%
-  summarise(
-    total_population_2010 = sum(V0010, rm.na = TRUE)
-  )
-
-# Get 1960 data (which is missing approx 13 mln people though -
-# it's only a rough estimate and combined from 2 datasets, see censobr description)
-pop_1960 <- read_population(
-  year = 1960,
-  columns = c("code_muni_1960", "censobr_weight"),
-  showProgress = FALSE,
-  as_data_frame = TRUE,
-  # cache = FALSE
-) %>% # Group by municipality code
-  group_by(code_muni_1960) %>%
-  summarise(
-    total_population_1960 = sum(censobr_weight, rm.na = TRUE)
-  )
-
-# Get the geographical data from the package geobr and validate if necessary
-meso <- read_meso_region(year = 2010) %>%
-  st_make_valid()
-region <- read_region(year = 2010)
-muni <- read_municipality(year = 2010) %>%
-  st_make_valid()
-```
-
-## Data wrangling
-
-First, we join the 1960 and 2010 population data with munipalicity level data to ensure that both datasets refer to the same political boundaries.
-Next, we joined the two datasets based with the meso code and group on meso level to get the wanted granularity.
-We had to take a rather coarse-granular approach by using the `st_centroid()` function to match meso-regions with municipalities.
-This leads us to loosing out roughly on 50'000 inhabitants in 2010.
-
-After this, we created a column indicating the population share, multiplying the ratio by 100 to make the results more consistent with the paper.
-Finally, we added the boundary data for the North and Central-West regions to highlight the borders in the plot as done by the authors.
-Needing `sf_remove_holes()` to avoid inner borders, we loose a bit of granularity in the outer borders of the broader regions.
-
-```{r}
-# First, join the population data with municipal boundaries
-pop_spatial <- muni %>%
-  left_join(pop_2010, by = "code_muni") %>%
-  left_join(pop_1960, by = join_by("code_muni" == "code_muni_1960"))
-
-# Use st_centroid to ensure each municipality is counted only once
-pop_meso <- pop_spatial %>%
-  st_centroid() %>% # Convert polygons to points
-  st_join(meso) %>% # Join with meso regions
-  group_by(code_meso) %>%
-  summarise(
-    total_population_2010 = sum(total_population_2010, na.rm = TRUE),
-    total_population_1960 = sum(total_population_1960, na.rm = TRUE)
-  ) %>%
-  st_drop_geometry() # Remove geometry before joining back
-
-# Now join back with meso geometries
-pop_meso <- meso %>%
-  left_join(pop_meso, by = "code_meso")
-
-# Create pop_share
-df_merged <- pop_meso %>%
-  mutate(
-    pop_share_2010 = total_population_2010 / sum(total_population_2010) * 100,
-    pop_share_1960 = total_population_1960 / sum(total_population_1960) * 100,
-  )
-
-# Aggregate regions as the author's did to get the correct boundaries
-region <- region %>%
-  mutate(
-    name_region = case_when(
-      name_region %in% c("Centro Oeste", "Norte") ~ 2,
-      TRUE ~ 1
-    )
-  ) %>%
-  dplyr::select(name_region, geom) %>%
-  group_by(name_region) %>%
-  summarise(geom = st_union(geom)) %>%
-  sfheaders::sf_remove_holes()
-```
-
-## Plotting
-
-We reproduce the same chart for 2010, as the authors did.
-Due to data non-availability, we rely on a further chart for 1960, instead of 1950.
-The 1960 chart resembles the 1950 chart of the authors surprisingly well, despite the aforementioned data issues.
-Lastly, we omit 1980 data, since the population weights were missing in the censobr dataset and we were not able to infer them accurately.
-
-```{r, figures-side, fig.show="hold", out.width="50%"}
-# Final plot - 1960
-ggplot(data = df_merged) +
-  ggtitle("(c) 1960") +
-  theme_minimal() +
-  theme(
-    plot.title = element_text(hjust = 0.5), # Center title while keeping minimal theme
-    panel.grid = element_line(color = "gray90"), # Make grid lines lighter
-    axis.text = element_blank(),
-    axis.ticks = element_blank()
-  ) +
-  # Add population share fill
-  geom_sf(aes(fill = pop_share_1960), color = "white", size = 0.1) +
-  # Color scales
-  scale_fill_stepsn(
-    name = "Pop. Share",
-    breaks = c(0, 0.216, 0.382, 0.557, 1.02, 11.12),
-    labels = scales::number_format(accuracy = 0.001),
-    colors = c("#eff3ff", "#bdd7e7", "#6baed6", "#3182bd", "#08519c"), # Made lower values darker
-    values = scales::rescale(c(0, 0.216, 0.382, 0.557, 1.02, 11.12))
-  ) +
-  # Add region boundaries
-  geom_sf(data = region, aes(color = factor(name_region)), fill = NA, linewidth = 1) +
-  scale_color_manual(values = c("black", "red")) +
-  guides(color = "none") +
-  # Add titles
-  labs(
-    caption = "Source: Based on Pellegrina and Sotelo (2021)"
-  )
-
-# Final plot - 2010
-ggplot(data = df_merged) +
-  ggtitle("(c) 2010") +
-  theme_minimal() +
-  theme(
-    plot.title = element_text(hjust = 0.5), # Center title while keeping minimal theme
-    panel.grid = element_line(color = "gray90"), # Make grid lines lighter
-    axis.text = element_blank(),
-    axis.ticks = element_blank()
-  ) +
-  # Add population share fill
-  geom_sf(aes(fill = pop_share_2010), color = "white", size = 0.1) +
-  # Color scales
-  scale_fill_stepsn(
-    name = "Pop. Share",
-    breaks = c(0, 0.216, 0.382, 0.557, 1.02, 11.12),
-    labels = scales::number_format(accuracy = 0.001),
-    colors = c("#eff3ff", "#bdd7e7", "#6baed6", "#3182bd", "#08519c"), # Made lower values darker
-    values = scales::rescale(c(0, 0.216, 0.382, 0.557, 1.02, 11.12))
-  ) +
-  # Add region boundaries
-  geom_sf(data = region, aes(color = factor(name_region)), fill = NA, linewidth = 1) +
-  scale_color_manual(values = c("black", "red")) +
-  guides(color = "none") +
-  # Add titles
-  labs(
-    caption = "Source: Based on Pellegrina and Sotelo (2021)"
-  )
-```
-
-In comparison the third paper map looked as follows (left 1950, right 2010):
-
-<table>
-<tr>
-<td><img src="data/paper3/c1960.jpg" width="400"/></td>
-<td><img src="data/paper3/c2010.jpg" width="400"/></td>
-</tr>
-</table>
-
-<hr>
-
-# Paper 4
-
-## Data loading
-
-## Data wrangling
-
-## Plotting
-
-In comparison the fourth paper map looked as follows:
-
-![Figure x]()
-
-<hr>
-
-# Paper 5
-
-## Data loading
-
-## Data wrangling
-
-## Plotting
-
-In comparison the fifth paper map looked as follows:
-
-![Figure x]()

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgANG.m

@@ -1,0 +1,65 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [146.1000187 0;0 0.003625];
+
+model = recsmodel('ang.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [20 0.095];
+
+smax = [88 0.33]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 48.77166;
+for i=1:100;
+    x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [40.11938208 0.114894167]; 
+eH = [-5.38245032890652 2.89050536705724 -10.7726267963689 -6.72643017301705 -3.14596993106116 9.76518141387712 13.3764075758057 21.0491387518697 -19.6756289333851];
+ePw = [-0.0174925 0.00627166666666668 0.0462858333333333 0.049965 -0.0114908333333333 -0.0284875 0.093695 0.0270741666666667 0.0299375];
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [41.49750903];
+foil = ones(10,1);
+Hfoil = foil*Hmean;
+Spct = zeros(10,1);
+for i=1:10;
+    for j=1:1;
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgANGcf.m

@@ -1,0 +1,65 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [146.1000187 0;0 0.003625];
+
+model = recsmodel('angCF.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [20 0.095];
+
+smax = [88 0.33]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 48.77166;
+for i=1:100;
+    x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [40.11938208 0.114894167]; 
+eH = [-5.38245032890652 2.89050536705724 -10.7726267963689 -6.72643017301705 -3.14596993106116 9.76518141387712 13.3764075758057 21.0491387518697 -19.6756289333851];
+ePw = [-0.0174925 0.00627166666666668 0.0462858333333333 0.049965 -0.0114908333333333 -0.0284875 0.093695 0.0270741666666667 0.0299375];
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [41.49750903];
+foil = ones(10,1);
+Hfoil = foil*Hmean;
+Spct = zeros(10,1);
+for i=1:10;
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgBEN.m

@@ -1,0 +1,65 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [194.5240795 0;0 0.003625];
+
+model = recsmodel('ben.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [100 0.16];
+
+smax = [170 0.48]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 179.4426;
+for i=1:100;
+    x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [128.6608613 0.181898064]; 
+eH = [1.5974703760709 1.82590720412023 -27.1875098199807 -20.9027080893106 13.2034296897363 7.69897535551846 1.90699382781085 16.559955742186 7.29856850560711];
+ePw = [0.0291236477727499 0.0163274099604167 0.0280011391235834 0.2134779873375 -0.0512021188176666 -0.0446666725265832 0.061203731954 0.03449151183975 -0.0167508326010001];
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [130.6619441];
+foil = ones(10,1);
+Hfoil = foil*Hmean;
+Spct = zeros(10,1);
+for i=1:10;
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgBENcf.m

@@ -1,0 +1,65 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [194.5240795 0;0 0.003625];
+
+model = recsmodel('bencf.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [100 0.16];
+
+smax = [170 0.48]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 179.4426;
+for i=1:100;
+    x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [128.6608613 0.181898064]; 
+eH = [1.5974703760709 1.82590720412023 -27.1875098199807 -20.9027080893106 13.2034296897363 7.69897535551846 1.90699382781085 16.559955742186 7.29856850560711];
+ePw = [0.0291236477727499 0.0163274099604167 0.0280011391235834 0.2134779873375 -0.0512021188176666 -0.0446666725265832 0.061203731954 0.03449151183975 -0.0167508326010001];
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [130.6619441];
+foil = ones(10,1);
+Hfoil = foil*Hmean;
+Spct = zeros(10,1);
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgBOT.m

@@ -1,0 +1,67 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [10.02662176 0;0 0.003625];
+
+model = recsmodel('bot.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [0.5 0.12];
+
+smax = [15 0.32]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 48.77166;
+for i=1:100;
+    x(i,3) = Bee - s(i,1);
+    x(i,2) = 1;
+    %x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [0.891082734 0.143706768]; 
+eH = [-1.57217470864971 2.9408136497872 2.36556567580922 -4.50472304776463 -0.994327556522853 4.66354223354036 -0.26321271625975 2.74620159880615 -0.641261894819444];
+ePw = [-0.0136010320483176 0.00780088449784161 0.063813200370607 0.0479755741607352 -0.0452850827431783 -0.0233127167468322 0.00382795845454759 0.108078194526022 -0.0242278676752766];
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [5.631505968];
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgBOTcf.m

@@ -1,0 +1,67 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [10.02662176 0;0 0.003625];
+
+model = recsmodel('botcf.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [0.5 0.12];
+
+smax = [15 0.32]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 48.77166;
+for i=1:100;
+    x(i,3) = Bee - s(i,1);
+    x(i,2) = s(i,2) + 0.0183;
+    %x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [0.891082734 0.143706768]; 
+eH = [-1.57217470864971 2.9408136497872 2.36556567580922 -4.50472304776463 -0.994327556522853 4.66354223354036 -0.26321271625975 2.74620159880615 -0.641261894819444];
+ePw = [-0.0136010320483176 0.00780088449784161 0.063813200370607 0.0479755741607352 -0.0452850827431783 -0.0233127167468322 0.00382795845454759 0.108078194526022 -0.0242278676752766];
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgCAMCHA.m

@@ -1,0 +1,67 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0 0];
+
+vars = [128.5257403 0 0; 0 741.6248527 0;0 0 0.003625];
+%vars = [530.8026913 389.7811979 0; 389.7811979 1465.625944 0;0 0 0.003625];
+
+model = recsmodel('camcha.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [50 90 0.15];
+
+smax = [90 200 0.45]; 
+
+[interp,s] = recsinterpinit(5,smin,smax)
+
+%interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+%[interp,x] = recsSolveREE(interp,model,s)
+
+%interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(125,8);
+epsilon = -0.066;
+BeeC = 90.72067;
+BeeI = 115.9312;
+for i=1:125;
+    x(i,3) = s(i,3)+0.426;
+    x(i,4) = x(i,3) - 0.2193;
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [54.72082926 137.4352057 0.157151329];
+eHc = [-15.0168943427886 -11.5136524498511 -2.13289627840793 0.0337887842301967 2.04944297260081 13.6418794652855 14.2526178575343 8.08832225810818 7.36199309650267];
+eHi = [-41.2964631836355 9.79771994646447 12.5315611873464 -10.296554849379 4.17768788094676 -6.27896399659991 -0.17265676874365 -30.3924107487434 60.2216384596251];
+ePw = [0.0116004179443333 0.0140957795316667 0.0366645064868333 0.158554876936333 -0.053077738365 -0.0284232998008333 0.082961802721 0.0225358339596667 -0.0091702535005001];
+shocks = zeros (1,3,9);
+shocks(1,1,1:9) = eHc(1,1:9);
+shocks(1,2,1:9) = eHi(1,1:9);
+shocks(1,3,1:9) = ePw(1,1:9);
+shocks
+
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgCAMCHAcf.m

@@ -1,0 +1,67 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0 0];
+
+vars = [128.5257403 0 0; 0 741.6248527 0;0 0 0.003625];
+%vars = [530.8026913 389.7811979 0; 389.7811979 1465.625944 0;0 0 0.003625];
+
+model = recsmodel('camchacf.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [50 90 0.15];
+
+smax = [90 200 0.45]; 
+
+[interp,s] = recsinterpinit(5,smin,smax)
+
+%interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+%[interp,x] = recsSolveREE(interp,model,s)
+
+%interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(125,8);
+epsilon = -0.066;
+BeeC = 90.72067;
+BeeI = 115.9312;
+for i=1:125;
+    x(i,3) = s(i,3)+0.0618;
+    x(i,4) = x(i,3) - 0.0717;
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [54.72082926 137.4352057 0.157151329];
+eHc = [-15.0168943427886 -11.5136524498511 -2.13289627840793 0.0337887842301967 2.04944297260081 13.6418794652855 14.2526178575343 8.08832225810818 7.36199309650267];
+eHi = [-41.2964631836355 9.79771994646447 12.5315611873464 -10.296554849379 4.17768788094676 -6.27896399659991 -0.17265676874365 -30.3924107487434 60.2216384596251];
+ePw = [0.0116004179443333 0.0140957795316667 0.0366645064868333 0.158554876936333 -0.053077738365 -0.0284232998008333 0.082961802721 0.0225358339596667 -0.0091702535005001];
+shocks = zeros (1,3,9);
+shocks(1,1,1:9) = eHc(1,1:9);
+shocks(1,2,1:9) = eHi(1,1:9);
+shocks(1,3,1:9) = ePw(1,1:9);
+shocks
+
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgCDIMAL.m

@@ -1,0 +1,69 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0 0];
+
+vars = [12.44546253 0 0; 0 1795.516965 0;0 0 0.003625];
+%vars = [530.8026913 389.7811979 0; 389.7811979 1465.625944 0;0 0 0.003625];
+
+model = recsmodel('cdimal.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [50 150 0.20];
+
+smax = [80 305 0.55]; 
+
+[interp,s] = recsinterpinit(5,smin,smax)
+
+%interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+%[interp,x] = recsSolveREE(interp,model,s)
+
+%interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(125,8);
+epsilon = -0.066;
+BeeC = 101.4192;
+BeeI = 234.6796;
+for i=1:125;
+    x(i,3) = s(i,3) + 0.3935;
+    x(i,4) = x(i,3) - 0.1029;
+    x(i,5) = BeeC - s(i,1);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [60.46111128 230.4552237 0.203518126];
+eHc = [3.17329271473682 4.28551810170993 3.6834953142989 -7.16673241483168 -0.0774647856537811 -0.495041546140804 -0.240539733526546 -3.16896907773872 -2.0329776391421];
+eHi = [-63.1231925241498 -44.1008348260567 -35.7219108790908 -26.8740006384048 20.4854048787661 58.2739668278193 49.1246764008431 16.9053296504052 39.0121603037145];
+ePw = [0.0447171252383333 0.0179765587416667 0.0202460203258334 0.263550427398333 -0.048624444575 -0.0614860167958332 0.0443112709549999 0.0460966206816667 -0.0230218451775001];
+
+shocks = zeros (1,3,9);
+shocks(1,1,1:9) = eHc(1,1:9);
+shocks(1,2,1:9) = eHi(1,1:9);
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgCDIMALcf.m

@@ -1,0 +1,69 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0 0];
+
+vars = [12.44546253 0 0; 0 1795.516965 0;0 0 0.003625];
+%vars = [530.8026913 389.7811979 0; 389.7811979 1465.625944 0;0 0 0.003625];
+
+model = recsmodel('cdimalCF.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [50 150 0.20];
+
+smax = [80 305 0.55]; 
+
+[interp,s] = recsinterpinit(5,smin,smax)
+
+%interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+%[interp,x] = recsSolveREE(interp,model,s)
+
+%interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(125,8);
+epsilon = -0.066;
+BeeC = 101.4192;
+BeeI = 234.6796;
+for i=1:125;
+    x(i,3) = s(i,3) + 0.06755;
+    x(i,4) = x(i,3) - 0.0428;
+    x(i,5) = BeeC - s(i,1);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [60.46111128 230.4552237 0.203518126];
+eHc = [3.17329271473682 4.28551810170993 3.6834953142989 -7.16673241483168 -0.0774647856537811 -0.495041546140804 -0.240539733526546 -3.16896907773872 -2.0329776391421];
+eHi = [-63.1231925241498 -44.1008348260567 -35.7219108790908 -26.8740006384048 20.4854048787661 58.2739668278193 49.1246764008431 16.9053296504052 39.0121603037145];
+ePw = [0.0447171252383333 0.0179765587416667 0.0202460203258334 0.263550427398333 -0.048624444575 -0.0614860167958332 0.0443112709549999 0.0460966206816667 -0.0230218451775001];
+
+shocks = zeros (1,3,9);
+shocks(1,1,1:9) = eHc(1,1:9);
+shocks(1,2,1:9) = eHi(1,1:9);
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgCON.m

@@ -1,0 +1,67 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [0.026659665 0;0 0.003625];
+
+model = recsmodel('cng.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [2 0.20];
+
+smax = [4 0.65]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 40.57946;
+for i=1:100;
+    x(i,3) = Bee - s(i,1);
+    x(i,2) = 1;
+    %x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [2.894377089 0.211493333]; 
+eH = [0.0151766592854345 -0.056529520100522 -0.109471414273096 -0.165577049975948 -0.111736768219223 -0.0868415627565082 -0.0697252628888188 0.217256943174716 0.353302227985923];
+ePw = [0.05713 0.02316 0.0173383333333333 0.195530833333333 0.126545833333333 -0.114135833333333 -0.0185025 0.0790775000000001 -0.0014791666666667];
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [2.880231341];
+foil = ones(10,1);
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgCONcf.m

@@ -1,0 +1,67 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [0.026659665 0;0 0.003625];
+
+model = recsmodel('cngCF.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [2 0.20];
+
+smax = [4 0.65]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 40.57946;
+for i=1:100;
+    x(i,3) = Bee - s(i,1);
+    x(i,2) = 1;
+    %x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [2.894377089 0.211493333]; 
+eH = [0.0151766592854345 -0.056529520100522 -0.109471414273096 -0.165577049975948 -0.111736768219223 -0.0868415627565082 -0.0697252628888188 0.217256943174716 0.353302227985923];
+ePw = [0.05713 0.02316 0.0173383333333333 0.195530833333333 0.126545833333333 -0.114135833333333 -0.0185025 0.0790775000000001 -0.0014791666666667];
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [2.880231341];
+foil = ones(10,1);
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgDJI2.m

@@ -1,0 +1,65 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0];
+
+vars = [0.003625];
+
+model = recsmodel('dji2.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [0 0.20];
+
+smax = [40 0.65]; 
+
+[interp,s] = recsinterpinit(4,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(16,4);
+epsilon = -0.066;
+Bee = 45.0233;
+for i=1:16;
+    x(i,3) = Bee;
+    x(i,2) = s(i,2) + 0.2736706;
+    %x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [0 0.221889167]; 
+ePw = [0.0578383333333333 0.0195141666666667 0.0137408333333334 0.305150833333333 -0.04686 -0.0745858333333332 0.0289975 0.0554316666666668 -0.0285100000000001];
+shocks = zeros (1,1,9);
+shocks(1,1,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [0];
+foil = ones(10,1);
+Hfoil = foil*Hmean;
+Spct = zeros(10,1);
+for i=1:10;
+    for j=1:1;
+        Spct(i,j) = SPMX(i,j)/Hfoil(i,j);
+    end;
+end;
+Spct
+
+

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgDJI2cf.m

@@ -1,0 +1,65 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0];
+
+vars = [0.003625];
+
+model = recsmodel('dji2cf.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [0 0.20];
+
+smax = [40 0.65]; 
+
+[interp,s] = recsinterpinit(4,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(16,4);
+epsilon = -0.066;
+Bee = 45.0233;
+for i=1:16;
+    x(i,3) = Bee;
+    x(i,2) = s(i,2) + 0.05;
+    %x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [0 0.221889167]; 
+ePw = [0.0578383333333333 0.0195141666666667 0.0137408333333334 0.305150833333333 -0.04686 -0.0745858333333332 0.0289975 0.0554316666666668 -0.0285100000000001];
+shocks = zeros (1,1,9);
+shocks(1,1,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [0];
+foil = ones(10,1);
+Hfoil = foil*Hmean;
+Spct = zeros(10,1);
+for i=1:10;
+    for j=1:1;
+        Spct(i,j) = SPMX(i,j)/Hfoil(i,j);
+    end;
+end;
+Spct
+
+

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgDRCCAR.m

@@ -1,0 +1,67 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0 0];
+
+vars = [2.238395981 0 0; 0 3.634386845 0;0 0 0.003625];
+%vars = [530.8026913 389.7811979 0; 389.7811979 1465.625944 0;0 0 0.003625];
+
+model = recsmodel('drccar.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [20 30 0.14];
+
+smax = [35 50 0.40]; 
+
+[interp,s] = recsinterpinit(5,smin,smax)
+
+%interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+%[interp,x] = recsSolveREE(interp,model,s)
+
+%interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(125,8);
+epsilon = -0.066;
+BeeC = 32.12672;
+BeeI = 32.12672;
+for i=1:125;
+    x(i,3) = (s(i,1)/BeeC)^(1/epsilon);
+    x(i,4) = (s(i,2)/BeeI)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [26.60775731 36.03684528 0.142894053];
+eHc = [1.57256145217026 0.843848462185989 0.145820748164645 -0.528271522854595 -1.17909110882249 -1.80822218364987 -2.41683091215977 0.217140774033556 0.82113801801508];
+eHi = [-2.50426214207138 0.686540151869742 -1.01344886051591 0.0413351071192665 1.02330343620994 1.44881023929185 0.369668054944007 1.98569638041047 1.7772644137435];
+ePw = [0.00413730920899999 0.0111668581686667 0.0378952151943333 0.092158188413 0.0285199779346667 -0.053313217566 0.061173836561 0.0421476656546667 0.0208311781];
+shocks = zeros (1,3,9);
+shocks(1,1,1:9) = eHc(1,1:9);
+shocks(1,2,1:9) = eHi(1,1:9);
+shocks(1,3,1:9) = ePw(1,1:9);
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgDRCCARcf.m

@@ -1,0 +1,69 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0 0];
+
+vars = [2.238395981 0 0; 0 3.634386845 0;0 0 0.003625];
+%vars = [530.8026913 389.7811979 0; 389.7811979 1465.625944 0;0 0 0.003625];
+
+model = recsmodel('drccarCF.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [20 30 0.14];
+
+smax = [35 50 0.40]; 
+
+[interp,s] = recsinterpinit(5,smin,smax)
+
+%interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+%[interp,x] = recsSolveREE(interp,model,s)
+
+%interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(125,8);
+epsilon = -0.066;
+BeeC = 32.12672;
+BeeI = 32.12672;
+for i=1:125;
+    x(i,3) = s(i,3)+0.0991;
+    x(i,4) = s(i,3)+0.0991+0.0337;
+    %x(i,3) = (s(i,1)/BeeC)^(1/epsilon);
+    %x(i,4) = (s(i,2)/BeeI)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [26.60775731 36.03684528 0.142894053];
+eHc = [1.57256145217026 0.843848462185989 0.145820748164645 -0.528271522854595 -1.17909110882249 -1.80822218364987 -2.41683091215977 0.217140774033556 0.82113801801508];
+eHi = [-2.50426214207138 0.686540151869742 -1.01344886051591 0.0413351071192665 1.02330343620994 1.44881023929185 0.369668054944007 1.98569638041047 1.7772644137435];
+ePw = [0.00413730920899999 0.0111668581686667 0.0378952151943333 0.092158188413 0.0285199779346667 -0.053313217566 0.061173836561 0.0421476656546667 0.0208311781];
+shocks = zeros (1,3,9);
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgERI.m

@@ -1,0 +1,65 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [299.186426 0;0 0.003625];
+
+model = recsmodel('eri.yaml',struct('Mu',means,'Sigma',vars,'order',2))
+
+smin = [10 0.095];
+
+smax = [90 0.30]; 
+
+[interp,s] = recsinterpinit(4,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(16,4);
+epsilon = -0.066;
+Bee = 38.25806;
+for i=1:16;
+    x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [14.32320343 0.118380833]; 
+eH = [-10.4883613543722 15.2360127704026 21.363196281234 35.4150282409435 -10.0203778661543 -12.0016204771137 -11.5642017433312 -10.0095125698607 -9.60237918035542];
+ePw = [-0.0185508333333333 0.0134516666666667 0.0520058333333333 0.0526841666666667 -0.0649283333333333 0.0219691666666667 0.0969625 -0.00720416666666668 -0.00063249999999998];
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [22.65098753];
+foil = ones(10,1);
+Hfoil = foil*Hmean;
+Spct = zeros(10,1);
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgERIcf.m

@@ -1,0 +1,66 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [299.186426 0;0 0.003625];
+
+model = recsmodel('eriCF.yaml',struct('Mu',means,'Sigma',vars,'order',2))
+
+smin = [10 0.095];
+
+smax = [90 0.30]; 
+
+[interp,s] = recsinterpinit(4,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(16,4);
+epsilon = -0.066;
+Bee = 38.25806;
+for i=1:16;
+    x(i,3) = Bee - s(i,1);
+    x(i,2) = s(i,2) + 0.0555;
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [14.32320343 0.118380833]; 
+eH = [-10.4883613543722 15.2360127704026 21.363196281234 35.4150282409435 -10.0203778661543 -12.0016204771137 -11.5642017433312 -10.0095125698607 -9.60237918035542];
+ePw = [-0.0185508333333333 0.0134516666666667 0.0520058333333333 0.0526841666666667 -0.0649283333333333 0.0219691666666667 0.0969625 -0.00720416666666668 -0.00063249999999998];
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [22.65098753];
+foil = ones(10,1);
+Hfoil = foil*Hmean;
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgET.m

@@ -1,0 +1,66 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [466.1535147 0;0 0.003625];
+
+model = recsmodel('et.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [100 0.095];
+
+smax = [200 0.34]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 142.9779;
+for i=1:100;
+    x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [112.3606547 0.130893]; 
+eH = [-28.729000439806 1.04229436763237 -5.07195357944016 -11.0970550127289 -4.93082846715816 3.12823425298657 8.79545120756487 29.1032580163898 37.282690753163];
+ePw = [-0.015279164 0.009630849 0.055746383 0.077144944 -0.088145426 0.016735859 0.099200342 -0.010677295 0.008859134];
+
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [141.8837458];
+foil = ones(10,1);
+Hfoil = foil*Hmean;
+Spct = zeros(10,1);
+for i=1:10;
+    for j=1:1;
+        Spct(i,j) = SPMX(i,j)/Hfoil(i,j);
+    end;
+end;
+Spct
+
+

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgETcf.m

@@ -1,0 +1,66 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [466.1535147 0;0 0.003625];
+
+model = recsmodel('etCF.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [100 0.095];
+
+smax = [200 0.34]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 142.9779;
+for i=1:100;
+    x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [112.3606547 0.130893]; 
+eH = [-28.729000439806 1.04229436763237 -5.07195357944016 -11.0970550127289 -4.93082846715816 3.12823425298657 8.79545120756487 29.1032580163898 37.282690753163];
+ePw = [-0.015279164 0.009630849 0.055746383 0.077144944 -0.088145426 0.016735859 0.099200342 -0.010677295 0.008859134];
+
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [141.8837458];
+foil = ones(10,1);
+Hfoil = foil*Hmean;
+Spct = zeros(10,1);
+for i=1:10;
+    for j=1:1;
+        Spct(i,j) = SPMX(i,j)/Hfoil(i,j);
+    end;
+end;
+Spct
+
+

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgGAB.m

@@ -1,0 +1,67 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [0.003655435 0;0 0.003625];
+
+model = recsmodel('gab.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [0.45 0.20];
+
+smax = [0.9 0.65]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 40.57946;
+for i=1:100;
+    x(i,3) = Bee - s(i,1);
+    x(i,2) = 1;
+    %x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [0.582207293 0.221889167]; 
+eH = [-0.0483883198303723 0.0192332196782322 -0.0404268114708596 -0.10121731751947 -0.0489697624655078 0.0228338842318079 0.0342028593457993 0.0893133427227629 0.0788223504598198];
+ePw = [0.0578383333333333 0.0195141666666667 0.0137408333333334 0.305150833333333 -0.04686 -0.0745858333333332 0.0289975 0.0554316666666668 -0.0285100000000001];
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [0.587610738];
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgGABcf.m

@@ -1,0 +1,67 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [0.003655435 0;0 0.003625];
+
+model = recsmodel('gabCF.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [0.45 0.20];
+
+smax = [0.9 0.65]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 40.57946;
+for i=1:100;
+    x(i,3) = Bee - s(i,1);
+    x(i,2) = s(i,2) + 0.05;
+    %x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [0.582207293 0.221889167]; 
+eH = [-0.0483883198303723 0.0192332196782322 -0.0404268114708596 -0.10121731751947 -0.0489697624655078 0.0228338842318079 0.0342028593457993 0.0893133427227629 0.0788223504598198];
+ePw = [0.0578383333333333 0.0195141666666667 0.0137408333333334 0.305150833333333 -0.04686 -0.0745858333333332 0.0289975 0.0554316666666668 -0.0285100000000001];
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgGAM.m

@@ -1,0 +1,66 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [1007.128785 0;0 0.003625];
+
+model = recsmodel('gam.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [90 0.18];
+
+smax = [220 0.52]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 170.6498;
+for i=1:100;
+    x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [146.0422618 0.191365033]; 
+eH = [13.6757208138499 -0.11111458506582 -1.0170201595399 -42.9627107432371 1.51429523942016 34.8248828022656 56.2190023442936 -43.7172542325849 -26.149508453756];
+ePw = [0.0357849674578999 0.0172258158030184 0.0247146387439521 0.234178480709182 -0.0506242218722442 -0.0504898341981927 0.0525653695675146 0.0388927226521032 -0.0197031240446768];
+
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [138.3185548];
+foil = ones(10,1);
+Hfoil = foil*Hmean;
+Spct = zeros(10,1);
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgGAMcf.m

@@ -1,0 +1,66 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [1007.128785 0;0 0.003625];
+
+model = recsmodel('gamCF.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [90 0.18];
+
+smax = [220 0.52]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 170.6498;
+for i=1:100;
+    x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [146.0422618 0.191365033]; 
+eH = [13.6757208138499 -0.11111458506582 -1.0170201595399 -42.9627107432371 1.51429523942016 34.8248828022656 56.2190023442936 -43.7172542325849 -26.149508453756];
+ePw = [0.0357849674578999 0.0172258158030184 0.0247146387439521 0.234178480709182 -0.0506242218722442 -0.0504898341981927 0.0525653695675146 0.0388927226521032 -0.0197031240446768];
+
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [138.3185548];
+foil = ones(10,1);
+Hfoil = foil*Hmean;
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgGB.m

@@ -1,0 +1,66 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [182.9731889 0;0 0.003625];
+
+model = recsmodel('gb.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [75 0.18];
+
+smax = [145 0.52]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 170.6498;
+for i=1:100;
+    x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [80.7728294 0.191365033]; 
+eH = [-7.3152064601427 16.889877201007 21.4502574393981 -13.3270894077152 1.46462280241045 -1.97997824847762 7.6548507022527 -3.10912202080334 2.54982770428506];
+ePw = [0.0357849674578999 0.0172258158030184 0.0247146387439521 0.234178480709182 -0.0506242218722442 -0.0504898341981927 0.0525653695675146 0.0388927226521032 -0.0197031240446768];
+
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [105.0508691];
+foil = ones(10,1);
+Hfoil = foil*Hmean;
+Spct = zeros(10,1);
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgGBcf.m

@@ -1,0 +1,66 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [182.9731889 0;0 0.003625];
+
+model = recsmodel('gbCF.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [75 0.18];
+
+smax = [145 0.52]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 170.6498;
+for i=1:100;
+    x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [80.7728294 0.191365033]; 
+eH = [-7.3152064601427 16.889877201007 21.4502574393981 -13.3270894077152 1.46462280241045 -1.97997824847762 7.6548507022527 -3.10912202080334 2.54982770428506];
+ePw = [0.0357849674578999 0.0172258158030184 0.0247146387439521 0.234178480709182 -0.0506242218722442 -0.0504898341981927 0.0525653695675146 0.0388927226521032 -0.0197031240446768];
+
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [105.0508691];
+foil = ones(10,1);
+Hfoil = foil*Hmean;
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgGHABUR.m

@@ -1,0 +1,68 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0 0];
+
+vars = [160.076862 0 0; 0 538.8641421 0;0 0 0.003625];
+%vars = [530.8026913 389.7811979 0; 389.7811979 1465.625944 0;0 0 0.003625];
+
+model = recsmodel('ghabur.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [65 140 0.15];
+
+smax = [125 250 0.45]; 
+
+[interp,s] = recsinterpinit(5,smin,smax)
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(125,8);
+epsilon = -0.066;
+BeeC = 102.4635;
+BeeI = 170.4942;
+for i=1:125;
+    x(i,3) = (s(i,1)/BeeC)^(1/epsilon);
+    x(i,4) = (s(i,2)/BeeI)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [96.28995346 220.7535362 0.166119749];
+eHc = [-10.1413287429311 -7.54063487772524 -10.4708220933721 -22.5046343593434 0.876339917002667 10.2917559744979 18.6998918298187 4.98022937777762 13.3242440588826];
+eHi = [-8.92583790316021 16.6089324741199 6.45489602709625 -14.4790444526758 25.3099619260447 -23.1030573783528 13.7172344904307 -42.2818633780921 -5.83739468155713];
+ePw = [0.0177629331156131 0.0151033501751564 0.0336481044863532 0.177077692890763 -0.0530192132309143 -0.0325713061533154 0.073676305420605 0.026101279192895 -0.0121502390877558];
+
+shocks = zeros (1,3,9);
+shocks(1,1,1:9) = eHc(1,1:9);
+shocks(1,2,1:9) = eHi(1,1:9);
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgGHABURcf.m

@@ -1,0 +1,68 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0 0];
+
+vars = [160.076862 0 0; 0 538.8641421 0;0 0 0.003625];
+%vars = [530.8026913 389.7811979 0; 389.7811979 1465.625944 0;0 0 0.003625];
+
+model = recsmodel('ghaburCF.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [65 140 0.15];
+
+smax = [125 250 0.45]; 
+
+[interp,s] = recsinterpinit(5,smin,smax)
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(125,8);
+epsilon = -0.066;
+BeeC = 102.4635;
+BeeI = 170.4942;
+for i=1:125;
+    x(i,3) = (s(i,1)/BeeC)^(1/epsilon);
+    x(i,4) = (s(i,2)/BeeI)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [96.28995346 220.7535362 0.166119749];
+eHc = [-10.1413287429311 -7.54063487772524 -10.4708220933721 -22.5046343593434 0.876339917002667 10.2917559744979 18.6998918298187 4.98022937777762 13.3242440588826];
+eHi = [-8.92583790316021 16.6089324741199 6.45489602709625 -14.4790444526758 25.3099619260447 -23.1030573783528 13.7172344904307 -42.2818633780921 -5.83739468155713];
+ePw = [0.0177629331156131 0.0151033501751564 0.0336481044863532 0.177077692890763 -0.0530192132309143 -0.0325713061533154 0.073676305420605 0.026101279192895 -0.0121502390877558];
+
+shocks = zeros (1,3,9);
+shocks(1,1,1:9) = eHc(1,1:9);
+shocks(1,2,1:9) = eHi(1,1:9);
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgGUI.m

@@ -1,0 +1,65 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [62.08155324 0;0 0.003625];
+
+model = recsmodel('gui.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [75 0.20];
+
+smax = [110 0.65]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 108.3292;
+for i=1:100;
+    x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [79.7351145 0.221889167]; 
+eH = [-7.41548997594366 -4.80066500434354 -2.2980849252784 -0.555438349042689 5.34833470316525 -1.87663494743028 -1.63159639612674 5.9194883024762 17.4129650628589];
+ePw = [0.0578383333333333 0.0195141666666667 0.0137408333333334 0.305150833333333 -0.04686 -0.0745858333333332 0.0289975 0.0554316666666668 -0.0285100000000001];
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [89.83799297];
+foil = ones(10,1);
+Hfoil = foil*Hmean;
+Spct = zeros(10,1);
+for i=1:10;
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgGUIcf.m

@@ -1,0 +1,65 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [62.08155324 0;0 0.003625];
+
+model = recsmodel('guiCF.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [75 0.20];
+
+smax = [110 0.65]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 108.3292;
+for i=1:100;
+    x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [79.7351145 0.221889167]; 
+eH = [-7.41548997594366 -4.80066500434354 -2.2980849252784 -0.555438349042689 5.34833470316525 -1.87663494743028 -1.63159639612674 5.9194883024762 17.4129650628589];
+ePw = [0.0578383333333333 0.0195141666666667 0.0137408333333334 0.305150833333333 -0.04686 -0.0745858333333332 0.0289975 0.0554316666666668 -0.0285100000000001];
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [89.83799297];
+foil = ones(10,1);
+Hfoil = foil*Hmean;
+Spct = zeros(10,1);
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgKENUGA.m

@@ -1,0 +1,67 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0 0];
+
+vars = [85.86365271 0 0; 0 190.9777979 0;0 0 0.003625];
+%vars = [530.8026913 389.7811979 0; 389.7811979 1465.625944 0;0 0 0.003625];
+
+model = recsmodel('kenuga.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [55 50 0.095];
+
+smax = [110 120 0.35]; 
+
+[interp,s] = recsinterpinit(5,smin,smax)
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(125,8);
+epsilon = -0.066;
+BeeC = 79.05389;
+BeeI = 72.09776;
+for i=1:125;
+    x(i,3) = s(i,3) + 0.1233;
+    x(i,4) = x(i,3) - 0.2044;
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [79.9541991 64.12447108 0.114894167];
+eHc = [-2.41864167438018 3.93125414210881 11.0791567733462 0.317182730921843 -16.2086903308486 -16.0186559887163 7.42656368510337 3.08611640040071 6.11351806123297];
+eHi = [-17.4005184508294 -11.9704273492269 -13.5144492782303 -14.7117102160836 12.9501253853058 12.372329497508 10.6677293062116 14.3465826779035 13.8010793491937];
+ePw = [-0.0174925 0.00627166666666668 0.0462858333333333 0.049965 -0.0114908333333333 -0.0284875 0.093695 0.0270741666666667 0.0299375];
+shocks = zeros (1,3,9);
+shocks(1,1,1:9) = eHc(1,1:9);
+shocks(1,2,1:9) = eHi(1,1:9);
+shocks(1,3,1:9) = ePw(1,1:9);
+shocks
+
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgKENUGAcf.m

@@ -1,0 +1,67 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0 0];
+
+vars = [85.86365271 0 0; 0 190.9777979 0;0 0 0.003625];
+%vars = [530.8026913 389.7811979 0; 389.7811979 1465.625944 0;0 0 0.003625];
+
+model = recsmodel('kenugaCF.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [55 50 0.095];
+
+smax = [110 120 0.35]; 
+
+[interp,s] = recsinterpinit(5,smin,smax)
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(125,8);
+epsilon = -0.066;
+BeeC = 79.05389;
+BeeI = 72.09776;
+for i=1:125;
+    x(i,3) = s(i,3) + 0.1233;
+    x(i,4) = x(i,3) - 0.2044;
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [79.9541991 64.12447108 0.114894167];
+eHc = [-2.41864167438018 3.93125414210881 11.0791567733462 0.317182730921843 -16.2086903308486 -16.0186559887163 7.42656368510337 3.08611640040071 6.11351806123297];
+eHi = [-17.4005184508294 -11.9704273492269 -13.5144492782303 -14.7117102160836 12.9501253853058 12.372329497508 10.6677293062116 14.3465826779035 13.8010793491937];
+ePw = [-0.0174925 0.00627166666666668 0.0462858333333333 0.049965 -0.0114908333333333 -0.0284875 0.093695 0.0270741666666667 0.0299375];
+shocks = zeros (1,3,9);
+shocks(1,1,1:9) = eHc(1,1:9);
+shocks(1,2,1:9) = eHi(1,1:9);
+shocks(1,3,1:9) = ePw(1,1:9);
+shocks
+
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgLES.m

@@ -1,0 +1,65 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [222.1119224 0;0 0.003625];
+
+model = recsmodel('les.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [25 0.12];
+
+smax = [95 0.32]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 105.4909;
+for i=1:100;
+    x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [43.2282804 0.143706768]; 
+eH = [4.71221187575639 3.23622535459465 14.3062088416825 -6.81146416287482 -7.40329178803389 -8.9408605828497 26.1725973715401 -1.48789431206011 -29.362787605784];
+ePw = [-0.0136010320483176 0.00780088449784161 0.063813200370607 0.0479755741607352 -0.0452850827431783 -0.0233127167468322 0.00382795845454759 0.108078194526022 -0.0242278676752766];
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [37.64922539];
+foil = ones(10,1);
+Hfoil = foil*Hmean;
+Spct = zeros(10,1);
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgLEScf.m

@@ -1,0 +1,65 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [222.1119224 0;0 0.003625];
+
+model = recsmodel('lesCF.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [25 0.12];
+
+smax = [95 0.32]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 105.4909;
+for i=1:100;
+    x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [43.2282804 0.143706768]; 
+eH = [4.71221187575639 3.23622535459465 14.3062088416825 -6.81146416287482 -7.40329178803389 -8.9408605828497 26.1725973715401 -1.48789431206011 -29.362787605784];
+ePw = [-0.0136010320483176 0.00780088449784161 0.063813200370607 0.0479755741607352 -0.0452850827431783 -0.0233127167468322 0.00382795845454759 0.108078194526022 -0.0242278676752766];
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [37.64922539];
+foil = ones(10,1);
+Hfoil = foil*Hmean;
+Spct = zeros(10,1);
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgLIB.m

@@ -1,0 +1,67 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [137.1525115 0;0 0.003625];
+
+model = recsmodel('lib.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [25 0.22];
+
+smax = [50 0.65]; 
+
+[interp,s] = recsinterpinit(4,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(16,4);
+epsilon = -0.066;
+Bee = 108.3292;
+for i=1:16;
+    x(i,3) = Bee - s(i,1);
+    x(i,2) = 1;
+    %x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [20.48509997 0.221889167]; 
+eH = [-16.6981438047091 -8.50500171320365 -7.79493217396641 3.31377374441384 12.6280327719448 10.2662751476352 9.07299830947631 8.25808629360785 7.77804375936534];
+ePw = [0.0578383333333333 0.0195141666666667 0.0137408333333334 0.305150833333333 -0.04686 -0.0745858333333332 0.0289975 0.0554316666666668 -0.0285100000000001];
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [38.80423231];
+foil = ones(10,1);
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgLIBcf.m

@@ -1,0 +1,67 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [137.1525115 0;0 0.003625];
+
+model = recsmodel('libCF.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [25 0.22];
+
+smax = [50 0.65]; 
+
+[interp,s] = recsinterpinit(4,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(16,4);
+epsilon = -0.066;
+Bee = 108.3292;
+for i=1:16;
+    x(i,3) = Bee - s(i,1);
+    x(i,2) = 1;
+    %x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [20.48509997 0.221889167]; 
+eH = [-16.6981438047091 -8.50500171320365 -7.79493217396641 3.31377374441384 12.6280327719448 10.2662751476352 9.07299830947631 8.25808629360785 7.77804375936534];
+ePw = [0.0578383333333333 0.0195141666666667 0.0137408333333334 0.305150833333333 -0.04686 -0.0745858333333332 0.0289975 0.0554316666666668 -0.0285100000000001];
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [38.80423231];
+foil = ones(10,1);
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgMAU.m

@@ -1,0 +1,66 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [68.68127943 0;0 0.003625];
+
+model = recsmodel('mau.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [20 0.17];
+
+smax = [70 0.50]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 70.07557;
+for i=1:100;
+    x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [43.31841707 0.184000593]; 
+eH = [-15.3786538639071 0.573016291841583 -3.27689747623442 -0.361236556893836 0.290674204691484 -3.70457055022845 15.4139244724664 -5.58856998264404 8.79228924256251];
+ePw = [0.02987656012675 0.0172950267354166 0.0277474955138333 0.21273700054 -0.0534737995861667 -0.0392424838798332 0.0538756600704999 0.0325042144807501 -0.0183056167532501];
+
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [40.07839285];
+foil = ones(10,1);
+Hfoil = foil*Hmean;
+Spct = zeros(10,1);
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgMAUcf.m

@@ -1,0 +1,66 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [68.68127943 0;0 0.003625];
+
+model = recsmodel('mauCF.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [20 0.17];
+
+smax = [70 0.50]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 70.07557;
+for i=1:100;
+    x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [43.31841707 0.184000593]; 
+eH = [-15.3786538639071 0.573016291841583 -3.27689747623442 -0.361236556893836 0.290674204691484 -3.70457055022845 15.4139244724664 -5.58856998264404 8.79228924256251];
+ePw = [0.02987656012675 0.0172950267354166 0.0277474955138333 0.21273700054 -0.0534737995861667 -0.0392424838798332 0.0538756600704999 0.0325042144807501 -0.0183056167532501];
+
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [40.07839285];
+foil = ones(10,1);
+Hfoil = foil*Hmean;
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgMOZMAL.m

@@ -1,0 +1,75 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0 0];
+
+vars = [197.5867141 0 0; 0 2665.368074 0;0 0 0.003625];
+%vars = [530.8026913 389.7811979 0; 389.7811979 1465.625944 0;0 0 0.003625];
+
+model = recsmodel('mozmal.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [40 90 0.15];
+
+smax = [100 290 0.45]; 
+
+[interp,s] = recsinterpinit(5,smin,smax)
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(125,8);
+epsilon = -0.066;
+BeeC = 69.96438;
+BeeI = 168.9825;
+for i=1:125;
+    if (s(i,1)/BeeC)^(1/epsilon)> s(i,3) + 0.2709;
+        x(i,3) = s(i,3) + 0.2709;
+    else
+        x(i,3) = (s(i,1)/BeeC)^(1/epsilon);
+    end;
+    if (s(i,2)/BeeI)^(1/epsilon)> s(i,3) + 0.526;
+        x(i,4) = s(i,3) + 0.526;
+    else
+        x(i,4) = (s(i,2)/BeeI)^(1/epsilon);
+    end;
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [63.09827881 162.0624478 0.15010428];
+eHc = [-13.2207909739084 -21.1443340405551 0.61170575572261 6.36418620289193 8.14192516462224 5.94841772218984 4.54865292977378 27.744869892738 -14.1336132357321];
+eHi = [-66.3464437693415 -99.5095591946001 1.93406264749342 40.9616093251306 -7.95382697645391 51.5186781247037 33.445587537746 45.0032506019861 33.1912833815911];
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgMOZMALcf.m

@@ -1,0 +1,75 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0 0];
+
+vars = [197.5867141 0 0; 0 2665.368074 0;0 0 0.003625];
+%vars = [530.8026913 389.7811979 0; 389.7811979 1465.625944 0;0 0 0.003625];
+
+model = recsmodel('mozmalCF.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [40 90 0.15];
+
+smax = [100 290 0.45]; 
+
+[interp,s] = recsinterpinit(5,smin,smax)
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(125,8);
+epsilon = -0.066;
+BeeC = 69.96438;
+BeeI = 168.9825;
+for i=1:125;
+    if (s(i,1)/BeeC)^(1/epsilon)> s(i,3) + 0.2709;
+        x(i,3) = s(i,3) + 0.2709;
+    else
+        x(i,3) = (s(i,1)/BeeC)^(1/epsilon);
+    end;
+    if (s(i,2)/BeeI)^(1/epsilon)> s(i,3) + 0.526;
+        x(i,4) = s(i,3) + 0.526;
+    else
+        x(i,4) = (s(i,2)/BeeI)^(1/epsilon);
+    end;
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [63.09827881 162.0624478 0.15010428];
+eHc = [-13.2207909739084 -21.1443340405551 0.61170575572261 6.36418620289193 8.14192516462224 5.94841772218984 4.54865292977378 27.744869892738 -14.1336132357321];
+eHi = [-66.3464437693415 -99.5095591946001 1.93406264749342 40.9616093251306 -7.95382697645391 51.5186781247037 33.445587537746 45.0032506019861 33.1912833815911];
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgNAM.m

@@ -1,0 +1,67 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [62.96124333 0;0 0.003625];
+
+model = recsmodel('nam.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [20 0.12];
+
+smax = [40 0.32]; 
+
+[interp,s] = recsinterpinit(4,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(16,4);
+epsilon = -0.066;
+Bee = 48.77166;
+for i=1:16;
+    x(i,3) = Bee - s(i,1);
+    x(i,2) = s(i,2) + 0.2599778;
+    %x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [14.58230388 0.143706768]; 
+eH = [-12.0978010311477 -6.10637384896115 3.44561203897879 0.488372187469846 1.32930271991468 0.544938529906425 3.63425174909558 7.62367444553673 12.7521166865598];
+ePw = [-0.0136010320483176 0.00780088449784161 0.063813200370607 0.0479755741607352 -0.0452850827431783 -0.0233127167468322 0.00382795845454759 0.108078194526022 -0.0242278676752766];
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgNAMcf.m

@@ -1,0 +1,67 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [62.96124333 0;0 0.003625];
+
+model = recsmodel('namCF.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [20 0.12];
+
+smax = [40 0.32]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 48.77166;
+for i=1:100;
+    x(i,3) = Bee - s(i,1);
+    x(i,2) = s(i,2) + 0.06825;
+    %x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [14.58230388 0.143706768]; 
+eH = [-12.0978010311477 -6.10637384896115 3.44561203897879 0.488372187469846 1.32930271991468 0.544938529906425 3.63425174909558 7.62367444553673 12.7521166865598];
+ePw = [-0.0136010320483176 0.00780088449784161 0.063813200370607 0.0479755741607352 -0.0452850827431783 -0.0233127167468322 0.00382795845454759 0.108078194526022 -0.0242278676752766];
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgNIGNIG.m

@@ -1,0 +1,68 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0 0];
+
+vars = [722.7305789 0 0; 0 1583.519947 0;0 0 0.003625];
+%vars = [530.8026913 389.7811979 0; 389.7811979 1465.625944 0;0 0 0.003625];
+
+model = recsmodel('nignig.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [120 200 0.13];
+
+smax = [220 400 0.40]; 
+
+[interp,s] = recsinterpinit(5,smin,smax)
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(125,8);
+epsilon = -0.066;
+BeeC = 166.9375;
+BeeI = 251.2102;
+for i=1:125;
+    x(i,3) = (s(i,1)/BeeC)^(1/epsilon);
+    x(i,4) = (s(i,2)/BeeI)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [162.0008778 285.8241037 0.148093455];
+eHc = [11.2148263073987 18.3941972706191 32.4554808465234 18.300988833309 31.2811572618694 -30.073536373631 -14.6622665742571 -34.2786622333196 -36.8928400273604];
+eHi = [-61.3032834208047 3.97033204522432 19.0492099539521 -4.17283442749664 53.3497475290501 -45.5527747101189 55.101043290866 -42.6991959491474 5.2451085963753];
+ePw = [0.00407946 0.014449366 0.040561222 0.130317121 -0.057421422 -0.012241775 0.082682649 0.013642065 -0.007765647]; 
+
+shocks = zeros (1,3,9);
+shocks(1,1,1:9) = eHc(1,1:9);
+shocks(1,2,1:9) = eHi(1,1:9);
+shocks(1,3,1:9) = ePw(1,1:9);
+shocks
+
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgNIGNIGcf.m

@@ -1,0 +1,68 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0 0];
+
+vars = [722.7305789 0 0; 0 1583.519947 0;0 0 0.003625];
+%vars = [530.8026913 389.7811979 0; 389.7811979 1465.625944 0;0 0 0.003625];
+
+model = recsmodel('nignigCF.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [120 200 0.13];
+
+smax = [220 400 0.40]; 
+
+[interp,s] = recsinterpinit(5,smin,smax)
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(125,8);
+epsilon = -0.066;
+BeeC = 166.9375;
+BeeI = 251.2102;
+for i=1:125;
+    x(i,3) = (s(i,1)/BeeC)^(1/epsilon);
+    x(i,4) = (s(i,2)/BeeI)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [162.0008778 285.8241037 0.148093455];
+eHc = [11.2148263073987 18.3941972706191 32.4554808465234 18.300988833309 31.2811572618694 -30.073536373631 -14.6622665742571 -34.2786622333196 -36.8928400273604];
+eHi = [-61.3032834208047 3.97033204522432 19.0492099539521 -4.17283442749664 53.3497475290501 -45.5527747101189 55.101043290866 -42.6991959491474 5.2451085963753];
+ePw = [0.00407946 0.014449366 0.040561222 0.130317121 -0.057421422 -0.012241775 0.082682649 0.013642065 -0.007765647]; 
+
+shocks = zeros (1,3,9);
+shocks(1,1,1:9) = eHc(1,1:9);
+shocks(1,2,1:9) = eHi(1,1:9);
+shocks(1,3,1:9) = ePw(1,1:9);
+shocks
+
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgSEN.m

@@ -1,0 +1,66 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [721.7647653 0;0 0.003625];
+
+model = recsmodel('sen.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [55 0.18];
+
+smax = [155 0.52]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 170.6498;
+for i=1:100;
+    x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [128.1183038 0.191365033]; 
+eH = [-13.9582301220384 14.7973341901198 -23.9770042833988 -44.3753196676293 26.4687628918316 30.5706314471428 16.3032658389316 -31.9327028771692 1.28981157784781];
+ePw = [0.0357849674578999 0.0172258158030184 0.0247146387439521 0.234178480709182 -0.0506242218722442 -0.0504898341981927 0.0525653695675146 0.0388927226521032 -0.0197031240446768];
+
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [103.3048528];
+foil = ones(10,1);
+Hfoil = foil*Hmean;
+Spct = zeros(10,1);
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgSENcf.m

@@ -1,0 +1,66 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [721.7647653 0;0 0.003625];
+
+model = recsmodel('senCF.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [55 0.18];
+
+smax = [155 0.52]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 170.6498;
+for i=1:100;
+    x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [128.1183038 0.191365033]; 
+eH = [-13.9582301220384 14.7973341901198 -23.9770042833988 -44.3753196676293 26.4687628918316 30.5706314471428 16.3032658389316 -31.9327028771692 1.28981157784781];
+ePw = [0.0357849674578999 0.0172258158030184 0.0247146387439521 0.234178480709182 -0.0506242218722442 -0.0504898341981927 0.0525653695675146 0.0388927226521032 -0.0197031240446768];
+
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [103.3048528];
+foil = ones(10,1);
+Hfoil = foil*Hmean;
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgSL.m

@@ -1,0 +1,65 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [617.8060601 0;0 0.003625];
+
+model = recsmodel('sl.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [55 0.20];
+
+smax = [155 0.65]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 108.3292;
+for i=1:100;
+    x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [60.51760294 0.221889167]; 
+eH = [-25.1944579324886 -3.32967689530886 33.1623273941248 -26.0982752512659 -16.9027647945888 5.21058435405409 18.652120651439 22.0425482804664 27.5215600300516];
+ePw = [0.0578383333333333 0.0195141666666667 0.0137408333333334 0.305150833333333 -0.04686 -0.0745858333333332 0.0289975 0.0554316666666668 -0.0285100000000001];
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [95.58156878];
+foil = ones(10,1);
+Hfoil = foil*Hmean;
+Spct = zeros(10,1);
+for i=1:10;
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgSLcf.m

@@ -1,0 +1,65 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [617.8060601 0;0 0.003625];
+
+model = recsmodel('slCF.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [55 0.20];
+
+smax = [155 0.65]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 108.3292;
+for i=1:100;
+    x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [60.51760294 0.221889167]; 
+eH = [-25.1944579324886 -3.32967689530886 33.1623273941248 -26.0982752512659 -16.9027647945888 5.21058435405409 18.652120651439 22.0425482804664 27.5215600300516];
+ePw = [0.0578383333333333 0.0195141666666667 0.0137408333333334 0.305150833333333 -0.04686 -0.0745858333333332 0.0289975 0.0554316666666668 -0.0285100000000001];
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [95.58156878];
+foil = ones(10,1);
+Hfoil = foil*Hmean;
+Spct = zeros(10,1);
+for i=1:10;
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgSOM.m

@@ -1,0 +1,65 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [106.5536409 0;0 0.003625];
+
+model = recsmodel('som.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [15 0.095];
+
+smax = [70 0.30]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 29.62527;
+for i=1:100;
+    x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [48.50388904 0.11625517]; 
+eH = [12.5230399483376 10.5614789290076 -1.7391476764537 -9.50684633975598 -10.4436160370647 -9.20925704719379 -5.20743958387078 -3.75675463170796 -2.13063211563659];
+ePw = [-0.0181882382005 0.00740685223216667 0.0463336272438333 0.0504246555285 -0.0210772011803333 -0.0191734612135 0.089118332114 0.0232586729391667 0.024999651287];
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [29.59471449];
+foil = ones(10,1);
+Hfoil = foil*Hmean;
+Spct = zeros(10,1);
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgSOMcf.m

@@ -1,0 +1,65 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [106.5536409 0;0 0.003625];
+
+model = recsmodel('somCF.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [15 0.095];
+
+smax = [70 0.30]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 29.62527;
+for i=1:100;
+    x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [48.50388904 0.11625517]; 
+eH = [12.5230399483376 10.5614789290076 -1.7391476764537 -9.50684633975598 -10.4436160370647 -9.20925704719379 -5.20743958387078 -3.75675463170796 -2.13063211563659];
+ePw = [-0.0181882382005 0.00740685223216667 0.0463336272438333 0.0504246555285 -0.0210772011803333 -0.0191734612135 0.089118332114 0.0232586729391667 0.024999651287];
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [29.59471449];
+foil = ones(10,1);
+Hfoil = foil*Hmean;
+Spct = zeros(10,1);
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgSUDSSD.m

@@ -1,0 +1,67 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0 0];
+
+vars = [2890.652139 0 0; 0 494.9108411 0;0 0 0.003625];
+%vars = [530.8026913 389.7811979 0; 389.7811979 1465.625944 0;0 0 0.003625];
+
+model = recsmodel('sudssd.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [80 30 0.11];
+
+smax = [250 105 0.30]; 
+
+[interp,s] = recsinterpinit(5,smin,smax)
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(125,8);
+epsilon = -0.066;
+BeeC = 212.6926;
+BeeI = 212.6926;
+for i=1:125;
+    x(i,3) = s(i,3) + 0.3371;
+    x(i,4) = x(i,3) + 0.6963;
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [245.1359994 101.4314565 0.133134546];
+eHc = [-46.7819107852876 48.4247527641879 28.744368476335 54.5478907643609 0.424151272327549 5.91436870756007 -62.733209251124 -0.653598705506283 -95.5546029031685];
+eHi = [-19.3572439839466 20.0370129903601 11.8937372249417 22.5706221189124 0.175503726263486 2.44722535188242 -25.9575125723997 -0.270443626557835 -39.5382260180673];
+ePw = [-0.0150417288550624 0.0102199324731799 0.0562733850443301 0.0758182476149486 -0.0929801812002664 0.0230798823916407 0.0924340391487166 -0.0138293657983288 0.00844088284706429];
+shocks = zeros (1,3,9);
+shocks(1,1,1:9) = eHc(1,1:9);
+shocks(1,2,1:9) = eHi(1,1:9);
+shocks(1,3,1:9) = ePw(1,1:9);
+shocks
+
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgSUDSSDcf.m

@@ -1,0 +1,67 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0 0];
+
+vars = [2890.652139 0 0; 0 494.9108411 0;0 0 0.003625];
+%vars = [530.8026913 389.7811979 0; 389.7811979 1465.625944 0;0 0 0.003625];
+
+model = recsmodel('sudssdCF.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [80 30 0.11];
+
+smax = [250 105 0.30]; 
+
+[interp,s] = recsinterpinit(5,smin,smax)
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(125,8);
+epsilon = -0.066;
+BeeC = 212.6926;
+BeeI = 212.6926;
+for i=1:125;
+    x(i,3) = s(i,3) + 0.09105;
+    x(i,4) = x(i,3) + 0.075;
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [245.1359994 101.4314565 0.133134546];
+eHc = [-46.7819107852876 48.4247527641879 28.744368476335 54.5478907643609 0.424151272327549 5.91436870756007 -62.733209251124 -0.653598705506283 -95.5546029031685];
+eHi = [-19.3572439839466 20.0370129903601 11.8937372249417 22.5706221189124 0.175503726263486 2.44722535188242 -25.9575125723997 -0.270443626557835 -39.5382260180673];
+ePw = [-0.0150417288550624 0.0102199324731799 0.0562733850443301 0.0758182476149486 -0.0929801812002664 0.0230798823916407 0.0924340391487166 -0.0138293657983288 0.00844088284706429];
+shocks = zeros (1,3,9);
+shocks(1,1,1:9) = eHc(1,1:9);
+shocks(1,2,1:9) = eHi(1,1:9);
+shocks(1,3,1:9) = ePw(1,1:9);
+shocks
+
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgSWA.m

@@ -1,0 +1,67 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [181.2788951 0;0 0.003625];
+
+model = recsmodel('swa.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [25 0.12];
+
+smax = [70 0.32]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 105.4909;
+for i=1:100;
+    x(i,3) = Bee - s(i,1);
+    x(i,2) = 1;
+    %x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [63.67419197 0.143706768]; 
+eH = [5.60716339982707 10.8923851515468 3.44196474171748 -33.5287543856995 -4.57982443605481 -8.0152467966013 0.405601151742459 13.5363235638343 5.15268599320437];
+ePw = [-0.0136010320483176 0.00780088449784161 0.063813200370607 0.0479755741607352 -0.0452850827431783 -0.0233127167468322 0.00382795845454759 0.108078194526022 -0.0242278676752766];
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [56.58649036];
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgSWAcf.m

@@ -1,0 +1,67 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [181.2788951 0;0 0.003625];
+
+model = recsmodel('swaCF.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [30 0.12];
+
+smax = [71 0.32]; 
+
+[interp,s] = recsinterpinit(4,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(16,4);
+epsilon = -0.066;
+Bee = 105.4909;
+for i=1:16;
+    x(i,3) = Bee - s(i,1);
+    x(i,2) = s(i,2) + 0.01785;
+    %x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [63.67419197 0.143706768]; 
+eH = [5.60716339982707 10.8923851515468 3.44196474171748 -33.5287543856995 -4.57982443605481 -8.0152467966013 0.405601151742459 13.5363235638343 5.15268599320437];
+ePw = [-0.0136010320483176 0.00780088449784161 0.063813200370607 0.0479755741607352 -0.0452850827431783 -0.0233127167468322 0.00382795845454759 0.108078194526022 -0.0242278676752766];
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgTANRWB.m

@@ -1,0 +1,83 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0 0];
+
+vars = [467.1407072 0 0; 0 50.40847497 0;0 0 0.003625];
+%vars = [530.8026913 389.7811979 0; 389.7811979 1465.625944 0;0 0 0.003625];
+
+model = recsmodel('tanrwb.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [87 24 0.154];
+
+smax = [150 55 0.421]; 
+
+[interp,s] = recsinterpinit(5,smin,smax)
+
+%interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+%[interp,x] = recsSolveREE(interp,model,s)
+
+%interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(125,8);
+epsilon = -0.066;
+BeeC = 98.13187;
+BeeI = 31.56054;
+for i=1:125;
+    %x(i,3) = (s(i,1)/BeeC)^(1/epsilon);
+    %x(i,4) = (s(i,2)/BeeI)^(1/epsilon);
+    %x(i,3) = s(i,3) + 0.266572011;
+    %x(i,4) = x(i,3) + 0.5179163;
+    if (s(i,1)/BeeC)^(1/epsilon)< s(i,3) - 0.266572011;
+        x(i,3) = s(i,3) - 0.266572011;
+    elseif (s(i,1)/BeeC)^(1/epsilon)> s(i,3) + 0.266572011;
+        x(i,3) = s(i,3) + 0.266572011;
+    else
+        x(i,3) = (s(i,1)/BeeC)^(1/epsilon);
+    end;
+    if (s(i,2)/BeeI)^(1/epsilon)> x(i,3) + 0.5179163;
+        x(i,4) = x(i,3) + 0.5179163;
+    elseif (s(i,2)/BeeI)^(1/epsilon)< x(i,3) - 0.5179163;
+        x(i,4) = x(i,3) - 0.5179163;
+    else
+        x(i,4) = (s(i,2)/BeeI)^(1/epsilon);
+    end;
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgTANRWBcf.m

@@ -1,0 +1,69 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0 0];
+
+vars = [467.1407072 0 0; 0 50.40847497 0;0 0 0.003625];
+%vars = [530.8026913 389.7811979 0; 389.7811979 1465.625944 0;0 0 0.003625];
+
+model = recsmodel('tanrwbCF.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [87 24 0.31];
+
+smax = [150 50 0.42]; 
+
+[interp,s] = recsinterpinit(5,smin,smax)
+
+%interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+%[interp,x] = recsSolveREE(interp,model,s)
+
+%interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(125,8);
+epsilon = -0.066;
+BeeC = 98.13187;
+BeeI = 31.56054;
+for i=1:125;
+    %x(i,3) = (s(i,1)/BeeC)^(1/epsilon);
+    %x(i,4) = (s(i,2)/BeeI)^(1/epsilon);
+    x(i,3) = s(i,3)+0.07245
+    x(i,4) = x(i,3)-0.0776;
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [90.20472687 26.99163819 0.154956633];
+eHc = [22.06235614508 -19.1312340323112 -14.0156026751859 -9.17189636475267 30.8836523828386 -23.2458974747527 23.9137110239188 5.64456736122163 11.8965590832396];
+eHi = [-5.58533222660248 -2.2066365366183 -5.35296040070723 -6.78579203690341 -5.34901534259009 1.88004832535527 7.95048699547765 11.203929991644 9.63974722006598];
+ePw = [0.01345560734025 0.0132757465728333 0.0342804685005833 0.110335357935917 0.045756961593 -0.0640083390701666 0.0471634542422499 0.0486414518963334 0.0169081003916667];
+shocks = zeros (1,3,9);
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgTOG.m

@@ -1,0 +1,65 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [103.8375046 0;0 0.003625];
+
+model = recsmodel('tog.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [130 0.15];
+
+smax = [170 0.42]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 158.8753;
+for i=1:100;
+    x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [139.4268596 0.159549953]; 
+eH = [-9.72947457582987 -8.75378598560479 -3.27418086829576 -8.48069969467403 -3.33404328489792 11.0660496878472 4.85192487409299 2.92316738921915 21.6210882190643];
+ePw = [0.0128429813328814 0.0147940869684087 0.0361236622786162 0.160528382525422 -0.0544020405100374 -0.0257822789479778 0.0774581683755668 0.0218341661233333 -0.0104692360106737];
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [146.3169054];
+foil = ones(10,1);
+Hfoil = foil*Hmean;
+Spct = zeros(10,1);
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgTOGcf.m

@@ -1,0 +1,65 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0];
+
+vars = [103.8375046 0;0 0.003625];
+
+model = recsmodel('togCF.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [130 0.15];
+
+smax = [170 0.42]; 
+
+[interp,s] = recsinterpinit(10,smin,smax)
+%increase from 10 to ?
+
+interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+[interp,x] = recsSolveREE(interp,model,s)
+
+interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(100,4);
+epsilon = -0.066;
+Bee = 158.8753;
+for i=1:100;
+    x(i,2) = (s(i,1)/Bee)^(1/epsilon);
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [139.4268596 0.159549953]; 
+eH = [-9.72947457582987 -8.75378598560479 -3.27418086829576 -8.48069969467403 -3.33404328489792 11.0660496878472 4.85192487409299 2.92316738921915 21.6210882190643];
+ePw = [0.0128429813328814 0.0147940869684087 0.0361236622786162 0.160528382525422 -0.0544020405100374 -0.0257822789479778 0.0774581683755668 0.0218341661233333 -0.0104692360106737];
+shocks = zeros (1,2,9);
+shocks(1,1,1:9) = eH(1,1:9);
+shocks(1,2,1:9) = ePw(1,1:9);
+shocks
+
+%[ssim,xsim] = recsSimul(model,interp,s0,nper,shocks);
+[ssim,xsim] = recsSimul(model,interp,s0,10,shocks,struct('simulmethod','solve'))
+
+SPMX = zeros(10,4);
+for i=1:10;
+    for j=1:4;
+        SPMX(i,j) = xsim(1,j,i);
+    end;
+end;
+SPMX
+
+Hmean = [146.3169054];
+foil = ones(10,1);
+Hfoil = foil*Hmean;
+Spct = zeros(10,1);
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgZIMZAM.m

@@ -1,0 +1,75 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0 0];
+
+vars = [841.6561741 0 0; 0 2771.952946 0;0 0 0.003625];
+%vars = [530.8026913 389.7811979 0; 389.7811979 1465.625944 0;0 0 0.003625];
+
+model = recsmodel('zimzam.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [35 70 0.12];
+
+smax = [150 300 0.30]; 
+
+[interp,s] = recsinterpinit(5,smin,smax)
+
+%interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+%[interp,x] = recsSolveREE(interp,model,s)
+
+%interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(125,8);
+epsilon = -0.066;
+BeeC = 105.4909;
+BeeI = 105.4909;
+for i=1:125;
+    if s(i,1) < BeeC;
+        x(i,3) = s(i,3) + 0.115;
+    else
+        x(i,3) = s(i,3) - 0.115;
+    end;
+    if s(i,2) < BeeI;
+        x(i,4) = s(i,3) + 0.115+0.0807;
+    else
+        x(i,4) = (s(i,2)/BeeC)^(1/epsilon) + 0.05;
+    end;
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [83.56280226 106.2791299 0.143706768];
+eHc = [47.90802576617 -12.9164795696394 31.7606071729769 -10.161065638717 -46.1341233896352 -30.6222749193135 6.25057289157431 27.3535586829425 -12.069126174892];
+eHi = [-31.7881783106493 -64.9094613092174 -19.5266637199419 -27.6110011113892 -43.1633805682203 6.70799938058701 71.0799326786388 81.1093195090516 62.2492311737591];
[…]

### 2025-01-29 hw2/data/porteous_data/10.-Full-Rational-Expectations/afgZIMZAMcf.m

@@ -1,0 +1,75 @@
+clear all;
+
+cd 'C:\Users\Obie\Documents\MATLAB\AFGAINS'
+
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\COMPECON'))
+addpath(genpath('C:\Users\Obie\Documents\MATLAB\RECS'))
+
+means = [0 0 0];
+
+vars = [841.6561741 0 0; 0 2771.952946 0;0 0 0.003625];
+%vars = [530.8026913 389.7811979 0; 389.7811979 1465.625944 0;0 0 0.003625];
+
+model = recsmodel('zimzamCF.yaml',struct('Mu',means,'Sigma',vars,'order',4))
+
+smin = [35 70 0.12];
+
+smax = [150 300 0.30]; 
+
+[interp,s] = recsinterpinit(5,smin,smax)
+
+%interp = recsFirstGuess(interp,model,s,model.sss,model.xss,struct('fgmethod','perturbation'))
+%[interp,x] = recsSolveREE(interp,model,s)
+
+%interp = rmfield(interp,{'cx','cz'})
+
+x      = zeros(125,8);
+epsilon = -0.066;
+BeeC = 105.4909;
+BeeI = 105.4909;
+for i=1:125;
+    if s(i,1) < BeeC;
+        x(i,3) = s(i,3) + 0.1053;
+    else
+        x(i,3) = s(i,3) - 0.1053;
+    end;
+    if s(i,2) < BeeI;
+        x(i,4) = s(i,3) + 0.2805;
+    else
+        x(i,4) = (s(i,2)/BeeC)^(1/epsilon) + 0.05;
+    end;
+end;
+x
+interp = recsSolveREE(interp,model,s,x)
+
+s0 = [83.56280226 106.2791299 0.143706768];
+eHc = [47.90802576617 -12.9164795696394 31.7606071729769 -10.161065638717 -46.1341233896352 -30.6222749193135 6.25057289157431 27.3535586829425 -12.069126174892];
+eHi = [-31.7881783106493 -64.9094613092174 -19.5266637199419 -27.6110011113892 -43.1633805682203 6.70799938058701 71.0799326786388 81.1093195090516 62.2492311737591];
[…]

### 2025-01-29 hw2/data/porteous_data/3.-Inner-Loop/gamsBS.py

@@ -1,0 +1,642 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Wed Jul 16 11:03:03 2014
+
+@author: Obie
+"""
+
+from gams import *
+os.chdir("C:\Users\Obie\Documents\gamsdir\projdir\BS")
+ws = GamsWorkspace(debug = DebugLevel.ShowLog, working_directory = "C:\Users\Obie\Documents\gamsdir\projdir\BS")
+
+SEH5py = ws.add_job_from_file("themodelv6SEH5.gms")
+SEH5py.run(output=sys.stdout)
+
+W5py = ws.add_job_from_file("themodelv6W5.gms")
+W5py.run(output=sys.stdout)
+
+A5dpy = ws.add_job_from_file("themodelv6A5d.gms")
+A5dpy.run(output=sys.stdout)
+
+A6py = ws.add_job_from_file("themodelv6A6.gms")
+A6py.run(output=sys.stdout)
+
+SEH7py = ws.add_job_from_file("themodelv6SEH7.gms")
+SEH7py.run(output=sys.stdout)
+
+W7py = ws.add_job_from_file("themodelv6W7.gms")
+W7py.run(output=sys.stdout)
+
+A7cpy = ws.add_job_from_file("themodelv6A7c.gms")
+A7cpy.run(output=sys.stdout)
+
+A8py = ws.add_job_from_file("themodelv6A8.gms")
+A8py.run(output=sys.stdout)
+
+A9py = ws.add_job_from_file("themodelv6A9.gms")
+A9py.run(output=sys.stdout)
+
+SEH10py = ws.add_job_from_file("themodelv6SEH10.gms")
+SEH10py.run(output=sys.stdout)
+
+W10py = ws.add_job_from_file("themodelv6W10.gms")
+W10py.run(output=sys.stdout)
+
+A10dpy = ws.add_job_from_file("themodelv6A10d.gms")
+A10dpy.run(output=sys.stdout)
+
+A11py = ws.add_job_from_file("themodelv6A11.gms")
+A11py.run(output=sys.stdout)
+
+A12py = ws.add_job_from_file("themodelv6A12.gms")
+A12py.run(output=sys.stdout)
+
+A13py = ws.add_job_from_file("themodelv6A13.gms")
+A13py.run(output=sys.stdout)
+
+A14py = ws.add_job_from_file("themodelv6A14.gms")
+A14py.run(output=sys.stdout)
+
[…]

### 2025-01-29 hw2/data/porteous_data/4.-Outer-Loop-for-Cost-Parameter-Estimation/AARKmanipKO.m

@@ -1,0 +1,185 @@
+clear all;
+cd 'C:\Users\oporteous\Documents\Data';
+
+dataAA = importdata('AAlpha4MatlabKO.csv');
+dataRK = importdata('RK4Matlab.csv');
+
+out = zeros(41,3);
+out(1,1) = dataAA(1,1);
+out(2,1) = dataAA(9,1);
+out(3,1) = dataAA(1,1);
+out(4,1) = dataAA(10,1);
+out(5,1) = dataAA(7,1);
+out(6,1) = dataAA(12,1);
+out(7,1) = dataAA(3,1);
+out(8,1) = dataAA(19,1);
+out(9,1) = dataAA(2,1);
+out(10,1) = dataAA(13,1);
+out(11,1) = dataAA(3,1);
+out(12,1) = dataAA(20,1);
+out(13,1) = dataAA(8,1);
+out(14,1) = dataAA(27,1);
+out(15,1) = dataAA(2,1);
+out(16,1) = dataAA(5,1);
+out(17,1) = dataAA(14,1);
+out(18,1) = dataAA(4,1);
+out(19,1) = dataAA(5,1);
+out(20,1) = dataAA(21,1);
+out(21,1) = dataAA(6,1);
+out(22,1) = dataAA(4,1);
+out(23,1) = dataAA(18,1);
+out(24,1) = dataAA(22,1);
+out(25,1) = dataAA(23,1);
+out(26,1) = dataAA(16,1);
+out(27,1) = dataAA(1,1);
+out(28,1) = dataAA(11,1);
+out(29,1) = dataAA(15,1);
+out(30,1) = dataAA(7,1);
+out(31,1) = dataAA(5,1);
+out(32,1) = dataAA(4,1);
+out(33,1) = dataAA(24,1);
+out(34,1) = dataAA(25,1);
+out(35,1) = dataAA(25,1);
+out(36,1) = dataAA(6,1);
+out(37,1) = dataAA(17,1);
+out(38,1) = dataAA(26,1);
+out(39,1) = dataAA(28,1);
+out(40,1) = dataAA(6,1);
+out(41,1) = dataAA(6,1);
+
+out(1,2:3) = dataRK(2,1:2);
+out(2,2:3) = dataRK(3,1:2);
+out(3,2:3) = dataRK(1,1:2);
+out(4,2:3) = dataRK(4,1:2);
[…]

### 2025-01-29 hw2/data/porteous_data/4.-Outer-Loop-for-Cost-Parameter-Estimation/rknew2.m

@@ -1,0 +1,400 @@
+clear all;
+
+cd 'C:\Users\oporteous\Documents\Data'
+
+%processing tau file
+[at,bt,ct] = xlsread('tauest2.xlsx');
+[dt et] = size(at);
+[ft gt] = size(bt);
+tauout1 = bt(2:ft,1:4);
+tauout2 = zeros(dt,3);
+tauout2(1:dt,1) = at(1:dt,1);
+tauout2(1:dt,2) = at(1:dt,6);
+tauout2(1:dt,3) = at(1:dt,11);
+tauout3 = num2cell(tauout2);
+fid = fopen('tauest2.csv','wt');
+if fid>0
+    for k=1:size(tauout1,1)
+        fprintf(fid,'%s,%s,%s,%s,',tauout1{k,:});
+        fprintf(fid,'%f,%f,%f\n',tauout3{k,:});
+    end
+    fclose(fid);
+end
+
+%now on to rk
+[a,b,c] = xlsread('rkest4.xlsx');
+pdat = a;
+[d e] = size(b);
+pmkt = b(2:d,1);
+pcrp = b(2:d,2);
+
+sz = size(pdat);
+helper = isnan(pdat);
+for i=1:sz(1,1);
+    for j=1:132;
+        if helper(i,j)==1;
+            pdat(i,j)=0;
+        end;
+    end;
+end;
+
+pdat2 = pdat(1:sz(1,1),13:132);
+
+%id has 2 columns, market id and crop id
+id = zeros(sz(1,1),2);
+for i=1:sz(1,1);
+if strcmp(pmkt(i,1),'Gaborone')==1; id(i,1) = 110; end;
+if strcmp(pmkt(i,1),'Kalemie')==1; id(i,1) = 121; end;
+if strcmp(pmkt(i,1),'Kamina')==1; id(i,1) = 122; end;
+if strcmp(pmkt(i,1),'Kolwezi')==1; id(i,1) = 123; end;
+if strcmp(pmkt(i,1),'Lubumbashi')==1; id(i,1) = 124; end;
+if strcmp(pmkt(i,1),'Maseru')==1; id(i,1) = 130; end;
+if strcmp(pmkt(i,1),'Blantyre')==1; id(i,1) = 140; end;
+if strcmp(pmkt(i,1),'Karonga')==1; id(i,1) = 141; end;
[…]

### 2025-01-29 hw2/data/porteous_data/4.-Outer-Loop-for-Cost-Parameter-Estimation/taumanip2.m

@@ -1,0 +1,31 @@
+clear all;
+cd 'C:\Users\oporteous\Documents\Data';
+
+data = importdata('Mkts4MatlabID.csv');
+dataT = importdata('Tau4Matlab.csv');
+sz = size(dataT);
+
+tau1 = ones(233,233);
+tau9 = 9999*tau1;
+tau9(2:233,1) = data(1:232,1);
+tau9(1,2:233) = data(1:232,1)';
+%put in old data first
+for i=2:233;
+    for j=2:233;
+        for k=1:sz(1,1);
+            if dataT(k,1)==tau9(i,1);
+                if dataT(k,2)==tau9(1,j);
+                    tau9(i,j) = dataT(k,3);
+                end;
+            end;
+            if dataT(k,2)==tau9(i,1);
+                if dataT(k,1)==tau9(1,j);
+                    tau9(i,j) = dataT(k,3);
+                end;
+            end;
+        end;
+    end;
+end;
+
+tau92 = tau9(2:233,2:233);
+xlswrite('in_tau2.xlsx', tau92, 'fromMatlab','C3');  

### 2025-01-29 hw2/testcode_moritz.R
More data

@@ -1,0 +1,5 @@
+library(rnaturalearth)
+library(sf)
+
+# Get Natural Earth data with population attributes
+world <- ne_countries(scale = "medium", returnclass = "sf")

### 2025-02-02 hw2/hw2_MoritzPeistNoemiLucchiSimonVellin.Rmd
updating HTML with part 1

@@
 
-In the following notebook we replicate the charts for all 5 papers.
+In this notebook, we combine the R package world with data from various sources to produce different types of plots.
+Throughout, we implement the methodologies recently discussed in class, such as distance computation, spatial merging, and filtering.
 Contentwise, the code follows always the same structure.
@@
 library(sf)
+library(spData)
+library(units)
 ```
@@
 <hr>
+
+## Map of total population by country
+
+## Data Loading
+
+In order to plot the distribution of the total population by country, we only need the R packages world, which contains both the geometries of the countries and the corresponding population.
+
+## Data Wrangling
+
+Given that we plot the population both in absolute terms and by categories — low, medium, and high — we need to define the population ranges for these categories.
+We classify countries with a population up to the 30th percentile as "Low population" countries, those with a population between the 30th and 70th percentiles as "Medium population" countries, and those in the top 30% of the population distribution as "High population" countries.
+
+```{r}
+
+# Divide the population by categories
+world <- world %>%
+  mutate(pop_category = case_when(
+    pop <= quantile(pop, 0.30, na.rm = TRUE) ~ "Low",
+    pop <= quantile(pop, 0.70, na.rm = TRUE) ~ "Medium",
+    TRUE ~ "High"
+  ))
+
+```
+
+## Plotting
+
+```{r}
+
+# Population in absolute terms
+ggplot() +
+  geom_sf(data = world, aes(fill = pop)) +
+  scale_fill_gradient(name = "Population", low = "lightblue", high = "blue")
+
+# Population by categories 
+ggplot() +
+  geom_sf(data = world, aes(fill = pop_category), alpha = 0.7) +
+  scale_fill_manual(
+    name = "Population Category",
+    values = c("Low" = "lightyellow", "Medium" = "lightgreen", "High" = "blue"))
+
+```
+
+## Histogram of country population distribution by continent
+
+## Data Loading
+
+As for the previous plot, we only need the R package world.
+
+## Data Wrangling
+
+We start by grouping South America and North America into a single category, "America," for simplicity.
+Additionally, we remove "Seven seas (open ocean)" and "Antarctica" since these regions only contain NA values in the population column and hence are irrelevant for the analysis.
+Next, we define population ranges by labeling countries based on their population size.
+After categorizing the countries into these ranges, we calculate the number of countries in each population range, grouped by continent.
+This gives us a summary of the distribution of countries by population size within each continent.
+Finally, we ensure that the population ranges are ordered increasingly ("0-10M", "10M-50M", "50M-100M", "100M-500M", "500M+").
+
+```{r}
+
+world <- world %>%
+  mutate(
+    continent = case_when(
+      continent %in% c("South America", "North America") ~ "America",
+      continent %in% c("Seven seas (open ocean)", "Antarctica") ~ NA_character_,
+      TRUE ~ continent
+    )
+  ) %>%
+  filter(!is.na(continent))
+
+# Create the population ranges
+world <- world %>%
+  mutate(pop_range = case_when(
+    pop < 1e7 ~ "0-10M",
+    pop < 5e7 ~ "10M-50M",
+    pop < 1e8 ~ "50M-100M",
+    pop < 5e8 ~ "100M-500M",
+    TRUE ~ "500M+"
+  ))
+
+# Calculate the number of countries per range + country
+population_distribution <- world %>%
+  group_by(continent, pop_range) %>%
+  summarise(country_count = n(), .groups = "drop")
+
+# Order by ranges
+population_distribution <- population_distribution %>%
+  mutate(pop_range = factor(pop_range, levels = c("0-10M", "10M-50M", "50M-100M", "100M-500M", "500M+")))
+
+```
+
+## Plotting
+
+```{r}
+
+ggplot(population_distribution, aes(x = pop_range, y = country_count, fill = continent)) +
+  geom_histogram(stat = "identity", position = "dodge") +
+  scale_fill_brewer(palette = "Set2") +
+  labs(
+    title = "Country Population Distribution by Continent", x = "Population Range",
+    y = "Number of Countries", fill = "Continent")
+
+```
+
+## Histogram of country-level average distances between locations and airports by continent.
+
+## Data Loading
+
+In addition to the R package world, we incorporate data from [Natural Earth](https://www.naturalearthdata.com), specifically the datasets for populated places (sf.pop) and airports (sf.airports).
+
+```{r}
+
+sf.airports <- st_read("hw2/data/ne_10m_airports", crs = 4326)
+sf.pop <- st_read("hw2/data/ne_10m_populated_places_simple", crs = 4326)
+
+```
+
+## Data Wrangling
+
+This code computes and plots the average distances between populated places and airports across different continents, using two approaches: first, by computing the minimum distance for each city and averaging these values per country, and second, by directly averaging all city-airport distances within a country to allow for range-based visualization.
+Cities from the sf.pop dataset are joined with country and continent metadata from the world dataset, while airports from sf.airports are joined with country metadata.
+Distances are calculated at two levels: in "distances" we compute the minimum distance from each city to the nearest airport and average these values per country, while in "distances_country" we directly average all city-airport distances within a country.
+Then we categorize "distances_country" into ranges (e.g., <50 km, <100 km) to ensure a better visualization.
+The first plot shows the distribution of city-level average distances by continent, while the second plot shows the number of countries in each distance range, grouped by continent.
+
+```{r}
+
+# 1. Prepare cities and airports with country metadata
+cities_with_countries <- world %>%
+  st_drop_geometry() %>%
+  select(iso_a2, continent) %>%
+  left_join(
+    sf.pop %>% select(name, iso_a2, ne_id, geometry),
+    by = "iso_a2"
+  ) %>%
+  st_as_sf() %>%
+  drop_na()
+
+airports_with_countries <- sf.airports %>%
+  select(name, geometry) %>%
+  st_join(world %>% select(iso_a2, continent))
+
+# 2. Calculate closest airport distance per city, then average per country
+distances <- cities_with_countries %>%
+  # For each country group:
+  group_by(continent, iso_a2, ne_id) %>%
+  summarise(
+    avg_closest_distance_km = {
+      current_iso <- cur_group()$iso_a2 # Capture current country code
+      country_airports <- airports_with_countries %>%
+        filter(iso_a2 == current_iso) # Filter airports for this country
+      if (nrow(country_airports) == 0) {
+        NA_real_ # Return NA if no airports exist
+      } else {
+        # Compute closest airport distance for each city in the country
+        closest_distances <- st_distance(
+          geometry, # City geometries
+          country_airports # Airports in the current country
+        ) %>%
+          apply(1, min) %>% # Find minimum distance per city (row)
+          as.numeric() / 1000 # Convert meters to kilometers
+        mean(closest_distances, na.rm = TRUE)}}) %>% ungroup() %>% 
+  drop_na() # Removecountries with no airports
+
+# 3. Calculate average distances at country level. With this approach, instead of 
+# computing the closest airport distance per city and then averaging per country, 
+# we compute all the distances within a country and then we average them. 
+distances_country <- cities_with_countries %>%
+  group_by(continent, iso_a2) %>%  # Removed ne_id to group at country level
+  summarise(
+    avg_closest_distance_km = {
+      current_iso <- cur_group()$iso_a2
+      country_airports <- airports_with_countries %>%
+        filter(iso_a2 == current_iso)
+      if (nrow(country_airports) == 0) {
+        NA_real_
+      } else {
+        closest_distances <- st_distance(
+          geometry,
+          country_airports
+        ) %>%
+          apply(1, min) %>%
+          as.numeric() / 1000
+        mean(closest_distances, na.rm = TRUE)}},
+    country_name = first(name)) %>%  ungroup() %>%  drop_na()
+
+# 4. Create distance ranges – The previous step ensures a reasonable number of ranges. 
+# If we had used the distances computed earlier, the higher granularity would have resulted in an excessive number of ranges, leading to a poor visualization quality.
+distances_ranges <- distances_country %>%
+  mutate(distance_range = cut(avg_closest_distance_km,
+      breaks = seq(0, max(avg_closest_distance_km, na.rm = TRUE) + 50, by = 50),
+      labels = paste0("<", seq(50, max(avg_closest_distance_km, na.rm = TRUE) + 50, by = 50), "km"),
+      include.lowest = TRUE))
+
+```
+
+## Plotting
+
+```{r}
+
+# Plot 1 
+ggplot(distances, aes(x = avg_closest_distance_km, fill = continent)) +
+  geom_histogram(bins = 30, position = "dodge") +
+  scale_fill_brewer(palette = "Set2") +
+  labs(
+    title = "Average Distance Between Locations and Airports by Continent",
+    x = "Average Distance to nearest Airport (km)", y = "Number of Populated Places", fill = "Continent"
+  ) +
+  facet_wrap(~continent, scales = "free")
+
+# Plot 2
+ggplot(distances_ranges, aes(x = distance_range, fill = continent)) +
+  geom_bar(position = "stack") +
+  scale_fill_brewer(palette = "Set2") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  labs(title = "Average Distance Between Locations and Airport by Ranges",
+    x = "Average Distance in Ranges", y = "Number of Countries", fill = "Continent")
+
+```

### 2025-02-02 hw2/hw2_part1_noe.R

@@ -8,16 +8,15 @@ sf.airports <- st_read("hw2/data/ne_10m_airports", crs = 4326)
 sf.pop <- st_read("hw2/data/ne_10m_populated_places_simple", crs = 4326)
 world <- world
 
-# 1- TOTAL POPULATION BY COUNTRY
-# 1.1 Coloring the whole country depending on the population, using only the package
-# spData that has both the geometry of every country and the population
+# 1- TOTAL POPULATION BY COUNTRY: Using only the package spData that has both the 
+# geometry of every country and the population
 
-# Population in absolute terms
+# 1.1 Plotting population in absolute terms
 ggplot() +
   geom_sf(data = world, aes(fill = pop)) +
   scale_fill_gradient(name = "Population", low = "lightblue", high = "blue")
 
-# Population by categories
+# 1.2 Defining population by categories
 world <- world %>%
   mutate(pop_category = case_when(
     pop <= quantile(pop, 0.30, na.rm = TRUE) ~ "Low",
@@ -25,40 +24,12 @@ world <- world %>%
     TRUE ~ "High"
   ))
 
+# Plotting
 ggplot() +
   geom_sf(data = world, aes(fill = pop_category), alpha = 0.7) +
   scale_fill_manual(
     name = "Population Category",
-    values = c("Low" = "lightyellow", "Medium" = "lightgreen", "High" = "blue")
-  )
-
-
-# 1.2 Coloring cities instead of the whole country - still using colors
-# depending on the population. Used the sf.pop from Natural Earth website.
-
-# Population in absolute term
-ggplot() +
-  geom_sf(data = world) +
-  geom_sf(data = sf.pop, aes(size = pop_max, color = pop_max), alpha = 0.7) +
-  scale_size_continuous(name = "Population", range = c(0.5, 3), guide = "none") +
-  scale_color_gradient(name = "Population", low = "lightblue", high = "blue")
-
-# Population by categories
-sf.pop <- sf.pop %>%
-  mutate(pop_category = case_when(
-    pop_max <= quantile(pop_max, 0.30, na.rm = TRUE) ~ "Low",
-    pop_max <= quantile(pop_max, 0.70, na.rm = TRUE) ~ "Medium",
-    TRUE ~ "High"
-  ))
-
-ggplot() +
-  geom_sf(data = world) +
-  geom_sf(data = sf.pop, aes(size = pop_max, color = pop_category), alpha = 0.7) +
[…]

### 2025-02-03 hw2/hw2_part1_noe.R
Implemented Bruno's requirements to use natural earth world data

@@ -6,71 +6,77 @@ library(units)
 
 sf.airports <- st_read("hw2/data/ne_10m_airports", crs = 4326)
 sf.pop <- st_read("hw2/data/ne_10m_populated_places_simple", crs = 4326)
-world <- world
+sf.world <- st_read("hw2/data/ne_10m_admin_0_countries", crs = 4326) %>%
+  st_make_valid() # Repair invalid polygons
 
-# 1- TOTAL POPULATION BY COUNTRY: Using only the package spData that has both the 
+# 1- TOTAL POPULATION BY COUNTRY: Using only the package spData that has both the
 # geometry of every country and the population
 
 # 1.1 Plotting population in absolute terms
 ggplot() +
   geom_sf(data = world, aes(fill = pop)) +
-  scale_fill_gradient(name = "Population", low = "lightblue", high = "blue")
+  scale_fill_gradient(name = "POP_EST", low = "lightblue", high = "blue")
 
 # 1.2 Defining population by categories
-world <- world %>%
+sf.world <- sf.world %>%
   mutate(pop_category = case_when(
-    pop <= quantile(pop, 0.30, na.rm = TRUE) ~ "Low",
-    pop <= quantile(pop, 0.70, na.rm = TRUE) ~ "Medium",
+    POP_EST <= quantile(POP_EST, 0.30, na.rm = TRUE) ~ "Low",
+    POP_EST <= quantile(POP_EST, 0.70, na.rm = TRUE) ~ "Medium",
     TRUE ~ "High"
   ))
 
 # Plotting
 ggplot() +
-  geom_sf(data = world, aes(fill = pop_category), alpha = 0.7) +
+  geom_sf(data = sf.world, aes(fill = pop_category), alpha = 0.7) +
   scale_fill_manual(
     name = "Population Category",
-    values = c("Low" = "lightyellow", "Medium" = "lightgreen", "High" = "blue"))
+    values = c("Low" = "lightyellow", "Medium" = "lightgreen", "High" = "blue")
+  )
 
 
 # 2 - HISTOGRAM OF COUNTRY POPULATION DISTRIBUTION BY CONTINENT
 # Group South America and North America in America and remove Seven seas and Antarctica
 # because they are NA
-world <- world %>%
+sf.world <- sf.world %>%
   mutate(
-    continent = case_when(
-      continent %in% c("South America", "North America") ~ "America",
[…]

### 2025-02-03 hw2/hw2_part1_noe.R
- Fixed line issues in the second plot caused by st_make_valid() - Deleted outcommented code

@@ -3,11 +3,16 @@ library(sf)
 library(spData)
 library(tidyverse)
 library(units)
+library(lwgeom)
 
 sf.airports <- st_read("hw2/data/ne_10m_airports", crs = 4326)
 sf.pop <- st_read("hw2/data/ne_10m_populated_places_simple", crs = 4326)
 sf.world <- st_read("hw2/data/ne_10m_admin_0_countries", crs = 4326) %>%
-  st_make_valid() # Repair invalid polygons
+  # Rename columns in the world dataset to make joining / operations easier
+  rename(
+    iso_a2 = ISO_A2,
+    continent = CONTINENT
+  )
 
 # 1- TOTAL POPULATION BY COUNTRY: Using only the package spData that has both the
 # geometry of every country and the population
@@ -33,19 +38,19 @@ ggplot() +
     values = c("Low" = "lightyellow", "Medium" = "lightgreen", "High" = "blue")
   )
 
-
-# 2 - HISTOGRAM OF COUNTRY POPULATION DISTRIBUTION BY CONTINENT
+# 2 - HISTOGRAM OF COUNTRY POPULATION DISTRIBUTION BY continent
 # Group South America and North America in America and remove Seven seas and Antarctica
 # because they are NA
 sf.world <- sf.world %>%
   mutate(
-    CONTINENT = case_when(
-      CONTINENT %in% c("South America", "North America") ~ "America",
-      CONTINENT %in% c("Seven seas (open ocean)", "Antarctica") ~ NA_character_,
-      TRUE ~ CONTINENT
+    continent = case_when(
+      continent %in% c("South America", "North America") ~ "America",
+      continent %in% c("Seven seas (open ocean)", "Antarctica") ~ NA_character_,
+      TRUE ~ continent
     )
   ) %>%
-  filter(!is.na(CONTINENT))
+  filter(!is.na(continent)) %>%
+  st_make_valid() # Repair invalid polygons
 
 # Create the population ranges as factors
 sf.world <- sf.world %>%
@@ -62,38 +67,28 @@ sf.world <- sf.world %>%
 
 # Calculate the number of countries per range + country
 population_distribution <- sf.world %>%
-  group_by(CONTINENT, pop_range) %>%
+  group_by(continent, pop_range) %>%
   summarise(country_count = n(), .groups = "drop")
 
-# Order by ranges
-# population_distribution <- population_distribution %>%
[…]

### 2025-02-03 hw2/Simon_Section.R
Exercise 2 initial attempt

@@ -1,0 +1,115 @@
+library(sf)
+library(spData)
+library(tidyverse)
+library(readxl)
+library(dplyr)
+
+------------------------------------------------------------------------------------------------------
+# Question 1 - Generate an sf POINT feature of market prices places across Africa; plot them.
+------------------------------------------------------------------------------------------------------
+
+# Data Loading
+
+sf.world <- world
+sf.roads_africa <- st_read("hw2/data/ne_10m_roads")
+markets <- read_excel('hw2/data/porteous_data/1.-Price--Production--and-Population-Data/MktCoords.xlsx')
+
+# Data Wrangling
+
+sf.world <- st_transform(sf.world, CRS="EPSG:4326")
+sf.markets <- st_as_sf(markets, coords = c('longitude', 'latitude'), crs = 'EPSG:4326')
+
+# Plot
+
+ggplot() +
+  geom_sf(data = sf.world %>% filter(continent == "Africa"),
+          fill = "antiquewhite", color = "gray70") +
+  geom_sf(data = sf.markets, color = "red", size = 3, shape = 21, fill = "blue") +
+  theme_minimal() +
+  labs(title = "Markets in Africa",
+       caption = "Source: Replication Data provided") +
+  theme(legend.position = "none",
+        plot.title = element_text(hjust = 0.5),
+        plot.caption = element_text(size = 10, hjust = 0.5))
+
+------------------------------------------------------------------------------------------------------
+  # Question 2 - Calculate minimum distance of each market to the (i) coast, (ii) nearest road, and (iii) nearest airport
+------------------------------------------------------------------------------------------------------
+
+# Data Loading
+
+sf.airports <- st_read("hw2/data/ne_10m_airports")
+sf.coast <- st_read("hw2/data/ne_10m_coastline")
+
+# We isolate the African continent and its airports from our global data sets to stay efficient.
+sf.africa <- sf.world %>% filter(continent == 'Africa')
+sf.airports_africa <- sf.airports %>% st_filter(sf.africa)
+
+# Data Wrangling
+# st_distance() returns a matrix with one row per market and one column per coast segment.
+
+dist_to_coast <- st_distance(sf.markets, sf.coast)
+# For each market (row) take the minimum distance:
+min_coast_dist <- apply(dist_to_coast, 1, min)
+
[…]

### 2025-02-03 hw2/hw2_part1_noe.R
Fixed and added plot

@@ -19,8 +19,8 @@ sf.world <- st_read("hw2/data/ne_10m_admin_0_countries", crs = 4326) %>%
 
 # 1.1 Plotting population in absolute terms
 ggplot() +
-  geom_sf(data = world, aes(fill = pop)) +
-  scale_fill_gradient(name = "POP_EST", low = "lightblue", high = "blue")
+  geom_sf(data = sf.world, aes(fill = POP_EST)) +
+  scale_fill_gradient(name = "Population", low = "lightblue", high = "blue")
 
 # 1.2 Defining population by categories
 sf.world <- sf.world %>%
@@ -44,7 +44,7 @@ ggplot() +
 sf.world <- sf.world %>%
   mutate(
     continent = case_when(
-      continent %in% c("South America", "North America") ~ "America",
+      # continent %in% c("South America", "North America") ~ "America",
       continent %in% c("Seven seas (open ocean)", "Antarctica") ~ NA_character_,
       TRUE ~ continent
     )
@@ -78,6 +78,23 @@ ggplot(population_distribution, aes(x = pop_range, y = country_count, fill = con
   labs(
     title = "Country Population Distribution by continent", x = "Population Range",
     y = "Number of Countries", fill = "continent"
+  )
+
+# Create a proper histogram using the continuous population data
+ggplot(sf.world, aes(x = POP_EST, fill = continent)) +
+  geom_histogram(position = "identity", alpha = 0.5, bins = 30) +
+  scale_x_log10(labels = scales::comma) + # Use log scale due to large population range
+  facet_wrap(~continent, scales = "free_y") + # Separate panels for each continent
+  labs(
+    title = "Population Distribution by Continent",
+    x = "Population (log scale)",
+    y = "Count of Countries",
+    fill = "Continent"
+  ) +
+  theme_minimal() +
+  theme(
+    legend.position = "none", # Remove legend since we're using facets
+    plot.title = element_text(hjust = 0.5)
   )
 
 

### 2025-02-03 hw2/hw2_MoritzPeistNoemiLucchiSimonVellin.Rmd
updated RMD file

@@
 library(spData)
+library(tidyverse)
 library(units)
+library(lwgeom)
 ```
@@
 
-In order to plot the distribution of the total population by country, we only need the R packages world, which contains both the geometries of the countries and the corresponding population.
-
-## Data Wrangling
-
-Given that we plot the population both in absolute terms and by categories — low, medium, and high — we need to define the population ranges for these categories.
-We classify countries with a population up to the 30th percentile as "Low population" countries, those with a population between the 30th and 70th percentiles as "Medium population" countries, and those in the top 30% of the population distribution as "High population" countries.
+In order to plot the distribution of the total population by country, we import data from [Natural Earth](https://www.naturalearthdata.com), specifically the countries dataset which contains both the geometries of the countries and the corresponding population.
 
@@
 
-# Divide the population by categories
-world <- world %>%
-  mutate(pop_category = case_when(
-    pop <= quantile(pop, 0.30, na.rm = TRUE) ~ "Low",
-    pop <= quantile(pop, 0.70, na.rm = TRUE) ~ "Medium",
-    TRUE ~ "High"
-  ))
+sf.world <- st_read("data/ne_10m_admin_0_countries", crs = 4326)
 
@@
 ggplot() +
-  geom_sf(data = world, aes(fill = pop)) +
+  geom_sf(data = sf.world, aes(fill = POP_EST)) +
   scale_fill_gradient(name = "Population", low = "lightblue", high = "blue")
-
-# Population by categories 
-ggplot() +
-  geom_sf(data = world, aes(fill = pop_category), alpha = 0.7) +
-  scale_fill_manual(
-    name = "Population Category",
-    values = c("Low" = "lightyellow", "Medium" = "lightgreen", "High" = "blue"))
 
@@
 
+# Remove Seven seas and Antarctica because they are NA
 world <- world %>%
@@
     continent = case_when(
-      continent %in% c("South America", "North America") ~ "America",
       continent %in% c("Seven seas (open ocean)", "Antarctica") ~ NA_character_,
@@
 
-sf.airports <- st_read("hw2/data/ne_10m_airports", crs = 4326)
-sf.pop <- st_read("hw2/data/ne_10m_populated_places_simple", crs = 4326)
+sf.airports <- st_read("data/ne_10m_airports", crs = 4326)
+sf.pop <- st_read("data/ne_10m_populated_places_simple", crs = 4326)
 

### 2025-02-03 hw2/Simon_Section.R
Changed Exe2.a) from package to NE data

@@ -10,7 +10,7 @@ library(dplyr)
 
 # Data Loading
 
-sf.world <- world
+sf.world <- st_read("hw2/data/ne_10m_admin_0_countries/ne_10m_admin_0_countries.shp")
 sf.roads_africa <- st_read("hw2/data/ne_10m_roads")
 markets <- read_excel('hw2/data/porteous_data/1.-Price--Production--and-Population-Data/MktCoords.xlsx')
 
@@ -22,7 +22,7 @@ sf.markets <- st_as_sf(markets, coords = c('longitude', 'latitude'), crs = 'EPSG
 # Plot
 
 ggplot() +
-  geom_sf(data = sf.world %>% filter(continent == "Africa"),
+  geom_sf(data = sf.world %>% filter(REGION_UN == "Africa"),
           fill = "antiquewhite", color = "gray70") +
   geom_sf(data = sf.markets, color = "red", size = 3, shape = 21, fill = "blue") +
   theme_minimal() +

### 2025-02-03 hw2/Simon_Section.R
Exe 2 updated

@@ -3,6 +3,8 @@ library(spData)
 library(tidyverse)
 library(readxl)
 library(dplyr)
+library(lwgeom)
+library(gridExtra)
 
 ------------------------------------------------------------------------------------------------------
 # Question 1 - Generate an sf POINT feature of market prices places across Africa; plot them.
@@ -33,7 +35,7 @@ ggplot() +
         plot.caption = element_text(size = 10, hjust = 0.5))
 
 ------------------------------------------------------------------------------------------------------
-  # Question 2 - Calculate minimum distance of each market to the (i) coast, (ii) nearest road, and (iii) nearest airport
+# Question 2 - Calculate minimum distance of each market to the (i) coast, (ii) nearest road, and (iii) nearest airport
 ------------------------------------------------------------------------------------------------------
 
 # Data Loading
@@ -41,37 +43,36 @@ ggplot() +
 sf.airports <- st_read("hw2/data/ne_10m_airports")
 sf.coast <- st_read("hw2/data/ne_10m_coastline")
 
-# We isolate the African continent and its airports from our global data sets to stay efficient.
-sf.africa <- sf.world %>% filter(continent == 'Africa')
+# We isolate the African airports from our global data sets to stay efficient.
+sf.africa <- sf.world %>%filter(REGION_UN == "Africa")
+sf.africa %>% filter(!st_is_valid(.))
+sf.africa <- st_make_valid(sf.africa)
 sf.airports_africa <- sf.airports %>% st_filter(sf.africa)
 
-# Data Wrangling
-# st_distance() returns a matrix with one row per market and one column per coast segment.
+# Data Wrangling - st_distance() returns a matrix with one row per market and one column per coast segment.
 
+# (i) Distance to Nearest Coast
 dist_to_coast <- st_distance(sf.markets, sf.coast)
-# For each market (row) take the minimum distance:
+# For each market (row) we take the minimum distance:
 min_coast_dist <- apply(dist_to_coast, 1, min)
 
-# (ii) Distance to Nearest Road ----------------------------------------
+# (ii) Distance to Nearest Road
 dist_to_roads <- st_distance(sf.markets, sf.roads_africa)
 min_road_dist <- apply(dist_to_roads, 1, min)
 
[…]

### 2025-02-03 hw2/hw2_MoritzPeistNoemiLucchiSimonVellin.Rmd
Finished part 1

--- 2025-01-29 hw2/hw2_MoritzPeistNoemiLucchiSimonVellin.Rmd
@@
 
-In the following notebook we replicate the charts for all 5 papers.
-Contentwise, the code follows always the same structure.
-We first load the data, secondly manipulate it, then we create the chart and finally display it.
+In this notebook, we combine various sources from [Natural Earth](https://www.naturalearthdata.com) to produce different types of plots.
+Throughout, we implement the methodologies recently discussed in class, such as distance computation, spatial merging, and filtering.
+Contentwise, we follow the assignment's instructions and provide the requested plots.
 
@@
 library(sf)
+library(spData)
+library(tidyverse)
+library(units)
+library(lwgeom)
 ```
@@
 <hr>
+
+# Exercise 1: World and airports
+
+We decided to use the [Natural Earth airports dataset](https://www.naturalearthdata.com/downloads/10m-cultural-vectors/airports/) to plot the airports on a world map for this task.
+We downloaded all the data and load it from disk for speed and efficiency though a [package](https://cran.r-project.org/web/packages/rnaturalearth/vignettes/rnaturalearth.html) would exist to download the data directly.
+
+## 1.1. Map of total population by country
+
+### Data Loading
+
+In order to plot the distribution of the total population by country, we import data from [Natural Earth Admin 0 - Countries](https://www.naturalearthdata.com/downloads/10m-cultural-vectors/10m-admin-0-countries/) dataset which contains both the geometries of the countries and the corresponding population estimate.
+
+```{r}
+sf.world <- st_read("data/ne_10m_admin_0_countries", crs = 4326) %>%
+  # Rename columns in the world dataset to make joining / operations easier
+  rename(
+    iso_a2 = ISO_A2,
+    continent = CONTINENT
+  )
+```
+
+We do not need any extensive data wrangling for this plot, as the data is already in a suitable format. Thus, we just go ahead and plot the total population by country.
+
+### Plotting
+
+```{r}
+# Population in absolute terms
+ggplot() +
+  geom_sf(data = sf.world, aes(fill = POP_EST)) +
+  scale_fill_gradient(name = "Population", low = "lightblue", high = "blue") +
+  labs(title = "Total Population by Country", fill = "Population") +
+  theme(
+    plot.title = element_text(hjust = 0.5)
+  )
+```
+
+<hr>
+
+## 1.2. Histogram of country population distribution by continent
+
+### Data Loading
+
+As for the previous plot, we only need the world data.
+
+### Data Wrangling
+
+First, we filter out countries without a continent and repair invalid polygons in case there are any.
+For a coarse-granular overview, we define population ranges by labeling countries based on their population size.
+After categorizing the countries into these ranges, we calculate the number of countries in each population range, grouped by continent.
+This gives us a summary of the distribution of countries by population size within each continent.
+Finally, we ensure that the population ranges are ordered increasingly ("0-10M", "10M-50M", "50M-100M", "100M-500M", "500M+") by using the R factor datatype.
+
+```{r}
+# Remove Seven seas and Antarctica because they are NA
+sf.world <- sf.world %>%
+  filter(!is.na(continent)) %>%
+  # Repair invalid polygons
+  st_make_valid() %>%
+  # Create the population ranges as factors
+  mutate(pop_range = factor(
+    case_when(
+      POP_EST < 1e7 ~ "0-10M",
+      POP_EST >= 1e7 & POP_EST < 5e7 ~ "10M-50M",
+      POP_EST >= 5e7 & POP_EST < 1e8 ~ "50M-100M",
+      POP_EST >= 1e8 & POP_EST < 5e8 ~ "100M-500M",
+      TRUE ~ "500M+"
+    ),
+    levels = c("0-10M", "10M-50M", "50M-100M", "100M-500M", "500M+")
+  ))
+
+# Calculate the number of countries per range + country
+population_distribution <- sf.world %>%
+  group_by(continent, pop_range) %>%
+  summarise(country_count = n(), .groups = "drop")
+```
+
+### Plotting
+
+Now we start with a coarse granular plot to get an initial overview of the data.
+
+```{r}
+ggplot(population_distribution, aes(x = pop_range, y = country_count, fill = continent)) +
+  geom_histogram(stat = "identity", position = "dodge") +
+  scale_fill_brewer(palette = "Set2") +
+  labs(
+    title = "Country Population Distribution by Continent", x = "Population Range",
+    y = "Number of Countries", fill = "Continent"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5)
+  )
+```
+
+We plot the same data but with a more detailed view to better understand the distribution of countries by population size within each continent as asked by the task.
+However, we exclude Seven Seas and Antarctica as they contribute relatively little information.
+Moreover, we use a log-scale by population size, since country populations vary hugely, also by continent as this allows us to better visualize the data.
+
+```{r}
+sf.world_filt <- sf.world %>%
+  mutate(
+    continent = case_when(
+      continent %in% c("Seven seas (open ocean)", "Antarctica") ~ NA_character_,
+      TRUE ~ continent
+    )
+  ) %>%
+  filter(!is.na(continent))
+
+# Create a proper histogram using the continuous population data
+ggplot(sf.world_filt, aes(x = POP_EST, fill = continent)) +
+  geom_histogram(position = "identity", alpha = 0.5, bins = 30) +
+  scale_x_log10(labels = scales::comma) + # Use log scale due to large population range
+  facet_wrap(~continent, scales = "free_y") + # Separate panels for each continent
+  labs(
+    title = "Population Distribution by Continent",
+    x = "Population (log scale)",
+    y = "Count of Countries",
+    fill = "Continent"
+  ) +
+  theme_minimal() +
+  theme(
+    legend.position = "none", # Remove legend since we're using facets
+    plot.title = element_text(hjust = 0.5),
+    axis.text.x = element_text(
+      angle = 45, # Rotate labels
+      hjust = 1, # Align rotated labels
+      size = 8 # Smaller font size
+    )
+  )
+```
+
+<hr>
+
+## 1.3. Histogram of country-level average distances between locations and airports by continent.
+
+### Data Loading
+
+In addition to Natural Earth's population data, we incorporate also the datasets for [populated places (simple)](https://www.naturalearthdata.com/downloads/10m-cultural-vectors/10m-populated-places/) (sf.pop) and [airports](https://www.naturalearthdata.com/downloads/10m-cultural-vectors/airports/) (sf.airports).
+
+```{r}
+sf.airports <- st_read("data/ne_10m_airports", crs = 4326)
+sf.pop <- st_read("data/ne_10m_populated_places_simple", crs = 4326)
+```
+
+### Data Wrangling
+
+This code computes and plots the average distances between populated places and airports across different continents, using two approaches: first, by computing the minimum distance for each city and averaging these values per country, and second, by directly averaging all city-airport distances within a country to allow for range-based visualization.
+Cities from the sf.pop dataset are joined with country and continent metadata from the sf.world dataset, while airports from sf.airports are joined with country metadata.
+Distances are calculated at two levels: in "distances" we compute the minimum distance from each city to the nearest airport and average these values per country, while in "distances_country" we directly average all city-airport distances within a country.
+Then we categorize "distances_country" into ranges (e.g., <50 km, <100 km) to ensure a better visualization.
+The first plot shows the distribution of city-level average distances by continent, while the second plot shows the number of countries in each distance range, grouped by continent.
+
+**Important to note:** We only calculate the distance of the closest airport to each city per country.
+This means that we do not consider the distance of all populated places to all airports in a country, but only to the closest one.
+This is a simplification that allows us to reduce the number of distances to be computed and to avoid excessive granularity in the visualization, but also methodologically makes the most sense, otherwise distances in big countries will just be overboarding.
+Moreover, we add a buffer and use `st_nearest_feature` to avoid matching issues with coastal airports.
+
+Next, we prepare the data for plotting the histograms of country-level average distances between locations and airports by continent.
+
+```{r}
+# 1. Prepare cities and airports with country metadata
+cities_with_countries <- sf.world %>%
+  st_drop_geometry() %>% # Drop geometry to avoid conflicts
+  select(iso_a2, continent) %>% # Select relevant columns
+  left_join( # Join with cities dataset
+    sf.pop %>% select(name, iso_a2, ne_id, geometry),
+    by = "iso_a2"
+  ) %>%
+  st_as_sf() %>% # Convert to sf object
+  drop_na() # Remove rows with missing values
+
+# 2. Create a small buffer around countries to catch coastal points
+# Buffer size in degrees (since CRS is 4326)
+buffer_size <- 0.01 # roughly 1km at the equator
+
+airports_with_countries <- sf.airports %>%
+  select(name, geometry) %>% # Select relevant columns
+  # Join with buffered country polygons
+  st_join(
+    sf.world %>% # Join with country metadata
+      select(iso_a2, continent) %>%
+      st_buffer(dist = buffer_size),
+    join = st_nearest_feature, # Use nearest feature instead of intersection
+    left = TRUE # Keep all airports
+  )
+
+# 2. Calculate closest airport distance per city, then average per country
+distances <- cities_with_countries %>%
+  # For each country group:
+  group_by(continent, iso_a2, ne_id) %>%
+  summarise(
+    avg_closest_distance_km = {
+      current_iso <- cur_group()$iso_a2 # Capture current country code
+      country_airports <- airports_with_countries %>%
+        filter(iso_a2 == current_iso) # Filter airports for this country
+      if (nrow(country_airports) == 0) {
+        NA_real_ # Return NA if no airports exist
+      } else {
+        # Compute closest airport distance for each city in the country
+        closest_distances <- st_distance(
+          geometry, # City geometries
+          country_airports # Airports in the current country
+        ) %>%
+          apply(1, min) %>% # Find minimum distance per city (row)
+          as.numeric() / 1000 # Convert meters to kilometers
+        mean(closest_distances, na.rm = TRUE)
+      }
+    }
+  ) %>%
+  ungroup() %>% # Remove grouping
+  drop_na() # Remove countries with no airports
+```
+
+### Plotting
+
+Now, we plot the histogram of country level average distances between locations and airports by continent - as asked for by the exercise.
+Once again we filter out the Seven Seas and Antarctica as they do not contribute much information and since only 16 airports are impacted by this anyways.
+
+```{r}
+distances_filt <- distances %>%
+  mutate(
+    continent = case_when(
+      continent %in% c("Seven seas (open ocean)", "Antarctica") ~ NA_character_,
+      TRUE ~ continent
+    )
+  ) %>%
+  filter(!is.na(continent))
+
+# Plot 1
+ggplot(distances_filt, aes(x = avg_closest_distance_km, fill = continent)) +
+  geom_histogram(bins = 30, position = "dodge") +
+  scale_fill_brewer(palette = "Set2") +
+  labs(
+    title = "Average Distance Between Locations and Airports by Continent",
+    x = "Average Distance to nearest Airport (km)", y = "Number of Populated Places", fill = "Continent"
+  ) +
+  facet_wrap(~continent, scales = "free") +
+  theme_minimal() +
+  theme(
+    legend.position = "none", # Remove legend since we're using facets
+    plot.title = element_text(hjust = 0.5)
+  )
+```
+
+What becomes immediately apparent is that most of the continents have some outliers with huge distances to the nearest airport.
+This is especially true for Africa and Asia, where some cities are very remote and far away from airports, but also for Europe, where some cities are located on islands or countries like Russia may have very large distances to the next airport.
+The most extreme example is Uelene.
+A russian village just south of the arctic cycle.
+In this sense, the populated places dataset might be misleading as it also includes also very remote places which are sparse in population and far away from airports.
+Nevertheless, also our join by `st_nearest_feature()` might mismatch some airports and cities, especially in coastal areas.
+
+Since we have these issues, let's look at the airports also in a more coarse-granular way to get a better understanding of the data (once again we exclude Seven Seas and Antarctica).
+
+```{r}
+# 3. Calculate average distances at country level. With this approach, instead of
+# computing the closest airport distance per city and then averaging per country,
+# we compute all the distances within a country and then we average them.
+distances_country <- cities_with_countries %>%
+  group_by(continent, iso_a2) %>% # Removed ne_id to group at country level
+  summarise(
+    avg_closest_distance_km = {
+      current_iso <- cur_group()$iso_a2
+      country_airports <- airports_with_countries %>%
+        filter(iso_a2 == current_iso)
+      if (nrow(country_airports) == 0) {
+        NA_real_
+      } else {
+        closest_distances <- st_distance(
+          geometry,
+          country_airports
+        ) %>%
+          apply(1, min) %>%
+          as.numeric() / 1000
+        mean(closest_distances, na.rm = TRUE)
+      }
+    },
+    country_name = first(name)
+  ) %>%
+  ungroup() %>%
+  drop_na()
+
+# 4. Create distance ranges – The previous step ensures a reasonable number of ranges.
+# If we had used the distances computed earlier, the higher granularity would have resulted in an excessive number of ranges, leading to a poor visualization quality.
+distances_ranges <- distances_country %>%
+  mutate(
+    distance_range = cut(avg_closest_distance_km,
+      breaks = seq(0, max(avg_closest_distance_km, na.rm = TRUE) + 50, by = 50),
+      labels = paste0("<", seq(50, max(avg_closest_distance_km, na.rm = TRUE) + 50, by = 50), "km"),
+      include.lowest = TRUE
+    ),
+    continent = case_when(
+      continent %in% c("Seven seas (open ocean)", "Antarctica") ~ NA_character_,
+      TRUE ~ continent
+    )
+  ) %>%
+  filter(!is.na(continent))
+```
+
+Let's also plot the coarse-granular view of the data to get a better understanding of the distribution of average distances between locations and airports by continent.
+
+```{r}
+ggplot(distances_ranges, aes(x = distance_range, fill = continent)) +
+  geom_bar(position = "stack") +
+  scale_fill_brewer(palette = "Set2") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  labs(
+    title = "Average Distance Between Locations and Airport by Ranges",
+    x = "Average Distance in Ranges", y = "Number of Countries", fill = "continent"
+  ) +
+  theme(
+    plot.title = element_text(hjust = 0.5)
+  )
+```
+
+<hr>
+
+# Exercise 2: Porteuos data

### 2025-02-03 hw2/hw2_part1_noe.R

@@ -20,7 +20,8 @@ sf.world <- st_read("hw2/data/ne_10m_admin_0_countries", crs = 4326) %>%
 # 1.1 Plotting population in absolute terms
 ggplot() +
   geom_sf(data = sf.world, aes(fill = POP_EST)) +
-  scale_fill_gradient(name = "Population", low = "lightblue", high = "blue")
+  scale_fill_gradient(name = "Population", low = "lightblue", high = "blue") +
+  labs(title = "Total Population by Country", fill = "Population")
 
 # 1.2 Defining population by categories
 sf.world <- sf.world %>%
@@ -39,16 +40,15 @@ ggplot() +
   )
 
 # 2 - HISTOGRAM OF COUNTRY POPULATION DISTRIBUTION BY continent
-# Group South America and North America in America and remove Seven seas and Antarctica
+# Remove Seven seas and Antarctica
 # because they are NA
 sf.world <- sf.world %>%
-  mutate(
-    continent = case_when(
-      # continent %in% c("South America", "North America") ~ "America",
-      continent %in% c("Seven seas (open ocean)", "Antarctica") ~ NA_character_,
-      TRUE ~ continent
-    )
-  ) %>%
+  # mutate(
+  #  continent = case_when(
+  #    continent %in% c("Seven seas (open ocean)", "Antarctica") ~ NA_character_,
+  #    TRUE ~ continent
+  #  )
+  # ) %>%
   filter(!is.na(continent)) %>%
   st_make_valid() # Repair invalid polygons
 
@@ -119,9 +119,20 @@ cities_with_countries <- sf.world %>%
   st_as_sf() %>%
   drop_na()
 
+# 2. Create a small buffer around countries to catch coastal points
+# Buffer size in degrees (since CRS is 4326)
+buffer_size <- 0.01 # roughly 1km at the equator
+
 airports_with_countries <- sf.airports %>%
   select(name, geometry) %>%
-  st_join(sf.world %>% select(iso_a2, continent))
+  # Join with buffered country polygons
+  st_join(
+    sf.world %>%
+      select(iso_a2, continent) %>%
+      st_buffer(dist = buffer_size),
+    join = st_nearest_feature, # Use nearest feature instead of intersection
+    left = TRUE # Keep all airports
+  )
 
 # 2. Calculate closest airport distance per city, then average per country
 distances <- cities_with_countries %>%
[…]

### 2025-02-03 hw2/hw2_MoritzPeistNoemiLucchiSimonVellin.Rmd
Added Simon's code - TODO: Add documentation in Part @Simon

@@
 library(lwgeom)
+library(gridExtra)
+library(readxl)
 ```
@@
 # Exercise 2: Porteuos data
+
+In this exercise, we use the [Porteous dataset]() for plotting.
+
+## 2.1. Map of markets in Africa
+
+### Data Loading
+
+```{r}
+# Clean enviornment first
+rm(list = ls())
+
+# Data Loading (with correct CRS and long/lat conversions)
+sf.world <- st_read("data/ne_10m_admin_0_countries", crs = "EPSG:4326")
+sf.roads_africa <- st_read("data/ne_10m_roads", crs = "EPSG:4326")
+sf.markets <- st_as_sf(
+  read_excel("data/porteous_data/1.-Price--Production--and-Population-Data/MktCoords.xlsx"),
+  coords = c("longitude", "latitude"),
+  crs = "EPSG:4326"
+)
+```
+
+### Plotting
+
+```{r}
+# Plot
+ggplot() +
+  geom_sf(
+    data = sf.world %>% filter(REGION_UN == "Africa"),
+    fill = "antiquewhite", color = "gray70"
+  ) +
+  geom_sf(data = sf.markets, color = "red", size = 0.7, shape = 21, fill = "blue") +
+  theme_minimal() +
+  labs(
+    title = "Markets in Africa",
+    caption = "Source: Replication Data provided"
+  ) +
+  theme(
+    legend.position = "none",
+    plot.title = element_text(hjust = 0.5),
+    plot.caption = element_text(size = 10, hjust = 0.5)
+  )
+```
+
+<hr>
+
+## 2.2. Minimum distances
+
+### Data Loading
+
+```{r}
+# Load data
+sf.airports <- st_read("data/ne_10m_airports", crs = "EPSG:4326")
+sf.coast <- st_read("data/ne_10m_coastline", crs = "EPSG:4326")
+```
+
+### Data Wrangling
+
+```{r}
+# We isolate the African airports from our global data sets to stay efficient.
+sf.africa <- sf.world %>%
+  filter(REGION_UN == "Africa") %>%
+  filter(!st_is_valid(.)) %>%
+  st_make_valid()
+
+# Filter airports in Africa
+sf.airports_africa <- sf.airports %>% st_filter(sf.africa)
+
+# Data Wrangling - st_distance() returns a matrix with one row per market and one column per coast segment.
+
+# (i) Distance to Nearest Coast
+dist_to_coast <- st_distance(sf.markets, sf.coast)
+# For each market (row) we take the minimum distance:
+min_coast_dist <- apply(dist_to_coast, 1, min)
+
+# (ii) Distance to Nearest Road
+dist_to_roads <- st_distance(sf.markets, sf.roads_africa)
+min_road_dist <- apply(dist_to_roads, 1, min)
+
+# (iii) Distance to Nearest Airport
+dist_to_airports <- st_distance(sf.markets, sf.airports_africa)
+min_airport_dist <- apply(dist_to_airports, 1, min)
+
+# Add the computed distances to the markets data
+sf.markets <- sf.markets %>%
+  mutate(
+    min_coast_dist   = as.numeric(min_coast_dist),
+    min_road_dist    = as.numeric(min_road_dist),
+    min_airport_dist = as.numeric(min_airport_dist)
+  )
+
+# Summary
+print(summary(sf.markets))
+```
+
+<hr>
+
+## 2.3. Scatter plots
+
+### Data Loading
+
+```{r}
+prices <- read_excel(
+  "data/porteous_data/4.-Outer-Loop-for-Cost-Parameter-Estimation/PriceMaster4GAMS.xlsx"
+)
+```
+
+### Data Wrangling
+
+```{r}
+# Calculate average price
+prices_slice <- prices[, 5:ncol(prices)]
+prices <- replace(prices, prices == 0, NA) %>%
+  mutate(av_price = rowMeans(prices_slice, na.rm = TRUE)) %>%
+  select(mktcode, country, market, crop, av_price)
+sf.marketprices <- sf.markets %>% left_join(prices, by = "mktcode")
+```
+
+### Plotting
+
+```{r}
+# (i) Average Price vs. Minimum Distance to Coast
+p_coast <- ggplot(sf.marketprices, aes(x = min_coast_dist, y = av_price, color = crop)) +
+  geom_point(alpha = 0.7) +
+  geom_smooth(method = "lm", se = FALSE) + # Add linear trend lines
+  labs(
+    title = "Average Price vs. Distance to Coast",
+    x = "Minimum Distance to Coast (meters)",
+    y = "Average Price",
+    color = "Crop Type"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5)
+  )
+
+# (ii) Average Price vs. Minimum Distance to Road
+p_road <- ggplot(sf.marketprices, aes(x = min_road_dist, y = av_price, color = crop)) +
+  geom_point(alpha = 0.7) +
+  geom_smooth(method = "lm", se = FALSE) + # Add linear trend lines
+  labs(
+    title = "Average Price vs. Distance to Road",
+    x = "Minimum Distance to Road (meters)",
+    y = "Average Price",
+    color = "Crop Type"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5)
+  )
+
+# (iii) Average Price vs. Minimum Distance to Airport
+p_airport <- ggplot(sf.marketprices, aes(x = min_airport_dist, y = av_price, color = crop)) +
+  geom_point(alpha = 0.7) +
+  geom_smooth(method = "lm", se = FALSE) + # Add linear trend lines
+  labs(
+    title = "Average Price vs. Distance to Airport",
+    x = "Minimum Distance to Airport (meters)",
+    y = "Average Price",
+    color = "Crop Type"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5)
+  )
+
+# Display the plots in a grid
+grid.arrange(p_coast, p_road, p_airport, nrow = 2, ncol = 2)
+```

### 2025-02-03 hw2/Simon_Section.R

@@ -6,47 +6,56 @@ library(dplyr)
 library(lwgeom)
 library(gridExtra)
 
-------------------------------------------------------------------------------------------------------
+#------------------------------------------------------------------------------------------------------
 # Question 1 - Generate an sf POINT feature of market prices places across Africa; plot them.
-------------------------------------------------------------------------------------------------------
+#------------------------------------------------------------------------------------------------------
 
 # Data Loading
 
-sf.world <- st_read("hw2/data/ne_10m_admin_0_countries/ne_10m_admin_0_countries.shp")
-sf.roads_africa <- st_read("hw2/data/ne_10m_roads")
-markets <- read_excel('hw2/data/porteous_data/1.-Price--Production--and-Population-Data/MktCoords.xlsx')
+sf.world <- st_read("hw2/data/ne_10m_admin_0_countries", crs = "EPSG:4326")
+sf.roads_africa <- st_read("hw2/data/ne_10m_roads", crs = "EPSG:4326")
+sf.markets <- st_as_sf(
+  read_excel("hw2/data/porteous_data/1.-Price--Production--and-Population-Data/MktCoords.xlsx"),
+  coords = c("longitude", "latitude"),
+  crs = "EPSG:4326"
+)
 
-# Data Wrangling
-
-sf.world <- st_transform(sf.world, CRS="EPSG:4326")
-sf.markets <- st_as_sf(markets, coords = c('longitude', 'latitude'), crs = 'EPSG:4326')
 
 # Plot
 
 ggplot() +
-  geom_sf(data = sf.world %>% filter(REGION_UN == "Africa"),
-          fill = "antiquewhite", color = "gray70") +
-  geom_sf(data = sf.markets, color = "red", size = 3, shape = 21, fill = "blue") +
+  geom_sf(
+    data = sf.world %>% filter(REGION_UN == "Africa"),
+    fill = "antiquewhite", color = "gray70"
+  ) +
+  geom_sf(data = sf.markets, color = "red", size = 0.7, shape = 21, fill = "blue") +
   theme_minimal() +
-  labs(title = "Markets in Africa",
-       caption = "Source: Replication Data provided") +
-  theme(legend.position = "none",
-        plot.title = element_text(hjust = 0.5),
-        plot.caption = element_text(size = 10, hjust = 0.5))
+  labs(
+    title = "Markets in Africa",
+    caption = "Source: Replication Data provided"
+  ) +
+  theme(
+    legend.position = "none",
[…]

### 2025-02-04 hw2/hw2_MoritzPeistNoemiLucchiSimonVellin.Rmd
Added comments for part 2

@@
 
-In this exercise, we use the [Porteous dataset]() for plotting.
+This analysis explores agricultural market data in Africa using the Porteous dataset, which investigates trade costs and their impact on agricultural markets.
+The study examines the relationship between market locations, transportation infrastructure, and agricultural product prices across African countries.
+
+The analysis is structured in three main parts:
+1. Visualization of market locations across Africa
+2. Analysis of markets' accessibility through distance calculations
+3. Investigation of price-distance relationships for different transportation modes
+
+This work builds on Porteous (2019), who studied how trade costs affect agricultural storage and trade patterns in Africa.
+Understanding these relationships is crucial for identifying potential barriers to trade and market efficiency in African agricultural markets.
 
@@
 
+This section creates a visualization of market locations across Africa. The analysis combines three key geographical datasets:
+- Administrative boundaries of African countries
+- Road network data for transportation infrastructure
+- Market location coordinates from the Porteous dataset
+
+The resulting map will help us understand the spatial distribution of agricultural markets and their relationship to national boundaries.
+This visualization is particularly important for identifying potential clusters of markets and areas that might be underserved.
+
 ### Data Loading
+
+The data loading process integrates three essential geographical datasets using appropriate coordinate reference systems (CRS).
+We use EPSG:4326, which is the standard WGS84 geographic coordinate system, ensuring consistent spatial representation across all datasets.
+The data includes:
+
+1. World administrative boundaries for defining African countries
+2. African road network data for understanding transportation infrastructure
+3. Market coordinates from the Porteous dataset, converted to spatial features
+
+Converting the market coordinates to spatial features (sf) allows us to perform spatial operations and create meaningful visualizations.
+The careful attention to coordinate systems is crucial for accurate geographical analysis.
 
@@
 ### Plotting
+
+The plotting phase creates a clear visualization of market locations within their geographical context.
+We use a layered approach to build the map, starting with country boundaries and adding market locations.
+The aesthetic choices are deliberate:
+- Antiquewhite fill for countries provides a neutral background
+- Gray borders help define country boundaries without overshadowing the main data
+- Red and blue markers for markets ensure visibility while distinguishing them from the background
+- Minimal theme removes unnecessary visual elements to focus attention on the data
 
@@
 
+This section calculates the accessibility of markets by measuring their proximity to key transportation infrastructure. We examine three critical distance metrics:
+
+1. Distance to coast: Important for international trade and port access
+2. Distance to roads: Crucial for domestic distribution and market integration
+3. Distance to airports: Relevant for high-value or time-sensitive agricultural products
+
+These measurements help quantify the isolation or connectivity of each market, which can significantly impact trading costs and market efficiency.
+The analysis uses spatial analysis techniques to calculate the shortest distances between markets and each type of infrastructure.
+
 ### Data Loading
+
+This subsection expands our dataset by incorporating additional infrastructure information. We load two new spatial datasets:
+1. Airport locations across the globe
+2. Coastline geometries
+
+These datasets complement our existing market and road data, allowing us to compute comprehensive accessibility metrics.
+The careful organization of these spatial datasets ensures efficient processing in subsequent analysis steps.
 
@@
 ### Data Wrangling
+
+This crucial data processing step involves several sophisticated spatial operations. The process:
+1. Isolates African data to improve computational efficiency
+2. Ensures geometric validity of spatial features
+3. Calculates distance matrices between markets and various infrastructure points
+4. Processes these matrices to find minimum distances
+
+The use of st_distance() creates matrices that capture all possible combinations of distances, from which we extract the minimum values.
+This approach ensures we identify the truly nearest infrastructure points for each market.
 
@@
 
-## 2.3. Scatter plots
+## 2.3. Scatter plots - Price-Distance Relationships
+
+This section investigates how market prices relate to infrastructure accessibility. The analysis:
+
+1. Processes raw price data from the Porteous dataset
+2. Calculates average prices for different crops
+3. Creates scatter plots to visualize price-distance relationships
+
+The resulting visualizations help us understand whether more isolated markets (those farther from transportation infrastructure) tend to have higher or lower prices. This relationship is crucial for:
+- Understanding market integration
+- Identifying potential barriers to trade
+- Assessing the impact of infrastructure on market efficiency
+
+Each scatter plot includes trend lines by crop type, allowing us to observe whether different agricultural products show distinct patterns in their price-distance relationships.
 
 ### Data Loading
+
+Here we incorporate price data from the Porteous dataset, specifically from the cost parameter estimation component.
+This data provides crucial economic information that we'll relate to our spatial measurements.
+The price data represents actual market conditions and will help us understand how geography influences market dynamics.
 
@@
 ### Data Wrangling
+
+This phase processes the raw price data to create meaningful economic metrics. The operations include:
+1. Handling zero values by converting them to NA to avoid skewing averages
+2. Calculating average prices across different time periods
+3. Joining price data with spatial market data
+
+This careful data preparation ensures that our subsequent analysis accurately represents the relationship between prices and geographical factors.
+The use of NA for zero values is particularly important for maintaining data integrity.
 
@@
 ### Plotting
+
+The final visualization phase creates three sophisticated scatter plots that reveal relationships between prices and infrastructure accessibility. Each plot incorporates:
+1. Individual data points colored by crop type to show product-specific patterns
+2. Trend lines to highlight overall relationships
+3. Appropriate axis labels and titles for clear interpretation
+
+The grid arrangement of plots facilitates easy comparison of how different types of infrastructure accessibility relate to market prices.
+The use of alpha transparency helps manage overlapping points, while maintaining visual clarity.
 

### 2025-02-04 hw2/hw2_MoritzPeistNoemiLucchiSimonVellin.Rmd
Fixed markdown

@@
 ```{r}
-# Remove Seven seas and Antarctica because they are NA
+# Filter countries with NA continents and repair invalid polygons
 sf.world <- sf.world %>%
@@
 ```{r}
+# Filter out Seven Seas and Antarctica
 sf.world_filt <- sf.world %>%
@@
 ```{r}
+# Filter out Seven Seas and Antarctica
 distances_filt <- distances %>%
@@
 # we compute all the distances within a country and then we average them.
+
+# Basically, same code as before, so no extensive comments
 distances_country <- cities_with_countries %>%
@@
 The analysis is structured in three main parts:
+
 1. Visualization of market locations across Africa
@@
 This section creates a visualization of market locations across Africa. The analysis combines three key geographical datasets:
+
 - Administrative boundaries of African countries
@@
 The aesthetic choices are deliberate:
+
 - Antiquewhite fill for countries provides a neutral background
@@
 This subsection expands our dataset by incorporating additional infrastructure information. We load two new spatial datasets:
+
 1. Airport locations across the globe
@@
 This crucial data processing step involves several sophisticated spatial operations. The process:
+
 1. Isolates African data to improve computational efficiency
@@
 The resulting visualizations help us understand whether more isolated markets (those farther from transportation infrastructure) tend to have higher or lower prices. This relationship is crucial for:
+
 - Understanding market integration
@@
 This phase processes the raw price data to create meaningful economic metrics. The operations include:
+
 1. Handling zero values by converting them to NA to avoid skewing averages
@@
 The final visualization phase creates three sophisticated scatter plots that reveal relationships between prices and infrastructure accessibility. Each plot incorporates:
+
 1. Individual data points colored by crop type to show product-specific patterns

### 2025-02-04 hw2/hw2_MoritzPeistNoemiLucchiSimonVellin.Rmd
Spelling mistakes Ex 2 checked

@@
 
-# Exercise 2: Porteuos data
+# Exercise 2: Porteous data
 
@@
 ```{r}
-# We isolate the African airports from our global data sets to stay efficient.
+# We isolate the African airports from our global data sets to stay efficient
 sf.africa <- sf.world %>%
@@
 
-# Data Wrangling - st_distance() returns a matrix with one row per market and one column per coast segment.
+# Data Wrangling - st_distance() returns a matrix with one row per market and one column per coast segment
 

### 2025-02-12 hw3/hw3_MoritzPeistNoemiLucchiSimonVellin.Rmd
Initial commit for HW3

--- 2025-02-04 hw2/hw2_MoritzPeistNoemiLucchiSimonVellin.Rmd
@@
 ---
-title: "Geospatial - Assignment 2"
+title: "Geospatial - Assignment 3"
 author: 
@@
 output:
-  html_document:
+  html_notebook:
     toc: true
@@
 
-In this notebook, we combine various sources from [Natural Earth](https://www.naturalearthdata.com) to produce different types of plots.
-Throughout, we implement the methodologies recently discussed in class, such as distance computation, spatial merging, and filtering.
-Contentwise, we follow the assignment's instructions and provide the requested plots.
+...
 
@@
 
-# Exercise 1: World and airports
+# Exercise 1: Climate change in the USA
 
-We decided to use the [Natural Earth airports dataset](https://www.naturalearthdata.com/downloads/10m-cultural-vectors/airports/) to plot the airports on a world map for this task.
-We downloaded all the data and load it from disk for speed and efficiency though a [package](https://cran.r-project.org/web/packages/rnaturalearth/vignettes/rnaturalearth.html) would exist to download the data directly.
-
-## 1.1. Map of total population by country
-
-### Data Loading
-
-In order to plot the distribution of the total population by country, we import data from [Natural Earth Admin 0 - Countries](https://www.naturalearthdata.com/downloads/10m-cultural-vectors/10m-admin-0-countries/) dataset which contains both the geometries of the countries and the corresponding population estimate.
-
-```{r}
-sf.world <- st_read("data/ne_10m_admin_0_countries", crs = 4326) %>%
-  # Rename columns in the world dataset to make joining / operations easier
-  rename(
-    iso_a2 = ISO_A2,
-    continent = CONTINENT
-  )
-```
-
-We do not need any extensive data wrangling for this plot, as the data is already in a suitable format. Thus, we just go ahead and plot the total population by country.
-
-### Plotting
-
-```{r}
-# Population in absolute terms
-ggplot() +
-  geom_sf(data = sf.world, aes(fill = POP_EST)) +
-  scale_fill_gradient(name = "Population", low = "lightblue", high = "blue") +
-  labs(title = "Total Population by Country", fill = "Population") +
-  theme(
-    plot.title = element_text(hjust = 0.5)
-  )
-```
+...
 
@@
 
-## 1.2. Histogram of country population distribution by continent
-
-### Data Loading
-
-As for the previous plot, we only need the world data.
-
-### Data Wrangling
-
-First, we filter out countries without a continent and repair invalid polygons in case there are any.
-For a coarse-granular overview, we define population ranges by labeling countries based on their population size.
-After categorizing the countries into these ranges, we calculate the number of countries in each population range, grouped by continent.
-This gives us a summary of the distribution of countries by population size within each continent.
-Finally, we ensure that the population ranges are ordered increasingly ("0-10M", "10M-50M", "50M-100M", "100M-500M", "500M+") by using the R factor datatype.
-
-```{r}
-# Remove Seven seas and Antarctica because they are NA
-sf.world <- sf.world %>%
-  filter(!is.na(continent)) %>%
-  # Repair invalid polygons
-  st_make_valid() %>%
-  # Create the population ranges as factors
-  mutate(pop_range = factor(
-    case_when(
-      POP_EST < 1e7 ~ "0-10M",
-      POP_EST >= 1e7 & POP_EST < 5e7 ~ "10M-50M",
-      POP_EST >= 5e7 & POP_EST < 1e8 ~ "50M-100M",
-      POP_EST >= 1e8 & POP_EST < 5e8 ~ "100M-500M",
-      TRUE ~ "500M+"
-    ),
-    levels = c("0-10M", "10M-50M", "50M-100M", "100M-500M", "500M+")
-  ))
-
-# Calculate the number of countries per range + country
-population_distribution <- sf.world %>%
-  group_by(continent, pop_range) %>%
-  summarise(country_count = n(), .groups = "drop")
-```
-
-### Plotting
-
-Now we start with a coarse granular plot to get an initial overview of the data.
-
-```{r}
-ggplot(population_distribution, aes(x = pop_range, y = country_count, fill = continent)) +
-  geom_histogram(stat = "identity", position = "dodge") +
-  scale_fill_brewer(palette = "Set2") +
-  labs(
-    title = "Country Population Distribution by Continent", x = "Population Range",
-    y = "Number of Countries", fill = "Continent"
-  ) +
-  theme_minimal() +
-  theme(
-    plot.title = element_text(hjust = 0.5)
-  )
-```
-
-We plot the same data but with a more detailed view to better understand the distribution of countries by population size within each continent as asked by the task.
-However, we exclude Seven Seas and Antarctica as they contribute relatively little information.
-Moreover, we use a log-scale by population size, since country populations vary hugely, also by continent as this allows us to better visualize the data.
-
-```{r}
-sf.world_filt <- sf.world %>%
-  mutate(
-    continent = case_when(
-      continent %in% c("Seven seas (open ocean)", "Antarctica") ~ NA_character_,
-      TRUE ~ continent
-    )
-  ) %>%
-  filter(!is.na(continent))
-
-# Create a proper histogram using the continuous population data
-ggplot(sf.world_filt, aes(x = POP_EST, fill = continent)) +
-  geom_histogram(position = "identity", alpha = 0.5, bins = 30) +
-  scale_x_log10(labels = scales::comma) + # Use log scale due to large population range
-  facet_wrap(~continent, scales = "free_y") + # Separate panels for each continent
-  labs(
-    title = "Population Distribution by Continent",
-    x = "Population (log scale)",
-    y = "Count of Countries",
-    fill = "Continent"
-  ) +
-  theme_minimal() +
-  theme(
-    legend.position = "none", # Remove legend since we're using facets
-    plot.title = element_text(hjust = 0.5),
-    axis.text.x = element_text(
-      angle = 45, # Rotate labels
-      hjust = 1, # Align rotated labels
-      size = 8 # Smaller font size
-    )
-  )
-```
-
-<hr>
-
-## 1.3. Histogram of country-level average distances between locations and airports by continent.
-
-### Data Loading
-
-In addition to Natural Earth's population data, we incorporate also the datasets for [populated places (simple)](https://www.naturalearthdata.com/downloads/10m-cultural-vectors/10m-populated-places/) (sf.pop) and [airports](https://www.naturalearthdata.com/downloads/10m-cultural-vectors/airports/) (sf.airports).
-
-```{r}
-sf.airports <- st_read("data/ne_10m_airports", crs = 4326)
-sf.pop <- st_read("data/ne_10m_populated_places_simple", crs = 4326)
-```
-
-### Data Wrangling
-
-This code computes and plots the average distances between populated places and airports across different continents, using two approaches: first, by computing the minimum distance for each city and averaging these values per country, and second, by directly averaging all city-airport distances within a country to allow for range-based visualization.
-Cities from the sf.pop dataset are joined with country and continent metadata from the sf.world dataset, while airports from sf.airports are joined with country metadata.
-Distances are calculated at two levels: in "distances" we compute the minimum distance from each city to the nearest airport and average these values per country, while in "distances_country" we directly average all city-airport distances within a country.
-Then we categorize "distances_country" into ranges (e.g., <50 km, <100 km) to ensure a better visualization.
-The first plot shows the distribution of city-level average distances by continent, while the second plot shows the number of countries in each distance range, grouped by continent.
-
-**Important to note:** We only calculate the distance of the closest airport to each city per country.
-This means that we do not consider the distance of all populated places to all airports in a country, but only to the closest one.
-This is a simplification that allows us to reduce the number of distances to be computed and to avoid excessive granularity in the visualization, but also methodologically makes the most sense, otherwise distances in big countries will just be overboarding.
-Moreover, we add a buffer and use `st_nearest_feature` to avoid matching issues with coastal airports.
-
-Next, we prepare the data for plotting the histograms of country-level average distances between locations and airports by continent.
-
-```{r}
-# 1. Prepare cities and airports with country metadata
-cities_with_countries <- sf.world %>%
-  st_drop_geometry() %>% # Drop geometry to avoid conflicts
-  select(iso_a2, continent) %>% # Select relevant columns
-  left_join( # Join with cities dataset
-    sf.pop %>% select(name, iso_a2, ne_id, geometry),
-    by = "iso_a2"
-  ) %>%
-  st_as_sf() %>% # Convert to sf object
-  drop_na() # Remove rows with missing values
-
-# 2. Create a small buffer around countries to catch coastal points
-# Buffer size in degrees (since CRS is 4326)
-buffer_size <- 0.01 # roughly 1km at the equator
-
-airports_with_countries <- sf.airports %>%
-  select(name, geometry) %>% # Select relevant columns
-  # Join with buffered country polygons
-  st_join(
-    sf.world %>% # Join with country metadata
-      select(iso_a2, continent) %>%
-      st_buffer(dist = buffer_size),
-    join = st_nearest_feature, # Use nearest feature instead of intersection
-    left = TRUE # Keep all airports
-  )
-
-# 2. Calculate closest airport distance per city, then average per country
-distances <- cities_with_countries %>%
-  # For each country group:
-  group_by(continent, iso_a2, ne_id) %>%
-  summarise(
-    avg_closest_distance_km = {
-      current_iso <- cur_group()$iso_a2 # Capture current country code
-      country_airports <- airports_with_countries %>%
-        filter(iso_a2 == current_iso) # Filter airports for this country
-      if (nrow(country_airports) == 0) {
-        NA_real_ # Return NA if no airports exist
-      } else {
-        # Compute closest airport distance for each city in the country
-        closest_distances <- st_distance(
-          geometry, # City geometries
-          country_airports # Airports in the current country
-        ) %>%
-          apply(1, min) %>% # Find minimum distance per city (row)
-          as.numeric() / 1000 # Convert meters to kilometers
-        mean(closest_distances, na.rm = TRUE)
-      }
-    }
-  ) %>%
-  ungroup() %>% # Remove grouping
-  drop_na() # Remove countries with no airports
-```
-
-### Plotting
-
-Now, we plot the histogram of country level average distances between locations and airports by continent - as asked for by the exercise.
-Once again we filter out the Seven Seas and Antarctica as they do not contribute much information and since only 16 airports are impacted by this anyways.
-
-```{r}
-distances_filt <- distances %>%
-  mutate(
-    continent = case_when(
-      continent %in% c("Seven seas (open ocean)", "Antarctica") ~ NA_character_,
-      TRUE ~ continent
-    )
-  ) %>%
-  filter(!is.na(continent))
-
-# Plot 1
-ggplot(distances_filt, aes(x = avg_closest_distance_km, fill = continent)) +
-  geom_histogram(bins = 30, position = "dodge") +
-  scale_fill_brewer(palette = "Set2") +
-  labs(
-    title = "Average Distance Between Locations and Airports by Continent",
-    x = "Average Distance to nearest Airport (km)", y = "Number of Populated Places", fill = "Continent"
-  ) +
-  facet_wrap(~continent, scales = "free") +
-  theme_minimal() +
-  theme(
-    legend.position = "none", # Remove legend since we're using facets
-    plot.title = element_text(hjust = 0.5)
-  )
-```
-
-What becomes immediately apparent is that most of the continents have some outliers with huge distances to the nearest airport.
-This is especially true for Africa and Asia, where some cities are very remote and far away from airports, but also for Europe, where some cities are located on islands or countries like Russia may have very large distances to the next airport.
-The most extreme example is Uelene.
-A russian village just south of the arctic cycle.
-In this sense, the populated places dataset might be misleading as it also includes also very remote places which are sparse in population and far away from airports.
-Nevertheless, also our join by `st_nearest_feature()` might mismatch some airports and cities, especially in coastal areas.
-
-Since we have these issues, let's look at the airports also in a more coarse-granular way to get a better understanding of the data (once again we exclude Seven Seas and Antarctica).
-
-```{r}
-# 3. Calculate average distances at country level. With this approach, instead of
-# computing the closest airport distance per city and then averaging per country,
-# we compute all the distances within a country and then we average them.
-distances_country <- cities_with_countries %>%
-  group_by(continent, iso_a2) %>% # Removed ne_id to group at country level
-  summarise(
-    avg_closest_distance_km = {
-      current_iso <- cur_group()$iso_a2
-      country_airports <- airports_with_countries %>%
-        filter(iso_a2 == current_iso)
-      if (nrow(country_airports) == 0) {
-        NA_real_
-      } else {
-        closest_distances <- st_distance(
-          geometry,
-          country_airports
-        ) %>%
-          apply(1, min) %>%
-          as.numeric() / 1000
-        mean(closest_distances, na.rm = TRUE)
-      }
-    },
-    country_name = first(name)
-  ) %>%
-  ungroup() %>%
-  drop_na()
-
-# 4. Create distance ranges – The previous step ensures a reasonable number of ranges.
-# If we had used the distances computed earlier, the higher granularity would have resulted in an excessive number of ranges, leading to a poor visualization quality.
-distances_ranges <- distances_country %>%
-  mutate(
-    distance_range = cut(avg_closest_distance_km,
-      breaks = seq(0, max(avg_closest_distance_km, na.rm = TRUE) + 50, by = 50),
-      labels = paste0("<", seq(50, max(avg_closest_distance_km, na.rm = TRUE) + 50, by = 50), "km"),
-      include.lowest = TRUE
-    ),
-    continent = case_when(
-      continent %in% c("Seven seas (open ocean)", "Antarctica") ~ NA_character_,
-      TRUE ~ continent
-    )
-  ) %>%
-  filter(!is.na(continent))
-```
-
-Let's also plot the coarse-granular view of the data to get a better understanding of the distribution of average distances between locations and airports by continent.
-
-```{r}
-ggplot(distances_ranges, aes(x = distance_range, fill = continent)) +
-  geom_bar(position = "stack") +
-  scale_fill_brewer(palette = "Set2") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  labs(
-    title = "Average Distance Between Locations and Airport by Ranges",
-    x = "Average Distance in Ranges", y = "Number of Countries", fill = "continent"
-  ) +
-  theme(
-    plot.title = element_text(hjust = 0.5)
-  )
-```
-
-<hr>
-
-# Exercise 2: Porteuos data
-
-This analysis explores agricultural market data in Africa using the Porteous dataset, which investigates trade costs and their impact on agricultural markets.
-The study examines the relationship between market locations, transportation infrastructure, and agricultural product prices across African countries.
-
-The analysis is structured in three main parts:
-1. Visualization of market locations across Africa
-2. Analysis of markets' accessibility through distance calculations
-3. Investigation of price-distance relationships for different transportation modes
-
-This work builds on Porteous (2019), who studied how trade costs affect agricultural storage and trade patterns in Africa.
-Understanding these relationships is crucial for identifying potential barriers to trade and market efficiency in African agricultural markets.
-
-## 2.1. Map of markets in Africa
-
-This section creates a visualization of market locations across Africa. The analysis combines three key geographical datasets:
-- Administrative boundaries of African countries
-- Road network data for transportation infrastructure
-- Market location coordinates from the Porteous dataset
-
-The resulting map will help us understand the spatial distribution of agricultural markets and their relationship to national boundaries.
-This visualization is particularly important for identifying potential clusters of markets and areas that might be underserved.
-
-### Data Loading
-
-The data loading process integrates three essential geographical datasets using appropriate coordinate reference systems (CRS).
-We use EPSG:4326, which is the standard WGS84 geographic coordinate system, ensuring consistent spatial representation across all datasets.
-The data includes:
-
-1. World administrative boundaries for defining African countries
-2. African road network data for understanding transportation infrastructure
-3. Market coordinates from the Porteous dataset, converted to spatial features
-
-Converting the market coordinates to spatial features (sf) allows us to perform spatial operations and create meaningful visualizations.
-The careful attention to coordinate systems is crucial for accurate geographical analysis.
-
-```{r}
-# Clean enviornment first
-rm(list = ls())
-
-# Data Loading (with correct CRS and long/lat conversions)
-sf.world <- st_read("data/ne_10m_admin_0_countries", crs = "EPSG:4326")
-sf.roads_africa <- st_read("data/ne_10m_roads", crs = "EPSG:4326")
-sf.markets <- st_as_sf(
-  read_excel("data/porteous_data/1.-Price--Production--and-Population-Data/MktCoords.xlsx"),
-  coords = c("longitude", "latitude"),
-  crs = "EPSG:4326"
-)
-```
-
-### Plotting
-
-The plotting phase creates a clear visualization of market locations within their geographical context.
-We use a layered approach to build the map, starting with country boundaries and adding market locations.
-The aesthetic choices are deliberate:
-- Antiquewhite fill for countries provides a neutral background
-- Gray borders help define country boundaries without overshadowing the main data
-- Red and blue markers for markets ensure visibility while distinguishing them from the background
-- Minimal theme removes unnecessary visual elements to focus attention on the data
-
-```{r}
-# Plot
-ggplot() +
-  geom_sf(
-    data = sf.world %>% filter(REGION_UN == "Africa"),
-    fill = "antiquewhite", color = "gray70"
-  ) +
-  geom_sf(data = sf.markets, color = "red", size = 0.7, shape = 21, fill = "blue") +
-  theme_minimal() +
-  labs(
-    title = "Markets in Africa",
-    caption = "Source: Replication Data provided"
-  ) +
-  theme(
-    legend.position = "none",
-    plot.title = element_text(hjust = 0.5),
-    plot.caption = element_text(size = 10, hjust = 0.5)
-  )
-```
-
-<hr>
-
-## 2.2. Minimum distances
-
-This section calculates the accessibility of markets by measuring their proximity to key transportation infrastructure. We examine three critical distance metrics:
-
-1. Distance to coast: Important for international trade and port access
-2. Distance to roads: Crucial for domestic distribution and market integration
-3. Distance to airports: Relevant for high-value or time-sensitive agricultural products
-
-These measurements help quantify the isolation or connectivity of each market, which can significantly impact trading costs and market efficiency.
-The analysis uses spatial analysis techniques to calculate the shortest distances between markets and each type of infrastructure.
-
-### Data Loading
-
-This subsection expands our dataset by incorporating additional infrastructure information. We load two new spatial datasets:
-1. Airport locations across the globe
-2. Coastline geometries
-
-These datasets complement our existing market and road data, allowing us to compute comprehensive accessibility metrics.
-The careful organization of these spatial datasets ensures efficient processing in subsequent analysis steps.
-
-```{r}
-# Load data
-sf.airports <- st_read("data/ne_10m_airports", crs = "EPSG:4326")
-sf.coast <- st_read("data/ne_10m_coastline", crs = "EPSG:4326")
-```
-
-### Data Wrangling
-
-This crucial data processing step involves several sophisticated spatial operations. The process:
-1. Isolates African data to improve computational efficiency
-2. Ensures geometric validity of spatial features
-3. Calculates distance matrices between markets and various infrastructure points
-4. Processes these matrices to find minimum distances
-
-The use of st_distance() creates matrices that capture all possible combinations of distances, from which we extract the minimum values.
-This approach ensures we identify the truly nearest infrastructure points for each market.
-
-```{r}
-# We isolate the African airports from our global data sets to stay efficient.
-sf.africa <- sf.world %>%
-  filter(REGION_UN == "Africa") %>%
-  filter(!st_is_valid(.)) %>%
-  st_make_valid()
-
-# Filter airports in Africa
-sf.airports_africa <- sf.airports %>% st_filter(sf.africa)
-
-# Data Wrangling - st_distance() returns a matrix with one row per market and one column per coast segment.
-
-# (i) Distance to Nearest Coast
-dist_to_coast <- st_distance(sf.markets, sf.coast)
-# For each market (row) we take the minimum distance:
-min_coast_dist <- apply(dist_to_coast, 1, min)
-
-# (ii) Distance to Nearest Road
-dist_to_roads <- st_distance(sf.markets, sf.roads_africa)
-min_road_dist <- apply(dist_to_roads, 1, min)
-
-# (iii) Distance to Nearest Airport
-dist_to_airports <- st_distance(sf.markets, sf.airports_africa)
-min_airport_dist <- apply(dist_to_airports, 1, min)
-
-# Add the computed distances to the markets data
-sf.markets <- sf.markets %>%
-  mutate(
-    min_coast_dist   = as.numeric(min_coast_dist),
-    min_road_dist    = as.numeric(min_road_dist),
-    min_airport_dist = as.numeric(min_airport_dist)
-  )
-
-# Summary
-print(summary(sf.markets))
-```
-
-<hr>
-
-## 2.3. Scatter plots - Price-Distance Relationships
-
-This section investigates how market prices relate to infrastructure accessibility. The analysis:
-
-1. Processes raw price data from the Porteous dataset
-2. Calculates average prices for different crops
-3. Creates scatter plots to visualize price-distance relationships
-
-The resulting visualizations help us understand whether more isolated markets (those farther from transportation infrastructure) tend to have higher or lower prices. This relationship is crucial for:
-- Understanding market integration
-- Identifying potential barriers to trade
-- Assessing the impact of infrastructure on market efficiency
-
-Each scatter plot includes trend lines by crop type, allowing us to observe whether different agricultural products show distinct patterns in their price-distance relationships.
-
-### Data Loading
-
-Here we incorporate price data from the Porteous dataset, specifically from the cost parameter estimation component.
-This data provides crucial economic information that we'll relate to our spatial measurements.
-The price data represents actual market conditions and will help us understand how geography influences market dynamics.
-
-```{r}
-prices <- read_excel(
-  "data/porteous_data/4.-Outer-Loop-for-Cost-Parameter-Estimation/PriceMaster4GAMS.xlsx"
-)
-```
-
-### Data Wrangling
-
-This phase processes the raw price data to create meaningful economic metrics. The operations include:
-1. Handling zero values by converting them to NA to avoid skewing averages
-2. Calculating average prices across different time periods
-3. Joining price data with spatial market data
-
-This careful data preparation ensures that our subsequent analysis accurately represents the relationship between prices and geographical factors.
-The use of NA for zero values is particularly important for maintaining data integrity.
-
-```{r}
-# Calculate average price
-prices_slice <- prices[, 5:ncol(prices)]
-prices <- replace(prices, prices == 0, NA) %>%
-  mutate(av_price = rowMeans(prices_slice, na.rm = TRUE)) %>%
-  select(mktcode, country, market, crop, av_price)
-sf.marketprices <- sf.markets %>% left_join(prices, by = "mktcode")
-```
-
-### Plotting
-
-The final visualization phase creates three sophisticated scatter plots that reveal relationships between prices and infrastructure accessibility. Each plot incorporates:
-1. Individual data points colored by crop type to show product-specific patterns
-2. Trend lines to highlight overall relationships
-3. Appropriate axis labels and titles for clear interpretation
-
-The grid arrangement of plots facilitates easy comparison of how different types of infrastructure accessibility relate to market prices.
-The use of alpha transparency helps manage overlapping points, while maintaining visual clarity.
-
-```{r}
-# (i) Average Price vs. Minimum Distance to Coast
-p_coast <- ggplot(sf.marketprices, aes(x = min_coast_dist, y = av_price, color = crop)) +
-  geom_point(alpha = 0.7) +
-  geom_smooth(method = "lm", se = FALSE) + # Add linear trend lines
-  labs(
-    title = "Average Price vs. Distance to Coast",
-    x = "Minimum Distance to Coast (meters)",
-    y = "Average Price",
-    color = "Crop Type"
-  ) +
-  theme_minimal() +
-  theme(
-    plot.title = element_text(hjust = 0.5)
-  )
-
-# (ii) Average Price vs. Minimum Distance to Road
-p_road <- ggplot(sf.marketprices, aes(x = min_road_dist, y = av_price, color = crop)) +
-  geom_point(alpha = 0.7) +
-  geom_smooth(method = "lm", se = FALSE) + # Add linear trend lines
-  labs(
-    title = "Average Price vs. Distance to Road",
-    x = "Minimum Distance to Road (meters)",
-    y = "Average Price",
-    color = "Crop Type"
-  ) +
-  theme_minimal() +
-  theme(
-    plot.title = element_text(hjust = 0.5)
-  )
-
-# (iii) Average Price vs. Minimum Distance to Airport
-p_airport <- ggplot(sf.marketprices, aes(x = min_airport_dist, y = av_price, color = crop)) +
-  geom_point(alpha = 0.7) +
-  geom_smooth(method = "lm", se = FALSE) + # Add linear trend lines
-  labs(
-    title = "Average Price vs. Distance to Airport",
-    x = "Minimum Distance to Airport (meters)",
-    y = "Average Price",
-    color = "Crop Type"
-  ) +
-  theme_minimal() +
-  theme(
-    plot.title = element_text(hjust = 0.5)
-  )
-
-# Display the plots in a grid
-grid.arrange(p_coast, p_road, p_airport, nrow = 2, ncol = 2)
-```
+# Exercise 2: Transportation Centrality

### 2025-02-12 hw3/part1_noe.R
adding code for plot 1

@@ -1,0 +1,79 @@
+rm(list=ls())
+library(terra)
+library(sf)
+library(tidyverse)
+library(exactextractr)
+
+# Import SPEI index 
+spei_index <- rast("hw3/data/spei01.nc")
+
+# Import US regions
+us_states <- spData::us_states
+
+# Get dates from SPEI dataset
+dates <- time(spei_index)
+
+# Filter only for years between 1965 and 2015
+valid_dates <- which(year(dates) >= 1965 & year(dates) <= 2015)
+
+# Extract the subset of SPEI data for our time period of interest
+spei_subset <- spei_index[[valid_dates]]
+
+# Remove from 'dates' the unnecessary dates 
+dates <- dates[valid_dates]
+
+# Zonal statistics to compute the mean (each row is a country and each column is a date)
+zonal_stats <- exact_extract(
+  spei_subset,
+  us_states,
+  fun = "mean")
+
+# Convert the wide format panel to long format panel and add state/region identifiers
+# This is necessary to compute the yearly mean per country and then the yearly mean per region
+spei_long <- zonal_stats %>%
+  as.data.frame() %>%
+  # Convert from wide to long format
+  pivot_longer(
+    cols = everything(),
+    names_to = "time_index",
+    values_to = "spei") %>%
+  # Add state and region identifiers, plus temporal information
+  mutate(
+    state = rep(us_states$NAME, each = length(dates)), 
+    region = rep(us_states$REGION, each = length(dates)),  
+    date = rep(dates, times = nrow(us_states)),  
+    year = year(date),
+    month = month(date)) %>%
+  select(state, region, year, month, spei)  
+
+# Compute the yearly mean for each state 
+spei_state_annual <- spei_long %>%
+  group_by(state, region, year) %>%
+  summarise(
+    mean_spei_state = mean(spei, na.rm = TRUE),  
+    .groups = "drop")
+
+# Compute the yearly mean for each region 
+spei_regional <- spei_state_annual %>%
+  group_by(region, year) %>%
+  summarise(
+    mean_spei_region = mean(mean_spei_state, na.rm = TRUE),
+    .groups = "drop")
+
+# Plotting
+ggplot(spei_regional, aes(x = year, y = mean_spei_region, color = region, group = region)) +
+  geom_line(linewidth = 0.8) +  
[…]

### 2025-02-12 hw3/hw3_MoritzPeistNoemiLucchiSimonVellin.Rmd
upatind .rmd file with part 1

@@
 library(readxl)
+library(terra)
+library(exactextractr)
 ```
@@
 
-...
+# Data Loading
+
+To analyze the SPEI (Standardized Precipitation Evapotranspiration Index) trends across US regions over the past 50 years, we need the NetCDF file (.nc) containing SPEI raster data and the US geographic data from spData package.
+The .nc file with data about the SPEI index has been downloaded from this [website] (https://digital.csic.es/handle/10261/268088).
+
+```{r}
+
+# Import SPEI index 
+spei_index <- rast("/Users/noemilucchi/Desktop/spei01.nc")
+
+# Import US regions
+us_states <- spData::us_states
+
+```
+
+# Data wrangling
+
+We first retrieve all dates from the SPEI index, then select only dates between 1965-2015 (our 50-year analysis period, limited by data availability).
+The zonal statistics step computes the average SPEI for each state, creating a wide-format panel where each row is a state and each column represents a specific date (month and year).
+We then transform this into a long-format panel, adding state and region identifiers along with separated year and month columns.
+This restructuring enables us to efficiently compute yearly means, first at the state level and then aggregated to the regional level.
+The separation of temporal components (year and month) facilitates these temporal aggregations while maintaining the geographic hierarchy (states nested within regions) for spatial analysis.
+
+```{r}
+
+# Get dates from SPEI dataset
+dates <- time(spei_index)
+
+# Filter only for years between 1965 and 2015
+valid_dates <- which(year(dates) >= 1965 & year(dates) <= 2015)
+
+# Extract the subset of SPEI data for our time period of interest
+spei_subset <- spei_index[[valid_dates]]
+
+# Remove from 'dates' the unnecessary dates 
+dates <- dates[valid_dates]
+
+# Zonal statistics to compute the mean (each row is a country and each column is a date)
+zonal_stats <- exact_extract(
+  spei_subset,
+  us_states,
+  fun = "mean")
+
+# Convert the wide format panel to long format panel and add state/region identifiers
+# This is necessary to compute the yearly mean per country and then the yearly mean per region
+spei_long <- zonal_stats %>%
+  as.data.frame() %>%
+  # Convert from wide to long format
+  pivot_longer(
+    cols = everything(),
+    names_to = "time_index",
+    values_to = "spei") %>%
+  # Add state and region identifiers, plus temporal information
+  mutate(
+    state = rep(us_states$NAME, each = length(dates)), 
+    region = rep(us_states$REGION, each = length(dates)),  
+    date = rep(dates, times = nrow(us_states)),  
+    year = year(date),
+    month = month(date)) %>%
+  select(state, region, year, month, spei)
+
+# Compute the yearly mean for each state 
+spei_state_annual <- spei_long %>%
+  group_by(state, region, year) %>%
+  summarise(
+    mean_spei_state = mean(spei, na.rm = TRUE),  
+    .groups = "drop")
+
+# Compute the yearly mean for each region 
+spei_regional <- spei_state_annual %>%
+  group_by(region, year) %>%
+  summarise(
+    mean_spei_region = mean(mean_spei_state, na.rm = TRUE),
+    .groups = "drop")
+
+```
+
+# Plotting
+
+After these manipulations, we can create a plot that shows the temporal trends of SPEI index for each region (colored lines) and the overall smoothed trend with confidence interval (blue line).
+
+```{r}
+
+# Plotting
+ggplot(spei_regional, aes(x = year, y = mean_spei_region, color = region, group = region)) +
+  geom_line(linewidth = 0.8) +  
+  geom_smooth(data = spei_regional, aes(x = year, y = mean_spei_region), 
+              method = "loess", color = "blue", se = TRUE, linewidth = 1.2, inherit.aes = FALSE) +
+  scale_color_brewer(palette = "Set2") +
+  scale_x_continuous(
+    breaks = seq(1970, 2010, by = 10),
+    expand = c(0.02, 0.02)) +
+  scale_y_continuous(
+    limits = c(-1, 1),  
+    breaks = seq(-1, 1, by = 1)) +
+  labs(
+    title = "SPEI Index by US Region (1965-2015)",
+    x = "Year",
+    y = "SPEI Index",
+    color = "Region")
+
+```
 

### 2025-02-12 hw3/hw3_MoritzPeistNoemiLucchiSimonVellin.Rmd
updating .rmd file with part 1

@@
 
-...
+Throughout this notebook, we will follow the same approach.
+We begin with Data Loading, where we import and prepare our necessary datasets.
+We then move to Data Manipulation, where we process and transform our data into the appropriate format for analysis.
+Finally, in the Visualization section, we create plots to effectively communicate our findings.
 
@@
 # Import SPEI index 
-spei_index <- rast("/Users/noemilucchi/Desktop/spei01.nc")
+spei_index <- rast("/hw3/data/spei01.nc")
 

### 2025-02-13 hw3/hw3_MoritzPeistNoemiLucchiSimonVellin.Rmd
Added pictures for comparisons
- Ignoring large geospatial files now - Rendering graphics in SVG and fixed notebook

@@
 ```{r setup, include=FALSE}
-# Disable warning messages
-knitr::opts_chunk$set(warning = FALSE, message = FALSE)
+# Load required package for SVG
+library(ragg)
+
+# Disable warning messages and SVG plots
+knitr::opts_chunk$set(
+  warning = FALSE,
+  message = FALSE,
+  # fig.ext = "svg",
+  fig.width = 16,
+  fig.height = 8,
+  fig.retina = 2,
+  dpi = 300,
+  dev = "svg"
+)
 ```
@@
 ```{r}
-
 # Import SPEI index
-spei_index <- rast("/hw3/data/spei01.nc")
+spei_index <- rast("data/ex1/spei01.nc")
 
@@
 us_states <- spData::us_states
-
 ```
@@
 ```{r}
-
 # Get dates from SPEI dataset
@@
   us_states,
-  fun = "mean")
+  fun = "mean"
+)
 
@@
     names_to = "time_index",
-    values_to = "spei") %>%
+    values_to = "spei"
+  ) %>%
   # Add state and region identifiers, plus temporal information
@@
     year = year(date),
-    month = month(date)) %>%
+    month = month(date)
+  ) %>%
   select(state, region, year, month, spei)
@@
     mean_spei_state = mean(spei, na.rm = TRUE),
-    .groups = "drop")
+    .groups = "drop"
+  )
 
@@
     mean_spei_region = mean(mean_spei_state, na.rm = TRUE),
-    .groups = "drop")
-
+    .groups = "drop"
+  )
 ```
@@
 ```{r}
-
 # Plotting
@@
   geom_line(linewidth = 0.8) +
-  geom_smooth(data = spei_regional, aes(x = year, y = mean_spei_region), 
-              method = "loess", color = "blue", se = TRUE, linewidth = 1.2, inherit.aes = FALSE) +
+  geom_smooth(
+    data = spei_regional, aes(x = year, y = mean_spei_region),
+    method = "loess", color = "blue", se = TRUE, linewidth = 1.2, inherit.aes = FALSE
+  ) +
   scale_color_brewer(palette = "Set2") +
@@
     breaks = seq(1970, 2010, by = 10),
-    expand = c(0.02, 0.02)) +
+    expand = c(0.02, 0.02)
+  ) +
   scale_y_continuous(
     limits = c(-1, 1),
-    breaks = seq(-1, 1, by = 1)) +
+    breaks = seq(-1, 1, by = 1)
+  ) +
   labs(
@@
     y = "SPEI Index",
-    color = "Region")
+    color = "Region"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5),
+  )
+```
 
-```
+In comparison the first graph looked as follows:
+
+![Figure 1](data/ex1/raster09.png)
 
@@
 # Exercise 2: Transportation Centrality
+
+...
+
+In comparison the first map looked as follows:
+
+![Map 1](data/ex2/raster07.png)
+
+...
+
+and the second graph as follows:
+
+![Figure 2](data/ex2/raster08.png)

### 2025-02-13 hw3/hw3_MoritzPeistNoemiLucchiSimonVellin.Rmd
Fixes

@@
 output:
+  html_document:
+    toc: true
+    df_print: paged
+  pdf_document:
+    toc: true
   html_notebook:
-    toc: true
-  pdf_document:
     toc: true
@@
   fig.retina = 2,
-  dpi = 300,
-  dev = "svg"
+  dpi = 300
+  # dev = "svg"
 )
@@
 
+<div style="text-align: center">
+
   ![Figure 1](data/ex1/raster09.png)
+
+</div>
 
@@
 
+<div style="text-align: center">
+
   ![Map 1](data/ex2/raster07.png)
+
+</div>
 
@@
 
-![Figure 2](data/ex2/raster08.png)
+<div style="text-align: center">
+
+  ![Map 2](data/ex2/raster08.png)
+
+</div>

### 2025-02-13 hw3/hw3_MoritzPeistNoemiLucchiSimonVellin.Rmd
Changed graph coloring slightly

@@
 # Plotting
-ggplot(spei_regional, aes(x = year, y = mean_spei_region, color = region, group = region)) +
-  geom_line(linewidth = 0.8) +
+ggplot(spei_regional, aes(x = year, y = mean_spei_region, color = region)) +
+  geom_line() +
   geom_smooth(
     data = spei_regional, aes(x = year, y = mean_spei_region),
-    method = "loess", color = "blue", se = TRUE, linewidth = 1.2, inherit.aes = FALSE
+    method = "loess", color = "blue", se = TRUE
   ) +
-  scale_color_brewer(palette = "Set2") +
+  scale_color_manual(values = c("red", "green", "turquoise", "purple")) +
   scale_x_continuous(

### 2025-02-13 hw3/part1_noe.R

@@ -1,11 +1,11 @@
-rm(list=ls())
+rm(list = ls())
 library(terra)
 library(sf)
 library(tidyverse)
 library(exactextractr)
 
-# Import SPEI index 
-spei_index <- rast("hw3/data/spei01.nc")
+# Import SPEI index
+spei_index <- rast("hw3/data/ex1/spei01.nc")
 
 # Import US regions
 us_states <- spData::us_states
@@ -19,14 +19,15 @@ valid_dates <- which(year(dates) >= 1965 & year(dates) <= 2015)
 # Extract the subset of SPEI data for our time period of interest
 spei_subset <- spei_index[[valid_dates]]
 
-# Remove from 'dates' the unnecessary dates 
+# Remove from 'dates' the unnecessary dates
 dates <- dates[valid_dates]
 
 # Zonal statistics to compute the mean (each row is a country and each column is a date)
 zonal_stats <- exact_extract(
   spei_subset,
   us_states,
-  fun = "mean")
+  fun = "mean"
+)
 
 # Convert the wide format panel to long format panel and add state/region identifiers
 # This is necessary to compute the yearly mean per country and then the yearly mean per region
@@ -36,44 +37,53 @@ spei_long <- zonal_stats %>%
   pivot_longer(
     cols = everything(),
     names_to = "time_index",
-    values_to = "spei") %>%
+    values_to = "spei"
+  ) %>%
   # Add state and region identifiers, plus temporal information
   mutate(
-    state = rep(us_states$NAME, each = length(dates)), 
-    region = rep(us_states$REGION, each = length(dates)),  
-    date = rep(dates, times = nrow(us_states)),  
+    state = rep(us_states$NAME, each = length(dates)),
+    region = rep(us_states$REGION, each = length(dates)),
+    date = rep(dates, times = nrow(us_states)),
     year = year(date),
-    month = month(date)) %>%
-  select(state, region, year, month, spei)  
+    month = month(date)
+  ) %>%
+  select(state, region, year, month, spei)
 
-# Compute the yearly mean for each state 
+# Compute the yearly mean for each state
 spei_state_annual <- spei_long %>%
   group_by(state, region, year) %>%
   summarise(
-    mean_spei_state = mean(spei, na.rm = TRUE),  
-    .groups = "drop")
+    mean_spei_state = mean(spei, na.rm = TRUE),
[…]

### 2025-02-13 hw3/hw3_MoritzPeistNoemiLucchiSimonVellin.Rmd
Added direct mean calculation

@@
 ```{r}
-# Main tidyverse (includes ggplot2)
-library(tidyverse)
+# Package Loading
+# Core data manipulation and visualization
+library(tidyverse) # Data manipulation and visualization suite
+library(ggplot2) # Advanced plotting (included in tidyverse)
 
-# Spatial packages
-library(sf)
-library(spData)
-library(tidyverse)
-library(units)
-library(lwgeom)
-library(gridExtra)
-library(readxl)
-library(terra)
-library(exactextractr)
+# Spatial data handling
+library(sf) # Simple Features for spatial vector data
+library(spData) # Spatial datasets
+library(terra) # Raster data handling
+library(exactextractr) # Precise spatial averaging
+
+# Additional utilities
+library(units) # Unit handling and conversion
+library(lwgeom) # Lightweight geometry operations
+library(gridExtra) # Arranging multiple plots
+library(readxl) # Excel file reading
 ```
@@
 
-# Data Loading
+## Data Description
+The analysis uses two main data sources:
+
+- SPEI (Standardized Precipitation Evapotranspiration Index) data:
+
+  - Source: digital.csic.es
+  - Format: NetCDF (.nc) file
+  - Resolution: Monthly data
+  - Variable: Drought index where negative values indicate dry conditions
+
+- US Geographic Data:
+
+  - Source: spData package
+  - Contains: State boundaries and regional classifications
+  - Regions: Northeast, Midwest, South, and West
+
+## Data Loading
 
 To analyze the SPEI (Standardized Precipitation Evapotranspiration Index) trends across US regions over the past 50 years, we need the NetCDF file (.nc) containing SPEI raster data and the US geographic data from spData package.
-The .nc file with data about the SPEI index has been downloaded from this [website] (https://digital.csic.es/handle/10261/268088).
+The .nc file with data about the SPEI index has been downloaded from this [website](https://digital.csic.es/handle/10261/268088).
 
@@
 
-# Data wrangling
+## Data Wrangling
 
 We first retrieve all dates from the SPEI index, then select only dates between 1965-2015 (our 50-year analysis period, limited by data availability).
-The zonal statistics step computes the average SPEI for each state, creating a wide-format panel where each row is a state and each column represents a specific date (month and year).
-We then transform this into a long-format panel, adding state and region identifiers along with separated year and month columns.
-This restructuring enables us to efficiently compute yearly means, first at the state level and then aggregated to the regional level.
-The separation of temporal components (year and month) facilitates these temporal aggregations while maintaining the geographic hierarchy (states nested within regions) for spatial analysis.
 
@@
 dates <- dates[valid_dates]
+```
 
+## Methodological Approaches
+
+We implement and compare two methodological approaches.
+The zonal statistics step computes the average SPEI for each state, creating a wide-format panel where each row is a state and each column represents a specific date (month and year).
+We then transform this into a long-format panel, adding state and region identifiers along with separated year and month columns.
+This restructuring enables us to efficiently compute yearly means, first at the state level and then aggregated to the regional level.
+However, this aggregation creates a `mean-of-means approach`.
+Below we present a second approach in which we calculate the mean for each region directly without the route via the states.
+The separation of temporal components (year and month) facilitates these temporal aggregations while maintaining the geographic hierarchy (states nested within regions) for spatial analysis.
+
+### State-to-Region Aggregation (Mean-of-means)
+
+This approach:
+
+- First calculates state-level means from SPEI grid cells
+- Then aggregates states to regions
+- Potential limitation: Gives equal weight to each state regardless of size
+
+```{r}
 # Zonal statistics to compute the mean (each row is a country and each column is a date)
@@
 
-# Plotting
+### Plotting
 
@@
     data = spei_regional, aes(x = year, y = mean_spei_region),
+    method = "loess", color = "blue", se = TRUE
+  ) +
+  scale_color_manual(values = c("red", "green", "turquoise", "purple")) +
+  scale_x_continuous(
+    breaks = seq(1970, 2010, by = 10),
+    expand = c(0.02, 0.02)
+  ) +
+  scale_y_continuous(
+    limits = c(-1, 1),
+    breaks = seq(-1, 1, by = 1)
+  ) +
+  labs(
+    title = "SPEI Index by US Region (1965-2015) - mean-of-means approach",
+    x = "Year",
+    y = "SPEI Index",
+    color = "Region"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5),
+  )
+```
+
+### Direct Regional Calculation
+
+Now we will use the direct mean approach by calculating means immediately on the region level across years. Key advantages:
+
+- Directly calculates regional means from SPEI grid cells
+- Maintains proper spatial weighting
+- More computationally efficient
+- Theoretically more sound for spatial averaging
+
+```{r}
+# Zonal statistics directly by region (grouping states by region first)
+regions_sf <- us_states %>%
+  group_by(REGION) %>%
+  summarise(geometry = st_union(geometry))
+
+# Calculate SPEI means directly for regions
+zonal_stats_regional <- exact_extract(
+  spei_subset,
+  regions_sf,
+  fun = "mean"
+)
+
+# Convert to long format with regional identifiers
+spei_regional_direct <- zonal_stats_regional %>%
+  as.data.frame() %>%
+  mutate(region = regions_sf$REGION) %>%
+  pivot_longer(
+    cols = -region,
+    names_to = "time_index",
+    values_to = "spei"
+  ) %>%
+  mutate(
+    date = rep(dates, times = nrow(regions_sf)),
+    year = year(date)
+  ) %>%
+  group_by(region, year) %>%
+  summarise(
+    mean_spei_region = mean(spei, na.rm = TRUE),
+    .groups = "drop"
+  )
+```
+
+### Plotting
+
+Next we plot the results and though the differences to our previous approach are minimal, the direct mean approach is more efficient and should be more what we are looking for:
+
+```{r}
+# Plotting
+ggplot(spei_regional_direct, aes(x = year, y = mean_spei_region, color = region)) +
+  geom_line() +
+  geom_smooth(
+    data = spei_regional_direct, aes(x = year, y = mean_spei_region),
     method = "loess", color = "blue", se = TRUE
@@
 
-In comparison the first graph looked as follows:
+## Results Interpretation
+
+The plots show:
+
+- SPEI trends for each region (colored lines)
+- Overall trend (blue smoothed line with confidence interval)
+- Time period: 1965-2015
+- Y-axis range: -1 to 1 (drought to wet conditions)
+
+While both approaches yield similar results due to the linear nature of averaging, the direct regional calculation is methodologically preferred for spatial climate data analysis.
+In comparison the first graph to be replicated looked as follows:
 

### 2025-02-13 hw3/part1_noe.R

@@ -22,54 +22,42 @@ spei_subset <- spei_index[[valid_dates]]
 # Remove from 'dates' the unnecessary dates
 dates <- dates[valid_dates]
 
-# Zonal statistics to compute the mean (each row is a country and each column is a date)
-zonal_stats <- exact_extract(
+# Zonal statistics directly by region (grouping states by region first)
+regions_sf <- us_states %>%
+  group_by(REGION) %>%
+  summarise(geometry = st_union(geometry))
+
+# Calculate SPEI means directly for regions
+zonal_stats_regional <- exact_extract(
   spei_subset,
-  us_states,
+  regions_sf,
   fun = "mean"
 )
 
-# Convert the wide format panel to long format panel and add state/region identifiers
-# This is necessary to compute the yearly mean per country and then the yearly mean per region
-spei_long <- zonal_stats %>%
+# Convert to long format with regional identifiers
+spei_regional_direct <- zonal_stats_regional %>%
   as.data.frame() %>%
-  # Convert from wide to long format
+  mutate(region = regions_sf$REGION) %>%
   pivot_longer(
-    cols = everything(),
+    cols = -region,
     names_to = "time_index",
     values_to = "spei"
   ) %>%
-  # Add state and region identifiers, plus temporal information
   mutate(
-    state = rep(us_states$NAME, each = length(dates)),
-    region = rep(us_states$REGION, each = length(dates)),
-    date = rep(dates, times = nrow(us_states)),
-    year = year(date),
-    month = month(date)
+    date = rep(dates, times = nrow(regions_sf)),
+    year = year(date)
   ) %>%
-  select(state, region, year, month, spei)
-
-# Compute the yearly mean for each state
-spei_state_annual <- spei_long %>%
-  group_by(state, region, year) %>%
-  summarise(
-    mean_spei_state = mean(spei, na.rm = TRUE),
-    .groups = "drop"
-  )
-
-# Compute the yearly mean for each region
-spei_regional <- spei_state_annual %>%
   group_by(region, year) %>%
   summarise(
-    mean_spei_region = mean(mean_spei_state, na.rm = TRUE),
+    mean_spei_region = mean(spei, na.rm = TRUE),
     .groups = "drop"
   )
 
 # Plotting
-ggplot(spei_regional, aes(x = year, y = mean_spei_region, color = region)) +
+ggplot(spei_regional_direct, aes(x = year, y = mean_spei_region, color = region)) +
   geom_line() +
[…]

### 2025-02-13 hw3/hw3_MoritzPeistNoemiLucchiSimonVellin.Rmd
first commit for part 2

@@
 library(exactextractr) # Precise spatial averaging
+library(gdistance) # Compute minimum distances on raster surfaces
 
@@
 
-...
+## Data Loading
+
+To calculate the distances between the 10 top populated cities in Spain, and represent the roads within the continental Spain, we used data from the [Natural Earth website](https://www.naturalearthdata.com/downloads/10m-cultural-vectors/).
+Precisely, we loaded:
+  - ne_10m_populated_places shapefile for the most populated cities 
+  - ne_10m_roads for the streets
+
+```{r}
+# Load populated cities and roads
+places <- st_read("/hw3/data/ne_10m_populated_places")
+roads <- st_read("/hw3/data/ne_10m_roads")
+```
+
+## Data Wrangling
+From the R package spData, we use the world dataset to extract the geometry of Spain, filtering only for continental Spain.
+Then, from the dataset containing the most populated cities, we extract only the cities in Spain, sort them in decreasing order of population, and retain only the top 10.
+Finally, from these top 10 cities, we extract Madrid and Vigo, which will be used in the plot.
+
+```{r}
+# Filter for continental Spain only 
+spain <- world %>% 
+  filter(name_long=='Spain') %>%
+  st_cast("POLYGON") %>%                   
+  slice_max(st_area(.))
+
+# Get top 10 most populated Spanish cities
+spain_top_10_cities <- places %>%
+  filter(ADM0NAME == "Spain") %>%
+  arrange(desc(POP_MAX)) %>%
+  slice_head(n = 10)
+
+# Filter Madrid and Vigo for the plot 
+madrid_vigo <- spain_top_10_cities %>%
+  filter(NAME %in% c("Madrid", "Vigo"))
+
+```
+
+## Methodological Approaches
+First, we transform the CRS from EPSG:4326 to EPSG:3035, as the former uses latitude and longitude, which would lead to incorrect distance calculations in kilometers.
+Next, we intersect the road network with Spain’s boundaries to restrict the analysis to roads that are actually within Spain.
+We then create an empty raster over Spain, set a high resolution, and define the CRS as EPSG:3035 for the same reason mentioned earlier.
+After that, we generate the friction surface by assigning a value of 1 to cells that contain roads and 100 to cells that do not.
+This ensures that areas without roads have higher friction, influencing the computation of the shortest path between cities.
+Subsequently, we compute the transition matrix and, finally, calculate and print the distance matrix, which includes all pairwise distances between the top 10 most populated Spanish cities.
+
+```{r}
+# Transform to planar EPSG:3035 for better distance calculation in km 
+spain_planar <- st_transform(spain, 3035)
+roads_planar <- st_transform(roads, 3035)
+cities_planar <- st_transform(spain_top_10_cities, 3035)
+
+# Intersect roads with Spain's boundaries 
+spain_roads <- st_intersection(roads_planar, spain_planar)
+
+# Create and prepare a raster on the top of Spain 
+raster <- rast(ext(spain_planar), resolution = 5000, crs = "EPSG:3035")  
+roads_vect <- vect(spain_roads)
+road_raster <- rasterize(roads_vect, raster, field = 1)
+
+# Create friction surface: we assign 100 to cells without streets and 1 with cells 
+# with streets. In this way we increase the friction for areas without streets, and so 
+# streets will be chosen when computing the distances between places. 
+road_values <- values(road_raster)
+road_values[is.na(road_values)] <- 100
+road_values[!is.na(road_values)] <- 1
+values(road_raster) <- road_values
+
+# Create transition matrix for distance calculation
+tr_matrix <- transition(raster(road_raster), 
+                        transitionFunction = mean, 
+                        directions = 8)
+
+# Calculate distances between all city pairs
+n_cities <- nrow(cities_planar)
+dist_matrix <- matrix(0, n_cities, n_cities)
+rownames(dist_matrix) <- cities_planar$NAME
+colnames(dist_matrix) <- cities_planar$NAME
+
+for(i in 1:n_cities) {
+  for(j in 1:n_cities) {
+    if(i < j) {  
+      path <- shortestPath(tr_matrix, 
+                           origin = as_Spatial(cities_planar[i,]), 
+                           goal = as_Spatial(cities_planar[j,]), 
+                           output = "SpatialLines")
+      path_sf <- st_as_sf(path)
+      st_crs(path_sf) <- st_crs(cities_planar)
+      # Calculate distance in kilometers
+      dist <- as.numeric(st_length(path_sf)) / 1000
+      dist_matrix[i,j] <- dist
+      dist_matrix[j,i] <- dist}}}
+
+# Print the distances
+print("Distances between cities (km):")
+print(round(dist_matrix, 1))
+```
+
+## Plotting 
+After these manipulations, we can plot continental Spain with its roads, highlighting Madrid and Vigo, as shown in the reference plot.
+
+```{r}
+# Plotting
+ggplot() +
+  geom_sf(data = st_transform(spain, 4326), fill = "lightgray") +
+  geom_sf(data = st_transform(spain_roads, 4326), color = "black", size = 0.3) +
+  geom_sf(data = st_transform(spain_top_10_cities, 4326), size = 1, color = "black") +
+  geom_sf(data = st_transform(madrid_vigo %>% filter(NAME == "Madrid"), 4326), 
+          aes(shape = "Madrid", color = "Madrid"), size = 3) +
+  geom_sf(data = st_transform(madrid_vigo %>% filter(NAME == "Vigo"), 4326), 
+          aes(shape = "Vigo", color = "Vigo"), size = 3) +
+  scale_color_manual(name = "Cities",
+                     values = c("Madrid" = "red", "Vigo" = "turquoise")) +
+  scale_shape_manual(name = "Cities",
+                     values = c("Madrid" = 16, "Vigo" = 17)) 
+```
 

### 2025-02-13 hw3/hw3_MoritzPeistNoemiLucchiSimonVellin.Rmd
updating .rmd for plot 1 - part 2

@@
 output:
+  html_notebook:
+    toc: true
+  pdf_document:
+    toc: true
   html_document:
@@
     df_print: paged
-  pdf_document:
-    toc: true
-  html_notebook:
-    toc: true
 ---
@@
 # Load populated cities and roads
-places <- st_read("/hw3/data/ne_10m_populated_places")
-roads <- st_read("/hw3/data/ne_10m_roads")
+places <- st_read("/hw3/data/ex2/ne_10m_populated_places")
+roads <- st_read("/hw3/data/ex2/ne_10m_roads")
 ```

### 2025-02-13 hw3/hw3_MoritzPeistNoemiLucchiSimonVellin.Rmd
knitted HTML

--- 2025-02-13 hw3/hw3_MoritzPeistNoemiLucchiSimonVellin.Rmd
@@
 # Package Loading
-# Core data manipulation and visualization
-library(tidyverse) # Data manipulation and visualization suite
-library(ggplot2) # Advanced plotting (included in tidyverse)
-
 # Spatial data handling
@@
 library(readxl) # Excel file reading
+
+# Core data manipulation and visualization
+library(tidyverse) # Data manipulation and visualization suite
+library(ggplot2) # Advanced plotting (included in tidyverse)
 ```
@@
 # Load populated cities and roads
-places <- st_read("/hw3/data/ne_10m_populated_places")
-roads <- st_read("/hw3/data/ne_10m_roads")
+places <- st_read("data/ex2/ne_10m_populated_places")
+roads <- st_read("data/ex2/ne_10m_roads")
 ```

### 2025-02-13 hw3/part2.R

@@ -1,0 +1,94 @@
+# Plot 2
+
+rm(list=ls())
+library(sf)
+library(terra)
+library(tidyverse)
+library(gdistance)
+library(spData)
+library(raster)
+
+# Load the geometry of Spain and filter for continental Spain only 
+spain <- world %>% 
+  filter(name_long=='Spain') %>%
+  st_cast("POLYGON") %>%                   
+  slice_max(st_area(.))  
+
+# Load populated cities and roads
+places <- st_read("/Users/noemilucchi/Desktop/CodeTerm2/Geospatial/ne_10m_populated_places")
+roads <- st_read("/Users/noemilucchi/Desktop/CodeTerm2/Geospatial/ne_10m_roads")
+
+# Get top 10 most populated Spanish cities
+spain_top_10_cities <- places %>%
+  filter(ADM0NAME == "Spain") %>%
+  arrange(desc(POP_MAX)) %>%
+  slice_head(n = 10)
+
+# Transform to planar EPSG:3035 for better distance calculation in km 
+spain_planar <- st_transform(spain, 3035)
+roads_planar <- st_transform(roads, 3035)
+cities_planar <- st_transform(spain_top_10_cities, 3035)
+
+# Intersect roads with Spain's boundaries 
+spain_roads <- st_intersection(roads_planar, spain_planar)
+
+# Create and prepare a raster on the top of Spain 
+raster <- rast(ext(spain_planar), resolution = 5000, crs = "EPSG:3035")  
+roads_vect <- vect(spain_roads)
+road_raster <- rasterize(roads_vect, raster, field = 1)
+
+# Create friction surface: we assign 100 to cells without streets and 1 with cells 
+# with streets. In this way we increase the friction for areas without streets, and so 
+# streets will be chosen when computing the distances between places. 
+road_values <- values(road_raster)
+road_values[is.na(road_values)] <- 100
+road_values[!is.na(road_values)] <- 1
+values(road_raster) <- road_values
+
+# Create transition matrix for distance calculation
+tr_matrix <- transition(raster(road_raster), 
+                        transitionFunction = mean, 
+                        directions = 8)
+
+# Calculate distances between all city pairs
+n_cities <- nrow(cities_planar)
+dist_matrix <- matrix(0, n_cities, n_cities)
+rownames(dist_matrix) <- cities_planar$NAME
[…]

### 2025-02-14 hw3/hw3_MoritzPeistNoemiLucchiSimonVellin.Rmd
Small formatting for ex2 part 1 and corrected file paths

@@
 ## Data Wrangling
+
 From the R package spData, we use the world dataset to extract the geometry of Spain, filtering only for continental Spain.
@@
 spain <- world %>%
-  filter(name_long=='Spain') %>%
+  filter(name_long == "Spain") %>%
   st_cast("POLYGON") %>%
@@
   filter(NAME %in% c("Madrid", "Vigo"))
-
 ```
 
-## Methodological Approaches
+## Methodological Approach
+
 First, we transform the CRS from EPSG:4326 to EPSG:3035, as the former uses latitude and longitude, which would lead to incorrect distance calculations in kilometers.
@@
   transitionFunction = mean,
-                        directions = 8)
+  directions = 8
+)
 
@@
 
-for(i in 1:n_cities) {
-  for(j in 1:n_cities) {
-    if(i < j) {  
+for (i in 1:n_cities) {
+  for (j in 1:n_cities) {
+    if (i < j) {
       path <- shortestPath(tr_matrix,
-                           origin = as_Spatial(cities_planar[i,]), 
-                           goal = as_Spatial(cities_planar[j,]), 
-                           output = "SpatialLines")
+        origin = as_Spatial(cities_planar[i, ]),
+        goal = as_Spatial(cities_planar[j, ]),
+        output = "SpatialLines"
+      )
       path_sf <- st_as_sf(path)
@@
       dist <- as.numeric(st_length(path_sf)) / 1000
-      dist_matrix[i,j] <- dist
-      dist_matrix[j,i] <- dist}}}
+      dist_matrix[i, j] <- dist
+      dist_matrix[j, i] <- dist
+    }
+  }
+}
 
@@
 ## Plotting
+
 After these manipulations, we can plot continental Spain with its roads, highlighting Madrid and Vigo, as shown in the reference plot.
@@
   geom_sf(data = st_transform(spain_top_10_cities, 4326), size = 1, color = "black") +
-  geom_sf(data = st_transform(madrid_vigo %>% filter(NAME == "Madrid"), 4326), 
-          aes(shape = "Madrid", color = "Madrid"), size = 3) +
-  geom_sf(data = st_transform(madrid_vigo %>% filter(NAME == "Vigo"), 4326), 
-          aes(shape = "Vigo", color = "Vigo"), size = 3) +
-  scale_color_manual(name = "Cities",
-                     values = c("Madrid" = "red", "Vigo" = "turquoise")) +
-  scale_shape_manual(name = "Cities",
-                     values = c("Madrid" = 16, "Vigo" = 17)) 
+  geom_sf(
+    data = st_transform(madrid_vigo %>% filter(NAME == "Madrid"), 4326),
+    aes(shape = "Madrid", color = "Madrid"), size = 3
+  ) +
+  geom_sf(
+    data = st_transform(madrid_vigo %>% filter(NAME == "Vigo"), 4326),
+    aes(shape = "Vigo", color = "Vigo"), size = 3
+  ) +
+  scale_color_manual(
+    name = "Cities",
+    values = c("Madrid" = "red", "Vigo" = "turquoise")
+  ) +
+  scale_shape_manual(
+    name = "Cities",
+    values = c("Madrid" = 16, "Vigo" = 17)
+  )
 ```

### 2025-02-14 hw3/part2.R

@@ -1,6 +1,6 @@
 # Plot 2
 
-rm(list=ls())
+rm(list = ls())
 library(sf)
 library(terra)
 library(tidyverse)
@@ -8,15 +8,15 @@ library(gdistance)
 library(spData)
 library(raster)
 
-# Load the geometry of Spain and filter for continental Spain only 
-spain <- world %>% 
-  filter(name_long=='Spain') %>%
-  st_cast("POLYGON") %>%                   
-  slice_max(st_area(.))  
+# Load the geometry of Spain and filter for continental Spain only
+spain <- world %>%
+  filter(name_long == "Spain") %>%
+  st_cast("POLYGON") %>%
+  slice_max(st_area(.))
 
 # Load populated cities and roads
-places <- st_read("/Users/noemilucchi/Desktop/CodeTerm2/Geospatial/ne_10m_populated_places")
-roads <- st_read("/Users/noemilucchi/Desktop/CodeTerm2/Geospatial/ne_10m_roads")
+places <- st_read("hw3/data/ex2/ne_10m_populated_places")
+roads <- st_read("hw3/data/ex2/ne_10m_roads")
 
 # Get top 10 most populated Spanish cities
 spain_top_10_cities <- places %>%
@@ -24,31 +24,32 @@ spain_top_10_cities <- places %>%
   arrange(desc(POP_MAX)) %>%
   slice_head(n = 10)
 
-# Transform to planar EPSG:3035 for better distance calculation in km 
+# Transform to planar EPSG:3035 for better distance calculation in km
 spain_planar <- st_transform(spain, 3035)
 roads_planar <- st_transform(roads, 3035)
 cities_planar <- st_transform(spain_top_10_cities, 3035)
 
-# Intersect roads with Spain's boundaries 
+# Intersect roads with Spain's boundaries
 spain_roads <- st_intersection(roads_planar, spain_planar)
 
-# Create and prepare a raster on the top of Spain 
-raster <- rast(ext(spain_planar), resolution = 5000, crs = "EPSG:3035")  
+# Create and prepare a raster on the top of Spain
+raster <- rast(ext(spain_planar), resolution = 5000, crs = "EPSG:3035")
 roads_vect <- vect(spain_roads)
 road_raster <- rasterize(roads_vect, raster, field = 1)
 
-# Create friction surface: we assign 100 to cells without streets and 1 with cells 
-# with streets. In this way we increase the friction for areas without streets, and so 
[…]

### 2025-02-28 hw3/hw3_MoritzPeistNoemiLucchiSimonVellin.Rmd
exercise 2 part b code and documentation done

@@
 
+# Part A
+
 ## Data Loading
@@
   ![Map 1](data/ex2/raster07.png)
-
 </div>
 
-...
+# Part B - Logical Approach
 
-and the second graph as follows:
+## Data Wrangling
+
+Using the distance matrix calculated in Part A, we now assess the isolation of Madrid and Vigo by analyzing their bilateral distances to the other top 10 most populated cities in Spain.
+We begin by extracting the distances for Madrid and Vigo.
+
+```{r}
+# Extract city names from the top 10 cities (excluding Madrid and Vigo for isolation analysis)
+other_cities <- spain_top_10_cities$NAME[!spain_top_10_cities$NAME %in% c("Madrid", "Vigo")]
+
+# Extract distances from Madrid to other cities
+madrid_distances <- dist_matrix["Madrid", other_cities]
+
+# Extract distances from Vigo to other cities
+vigo_distances <- dist_matrix["Vigo", other_cities]
+```
+
+## Analysis
+
+Isolation is determined by comparing the average distances from each city to the others, where a higher average distance indicates greater isolation.
+We compute summary statistics, and calculate the average distance for each city to make a quantitative comparison.
+
+```{r}
+# Calculate summary statistics for isolation
+madrid_summary <- summary(madrid_distances)
+vigo_summary <- summary(vigo_distances)
+
+# Print summary statistics to compare isolation
+cat("\nSummary of distances from Madrid to other top 10 cities (km):\n")
+print(round(madrid_summary, 1))
+cat("\nSummary of distances from Vigo to other top 10 cities (km):\n")
+print(round(vigo_summary, 1))
+
+# Calculate average distances to assess isolation
+madrid_avg_distance <- mean(madrid_distances)
+vigo_avg_distance <- mean(vigo_distances)
+
+cat("\nAverage distance from Madrid to other cities:", round(madrid_avg_distance, 1), "km\n")
+cat("Average distance from Vigo to other cities:", round(vigo_avg_distance, 1), "km\n")
+
+# Determine which city is more isolated (higher average distance indicates more isolation)
+if (madrid_avg_distance > vigo_avg_distance) {
+  cat("Madrid appears more isolated based on average distance.\n")
+} else if (vigo_avg_distance > madrid_avg_distance) {
+  cat("Vigo appears more isolated based on average distance.\n")
+} else {
+  cat("Madrid and Vigo have similar isolation based on average distance.\n")
+}
+```
+
+## Plotting
+
+To further explore the isolation of Madrid and Vigo, we create smoothed density distributions of their distances to the other top 10 cities.
+This visualization helps us understand the spread and concentration of distances, where a distribution with a wider spread or higher density at greater distances suggests greater isolation.
+We use geom_density() to plot the distributions, with Madrid in red and Vigo in turquoise, matching the style of the reference plot.
+
+```{r}
+# Create a data frame with distances for plotting
+density_data <- data.frame(
+  City = c(rep("Madrid", length(madrid_distances)), rep("Vigo", length(vigo_distances))),
+  Distance = c(madrid_distances, vigo_distances)
+)
+
+# Plot smoothed density distributions using geom_density()
+density_plot <- ggplot(density_data, aes(x = Distance, color = City, fill = City)) +
+  geom_density(alpha = 0.3) +  # Add transparency for overlap
+  scale_color_manual(values = c("Madrid" = "red", "Vigo" = "turquoise")) +
+  scale_fill_manual(values = c("Madrid" = "red", "Vigo" = "turquoise")) +
+  labs(
+    title = "Smoothed Density Distributions of Distances",
+    x = "Distance (km)",
+    y = "Density",
+    color = "City",
+    fill = "City"
+  ) +
+  theme_minimal()
+
+# Display the plot
+print(density_plot)
+```
+
+In comparison the second graph looked as follows:
 
@@
 </div>
+
+# Part B - Alternative Approach
+
+Although we think it seems logical to exclude Madrid and Virgo from the distance calculations, the reference plot seem to include them, as displayed by the noticeable density at x=0 for both cities.
+In this section, we computed the same plot but with "self-distances" included.
+
+```{r}
+# Extract distances for Madrid and Vigo to all cities, including themselves (self-distance = 0)
+madrid_distances_all <- dist_matrix["Madrid", ]
+vigo_distances_all <- dist_matrix["Vigo", ]
+
+# Create a data frame with distances for plotting, including self-distances
+density_data_all <- data.frame(
+  City = c(rep("Madrid", length(madrid_distances_all)), rep("Vigo", length(vigo_distances_all))),
+  Distance = c(madrid_distances_all, vigo_distances_all)
+)
+
+# Plot smoothed density distributions using geom_density(), including self-distances
+density_plot_all <- ggplot(density_data_all, aes(x = Distance, color = City, fill = City)) +
+  geom_density(alpha = 0.3) +  # Add transparency for overlap
+  scale_color_manual(values = c("Madrid" = "red", "Vigo" = "turquoise"), labels = c("Madrid", "Vigo")) +
+  scale_fill_manual(values = c("Madrid" = "red", "Vigo" = "turquoise"), labels = c("Madrid", "Vigo")) +
+  labs(
+    x = "distance",
+    y = "density",
+    color = "origin",
+    fill = "origin"
+  ) +
+  theme_minimal() +
+  theme(
+    legend.title = element_text(size = 10),
+    legend.text = element_text(size = 8)
+  )
+
+# Display the plot
+print(density_plot_all)
+```
+
+</div>

### 2025-02-28 hw3/part2.R

@@ -118,7 +118,7 @@ ggplot() +
     values = c("Madrid" = 16, "Vigo" = 17)
   )
 ---------
-# PART B
+# PART B - Logical Approach
 
 # Extract city names from the top 10 cities (excluding Madrid and Vigo themselves for isolation analysis)
 other_cities <- spain_top_10_cities$NAME[!spain_top_10_cities$NAME %in% c("Madrid", "Vigo")]
@@ -175,3 +175,36 @@ density_plot <- ggplot(density_data, aes(x = Distance, color = City, fill = City
 
 # Display the plot
 print(density_plot)
+
+---------
+# PART B - Alternative Approach
+
+# Extract distances for Madrid and Vigo to all cities, including themselves (self-distance = 0)
+madrid_distances_all <- dist_matrix["Madrid", ]
+vigo_distances_all <- dist_matrix["Vigo", ]
+
+# Create a data frame with distances for plotting, including self-distances
+density_data_all <- data.frame(
+  City = c(rep("Madrid", length(madrid_distances_all)), rep("Vigo", length(vigo_distances_all))),
+  Distance = c(madrid_distances_all, vigo_distances_all)
+)
+
+# Plot smoothed density distributions using geom_density(), including self-distances
+density_plot_all <- ggplot(density_data_all, aes(x = Distance, color = City, fill = City)) +
+  geom_density(alpha = 0.3) +  # Add transparency for overlap
+  scale_color_manual(values = c("Madrid" = "red", "Vigo" = "turquoise"), labels = c("Madrid", "Vigo")) +
+  scale_fill_manual(values = c("Madrid" = "red", "Vigo" = "turquoise"), labels = c("Madrid", "Vigo")) +
+  labs(
+    x = "distance",
+    y = "density",
+    color = "origin",
+    fill = "origin"
+  ) +
+  theme_minimal() +
+  theme(
+    legend.title = element_text(size = 10),
+    legend.text = element_text(size = 8)
+  )
+
+# Display the plot
+print(density_plot_all)

### 2025-03-02 hw3/hw3_MoritzPeistNoemiLucchiSimonVellin.Rmd
Refined 2nd part

@@
     df_print: paged
+  html_notebook:
+    toc: true
   pdf_document:
-    toc: true
-  html_notebook:
     toc: true
@@
   dpi = 300
-  # dev = "svg"
+  #dev = "svg"
 )
@@
 
-To calculate the distances between the 10 top populated cities in Spain, and represent the roads within the continental Spain, we used data from the [Natural Earth website](https://www.naturalearthdata.com/downloads/10m-cultural-vectors/).
+To calculate the distances between the 10 top populated cities in Spain, and represent the roads within continental Spain.
+We used data from the [Natural Earth website](https://www.naturalearthdata.com/downloads/10m-cultural-vectors/).
 Precisely, we loaded:
+
 - ne_10m_populated_places shapefile for the most populated cities 
@@
 
-From the R package spData, we use the world dataset to extract the geometry of Spain, filtering only for continental Spain.
+From the R package spData, we use the world dataset to extract the geometry of Spain, filtering only for continental Spain i.e., we exclude roads outside continental Spain which seems to not have happened in the initial graph were roads overflow frontiers.
 Then, from the dataset containing the most populated cities, we extract only the cities in Spain, sort them in decreasing order of population, and retain only the top 10.
@@
     name = "Cities",
-    values = c("Madrid" = "red", "Vigo" = "turquoise")
+    values = c("Madrid" = "red", "Vigo" = "blue")
   ) +
@@
 We begin by extracting the distances for Madrid and Vigo.
+Note, however, that for the average distance calculation, that we exclude the cities from the calculations with themselves to exclude the self-distance of 0.
 
@@
 # Extract city names from the top 10 cities (excluding Madrid and Vigo for isolation analysis)
-other_cities <- spain_top_10_cities$NAME[!spain_top_10_cities$NAME %in% c("Madrid", "Vigo")]
+df_dist <- as.data.frame(dist_matrix) %>%
+  rownames_to_column(var = "origin") %>%
+  # Set up long format of the matrix
+  gather(key = "destination", value = "distance", -origin) %>%
+  # Exclude all other cities
+  filter(origin %in% c("Madrid", "Vigo")) %>% 
+  # Order first by origin, then by destination
+  arrange(origin, destination)
+
+# Now exclude the self-distances
+df_dist_wo_self <- df_dist %>%
+  filter(origin != destination)
 
 # Extract distances from Madrid to other cities
-madrid_distances <- dist_matrix["Madrid", other_cities]
+madrid_distances <- df_dist_wo_self %>%
+  pull(distance)
 
 # Extract distances from Vigo to other cities
-vigo_distances <- dist_matrix["Vigo", other_cities]
+vigo_distances <- df_dist_wo_self %>%
+  pull(distance)
 ```
@@
 ```{r}
-# Create a data frame with distances for plotting
-density_data <- data.frame(
-  City = c(rep("Madrid", length(madrid_distances)), rep("Vigo", length(vigo_distances))),
-  Distance = c(madrid_distances, vigo_distances)
-)
-
 # Plot smoothed density distributions using geom_density()
-density_plot <- ggplot(density_data, aes(x = Distance, color = City, fill = City)) +
+density_plot <- ggplot(df_dist_wo_self, aes(x = distance, color = origin, fill = origin)) +
   geom_density(alpha = 0.3) +  # Add transparency for overlap
@@
 ```{r}
-# Extract distances for Madrid and Vigo to all cities, including themselves (self-distance = 0)
-madrid_distances_all <- dist_matrix["Madrid", ]
-vigo_distances_all <- dist_matrix["Vigo", ]
-
-# Create a data frame with distances for plotting, including self-distances
-density_data_all <- data.frame(
-  City = c(rep("Madrid", length(madrid_distances_all)), rep("Vigo", length(vigo_distances_all))),
-  Distance = c(madrid_distances_all, vigo_distances_all)
-)
-
 # Plot smoothed density distributions using geom_density(), including self-distances
-density_plot_all <- ggplot(density_data_all, aes(x = Distance, color = City, fill = City)) +
+density_plot_all <- ggplot(df_dist, aes(x = distance, color = origin, fill = origin)) +
   geom_density(alpha = 0.3) +  # Add transparency for overlap
@@
 ```
-
-</div>

### 2025-03-03 hw3/hw3_MoritzPeistNoemiLucchiSimonVellin.Rmd
Finalized submission 3: - notebook comments added - fixed isolation analysis

@@
 
-Next we plot the results and though the differences to our previous approach are minimal, the direct mean approach is more efficient and should be more what we are looking for:
+Next we plot the results and though the differences to our previous approach are minimal. However, the direct mean approach is more efficient and should be more what we are looking for:
 
@@
 
-While both approaches yield similar results due to the linear nature of averaging, the direct regional calculation is methodologically preferred for spatial climate data analysis.
+While both approaches yield similar results due to the linear nature of averaging, the direct regional calculation is methodologically preferred for spatial climate data analysis for several key reasons:
+
+1.
+Spatial Weighting: The mean-of-means approach gives equal weight to each state regardless of its size, potentially overrepresenting smaller states in regional averages.
+The direct approach properly weights each grid cell based on its actual area contribution to the region.
+2.
+Boundary Effects: State boundaries are political, not climatological.
+The mean-of-means approach artificially segments climate patterns along state lines before aggregation, while the direct approach preserves natural climate continuity across regions.
+3. Statistical Accuracy: Multiple averaging steps in the mean-of-means approach can introduce compounding biases. The direct calculation minimizes this by reducing the number of aggregation steps.
+4. Computational Efficiency: The direct approach requires fewer calculation steps and less memory, making it more efficient for large datasets.
+
+While both methods yield similar results in this case (due to the linear nature of averaging and relatively uniform spatial distribution), the direct regional calculation better represents the true spatial characteristics of climate data.
+
 In comparison the first graph to be replicated looked as follows:
@@
 
+## Discussion of Plot Differences
+
+After comparing both visualizations, the most likely reasons for the differences between our plot and the reference plot are:
+
+- Different SPEI Timescale: Our code specifically uses spei01.nc (1-month SPEI), while the original plot likely uses a longer timescale index such as SPEI-12 or SPEI-24.
+Longer timescales produce smoother lines as they capture longer-term drought conditions rather than monthly fluctuations.
+- Temporal Aggregation Method: Our code computes simple annual averages from monthly data, whereas the original plot might use a different aggregation approach (e.g., weighted seasonal averages or rolling averages) that reduces volatility.
+- Smoothing Parameters: The original plot appears to use more aggressive smoothing for both regional lines and the trend line.
+Our geom_smooth() with default LOESS parameters may use different span or degree settings than the original.
+- Data Pre-processing: The original may have applied additional data cleaning steps such as outlier removal or drought-specific thresholding that aren't present in our implementation.
+- Spatial Aggregation Differences: The original plot might use different regional boundary definitions or apply population-weighted rather than area-weighted averaging.
+- Year Range Subsetting: Despite both claiming 1965-2015, the exact start/end dates might differ slightly, affecting trend visualization, especially at the endpoints.
+- Different SPEI Version: The SPEI calculation methodology itself has evolved over time, and different versions of the dataset might be used.
+
+The most significant factor is likely the SPEI timescale difference, as this would fundamentally change the volatility patterns visible in the visualization.
+
 <hr>
@@
 
-# Part A
+## Part A
 
-## Data Loading
+### Data Loading
 
@@
 
-- ne_10m_populated_places shapefile for the most populated cities 
-- ne_10m_roads for the streets
+- ne_10m_populated_places: for the most populated cities 
+- ne_10m_roads: for the streets
 
@@
 
-## Data Wrangling
+### Data Wrangling
 
@@
 
-## Methodological Approach
+### Methodological Approach
 
@@
   for (j in 1:n_cities) {
-    if (i < j) {
+    if (i < j) { # Only calculate the upper triangle and mirror
       path <- shortestPath(tr_matrix,
@@
       dist_matrix[i, j] <- dist
+      # Mirror the distance in lower triangle
       dist_matrix[j, i] <- dist
@@
 
-## Plotting
+### Plotting - Map of Spain with Roads and Cities
 
@@
     values = c("Madrid" = 16, "Vigo" = 17)
+  ) +
+  theme(
+    # Make background white
+    panel.background = element_rect(fill = "white")
   )
@@
   ![Map 1](data/ex2/raster07.png)
+
 </div>
 
-# Part B - Logical Approach
+## Part B - Distance Density Plots
 
-## Data Wrangling
+### Data Wrangling
 
@@
 ```{r}
-# Extract city names from the top 10 cities (excluding Madrid and Vigo for isolation analysis)
+# Extract city names from the top 10 cities in long format
 df_dist <- as.data.frame(dist_matrix) %>%
@@
   gather(key = "destination", value = "distance", -origin) %>%
-  # Exclude all other cities
+  # Exclude all other cities from origins
   filter(origin %in% c("Madrid", "Vigo")) %>% 
@@
 madrid_distances <- df_dist_wo_self %>%
+  filter(origin == "Madrid") %>%
   pull(distance)
@@
 vigo_distances <- df_dist_wo_self %>%
+  filter(origin == "Vigo") %>%
   pull(distance)
@@
 
-## Analysis
+### Analysis
 
@@
 
-## Plotting
+### Plotting - Distance Density Plot (excluding self-distances)
 
@@
 We use geom_density() to plot the distributions, with Madrid in red and Vigo in turquoise, matching the style of the reference plot.
+However, we exclude the self-distance of 0.
 
@@
 
-In comparison the second graph looked as follows:
+In comparison the second reference graph looked as follows:
 
@@
 
-# Part B - Alternative Approach
+### Plotting - Distance Density Plot (including self-distances)
 
-Although we think it seems logical to exclude Madrid and Virgo from the distance calculations, the reference plot seem to include them, as displayed by the noticeable density at x=0 for both cities.
-In this section, we computed the same plot but with "self-distances" included.
+Although we think it seems logical to exclude the self-distances of Madrid and Vigo from the distance calculations, the reference plot seems to include them, as displayed by the noticeable density at x=0 for both cities.
+Thus, in this section, we compute the same plot but with "self-distances" included.
 
@@
 ```
+
+## Discussion of Plot Differences: Self-Distances and Calculation Parameters
+
+### Self-Distances Methodological Considerations
+
+The decision to include or exclude self-distances (city to itself) significantly impacts isolation analysis:
+
+- **Excluding self-distances** (preferred for isolation studies) focuses on inter-city relationships, avoids artificial density spikes, and aligns with network analysis conventions.
+
+- **Including self-distances** might be justified when measuring overall accessibility rather than isolation specifically, or when considering all possible origin-destination pairs in a transportation system.
+
+### Geospatial Parameters Affecting Results
+
+The differences between our plot and the reference are likely due to our choice of calculation parameters:
+
+1. **Raster Resolution** (5000 units): Finer resolution would capture road networks more precisely but at computational cost.
+
+2. **Friction Surface Values** (1:100 ratio): This strongly influences path selection and could dramatically change distances if the reference used different values.
+
+3. **Road Network Data**: Differences in the Natural Earth road dataset completeness or classification.
+
+4. **Transition Matrix Configuration**: The 8-direction setting affects possible path trajectories.
+
+5. **Continental Spain Definition**: Different methods for isolating mainland Spain could affect which roads are included.
+
+### Density Visualization Differences
+
+Additional technical factors in the visualization itself:
+
+1. **Bandwidth Parameter**: Different smoothing parameters in `geom_density()` would change curve shapes.
+
+2. **Alpha Transparency and Color Settings**: Affect visual appearance of overlapping areas.
+
+3. **Kernel Function**: Different mathematical functions for density estimation produce different curves.
+
+Despite these differences, both plots support the same conclusion that Vigo is more isolated than Madrid, with Vigo's distribution showing consistently greater distances to other population centers.

### 2025-03-15 final/final_MoritzPeistNoemiLucchiSimonVellin.Rmd
Initial commit final assignment

---
title: "R Notebook"
output: html_notebook
---

Load packages

```{r Packages}
library(data.table)   # For faster data manipulation
library(tidyverse)    # For data manipulation and visualization
library(sf)           # For spatial data handling
library(leaflet)      # For interactive maps
library(leaflet.extras) # For additional leaflet features
library(mapview)      # For easier map visualization
library(tmap)         # For thematic maps
library(tigris)       # For US road networks
library(future)       # For parallel processing
library(future.apply) # For parallel processing with apply functions

# Create directories if they don't exist
if (!dir.exists("./data/tigris")) {
  dir.create("./data/tigris", recursive = TRUE)
}

# Set custom cache directory (optional)
options(tigris_cache_dir = "./data/tigris")
# Configure tigris to use caching
options(tigris_use_cache = TRUE)
```

# Set-up convenience functions for parallel processing (as data is quite large)
```{r Parallel Buffering}
# Function to create buffers in batches with proper projection
create_buffers_in_batches <- function(sf_object, buffer_dist, batch_size = 500) {
  # First, reproject to an appropriate projected CRS for the region
  # For California, UTM Zone 10N (EPSG:26910) or 11N (EPSG:26911) works well
  # If we're unsure about your region, a web mercator projection (3857) is a reasonable default
  message("Reprojecting data to a meter-based CRS...")
  sf_object_projected <- st_transform(sf_object, 3310)  # Web Mercator

  n_features <- nrow(sf_object_projected)
  n_batches <- ceiling(n_features / batch_size)

  # Create empty list to store results
  buffer_list <- vector("list", n_batches)

  # Process in batches
  for (i in 1:n_batches) {
    start_idx <- (i-1) * batch_size + 1
    end_idx <- min(i * batch_size, n_features)

    cat(sprintf("Processing batch %d of %d (features %d to %d)\n", 
                i, n_batches, start_idx, end_idx))

    # Extract batch
    batch <- sf_object_projected[start_idx:end_idx, ]

    # Create buffer (with parallel processing within sf)
    buffer_list[[i]] <- st_buffer(batch, dist = buffer_dist)
  }

  # Combine results
  result <- do.call(rbind, buffer_list)

  # Reproject back to original CRS if needed
  message("Reprojecting results back to original CRS...")
  st_transform(result, st_crs(sf_object))
}
```

Load data (filter for California for now to limit data set size)

```{r Data Load Accidents}
# Efficient approach
df.acc <- fread("data/us_accidents/US_accidents_March23.csv")[
  # Filter date range to 2016-2021
  as.Date(Start_Time) >= as.Date("2016-01-01") & 
  as.Date(Start_Time) <= as.Date("2021-12-31") & 
  # And California
  State == "CA"
][, `:=`(
  # Add year, quarter, month columns
  year = data.table::year(Start_Time),
  quarter = data.table::quarter(Start_Time),
  month = data.table::month(Start_Time),
  # Calculate duration (assuming End_Time exists in the dataset)
  duration = as.numeric(difftime(End_Time, Start_Time, units = "days"))
)] %>% 
  as_tibble()  # Convert to tibble only at the end for performance
```

```{r Data Load Constructions}
df.const <- fread("data/us_constructions/US_constructions_Dec21.csv")[
  # Filter date range to 2016-2021
  as.Date(Start_Time) >= as.Date("2016-01-01") & 
  as.Date(Start_Time) <= as.Date("2021-12-31") & 
    # And California
  State == "CA"
][, `:=`(
  # Add year, quarter, month columns
  year = year(Start_Time),
  quarter = quarter(Start_Time),
  month = month(Start_Time),
  # Calculate duration (assuming End_Time exists in the dataset)
  duration = as.numeric(difftime(End_Time, Start_Time, units = "days"))
)] %>% 
  as_tibble()  # Convert to tibble only at the end for performance
```

Convert dataframes to sf objects

```{r Geospatial Data}
# Convert accident data to sf object
df.acc.sf <- df.acc %>%
  filter(!is.na(Start_Lat) & !is.na(Start_Lng)) %>%
  st_as_sf(coords = c("Start_Lng", "Start_Lat"), crs = 4326)

# Convert construction data to sf object
df.const.sf <- df.const %>%
  filter(!is.na(Start_Lat) & !is.na(Start_Lng)) %>%
  st_as_sf(coords = c("Start_Lng", "Start_Lat"), crs = 4326)
```

```{r Data Load Roads}
# Get US roads (this can be slow for the entire US, so we might want to limit by state)

# First get all counties in California
ca_counties <- counties("CA", year = 2021)

# Then get roads for each county
ca_roads_list <- lapply(ca_counties$COUNTYFP, function(county_code) {
  roads("CA", county = county_code, year = 2021)
})

# Optionally combine all county road data into one object
ca_roads <- do.call(rbind, ca_roads_list)
```

```{r Static Maps}
# Static map for Accidents with year coloring
# First, store your plot in a variable
ca_plot <- ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  geom_sf(data = df.acc.sf %>% filter(State == "CA"), 
          aes(color = factor(year)), alpha = 0.7, size = 0.5) +
  scale_color_viridis_d(name = "Year") +  # Use viridis color palette for discrete values
  theme_minimal() +
  labs(title = "US Road Accidents by Year (2016-2021)") +
  theme(legend.position = "bottom")
print(ca_plot)
# Then save it using ggsave
#ggsave(filename = "imgs/california_accidents.png", 
#       plot = ca_plot,
#       width = 10, # width in inches
#       height = 8, # height in inches
#       dpi = 300)  # resolution

# Construction sites map with year coloring
# First, store our plot in a variable
ca_const_plot <- ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  geom_sf(data = df.const.sf %>% filter(State == "CA"), 
          aes(color = factor(year)), alpha = 0.7, size = 0.5) +
  scale_color_viridis_d(name = "Year") +  # Use viridis color palette for discrete values
  theme_minimal() +
  labs(title = "US Construction Sites by Year (2016-2021)") +
  theme(legend.position = "bottom")
print(ca_const_plot)
# Then save it using ggsave
#ggsave(filename = "imgs/california_construction.png", 
#       plot = ca_const_plot,
#       width = 10, # width in inches
#       height = 8, # height in inches
#       dpi = 300)  # resolution
```

## Heatmaps - interactive

```{r Interactive Maps I}
# For interactive maps (often better for large datasets)
# Accidents map
accident_map <- leaflet() %>%
  addTiles() %>%
  addHeatmap(data = df.acc.sf, 
             intensity = ~1,
             radius = 8, 
             blur = 10) %>%
  setView(lng = -119.4179, lat = 36.7783, zoom = 6) %>% # Center to CA
  setMaxBounds(lng1 = -124.6, lat1 = 42.0,    # Northwest corner of CA
               lng2 = -114.1, lat2 = 32.5)    # Southeast corner of CA

# Display maps
accident_map
```

```{r Interactive Maps II}
# Construction sites map with California focus and bounds constraints
construction_map <- leaflet() %>%
  addTiles() %>%
  addHeatmap(data = df.const.sf, 
             intensity = ~1,
             radius = 8, 
             blur = 10,
             gradient = c("yellow", "orange", "red")) %>%
  setView(lng = -119.4179, lat = 36.7783, zoom = 6) %>%  # Center on California
  setMaxBounds(lng1 = -124.6, lat1 = 42.0,    # Northwest corner of CA
               lng2 = -114.1, lat2 = 32.5)    # Southeast corner of CA

# Display the map
construction_map
```

## Buffer zone analysis

```{r Buffer Zone Analysis}
# Set up parallel processing
future::plan(future::multisession, workers = parallel::detectCores() - 1)

# Set sf to use parallel processing
sf_use_s2(FALSE)  # Disable S2 spherical geometry
cores <- parallel::detectCores() - 1

# Process each dataset with batching
cat("Creating point buffers...\n")
const_buffers <- create_buffers_in_batches(df.const.sf, buffer_dist = 1000, batch_size = 500)

# Return to sequential processing
future::plan(future::sequential)
sf_use_s2(TRUE)  # Re-enable S2 spherical geometry

# Find accidents within construction zones
# Use a spatial index and parallel processing
plan(multisession, workers = parallel::detectCores() - 1)  # Use parallel processing

accidents_in_construction <- df.acc.sf %>%
  st_join(ca_const_buffers, join = st_within, prepared = TRUE)

# Reset to sequential processing when done
plan(sequential)

# Count accidents per construction zone
accidents_per_zone <- accidents_in_construction %>%
  group_by(ID) %>%  # Grouping by the ID column
  summarize(accident_count = n())

# Visualize construction zones with accident counts
# First, store our plot in a variable
accident_zones_plot <- ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  geom_sf(data = accidents_per_zone, aes(fill = accident_count), alpha = 0.7) +
  scale_fill_viridis_c() +
  theme_minimal() +
  labs(title = "Accident Counts within Construction Zones (2016-2021)")

print(accident_zones_plot)

# Then save it using ggsave
#ggsave(filename = "imgs/accident_counts_construction_zones.png", 
#       plot = accident_zones_plot,
#       width = 10, # width in inches
#       height = 8, # height in inches
#       dpi = 300)  # resolution
```

Advanced construction sites:

```{r SF linestring}
# Convert construction data to linestring sf object
df.const.lines <- df.const %>%
  filter(!is.na(Start_Lat) & !is.na(Start_Lng) & !is.na(End_Lat) & !is.na(End_Lng)) %>%
  mutate(geometry = pmap(list(Start_Lng, Start_Lat, End_Lng, End_Lat), 
                         function(start_lng, start_lat, end_lng, end_lat) {
                           coords <- matrix(c(start_lng, start_lat, 
                                             end_lng, end_lat), 
                                           ncol = 2, byrow = TRUE)
                           st_linestring(coords)
                         })) %>%
  st_sf(crs = 4326)
```

```{r Maps with Construction Lines}
# Static map with construction line segments
ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  geom_sf(data = df.const.lines %>% filter(State == "CA"), 
          color = "orange", size = 1) +
  theme_minimal() +
  labs(title = "US Construction Segments (2016-2021)")

# Interactive map with construction segments focused on California
construction_segment_map <- leaflet() %>%
  addTiles() %>%
  addPolylines(data = df.const.lines, 
               color = "orange",
               weight = 3,
               opacity = 0.7,
               popup = ~paste("ID:", ID, "<br>Duration:", Duration)) %>%
  setView(lng = -119.4179, lat = 36.7783, zoom = 6) %>%  # Center on California
  setMaxBounds(lng1 = -124.6, lat1 = 42.0,    # Northwest corner of CA
               lng2 = -114.1, lat2 = 32.5)    # Southeast corner of CA

# Display the map
construction_segment_map
```

```{r Construction Impact Analysis with Buffers}
# Set up parallel processing
future::plan(future::multisession, workers = parallel::detectCores() - 1)

# Set sf to use parallel processing
sf_use_s2(FALSE)  # Disable S2 spherical geometry
cores <- parallel::detectCores() - 1

cat("Creating line segment buffers...\n")
const_segment_buffers <- create_buffers_in_batches(df.const.lines, buffer_dist = 1000, batch_size = 250)

# Return to sequential processing
future::plan(future::sequential)
sf_use_s2(TRUE)  # Re-enable S2 spherical geometry

# Find accidents near construction segments
accidents_near_construction <- df.acc.sf %>%
  st_join(const_segment_buffers, join = st_intersects)

# Analyze accident rate per construction segment length
accidents_per_segment <- accidents_near_construction %>%
  group_by(ID) %>%  # Assuming there's an ID column for construction segments
  summarize(
    accident_count = n(),
    segment_length = as.numeric(st_length(geometry[1])),  # Length in meters
    accidents_per_km = accident_count / (segment_length/1000)
  )

# Visualize segments by accident rate
ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  geom_sf(data = accidents_per_segment, aes(color = accidents_per_km), size = 1.5) +
  scale_color_viridis_c() +
  theme_minimal() +
  labs(title = "Accident Rate per Construction Segment (2016-2021)",
       color = "Accidents/km")
```

```{r Temporal Analysis and Severity Comparison with Buffers}
# Temporal analysis - construction impact over time
accidents_by_time <- accidents_near_construction %>%
  mutate(month = floor_date(as.Date(Start_Time), "month")) %>%
  group_by(month) %>%
  summarize(accident_count = n())

# Compare accident severity near vs. away from construction
severity_comparison <- bind_rows(
  accidents_near_construction %>% 
    mutate(location = "Near construction") %>% 
    select(Severity, location),

  anti_join(df.acc.sf, accidents_near_construction, by = "ID") %>% 
    mutate(location = "Away from construction") %>% 
    select(Severity, location)
)

# Visualize severity comparison
ggplot(severity_comparison, aes(x = Severity, fill = location)) +
  geom_bar(position = "dodge") +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal() +
  labs(title = "Accident Severity: Near vs. Away from Construction",
       x = "Severity Level",
       y = "Count",
       fill = "Location")
```

### 2025-03-15 final/BackupCode.R

@@ -1,0 +1,188 @@
+## Heatmaps - static
+``
+`{r Heatmap static - accidents}
+
+# Create heatmap of accidents using ggplot2
+
+accident_heatmap <- ggplot() +
+
+  # Add California roads as base layer
+
+  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
+
+  # Create density heatmap
+
+  stat_density_2d(
+
+    data = df.acc %>% filter(!is.na(Start_Lat) & !is.na(Start_Lng)),
+
+    aes(x = Start_Lng, y = Start_Lat, fill = after_stat(density)),
+
+    geom = "tile", 
+
+    contour = FALSE,
+
+    alpha = 0.7,
+
+    h = 0.1,  # Bandwidth - adjust to match leaflet radius
+
+    n = 200   # Resolution - higher values create smoother heatmap
+
+  ) +
+
+  # Use color scheme similar to leaflet's default heatmap
+
+  scale_fill_gradientn(
+
+    colors = c("#0000FF", "#00FFFF", "#00FF00", "#FFFF00", "#FF0000"),
+
+    name = "Accident\nDensity"
+
+  ) +
+
+  # Set appropriate boundaries for California
+
+  coord_sf(
+
+    xlim = c(-124.5, -114.5),
+
+    ylim = c(32.5, 42.5)
+
+  ) +
+
+  theme_minimal() +
+
+  labs(
+
+    title = "Accident Density Heatmap in California (2016-2021)",
+
+    x = NULL,
+
+    y = NULL
+
+  ) +
+
+  theme(
+
+    legend.position = "right",
+
+    plot.title = element_text(face = "bold")
+
+  )
+
+
+
+# Display the plot
+
+print(accident_heatmap)
+
+
+
+# Save the heatmap
+
+ggsave(filename = "imgs/california_accidents_heatmap.png", 
+
+       plot = accident_heatmap,
+
+       width = 10, # width in inches
+
+       height = 8, # height in inches
+
+       dpi = 300)  # resolution
+
+`
+``
+
+``
+`{r Heatmap static - construction}
+
+# Create heatmap of construction sites using ggplot2
+
+construction_heatmap <- ggplot() +
+
+  # Add California roads as base layer
+
+  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
+
+  # Create density heatmap
+
+  stat_density_2d(
+
[…]

### 2025-03-16 final/final_MoritzPeistNoemiLucchiSimonVellin.Rmd
Added PostGIS performance improvement and switched to primary and secondary roads only

@@
 title: "R Notebook"
-output: html_notebook
+output:
+  html_document:
+    df_print: paged
 ---
@@
 library(future.apply) # For parallel processing with apply functions
+library(sf)           # For spatial data handling
+library(RPostgres)    # For PostgreSQL connection
+library(DBI)          # For database interface
 
@@
 # Get US roads (this can be slow for the entire US, so we might want to limit by state)
+# We also only get Primary and Secondary roads as they are the most important (i.e., where most constructions and accidents happen). Otherwise, the dataset would become to large (1.1GB)
+
+ca_roads <- primary_secondary_roads("CA", year = 2021)
+
+# Below code to get ALL roads for California (not recommended due to size)
 
 # First get all counties in California
-ca_counties <- counties("CA", year = 2021)
+#ca_counties <- counties("CA", year = 2021)
 
 # Then get roads for each county
-ca_roads_list <- lapply(ca_counties$COUNTYFP, function(county_code) {
-  roads("CA", county = county_code, year = 2021)
-})
+#ca_roads_list <- lapply(ca_counties$COUNTYFP, function(county_code) {
+#  roads("CA", county = county_code, year = 2021)
+#})
 
 # Optionally combine all county road data into one object
-ca_roads <- do.call(rbind, ca_roads_list)
+#ca_roads <- do.call(rbind, ca_roads_list)
 ```
@@
 # Then save it using ggsave
-#ggsave(filename = "imgs/california_accidents.png", 
-#       plot = ca_plot,
-#       width = 10, # width in inches
-#       height = 8, # height in inches
-#       dpi = 300)  # resolution
+ggsave(filename = "imgs/california_accidents.png", 
+       plot = ca_plot,
+       width = 10, # width in inches
+       height = 8, # height in inches
+       dpi = 300)  # resolution
 
@@
 # Then save it using ggsave
-#ggsave(filename = "imgs/california_construction.png", 
-#       plot = ca_const_plot,
-#       width = 10, # width in inches
-#       height = 8, # height in inches
-#       dpi = 300)  # resolution
+ggsave(filename = "imgs/california_construction.png", 
+       plot = ca_const_plot,
+       width = 10, # width in inches
+       height = 8, # height in inches
+       dpi = 300)  # resolution
+```
+
+```{r Heatmap static}
+# Create heatmap of accidents using ggplot2
+accident_heatmap <- ggplot() +
+  # Add California roads as base layer
+  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
+  # Create density heatmap
+  stat_density_2d(
+    data = df.acc %>% filter(!is.na(Start_Lat) & !is.na(Start_Lng)),
+    aes(x = Start_Lng, y = Start_Lat, fill = after_stat(density)),
+    geom = "tile", 
+    contour = FALSE,
+    alpha = 0.7,
+    h = 0.1,  # Bandwidth - adjust to match leaflet radius
+    n = 200   # Resolution - higher values create smoother heatmap
+  ) +
+  # Use color scheme similar to leaflet's default heatmap
+  scale_fill_gradientn(
+    colors = c("#0000FF", "#00FFFF", "#00FF00", "#FFFF00", "#FF0000"),
+    name = "Accident\nDensity"
+  ) +
+  # Set appropriate boundaries for California
+  coord_sf(
+    xlim = c(-124.5, -114.5),
+    ylim = c(32.5, 42.5)
+  ) +
+  theme_minimal() +
+  labs(
+    title = "Accident Density Heatmap in California (2016-2021)",
+    x = NULL,
+    y = NULL
+  ) +
+  theme(
+    legend.position = "right",
+    plot.title = element_text(face = "bold")
+  )
+# Display the plot
+print(accident_heatmap)
+```
+
+```{r Construction Heatmap static}
+# Create heatmap of construction sites using ggplot2
+construction_heatmap <- ggplot() +
+  # Add California roads as base layer
+  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
+  # Create density heatmap
+  stat_density_2d(
+    data = df.const %>% filter(!is.na(Start_Lat) & !is.na(Start_Lng)),
+    aes(x = Start_Lng, y = Start_Lat, fill = after_stat(density)),
+    geom = "tile", 
+    contour = FALSE,
+    alpha = 0.7,
+    h = 0.1,  # Bandwidth
+    n = 200   # Resolution
+  ) +
+  # Use color scheme matching your leaflet construction map
+  scale_fill_gradientn(
+    colors = c("yellow", "orange", "red"),
+    name = "Construction\nDensity"
+  ) +
+  # Set appropriate boundaries for California
+  coord_sf(
+    xlim = c(-124.5, -114.5),
+    ylim = c(32.5, 42.5)
+  ) +
+  theme_minimal() +
+  labs(
+    title = "Construction Density Heatmap in California (2016-2021)",
+    x = NULL,
+    y = NULL
+  ) +
+  theme(
+    legend.position = "right",
+    plot.title = element_text(face = "bold")
+  )
+# Display the plot
+print(construction_heatmap)
 ```
@@
 
-## Buffer zone analysis
+We construct construction sites now also as line objects since they have a start and end point
 
-```{r Buffer Zone Analysis}
-# Set up parallel processing
-future::plan(future::multisession, workers = parallel::detectCores() - 1)
-
-# Set sf to use parallel processing
-sf_use_s2(FALSE)  # Disable S2 spherical geometry
-cores <- parallel::detectCores() - 1
-
-# Process each dataset with batching
-cat("Creating point buffers...\n")
-const_buffers <- create_buffers_in_batches(df.const.sf, buffer_dist = 1000, batch_size = 500)
-
-# Return to sequential processing
-future::plan(future::sequential)
-sf_use_s2(TRUE)  # Re-enable S2 spherical geometry
-
-# Find accidents within construction zones
-# Use a spatial index and parallel processing
-plan(multisession, workers = parallel::detectCores() - 1)  # Use parallel processing
-
-accidents_in_construction <- df.acc.sf %>%
-  st_join(ca_const_buffers, join = st_within, prepared = TRUE)
-
-# Reset to sequential processing when done
-plan(sequential)
-
-# Count accidents per construction zone
-accidents_per_zone <- accidents_in_construction %>%
-  group_by(ID) %>%  # Grouping by the ID column
-  summarize(accident_count = n())
-
-# Visualize construction zones with accident counts
-# First, store our plot in a variable
-accident_zones_plot <- ggplot() +
-  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
-  geom_sf(data = accidents_per_zone, aes(fill = accident_count), alpha = 0.7) +
-  scale_fill_viridis_c() +
-  theme_minimal() +
-  labs(title = "Accident Counts within Construction Zones (2016-2021)")
-
-print(accident_zones_plot)
-
-# Then save it using ggsave
-#ggsave(filename = "imgs/accident_counts_construction_zones.png", 
-#       plot = accident_zones_plot,
-#       width = 10, # width in inches
-#       height = 8, # height in inches
-#       dpi = 300)  # resolution
-```
-
-Advanced construction sites:
-
-```{r SF linestring}
+```{r Construction linestrings}
 # Convert construction data to linestring sf object
@@
 
-```{r Maps with Construction Lines}
+## Spatial operations in PostgreSQL
+
+Since file sizes are so large and Spatial operations in R incredibly slow, we retreat to a PostgreSQL database for the following operations (PostGIS also uses the SF framework, thus R are almost identical to the following SQL functions used).
+
+```{r DB ops}
+# Create a connection to your PostgreSQL database
+con <- dbConnect(
+  Postgres(),
+  dbname = "geospatial_final",
+  host = "localhost",
+  port = 5432,
+  # Uncomment and add your credentials if needed
+  user = "postgres",
+  password = Sys.getenv("POSTGRES_PWD")
+)
+
+# Set connection parameters for performance optimization
+dbExecute(con, "SET work_mem = '128MB'")           # Increase working memory
+dbExecute(con, "SET maintenance_work_mem = '256MB'") # For index creation
+dbExecute(con, "SET random_page_cost = 1.1")       # For SSD storage
+dbExecute(con, "SET effective_cache_size = '2GB'") # Estimate of memory available
+dbExecute(con, "SET max_parallel_workers_per_gather = 4") # Use multiple cores
+
+# Check if PostGIS extension is enabled
+postgis_check <- dbGetQuery(con, "SELECT PostGIS_version()")
+cat("PostGIS version:", postgis_check[[1]], "\n")
+
+# Check if SRID 3310 is available and add if not
+dbExecute(con, "
+  DO $$
+  BEGIN
+    IF NOT EXISTS (SELECT 1 FROM spatial_ref_sys WHERE srid = 3310) THEN
+      INSERT INTO spatial_ref_sys (srid, auth_name, auth_srid, proj4text, srtext)
+      VALUES (3310, 'EPSG', 3310, 
+      '+proj=aea +lat_1=34 +lat_2=40.5 +lat_0=0 +lon_0=-120 +x_0=0 +y_0=-4000000 +datum=NAD83 +units=m +no_defs',
+      'PROJCS[\"NAD83 / California Albers\",GEOGCS[\"NAD83\",DATUM[\"North_American_Datum_1983\",SPHEROID[\"GRS 1980\",6378137,298.257222101,AUTHORITY[\"EPSG\",\"7019\"]],AUTHORITY[\"EPSG\",\"6269\"]],PRIMEM[\"Greenwich\",0,AUTHORITY[\"EPSG\",\"8901\"]],UNIT[\"degree\",0.01745329251994328,AUTHORITY[\"EPSG\",\"9122\"]],AUTHORITY[\"EPSG\",\"4269\"]],PROJECTION[\"Albers_Conic_Equal_Area\"],PARAMETER[\"standard_parallel_1\",34],PARAMETER[\"standard_parallel_2\",40.5],PARAMETER[\"latitude_of_center\",0],PARAMETER[\"longitude_of_center\",-120],PARAMETER[\"false_easting\",0],PARAMETER[\"false_northing\",-4000000],UNIT[\"metre\",1,AUTHORITY[\"EPSG\",\"9001\"]],AUTHORITY[\"EPSG\",\"3310\"]]');
+    END IF;
+  END
+  $$;
+")
+
+# Check if tables exist and only create them if they don't
+# Begin transaction for data processing
+dbExecute(con, "BEGIN")
+
+# Check if accidents table exists, if not create it
+has_accidents <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'accidents')")$exists
+if (!has_accidents) {
+  # Write accidents data to database
+  sf::st_write(df.acc.sf, con, "accidents", 
+               layer_options = c("GEOM_TYPE=GEOMETRY", 
+                                 "GEOMETRY_NAME=geometry",
+                                 "SRID=4326"))
+  cat("Written accidents data to database\n")
+
+  # Create spatial index for accidents
+  dbExecute(con, "CREATE INDEX idx_accidents_geometry ON accidents USING GIST(geometry)")
+  dbExecute(con, "ANALYZE accidents")
+  cat("Created spatial index for accidents table\n")
+}
+
+# Check if construction table exists, if not create it
+has_construction <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'construction')")$exists
+if (!has_construction) {
+  # Write construction data to database
+  sf::st_write(df.const.sf, con, "construction", 
+               layer_options = c("GEOM_TYPE=GEOMETRY", 
+                                 "GEOMETRY_NAME=geometry",
+                                 "SRID=4326"))
+  cat("Written construction data to database\n")
+
+  # Create spatial index for construction
+  dbExecute(con, "CREATE INDEX idx_construction_geometry ON construction USING GIST(geometry)")
+  dbExecute(con, "ANALYZE construction")
+  cat("Created spatial index for construction table\n")
+}
+
+# Check if construction table with lines exists, if not create it
+has_construction_lines <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'construction_lines')")$exists
+if (!has_construction_lines) {
+  # Write construction data to database
+  sf::st_write(df.const.lines, con, "construction_lines", 
+               layer_options = c("GEOM_TYPE=GEOMETRY", 
+                                 "GEOMETRY_NAME=geometry",
+                                 "SRID=4326"))
+  cat("Written construction_lines data to database\n")
+
+  # Create spatial index for construction
+  dbExecute(con, "CREATE INDEX idx_construction_lines_geometry ON construction_lines USING GIST(geometry)")
+  dbExecute(con, "ANALYZE construction_lines")
+  cat("Created spatial index for construction_lines table\n")
+}
+
+# Commit transaction for base tables
+dbExecute(con, "COMMIT")
+
+# Begin transaction for buffer creation
+dbExecute(con, "BEGIN")
+
+# Check if construction_buffers table exists, if not create it
+has_construction_buffers <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'construction_buffers')")$exists
+if (!has_construction_buffers) {
+  # Create construction buffers table
+  dbExecute(con, "
+    CREATE TABLE construction_buffers AS
+    SELECT 
+      \"ID\" as id,
+      ST_Transform(
+        ST_Buffer(
+          ST_Transform(geometry, 3310), 
+          300
+        ), 
+        4326
+      ) AS geometry
+    FROM construction
+  ")
+
+  # Add primary key and spatial index
+  dbExecute(con, "ALTER TABLE construction_buffers ADD PRIMARY KEY (id)")
+  dbExecute(con, "CREATE INDEX idx_construction_buffers_geometry ON construction_buffers USING GIST(geometry)")
+  dbExecute(con, "SELECT UpdateGeometrySRID('construction_buffers', 'geometry', 4326)")
+  dbExecute(con, "ANALYZE construction_buffers")
+  cat("Created construction_buffers table with spatial index\n")
+}
+
+# Commit buffer transaction
+dbExecute(con, "COMMIT")
+
+# Begin transaction for join table
+dbExecute(con, "BEGIN")
+
+# Check if accidents_in_construction table exists, if not create it
+has_accidents_in_construction <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'accidents_in_construction')")$exists
+# Check if accidents_in_construction table exists, if not create it
+has_accidents_in_construction <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'accidents_in_construction')")$exists
+if (!has_accidents_in_construction) {
+  # Create join table with spatial and temporal constraints
+  dbExecute(con, "
+    CREATE TABLE accidents_in_construction AS
+    SELECT 
+      a.\"ID\" AS accident_id,
+      b.id AS construction_id
+    FROM 
+      accidents a
+    JOIN 
+      construction_buffers b ON ST_Within(a.geometry, b.geometry)
+    JOIN 
+      construction c ON b.id = c.\"ID\"
+    WHERE
+      -- Join only when accident date overlaps with construction date period
+      (a.\"Start_Time\" BETWEEN c.\"Start_Time\" AND c.\"End_Time\")
+      OR 
+      (a.\"End_Time\" BETWEEN c.\"Start_Time\" AND c.\"End_Time\")
+      OR 
+      (a.\"Start_Time\" <= c.\"Start_Time\" AND a.\"End_Time\" >= c.\"End_Time\")
+  ")
+
+  # Add primary key and indices
+  dbExecute(con, "ALTER TABLE accidents_in_construction ADD PRIMARY KEY (accident_id, construction_id)")
+  dbExecute(con, "CREATE INDEX idx_aic_accident_id ON accidents_in_construction(accident_id)")
+  dbExecute(con, "CREATE INDEX idx_aic_construction_id ON accidents_in_construction(construction_id)")
+  dbExecute(con, "ANALYZE accidents_in_construction")
+  cat("Created accidents_in_construction table with indices\n")
+}
+
+# Commit join transaction
+dbExecute(con, "COMMIT")
+
+# Read the joined data back into R
+accidents_with_construction <- st_read(con, query = "
+  SELECT 
+    a.*, 
+    c.\"ID\" AS construction_id
+  FROM 
+    accidents a
+  LEFT JOIN 
+    accidents_in_construction b ON a.\"ID\" = b.accident_id
+  LEFT JOIN 
+    construction c ON b.construction_id = c.\"ID\"
+")
+
+accidents_per_zone <- st_read(con, query = "
+  SELECT 
+    c.\"ID\" AS construction_id,
+    COUNT (*) AS accident_count,
+    c.geometry
+  FROM 
+    accidents a
+  LEFT JOIN 
+    accidents_in_construction b ON a.\"ID\" = b.accident_id
+  LEFT JOIN 
+    construction c ON b.construction_id = c.\"ID\"
+  GROUP BY c.\"ID\", c.geometry
+")
+
+# Disconnect from database
+dbDisconnect(con)
+
+# Example of additional analysis with the joined data
+cat("Total accidents:", nrow(accidents_with_construction), "\n")
+cat("Accidents within construction areas:", sum(!is.na(accidents_with_construction$construction_id)), "\n")
+```
+
+## Buffer zone analysis
+
+In theory, the commented code below would achieve the same as our operation PostGIS but way slower. We will not run it here, but keep it to demonstrate that we could implement in R as well.
+
+```{r Buffer Zone Analysis}
+# Set up parallel processing
+#future::plan(future::multisession, workers = parallel::detectCores() - 1)
+
+# Set sf to use parallel processing
+#sf_use_s2(FALSE)  # Disable S2 spherical geometry
+#cores <- parallel::detectCores() - 1
+
+# Process each dataset with batching
+#cat("Creating point buffers...\n")
+#const_buffers <- create_buffers_in_batches(df.const.sf, buffer_dist = 1000, batch_size = 500)
+
+# Compute accidents within construction zones
+#accidents_in_construction <- df.acc.sf %>%
+#  st_join(const_buffers, join = st_within, prepared = TRUE)
+
+# Return to sequential processing
+#future::plan(future::sequential)
+#sf_use_s2(TRUE)  # Re-enable S2 spherical geometry
+
+# Count accidents per construction zone
+#accidents_per_zone <- accidents_with_construction %>%
+#  group_by(ID) %>%  # Grouping by the ID column
+#  summarize(accident_count = n())
+
+accidents_per_zone <- accidents_per_zone %>% drop_na()
+
+# Visualize construction zones with accident counts
+# First, store our plot in a variable
+accident_zones_plot <- ggplot() +
+  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
+  geom_sf(data = accidents_per_zone %>% arrange(accident_count), 
+          aes(color = accident_count), size = 1.5) +
+  scale_color_viridis_c() +
+  theme_minimal() +
+  labs(title = "Accident Counts within Construction Zones (2016-2021)")
+
+print(accident_zones_plot)
+
+# Then save it using ggsave
+ggsave(filename = "imgs/accident_counts_construction_zones.png", 
+       plot = accident_zones_plot,
+       width = 10, # width in inches
+       height = 8, # height in inches
+       dpi = 300)  # resolution
+```
+
+## Advanced construction sit analysis (line objects):
+
+```{r Maps with Construction Lines - static}
 # Static map with construction line segments
@@
   labs(title = "US Construction Segments (2016-2021)")
+```
 
+```{r Maps with Construction Lines - dynamic}
 # Interactive map with construction segments focused on California
@@
                opacity = 0.7,
-               popup = ~paste("ID:", ID, "<br>Duration:", Duration)) %>%
+               popup = ~paste("ID:", ID, "<br>Duration:", duration)) %>%
   setView(lng = -119.4179, lat = 36.7783, zoom = 6) %>%  # Center on California
@@
 
+Again, we execute computationally heavy operations in PostGIS but have the commented R code below to showcase a theoretical, but slow implementatiojn in R.
+
 ```{r Construction Impact Analysis with Buffers}
 # Set up parallel processing
-future::plan(future::multisession, workers = parallel::detectCores() - 1)
+#future::plan(future::multisession, workers = parallel::detectCores() - 1)
 
 # Set sf to use parallel processing
-sf_use_s2(FALSE)  # Disable S2 spherical geometry
-cores <- parallel::detectCores() - 1
+#f_use_s2(FALSE)  # Disable S2 spherical geometry
+#cores <- parallel::detectCores() - 1
 
-cat("Creating line segment buffers...\n")
-const_segment_buffers <- create_buffers_in_batches(df.const.lines, buffer_dist = 1000, batch_size = 250)
+#cat("Creating line segment buffers...\n")
+#const_segment_buffers <- create_buffers_in_batches(df.const.lines, buffer_dist = 1000, batch_size = 250)
+
+# Find accidents near construction segments
+#accidents_near_construction <- df.acc.sf %>%
+#  st_join(const_segment_buffers, join = st_intersects, prepared = TRUE)
 
 # Return to sequential processing
-future::plan(future::sequential)
-sf_use_s2(TRUE)  # Re-enable S2 spherical geometry
-
-# Find accidents near construction segments
-accidents_near_construction <- df.acc.sf %>%
-  st_join(const_segment_buffers, join = st_intersects)
+#future::plan(future::sequential)
+#sf_use_s2(TRUE)  # Re-enable S2 spherical geometry
 

### 2025-03-16 final/BackupCode.R

@@ -2,93 +2,363 @@
 ``
 `{r Heatmap static - accidents}
 
+
+
+
+
+
+
 # Create heatmap of accidents using ggplot2
+
+
+
+
+
+
 
 accident_heatmap <- ggplot() +
 
+
+
+
+
+
+
   # Add California roads as base layer
+
+
+
+
+
+
 
   geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
 
+
+
+
+
+
+
   # Create density heatmap
+
+
+
+
+
+
 
   stat_density_2d(
 
+
+
+
+
+
+
     data = df.acc %>% filter(!is.na(Start_Lat) & !is.na(Start_Lng)),
+
+
+
+
+
+
 
     aes(x = Start_Lng, y = Start_Lat, fill = after_stat(density)),
 
+
+
+
+
+
+
     geom = "tile", 
+
+
+
+
+
+
 
     contour = FALSE,
 
+
+
+
+
+
+
     alpha = 0.7,
+
+
+
+
+
+
 
     h = 0.1,  # Bandwidth - adjust to match leaflet radius
 
+
+
+
+
+
+
     n = 200   # Resolution - higher values create smoother heatmap
 
+
+
+
+
+
+
   ) +
+
+
+
+
+
+
 
   # Use color scheme similar to leaflet's default heatmap
 
+
+
+
+
+
+
   scale_fill_gradientn(
+
+
+
+
+
+
 
     colors = c("#0000FF", "#00FFFF", "#00FF00", "#FFFF00", "#FF0000"),
 
+
+
+
+
+
+
     name = "Accident\nDensity"
 
+
+
+
+
+
+
   ) +
+
+
+
+
+
+
 
   # Set appropriate boundaries for California
 
+
+
+
+
+
+
   coord_sf(
+
+
+
+
+
+
 
     xlim = c(-124.5, -114.5),
 
+
+
+
+
+
+
     ylim = c(32.5, 42.5)
 
+
+
+
+
+
+
   ) +
+
+
+
+
+
+
 
   theme_minimal() +
 
+
+
+
+
+
+
   labs(
+
+
+
+
+
+
 
     title = "Accident Density Heatmap in California (2016-2021)",
 
+
+
+
+
+
+
     x = NULL,
+
+
+
+
+
+
 
     y = NULL
 
+
+
+
+
+
+
   ) +
+
+
+
+
+
+
 
   theme(
 
+
+
+
+
+
+
[…]

### 2025-03-16 final/final_MoritzPeistNoemiLucchiSimonVellin.Rmd
Added styling for submission file

@@
 ---
-title: "R Notebook"
+title: "Geospatial Data Science - Final project"
+subtitle: "Analysis of US road accidents and construction sites"
+author: 
+- Moritz Peist
+- Noemi Lucchi
+- Simon Vellin
 output:
-  html_document:
+  html_notebook:
     df_print: paged
+    toc: true
 ---
+
+<hr>
+
+```{css, echo=FALSE}
+h1.title {
+  text-align: center;
+}
+
+h4.author, h4.date {
+  text-align: center;
+}
+
+.author {
+  text-align: center;
+}
+
+.date {
+  text-align: center;
+}
+
+/* Add styling for subtitle */
+h3.subtitle {
+  text-align: center;
+  font-style: italic;
+  margin-top: -10px;
+  color: #606060;
+}
+```
+
+<div style="text-align: center;">
+  <img src="../Barcelona_School_of_Economics_logo.svg" alt="BSE logo">
+</div>
+
+<hr>
 
@@
   geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
+  # Arrange data in descending order so high frequency zones are plotted on top
   geom_sf(data = accidents_per_zone %>% arrange(accident_count), 

### 2025-03-17 final/Test.Rmd
Added version wo PostGIS

---
title: "Geospatial Data Science - Final project"
author:
- Moritz Peist
- Noemi Lucchi
- Simon Vellin
subtitle: Analysis of US road accidents and construction sites
output:
  html_document:
    toc: true
    df_print: paged
  html_notebook:
    df_print: paged
    toc: true
---

<hr>

```{css, echo=FALSE}
h1.title {
  text-align: center;
}

h4.author, h4.date {
  text-align: center;
}

.author {
  text-align: center;
}

.date {
  text-align: center;
}

/* Add styling for subtitle */
h3.subtitle {
  text-align: center;
  font-style: italic;
  margin-top: -10px;
  color: #606060;
}
```

<div style="text-align: center;">
  <img src="../Barcelona_School_of_Economics_logo.svg" alt="BSE logo">
</div>

<hr>

Load packages

```{r Packages}
library(data.table)   # For faster data manipulation
library(tidyverse)    # For data manipulation and visualization
library(sf)           # For spatial data handling
library(leaflet)      # For interactive maps
library(leaflet.extras) # For additional leaflet features
library(mapview)      # For easier map visualization
library(tmap)         # For thematic maps
library(tigris)       # For US road networks
library(future)       # For parallel processing
library(future.apply) # For parallel processing with apply functions
library(sf)           # For spatial data handling
# Remove PostgreSQL packages as we're not using them anymore
# library(RPostgres)    # For PostgreSQL connection
# library(DBI)          # For database interface

# Create directories if they don't exist
if (!dir.exists("./data/tigris")) {
  dir.create("./data/tigris", recursive = TRUE)
}

# Set custom cache directory (optional)
options(tigris_cache_dir = "./data/tigris")
# Configure tigris to use caching
options(tigris_use_cache = TRUE)
```

# Set-up convenience functions for parallel processing (as data is quite large)
```{r Parallel Buffering}
# Function to create buffers in batches with proper projection
create_buffers_in_batches <- function(sf_object, buffer_dist, batch_size = 500) {
  # First, reproject to an appropriate projected CRS for the region
  # For California, UTM Zone 10N (EPSG:26910) or 11N (EPSG:26911) works well
  # If we're unsure about your region, a web mercator projection (3857) is a reasonable default
  message("Reprojecting data to a meter-based CRS...")
  sf_object_projected <- st_transform(sf_object, 3310)  # California Albers

  n_features <- nrow(sf_object_projected)
  n_batches <- ceiling(n_features / batch_size)

  # Create empty list to store results
  buffer_list <- vector("list", n_batches)

  # Process in batches
  for (i in 1:n_batches) {
    start_idx <- (i-1) * batch_size + 1
    end_idx <- min(i * batch_size, n_features)

    cat(sprintf("Processing batch %d of %d (features %d to %d)\n", 
                i, n_batches, start_idx, end_idx))

    # Extract batch
    batch <- sf_object_projected[start_idx:end_idx, ]

    # Create buffer (with parallel processing within sf)
    buffer_list[[i]] <- st_buffer(batch, dist = buffer_dist)
  }

  # Combine results
  result <- do.call(rbind, buffer_list)

  # Reproject back to original CRS if needed
  message("Reprojecting results back to original CRS...")
  st_transform(result, st_crs(sf_object))
}
```

Load data (filter for California for now to limit data set size)

```{r Data Load Accidents}
# Efficient approach
df.acc <- fread("data/us_accidents/US_accidents_March23.csv")[
  # Filter date range of 2021
  lubridate::year(as.Date(Start_Time)) == 2021 & 
  # And California
  State == "CA"
][, `:=`(
  # Add year, quarter, month columns
  year = data.table::year(Start_Time),
  quarter = data.table::quarter(Start_Time),
  month = data.table::month(Start_Time),
  # Calculate duration (assuming End_Time exists in the dataset)
  duration = as.numeric(difftime(End_Time, Start_Time, units = "days"))
)] %>% 
  as_tibble()  # Convert to tibble only at the end for performance
```

```{r Data Load Constructions}
df.const <- fread("data/us_constructions/US_constructions_Dec21.csv")[
  # Filter date range of 2021
  lubridate::year(as.Date(Start_Time)) == 2021 & 
  # And California
  State == "CA"
][, `:=`(
  # Add year, quarter, month columns
  year = year(Start_Time),
  quarter = quarter(Start_Time),
  month = month(Start_Time),
  # Calculate duration (assuming End_Time exists in the dataset)
  duration = as.numeric(difftime(End_Time, Start_Time, units = "days"))
)] %>% 
  as_tibble()  # Convert to tibble only at the end for performance
```

Convert dataframes to sf objects

```{r Geospatial Data}
# Convert accident data to sf object
df.acc.sf <- df.acc %>%
  filter(!is.na(Start_Lat) & !is.na(Start_Lng)) %>%
  st_as_sf(coords = c("Start_Lng", "Start_Lat"), crs = 4326)

# Convert construction data to sf object
df.const.sf <- df.const %>%
  filter(!is.na(Start_Lat) & !is.na(Start_Lng)) %>%
  st_as_sf(coords = c("Start_Lng", "Start_Lat"), crs = 4326)
```

```{r Data Load Roads}
# Get US roads (this can be slow for the entire US, so we might want to limit by state)
# We also only get Primary and Secondary roads as they are the most important (i.e., where most constructions and accidents happen). Otherwise, the dataset would become to large (1.1GB)

ca_roads <- primary_secondary_roads("CA", year = 2021)

# Below code to get ALL roads for California (not recommended due to size)

# First get all counties in California
#ca_counties <- counties("CA", year = 2021)

# Then get roads for each county
#ca_roads_list <- lapply(ca_counties$COUNTYFP, function(county_code) {
#  roads("CA", county = county_code, year = 2021)
#})

# Optionally combine all county road data into one object
#ca_roads <- do.call(rbind, ca_roads_list)
```

```{r Static Maps}
# Static map for Accidents with year coloring
# First, store your plot in a variable
ca_plot <- ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  geom_sf(data = df.acc.sf %>% filter(State == "CA"), 
          aes(color = factor(year)), alpha = 0.7, size = 0.5) +
  scale_color_viridis_d(name = "Year") +  # Use viridis color palette for discrete values
  theme_minimal() +
  labs(title = "US Road Accidents by Year (2016-2021)") +
  theme(legend.position = "bottom")
print(ca_plot)
# Then save it using ggsave
ggsave(filename = "imgs/california_accidents.png", 
       plot = ca_plot,
       width = 10, # width in inches
       height = 8, # height in inches
       dpi = 300)  # resolution

# Construction sites map with year coloring
# First, store our plot in a variable
ca_const_plot <- ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  geom_sf(data = df.const.sf %>% filter(State == "CA"), 
          aes(color = factor(year)), alpha = 0.7, size = 0.5) +
  scale_color_viridis_d(name = "Year") +  # Use viridis color palette for discrete values
  theme_minimal() +
  labs(title = "US Construction Sites by Year (2016-2021)") +
  theme(legend.position = "bottom")
print(ca_const_plot)
# Then save it using ggsave
ggsave(filename = "imgs/california_construction.png", 
       plot = ca_const_plot,
       width = 10, # width in inches
       height = 8, # height in inches
       dpi = 300)  # resolution
```

```{r Heatmap static}
# Create heatmap of accidents using ggplot2
# Modified accident heatmap with transparency for low density areas
accident_heatmap <- ggplot() +
  # Add California roads as base layer
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  # Create density heatmap with transparency threshold
  stat_density_2d(
    data = df.acc %>% filter(!is.na(Start_Lat) & !is.na(Start_Lng)),
    aes(x = Start_Lng, y = Start_Lat, fill = after_stat(density), alpha = after_stat(density)),
    geom = "tile", 
    contour = FALSE,
    h = 0.1,
    n = 200
  ) +
  # Set alpha scaling with a minimum threshold
  scale_alpha_continuous(range = c(0, 0.9), guide = "none") +
  # Use color scheme similar to leaflet's default heatmap
  scale_fill_gradientn(
    colors = c("#0000FF", "#00FFFF", "#00FF00", "#FFFF00", "#FF0000"),
    name = "Accident\nDensity"
  ) +
  coord_sf(
    xlim = c(-124.5, -114.5),
    ylim = c(32.5, 42.5)
  ) +
  theme_minimal() +
  labs(
    title = "Accident Density Heatmap in California (2016-2021)",
    x = NULL,
    y = NULL
  ) +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold")
  )
# Display the plot
print(accident_heatmap)

ggsave(filename = "imgs/accident_heatmap.png", 
       plot = accident_heatmap,
       width = 10, # width in inches
       height = 8, # height in inches
       dpi = 300)  # resolution

```

```{r Construction Heatmap static}
# Create heatmap of construction sites using ggplot2
construction_heatmap <- ggplot() +
  # Add California roads as base layer
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  # Create density heatmap
  stat_density_2d(
    data = df.const %>% filter(!is.na(Start_Lat) & !is.na(Start_Lng)),
    aes(x = Start_Lng, y = Start_Lat, fill = after_stat(density)),
    geom = "tile", 
    contour = FALSE,
    alpha = 0.7,
    h = 0.1,  # Bandwidth
    n = 200   # Resolution
  ) +
  # Set alpha scaling with a minimum threshold
  scale_alpha_continuous(range = c(0, 0.9), guide = "none") +
  # Use color scheme matching your leaflet construction map
  scale_fill_gradientn(
    colors = c("yellow", "orange", "red"),
    name = "Construction\nDensity"
  ) +
  # Set appropriate boundaries for California
  coord_sf(
    xlim = c(-124.5, -114.5),
    ylim = c(32.5, 42.5)
  ) +
  theme_minimal() +
  labs(
    title = "Construction Density Heatmap in California (2016-2021)",
    x = NULL,
    y = NULL
  ) +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold")
  )
# Display the plot
print(construction_heatmap)
```

## Heatmaps - interactive

```{r Interactive Maps I}
# For interactive maps (often better for large datasets)
# Accidents map
accident_map <- leaflet() %>%
  addTiles() %>%
  addHeatmap(data = df.acc.sf, 
             intensity = ~1,
             radius = 8, 
             blur = 10) %>%
  setView(lng = -119.4179, lat = 36.7783, zoom = 6) %>% # Center to CA
  setMaxBounds(lng1 = -124.6, lat1 = 42.0,    # Northwest corner of CA
               lng2 = -114.1, lat2 = 32.5)    # Southeast corner of CA

# Display maps
accident_map
```

```{r Interactive Maps II}
# Construction sites map with California focus and bounds constraints
construction_map <- leaflet() %>%
  addTiles() %>%
  addHeatmap(data = df.const.sf, 
             intensity = ~1,
             radius = 8, 
             blur = 10,
             gradient = c("yellow", "orange", "red")) %>%
  setView(lng = -119.4179, lat = 36.7783, zoom = 6) %>%  # Center on California
  setMaxBounds(lng1 = -124.6, lat1 = 42.0,    # Northwest corner of CA
               lng2 = -114.1, lat2 = 32.5)    # Southeast corner of CA

# Display the map
construction_map
```

We construct construction sites now also as line objects since they have a start and end point

```{r Construction linestrings}
# Convert construction data to linestring sf object
df.const.lines <- df.const %>%
  filter(!is.na(Start_Lat) & !is.na(Start_Lng) & !is.na(End_Lat) & !is.na(End_Lng)) %>%
  mutate(geometry = pmap(list(Start_Lng, Start_Lat, End_Lng, End_Lat), 
                         function(start_lng, start_lat, end_lng, end_lat) {
                           coords <- matrix(c(start_lng, start_lat, 
                                             end_lng, end_lat), 
                                           ncol = 2, byrow = TRUE)
                           st_linestring(coords)
                         })) %>%
  st_sf(crs = 4326)
```

## Spatial operations in R (replacing PostGIS)

```{r R Spatial Operations}
# Set up parallel processing to speed up computations
future::plan(future::multisession, workers = parallel::detectCores() - 1)

cat("Creating construction buffers...\n")

# Step 1A: Create construction point buffers
# Transform to meter-based CRS, buffer, and transform back
construction_buffers <- df.const.sf %>%
  # Add ID as a primary key if it doesn't exist
  mutate(construction_id = if("ID" %in% names(.)) ID else row_number()) %>%
  # Select only needed columns
  select(construction_id, Start_Time, End_Time, geometry) %>%
  # Rename columns to match our schema
  rename(start_time = Start_Time, end_time = End_Time) %>%
  # Create the buffer using our batch function
  create_buffers_in_batches(buffer_dist = 300, batch_size = 500)

cat("Creating construction line buffers...\n")

# Calculate segment length in meters
df.const.lines_with_length <- df.const.lines %>%
  # Transform to meter-based CRS
  st_transform(3310) %>%
  # Calculate length in meters
  mutate(segment_length_m = as.numeric(st_length(geometry))) %>%
  # Transform back to WGS84
  st_transform(4326)

# Step 1A-L: Create construction line buffers
construction_line_buffers <- df.const.lines_with_length %>%
  # Add ID as a primary key if it doesn't exist
  mutate(construction_id = if("ID" %in% names(.)) ID else row_number()) %>%
  # Select only needed columns
  select(construction_id, Start_Time, End_Time, segment_length_m, geometry) %>%
  # Rename columns to match our schema
  rename(start_time = Start_Time, end_time = End_Time) %>%
  # Create the buffer using our batch function
  create_buffers_in_batches(buffer_dist = 300, batch_size = 250)

# Step 1B: Spatial join to find accidents within construction buffers
cat("Finding accidents within construction buffers...\n")
spatial_matches <- df.acc.sf %>%
  # Prepare accident data
  mutate(accident_id = if("ID" %in% names(.)) ID else row_number(),
         accident_severity = if("Severity" %in% names(.)) Severity else NA_integer_,
         accident_time = Start_Time) %>%
  # Select only needed columns
  select(accident_id, accident_severity, accident_time, geometry) %>%
  # Spatial join with construction buffers
  st_join(construction_buffers, join = st_within) %>%
  # Filter for accidents within temporal range of construction
  filter(!is.na(construction_id) & 
         accident_time >= start_time & 
         accident_time <= end_time) %>%
  # Select final columns
  select(accident_id, accident_severity, accident_time, construction_id)

# Step 1B-L: Spatial join to find accidents within construction line buffers
cat("Finding accidents within construction line buffers...\n")
line_spatial_matches <- df.acc.sf %>%
  # Prepare accident data
  mutate(accident_id = if("ID" %in% names(.)) ID else row_number(),
         accident_severity = if("Severity" %in% names(.)) Severity else NA_integer_,
         accident_time = Start_Time) %>%
  # Select only needed columns
  select(accident_id, accident_severity, accident_time, geometry) %>%
  # Spatial join with construction line buffers
  st_join(construction_line_buffers, join = st_within) %>%
  # Filter for accidents within temporal range of construction
  filter(!is.na(construction_id) & 
         accident_time >= start_time & 
         accident_time <= end_time) %>%
  # Select final columns
  select(accident_id, accident_severity, accident_time, construction_id, segment_length_m)

# Step 2: Count construction sites per accident
cat("Counting construction sites per accident...\n")
construction_counts <- spatial_matches %>%
  # Convert to regular dataframe (don't need geometry anymore)
  st_set_geometry(NULL) %>%
  # Group and count
  group_by(accident_id) %>%
  summarize(nearby_construction_count = n())

# Step 2-L: Count construction line sites per accident
cat("Counting construction line sites per accident...\n")
line_construction_counts <- line_spatial_matches %>%
  # Convert to regular dataframe (don't need geometry anymore)
  st_set_geometry(NULL) %>%
  # Group and count
  group_by(accident_id) %>%
  summarize(nearby_line_construction_count = n())

# Step 3: Final analysis table - one construction site per accident
cat("Creating final analysis tables...\n")
# Create a function to randomly select one construction site per accident
select_random_construction <- function(df) {
  df %>%
    group_by(accident_id) %>%
    slice_sample(n = 1) %>%
    ungroup()
}

# Apply for point construction sites
accident_construction_analysis <- spatial_matches %>%
  st_set_geometry(NULL) %>%
  # Join with counts
  left_join(construction_counts, by = "accident_id") %>%
  # Add flags
  mutate(
    construction_type = 'POINT',
    temporal_overlap = 1,
    is_near_active_construction = 1
  ) %>%
  # Select one construction site per accident randomly
  select_random_construction()

# Apply for line construction sites
line_accident_construction_analysis <- line_spatial_matches %>%
  st_set_geometry(NULL) %>%
  # Join with counts
  left_join(line_construction_counts, by = "accident_id") %>%
  # Add flags
  mutate(
    temporal_overlap = 1,
    is_near_active_construction = 1
  ) %>%
  # Select one construction site per accident randomly
  select_random_construction()

# Step 4-L: Calculate accident rates by segment length
cat("Calculating accident rates by segment length...\n")
construction_segment_accident_rates <- df.const.lines_with_length %>%
  # Ensure we have an ID column
  mutate(construction_id = if("ID" %in% names(.)) ID else row_number()) %>%
  # Left join with accident analysis
  left_join(
    line_accident_construction_analysis %>%
      group_by(construction_id) %>%
      summarize(accident_count = n()),
    by = "construction_id"
  ) %>%
  # Replace NA counts with 0
  mutate(accident_count = replace_na(accident_count, 0)) %>%
  # Calculate accidents per km
  mutate(accidents_per_km = accident_count / (segment_length_m / 1000))

# Return to sequential processing
future::plan(future::sequential)

# Total accidents summary - for comparison
total_accidents <- nrow(df.acc.sf)
accidents_near_construction <- nrow(accident_construction_analysis)

cat("Total accidents in dataset:", total_accidents, "\n")
cat("Accidents near active construction:", accidents_near_construction, "\n")
cat("Percentage of accidents near active construction: ", 
    round(accidents_near_construction/total_accidents*100, 2), "%\n")

# Get statistics on construction zones
accidents_per_zone <- construction_segment_accident_rates %>%
  filter(accident_count > 0) %>%
  select(construction_id, accident_count, geometry)

cat("Construction zones with accidents:", nrow(accidents_per_zone), "\n")
cat("Maximum accidents in a single construction zone:", max(accidents_per_zone$accident_count), "\n")
cat("Average accidents per construction zone:", round(mean(accidents_per_zone$accident_count), 2), "\n")

# Join accident data with construction info
accidents_with_construction <- df.acc.sf %>%
  mutate(accident_id = if("ID" %in% names(.)) ID else row_number()) %>%
  left_join(
    accident_construction_analysis %>%
      select(accident_id, construction_id, nearby_construction_count, construction_type),
    by = "accident_id"
  ) %>%
  mutate(near_construction = ifelse(is.na(construction_id), 0, 1))

# Construction density statistics
construction_density_stats <- accident_construction_analysis %>%
  group_by(nearby_construction_count) %>%
  summarize(accident_count = n()) %>%
  arrange(nearby_construction_count)

print(construction_density_stats)

# Severity analysis near line construction
severity_by_construction <- line_accident_construction_analysis %>%
  group_by(accident_severity) %>%
  summarize(count = n()) %>%
  arrange(accident_severity)

# Comparison of accident severity near vs. away from construction
accidents_near <- accidents_with_construction %>%
  filter(near_construction == 1) %>%
  st_set_geometry(NULL) %>%
  mutate(location = "Near construction") %>%
  select(Severity, location)

accidents_away <- accidents_with_construction %>%
  filter(near_construction == 0) %>%
  st_set_geometry(NULL) %>%
  mutate(location = "Away from construction") %>%
  select(Severity, location)

severity_comparison <- bind_rows(accidents_near, accidents_away) %>%
  group_by(Severity, location) %>%
  summarize(count = n(), .groups = "drop") %>%
  arrange(Severity, location)

# Time-based analysis
temporal_analysis <- line_accident_construction_analysis %>%
  mutate(month = floor_date(accident_time, "month")) %>%
  group_by(month) %>%
  summarize(accident_count = n()) %>%
  arrange(month)
```

## Buffer zone analysis

The code below implements the PostGIS functionality directly in R.

```{r Buffer Zone Analysis}
# Visualize construction zones with accident counts
# First, store our plot in a variable
accident_zones_plot <- ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  # Arrange data in descending order so high frequency zones are plotted on top
  geom_sf(data = accidents_per_zone %>% arrange(accident_count), 
          aes(color = accident_count), size = 1.5) +
  scale_color_viridis_c() +
  theme_minimal() +
  labs(title = "Accident Counts within Construction Zones (2021)")

print(accident_zones_plot)

# Then save it using ggsave
ggsave(filename = "imgs/accident_counts_construction_zones.png", 
       plot = accident_zones_plot,
       width = 10, # width in inches
       height = 8, # height in inches
       dpi = 300)  # resolution
```

## Advanced construction sit analysis (line objects):

```{r Maps with Construction Lines - static}
# Static map with construction line segments
ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  geom_sf(data = df.const.lines %>% filter(State == "CA"), 
          color = "orange", size = 1) +
  theme_minimal() +
  labs(title = "US Construction Segments (2021)")
```

```{r Maps with Construction Lines - dynamic}
# Interactive map with construction segments focused on California
construction_segment_map <- leaflet() %>%
  addTiles() %>%
  addPolylines(data = df.const.lines, 
               color = "orange",
               weight = 3,
               opacity = 0.7,
               popup = ~paste("ID:", ID, "<br>Duration:", duration)) %>%
  setView(lng = -119.4179, lat = 36.7783, zoom = 6) %>%  # Center on California
  setMaxBounds(lng1 = -124.6, lat1 = 42.0,    # Northwest corner of CA
               lng2 = -114.1, lat2 = 32.5)    # Southeast corner of CA

# Display the map
construction_segment_map
```

```{r Construction Impact Analysis with Buffers}
# Visualize segments by accident rate
ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  geom_sf(data = construction_segment_accident_rates, aes(color = accidents_per_km), size = 1.5) +
  scale_color_viridis_c() +
  theme_minimal() +
  labs(title = "Accident Rate per Construction Segment (2016-2021)",
       color = "Accidents/km")
```

```{r Temporal Analysis and Severity Comparison with Buffers}
# Visualize temporal analysis
ggplot(temporal_analysis, aes(x = month, y = accident_count)) +
  geom_line() +
  geom_point() +
  #scale_x_date(date_labels = "%b %Y", date_breaks = "1 month") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Accidents Near Construction Over Time",
       x = "Month",
       y = "Accident Count")

# Visualize severity comparison
ggplot(severity_comparison, aes(x = Severity, y = count, fill = location)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal() +
  labs(title = "Accident Severity: Near vs. Away from Construction",
       x = "Severity Level",
       y = "Count",
       fill = "Location")
```

### 2025-03-18 final/final_wo_PostGIS.Rmd
Added a streamlined version

--- 2025-03-17 final/Test.Rmd
@@
 ```
+
+## Regressions
+
+### 1. Panel Data Set
+
+First, we set-up a panel data set for our regressions.
+```{r}
+# Create road segment panel
+# Step 1: Define time periods (monthly for 2021)
+time_periods <- seq(from = as.Date("2021-01-01"), 
+                    to = as.Date("2021-12-01"), 
+                    by = "month")
+
+# Step 2: Create road segments (either use existing or create grid cells)
+# Option A: Use existing road segments from ca_roads
+road_segments <- ca_roads %>%
+  mutate(segment_id = row_number()) %>%
+  select(segment_id, LINEARID, FULLNAME, geometry)
+
+# Step 3: Generate panel by creating all combinations of segments and time periods
+panel_data <- expand.grid(
+  segment_id = road_segments$segment_id,
+  month = time_periods
+) %>%
+  as_tibble()
+
+# Step 4: Join spatial information
+panel_data <- panel_data %>%
+  left_join(road_segments %>% st_set_geometry(NULL) %>% select(segment_id), 
+            by = "segment_id")
+```
+
+### 2. Treatment variables
+
+We identify which segments had construction during which periods:
+```{r}
+# Identify construction treatment status for each segment-month
+construction_panel <- df.const.sf %>%
+  st_transform(st_crs(road_segments)) %>%
+  # Create month variable
+  mutate(month = floor_date(Start_Time, "month")) %>%
+  # Spatial join with road segments
+  st_join(road_segments) %>%
+  # Group by segment and month
+  group_by(segment_id, month) %>%
+  # Create treatment indicator
+  summarize(
+    has_construction = 1,
+    construction_count = n(),
+    .groups = "drop"
+  ) %>%
+  st_set_geometry(NULL)
+
+# Join treatment to panel
+panel_data <- panel_data %>%
+  left_join(construction_panel, by = c("segment_id", "month")) %>%
+  mutate(has_construction = ifelse(is.na(has_construction), 0, 1))
+```
+
+### 3. Outcome variables
+
+We count accidents by segment and time period:
+
+```{r}
+# Count accidents by segment and month
+accident_counts <- df.acc.sf %>%
+  st_transform(st_crs(road_segments)) %>%
+  # Create month variable
+  mutate(month = floor_date(Start_Time, "month")) %>%
+  # Spatial join with road segments 
+  st_join(road_segments) %>%
+  # Group by segment and month
+  group_by(segment_id, month) %>%
+  # Count accidents and summarize severity
+  summarize(
+    accident_count = n(),
+    severe_accident_count = sum(Severity >= 3, na.rm = TRUE),
+    avg_severity = mean(Severity, na.rm = TRUE),
+    .groups = "drop"
+  ) %>%
+  st_set_geometry(NULL)
+
+# Join accidents to panel
+panel_data <- panel_data %>%
+  left_join(accident_counts, by = c("segment_id", "month")) %>%
+  mutate(
+    accident_count = replace_na(accident_count, 0),
+    severe_accident_count = replace_na(severe_accident_count, 0),
+    avg_severity = replace_na(avg_severity, 0)
+  )
+```
+
+### 4. Feature Engineering
+
+We add several control variables:
+
+```{r}
+# 1. Add urban/rural classification (crucial for your research question)
+# Get Census urban areas
+urban_areas <- urban_areas(year = 2021)
+
+# Identify if road segments are in urban areas
+road_segments_urban <- road_segments %>%
+  st_transform(st_crs(urban_areas)) %>%
+  st_join(urban_areas %>% select(UACE, NAME, UATYP), join = st_intersects) %>%
+  mutate(
+    is_urban = !is.na(UACE),
+    urban_type = ifelse(is.na(UATYP), "Rural", UATYP),
+    urban_name = NAME
+  ) %>%
+  select(segment_id, is_urban, urban_type, urban_name)
+
+# Join urban classification to panel
+panel_data <- panel_data %>%
+  left_join(road_segments_urban %>% st_set_geometry(NULL) %>% 
+              select(segment_id, is_urban, urban_type), 
+            by = "segment_id")
+
+# 2. Add segment characteristics from road data
+segment_chars <- road_segments %>%
+  mutate(
+    road_type = case_when(
+      grepl("Interstate", RTTYP) ~ "Interstate",
+      grepl("U.S. Highway", RTTYP) ~ "US Highway",
+      grepl("State Highway", RTTYP) ~ "State Highway",
+      TRUE ~ "Other"
+    ),
+    segment_length_km = as.numeric(st_length(geometry))/1000
+  ) %>%
+  select(segment_id, road_type, segment_length_km) %>%
+  st_set_geometry(NULL)
+
+# Join segment characteristics
+panel_data <- panel_data %>%
+  left_join(segment_chars, by = "segment_id")
+
+# 3. Add time-based features
+panel_data <- panel_data %>%
+  mutate(
+    year = year(month),
+    month_num = month(month),
+    quarter = quarter(month),
+    is_holiday_month = month_num %in% c(7, 11, 12),  # July, Nov, Dec
+    is_winter = month_num %in% c(12, 1, 2),
+    is_summer = month_num %in% c(6, 7, 8)
+  )
+
+# 4. Normalize accident counts by segment length
+panel_data <- panel_data %>%
+  mutate(
+    accidents_per_km = accident_count / segment_length_km,
+    severe_accidents_per_km = severe_accident_count / segment_length_km
+  )
+```
+
+And finally we also add the weather data from the initial accidents
+
+```{r}
+# Extract weather data from accidents dataset
+weather_by_segment_month <- df.acc.sf %>%
+  st_transform(st_crs(road_segments)) %>%
+  mutate(month = floor_date(Start_Time, "month")) %>%
+  st_join(road_segments) %>%
+  group_by(segment_id, month) %>%
+  summarize(
+    avg_temperature = mean(Temperature, na.rm = TRUE),
+    avg_precipitation = mean(Precipitation, na.rm = TRUE),
+    pct_adverse_weather = mean(Weather_Condition %in% 
+                              c("Rain", "Snow", "Fog", "Thunderstorm"), na.rm = TRUE),
+    .groups = "drop"
+  ) %>%
+  st_set_geometry(NULL)
+
+# Join weather data
+panel_data <- panel_data %>%
+  left_join(weather_by_segment_month, by = c("segment_id", "month"))
+```
+
+### 5. Regression time
+
+Now we can run the regression models based on our DiD design
+
+```{r}
+library(fixest) # For efficient fixed effects estimation
+
+# 1. Base DiD Model
+model1 <- feols(
+  accident_count ~ has_construction | segment_id + month, 
+  data = panel_data
+)
+
+# 2. Extended Model with Urban/Rural Heterogeneity
+model2 <- feols(
+  accident_count ~ has_construction + has_construction:is_urban | segment_id + month, 
+  data = panel_data
+)
+
+# 3. Full Model with additional controls
+model3 <- feols(
+  accident_count ~ has_construction + has_construction:is_urban + 
+                  has_construction:is_winter + has_construction:is_summer +
+                  avg_temperature + avg_precipitation + pct_adverse_weather | 
+                  segment_id + month,
+  data = panel_data
+)
+
+# 4. Accident rate model (normalized by segment length)
+model4 <- feols(
+  accidents_per_km ~ has_construction + has_construction:is_urban | segment_id + month, 
+  data = panel_data
+)
+
+# 5. Severity model
+model5 <- feols(
+  avg_severity ~ has_construction + has_construction:is_urban | segment_id + month, 
+  data = panel_data
+)
+
+# Display results in a regression table
+library(modelsummary)
+modelsummary(list(model1, model2, model3, model4, model5),
+             stars = TRUE,
+             title = "DiD Estimates of Construction Impacts on Accidents")
+```
+
+### 6. Robustness Checks
+
+We can also run some robustness checks to test the sensitivity of our results
+
+```{r}
+### 6. Robustness Checks
+
+# 1. EVENT STUDY APPROACH - Analyze dynamic effects
+# Create relative time variables to construction start
+panel_data_event <- panel_data %>%
+  group_by(segment_id) %>%
+  # Find first month with construction for each segment
+  mutate(
+    first_construction = if(any(has_construction == 1)) min(month[has_construction == 1]) else NA,
+    # Calculate months relative to construction start
+    rel_month = as.numeric(difftime(month, first_construction, units = "days")) / 30,
+    # Create factor for event study
+    rel_period = case_when(
+      is.na(rel_month) ~ NA_character_,
+      rel_month <= -3 ~ "t-3+",
+      rel_month <= -2 ~ "t-2",
+      rel_month <= -1 ~ "t-1",
+      rel_month < 0 ~ "t-0",
+      rel_month < 1 ~ "t+0",
+      rel_month < 2 ~ "t+1",
+      rel_month < 3 ~ "t+2",
+      TRUE ~ "t+3+"
+    )
+  ) %>%
+  ungroup() %>%
+  # Create dummy variables for each relative period
+  mutate(
+    rel_period = factor(rel_period, 
+                        levels = c("t-3+", "t-2", "t-1", "t-0", "t+0", "t+1", "t+2", "t+3+"))
+  )
+
+# Run event study regression
+event_model <- feols(
+  accident_count ~ i(rel_period, ref = "t-1") | segment_id + month, 
+  data = panel_data_event %>% filter(!is.na(rel_period))
+)
+
+# Visualize event study results
+coef_plot <- iplot(event_model, main = "Event Study: Effect of Construction on Accidents",
+      xlab = "Months Relative to Construction Start", ylab = "Effect on Accident Count")
+
+# 2. PLACEBO TESTS - Randomly assign treatment
+set.seed(123)
+placebo_data <- panel_data %>%
+  # Randomly shuffle the has_construction variable
+  mutate(placebo_construction = sample(has_construction))
+
+# Run placebo model
+placebo_model <- feols(
+  accident_count ~ placebo_construction | segment_id + month, 
+  data = placebo_data
+)
+
+# Run multiple placebo tests
+placebo_results <- data.frame(iteration = 1:100, estimate = NA, std.error = NA)
+for(i in 1:100) {
+  set.seed(i)
+  # Create new placebo treatment
+  temp_data <- panel_data %>%
+    mutate(placebo_construction = sample(has_construction))
+
+  # Run model
+  temp_model <- feols(
+    accident_count ~ placebo_construction | segment_id + month, 
+    data = temp_data
+  )
+
+  # Store results
+  placebo_results$estimate[i] <- coef(temp_model)[1]
+  placebo_results$std.error[i] <- se(temp_model)[1]
+}
+
+# Plot placebo distribution vs. actual effect
+placebo_plot <- ggplot(placebo_results, aes(x = estimate)) +
+  geom_histogram(bins = 30, fill = "lightblue", color = "black") +
+  geom_vline(xintercept = coef(model1)[1], color = "red", linewidth = 1.5) +
+  theme_minimal() +
+  labs(title = "Distribution of Placebo Effects vs. Actual Effect",
+       x = "Estimated Effect", y = "Count",
+       caption = "Red line shows actual estimated effect")
+
+# 3. BUFFER DISTANCE SENSITIVITY - Test different buffer definitions
+# This requires going back to your spatial join code and creating different versions
+# Here's a simplified approach (assuming you can regenerate panel data with different buffers)
+buffer_results <- data.frame(
+  buffer_dist = c(200, 300, 400, 500),
+  estimate = c(0.12, 0.15, 0.11, 0.09),  # Example values - replace with actual results
+  lower_ci = c(0.05, 0.08, 0.04, 0.01),  # from running your models with different buffer distances
+  upper_ci = c(0.19, 0.22, 0.18, 0.17)
+)
+
+# Plot buffer sensitivity results
+buffer_plot <- ggplot(buffer_results, aes(x = factor(buffer_dist), y = estimate)) +
+  geom_point(size = 3) +
+  geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2) +
+  theme_minimal() +
+  labs(title = "Sensitivity to Buffer Distance",
+       x = "Buffer Distance (meters)", y = "Estimated Effect on Accident Count")
+
+# 4. ALTERNATIVE FIXED EFFECTS - Test different specifications
+alt_fe_model1 <- feols(accident_count ~ has_construction, data = panel_data)  # No FE
+alt_fe_model2 <- feols(accident_count ~ has_construction | segment_id, data = panel_data) # Only segment FE
+alt_fe_model3 <- feols(accident_count ~ has_construction | month, data = panel_data) # Only time FE
+alt_fe_model4 <- feols(accident_count ~ has_construction | segment_id + month, data = panel_data) # Both FE (baseline)
+
+# Compare models
+fe_comparison <- modelsummary(list("No FE" = alt_fe_model1, 
+                                  "Segment FE" = alt_fe_model2,
+                                  "Time FE" = alt_fe_model3,
+                                  "Both FE" = alt_fe_model4),
+                            stars = TRUE)
+
+# 5. DIFFERENT OUTCOME TRANSFORMATIONS - Test sensitivity to functional form
+transform_model1 <- feols(accident_count ~ has_construction | segment_id + month, 
+                       data = panel_data) # Linear (baseline)
+transform_model2 <- feols(log(accident_count + 1) ~ has_construction | segment_id + month, 
+                       data = panel_data) # Log transform
+transform_model3 <- fepois(accident_count ~ has_construction | segment_id + month, 
+                       data = panel_data) # Poisson
+
+# Compare transformations
+transform_comparison <- modelsummary(list("Linear" = transform_model1,
+                                        "Log" = transform_model2,
+                                        "Poisson" = transform_model3),
+                                   stars = TRUE)
+
+# 6. Combine robustness visualizations into a panel
+library(patchwork)
+robustness_panel <- (coef_plot / placebo_plot) | (buffer_plot / fe_comparison)
+robustness_panel + plot_annotation(
+  title = "Robustness Checks for Construction Safety Analysis",
+  theme = theme(plot.title = element_text(size = 16, hjust = 0.5))
+)
+```
+
+Additional validation, propensity score matching:
+
+```{r}
+library(MatchIt)
+matched_segments <- matchit(
+  has_construction ~ road_type + segment_length_km + is_urban,
+  data = panel_data %>% filter(month == min(month)),
+  method = "nearest"
+)
+```

### 2025-03-18 final/streamlined_nb.Rmd
Added a streamlined version

---
title: "Geospatial Data Science - Final project"
author:
- Moritz Peist
- Noemi Lucchi
- Simon Vellin
subtitle: Analysis of US road accidents and construction sites
output:
  html_notebook:
    df_print: paged
    toc: true
  html_document:
    toc: true
    df_print: paged
---

<hr>

```{css, echo=FALSE}
h1.title {
  text-align: center;
}

h4.author, h4.date {
  text-align: center;
}

.author {
  text-align: center;
}

.date {
  text-align: center;
}

/* Add styling for subtitle */
h3.subtitle {
  text-align: center;
  font-style: italic;
  margin-top: -10px;
  color: #606060;
}
```

<div style="text-align: center;">
  <img src="../Barcelona_School_of_Economics_logo.svg" alt="BSE logo">
</div>

<hr>

## 1. Setup and Packages

```{r setup, message=FALSE, warning=FALSE}
# Load all necessary packages
library(data.table)   # For faster data manipulation
library(tidyverse)    # For data manipulation and visualization
library(sf)           # For spatial data handling
library(leaflet)      # For interactive maps
library(leaflet.extras) # For additional leaflet features
library(mapview)      # For easier map visualization
library(tmap)         # For thematic maps
library(tigris)       # For US road networks
library(future)       # For parallel processing
library(future.apply) # For parallel processing with apply functions
library(fixest)       # For efficient fixed effects estimation
library(modelsummary) # For regression tables
library(patchwork)    # For combining plots
library(MatchIt)      # For matching
library(lubridate)    # For date handling

# Create directories if they don't exist
if (!dir.exists("./data/tigris")) {
  dir.create("./data/tigris", recursive = TRUE)
}
if (!dir.exists("./data/processed")) {
  dir.create("./data/processed", recursive = TRUE)
}
if (!dir.exists("./imgs")) {
  dir.create("./imgs", recursive = TRUE)
}

# Set custom cache directory
options(tigris_cache_dir = "./data/tigris")
# Configure tigris to use caching
options(tigris_use_cache = TRUE)
```

## 2. Helper Functions for Spatial Operations

```{r helper_functions}
# Function to create buffers in batches with proper projection
create_buffers_in_batches <- function(sf_object, buffer_dist, batch_size = 500) {
  message("Reprojecting data to a meter-based CRS...")
  sf_object_projected <- st_transform(sf_object, 3310)  # California Albers

  n_features <- nrow(sf_object_projected)
  n_batches <- ceiling(n_features / batch_size)

  # Create empty list to store results
  buffer_list <- vector("list", n_batches)

  # Process in batches
  for (i in 1:n_batches) {
    start_idx <- (i-1) * batch_size + 1
    end_idx <- min(i * batch_size, n_features)

    cat(sprintf("Processing batch %d of %d (features %d to %d)\n", 
                i, n_batches, start_idx, end_idx))

    # Extract batch
    batch <- sf_object_projected[start_idx:end_idx, ]

    # Create buffer
    buffer_list[[i]] <- st_buffer(batch, dist = buffer_dist)
  }

  # Combine results
  result <- do.call(rbind, buffer_list)

  # Reproject back to original CRS
  message("Reprojecting results back to original CRS...")
  st_transform(result, st_crs(sf_object))
}

# Unified function to create both point and line buffers
create_all_buffers <- function(point_sf, line_sf, buffer_dist = 300) {
  # Process point buffers
  point_buffers <- point_sf %>%
    mutate(construction_id = if("ID" %in% names(.)) ID else row_number(),
           construction_type = "point") %>%
    select(construction_id, construction_type, Start_Time, End_Time, geometry) %>%
    rename(start_time = Start_Time, end_time = End_Time) %>%
    create_buffers_in_batches(buffer_dist = buffer_dist, batch_size = 500)

  # Process line buffers if provided
  if(!missing(line_sf) && nrow(line_sf) > 0) {
    # Calculate segment length once
    line_sf_with_length <- line_sf %>%
      st_transform(3310) %>%
      mutate(segment_length_m = as.numeric(st_length(geometry))) %>%
      st_transform(4326)

    line_buffers <- line_sf_with_length %>%
      mutate(construction_id = if("ID" %in% names(.)) ID else row_number() + nrow(point_sf),
             construction_type = "line") %>%
      select(construction_id, construction_type, Start_Time, End_Time, 
             segment_length_m, geometry) %>%
      rename(start_time = Start_Time, end_time = End_Time) %>%
      create_buffers_in_batches(buffer_dist = buffer_dist, batch_size = 250)

    # Combine buffers
    return(list(
      combined_buffers = bind_rows(
        select(point_buffers, -construction_type),
        select(line_buffers, -construction_type, -segment_length_m)
      ),
      point_buffers = point_buffers,
      line_buffers = line_buffers,
      line_data = line_sf_with_length
    ))
  }

  # Return just point buffers if no lines
  return(list(
    combined_buffers = point_buffers,
    point_buffers = point_buffers
  ))
}

# Helper function to find accidents near constructions
find_accidents_near_construction <- function(accident_sf, buffer_results) {
  # For point buffers
  point_spatial_matches <- accident_sf %>%
    mutate(accident_id = if("ID" %in% names(.)) ID else row_number(),
           accident_severity = if("Severity" %in% names(.)) Severity else NA_integer_,
           accident_time = Start_Time) %>%
    select(accident_id, accident_severity, accident_time, geometry) %>%
    st_join(buffer_results$point_buffers, join = st_within) %>%
    filter(!is.na(construction_id) & 
           accident_time >= start_time & 
           accident_time <= end_time) %>%
    select(accident_id, accident_severity, accident_time, construction_id, construction_type)

  # For line buffers if they exist
  if ("line_buffers" %in% names(buffer_results)) {
    line_spatial_matches <- accident_sf %>%
      mutate(accident_id = if("ID" %in% names(.)) ID else row_number(),
             accident_severity = if("Severity" %in% names(.)) Severity else NA_integer_,
             accident_time = Start_Time) %>%
      select(accident_id, accident_severity, accident_time, geometry) %>%
      st_join(buffer_results$line_buffers, join = st_within) %>%
      filter(!is.na(construction_id) & 
             accident_time >= start_time & 
             accident_time <= end_time) %>%
      select(accident_id, accident_severity, accident_time, construction_id, 
             construction_type, segment_length_m)

    # Combine both types of matches
    return(list(
      point_matches = point_spatial_matches,
      line_matches = line_spatial_matches,
      all_matches = bind_rows(
        point_spatial_matches,
        select(line_spatial_matches, -segment_length_m)
      )
    ))
  }

  # Return just point matches if no line buffers
  return(list(
    point_matches = point_spatial_matches,
    all_matches = point_spatial_matches
  ))
}

# Helper function to analyze accidents near construction
analyze_accidents_near_construction <- function(spatial_matches, accident_sf, line_data) {
  # Count construction sites per accident
  construction_counts <- spatial_matches$all_matches %>%
    st_set_geometry(NULL) %>%
    group_by(accident_id) %>%
    summarize(nearby_construction_count = n())

  # Randomly select one construction site per accident
  accident_construction_analysis <- spatial_matches$all_matches %>%
    st_set_geometry(NULL) %>%
    left_join(construction_counts, by = "accident_id") %>%
    mutate(
      temporal_overlap = 1,
      is_near_active_construction = 1
    ) %>%
    group_by(accident_id) %>%
    slice_sample(n = 1) %>%
    ungroup()

  # Calculate accident rates for line segments
  if ("line_matches" %in% names(spatial_matches) && !is.null(line_data)) {
    construction_segment_accident_rates <- line_data %>%
      mutate(construction_id = if("ID" %in% names(.)) ID else row_number()) %>%
      left_join(
        spatial_matches$line_matches %>%
          st_set_geometry(NULL) %>%
          group_by(construction_id) %>%
          summarize(accident_count = n()),
        by = "construction_id"
      ) %>%
      mutate(
        accident_count = replace_na(accident_count, 0),
        accidents_per_km = accident_count / (segment_length_m / 1000)
      )
  } else {
    construction_segment_accident_rates <- NULL
  }

  # Add construction info to all accidents
  accidents_with_construction <- accident_sf %>%
    mutate(accident_id = if("ID" %in% names(.)) ID else row_number()) %>%
    left_join(
      accident_construction_analysis %>%
        select(accident_id, construction_id, nearby_construction_count, construction_type),
      by = "accident_id"
    ) %>%
    mutate(near_construction = ifelse(is.na(construction_id), 0, 1))

  # Return results
  return(list(
    accident_construction_analysis = accident_construction_analysis,
    construction_segment_accident_rates = construction_segment_accident_rates,
    accidents_with_construction = accidents_with_construction,
    total_accidents = nrow(accident_sf),
    accidents_near_construction = nrow(accident_construction_analysis),
    pct_near_construction = round(nrow(accident_construction_analysis)/nrow(accident_sf)*100, 2)
  ))
}

# NEW FUNCTION: Create analysis panel with improved spatial matching
create_analysis_panel <- function(df.const.sf, df.acc.sf, ca_roads, buffer_dist = 300) {
  # Step 1: Create a more appropriate spatial unit - larger road chunks
  # Dissolve/merge road segments by road name and type to create longer segments
  cat("Simplifying road network...\n")
  simplified_roads <- ca_roads %>%
    group_by(FULLNAME, RTTYP) %>%
    summarize(geometry = st_union(geometry), .groups = "drop") %>%
    mutate(road_id = row_number(),
           road_length_km = as.numeric(st_length(st_transform(geometry, 3310)))/1000)

  # Step 2: Create construction buffers and find which roads they intersect
  cat("Creating construction buffers and finding affected roads...\n")
  const_buffers <- st_buffer(st_transform(df.const.sf, 3310), buffer_dist) %>%
    st_transform(st_crs(ca_roads))

  road_construction <- simplified_roads %>%
    st_join(const_buffers %>% 
              select(ID, Start_Time, End_Time), join = st_intersects) %>%
    filter(!is.na(ID)) %>%
    mutate(
      const_start = floor_date(Start_Time, "month"),
      const_end = floor_date(End_Time, "month")
    ) %>%
    st_set_geometry(NULL) %>%
    select(road_id, FULLNAME, RTTYP, const_start, const_end, ID)

  # Step 3: Create a panel with monthly observations for all roads
  cat("Creating panel dataset...\n")
  months <- seq(as.Date("2021-01-01"), as.Date("2021-12-01"), by = "month")

  # Create base panel with all road-month combinations
  panel_base <- expand.grid(
    road_id = simplified_roads$road_id,
    month = months
  ) %>% as_tibble()

  # Add construction activity flags
  cat("Adding construction treatment indicators...\n")
  panel_with_construction <- panel_base %>%
    left_join(
      road_construction %>%
        # Expand each construction to all months it spans
        group_by(road_id, ID) %>%
        do({
          months_seq <- seq(from = .$const_start[1], to = .$const_end[1], by = "month")
          data.frame(
            road_id = .$road_id[1],
            month = months_seq,
            has_construction = 1,
            const_id = .$ID[1]
          )
        }) %>%
        ungroup() %>%
        # Consolidate multiple constructions on same road-month
        group_by(road_id, month) %>%
        summarize(
          has_construction = 1,
          construction_count = n(),
          .groups = "drop"
        ),
      by = c("road_id", "month")
    ) %>%
    mutate(has_construction = ifelse(is.na(has_construction), 0, has_construction))

  # Step 4: Count accidents for each road segment by month
  cat("Matching accidents to roads...\n")
  # Buffer roads slightly to catch nearby accidents
  road_buffers <- st_buffer(st_transform(simplified_roads, 3310), 50) %>%
    st_transform(st_crs(df.acc.sf))

  accident_counts <- df.acc.sf %>%
    mutate(month = floor_date(Start_Time, "month")) %>%
    st_join(road_buffers %>% select(road_id), join = st_intersects) %>%
    filter(!is.na(road_id)) %>%
    st_set_geometry(NULL) %>%
    group_by(road_id, month) %>%
    summarize(
      accident_count = n(),
      severe_accident_count = sum(Severity >= 3, na.rm = TRUE),
      avg_severity = mean(Severity, na.rm = TRUE),
      .groups = "drop"
    )

  # Step 5: Create final panel with all variables
  cat("Finalizing panel dataset...\n")
  panel_final <- panel_with_construction %>%
    left_join(accident_counts, by = c("road_id", "month")) %>%
    left_join(
      simplified_roads %>%
        st_set_geometry(NULL) %>%
        select(road_id, road_length_km, FULLNAME, RTTYP),
      by = "road_id"
    ) %>%
    # Add road type classification
    mutate(
      road_type = case_when(
        grepl("I", RTTYP) ~ "Interstate",
        grepl("U", RTTYP) ~ "US Highway",
        grepl("S", RTTYP) ~ "State Highway",
        TRUE ~ "Other"
      ),
      # Fill in missing accident counts with 0
      accident_count = replace_na(accident_count, 0),
      severe_accident_count = replace_na(severe_accident_count, 0),
      avg_severity = replace_na(avg_severity, 0),
      # Calculate per-km rates
      accidents_per_km = accident_count / road_length_km,
      severe_accidents_per_km = severe_accident_count / road_length_km,
      # Extract time variables
      year = year(month),
      month_num = month(month),
      is_winter = month_num %in% c(12, 1, 2),
      is_summer = month_num %in% c(6, 7, 8)
    )

  # Add urban classification
  cat("Adding urban/rural classification...\n")
  urban_areas <- urban_areas(year = 2021)

  # Check which roads intersect urban areas
  urban_roads <- simplified_roads %>%
    st_transform(st_crs(urban_areas)) %>%
    st_join(urban_areas %>% select(UACE10, NAME10, UATYP10), join = st_intersects) %>%
    mutate(
      is_urban = !is.na(UACE10),
      urban_type = ifelse(is.na(UATYP10), "Rural", UATYP10)
    ) %>%
    st_set_geometry(NULL) %>%
    select(road_id, is_urban, urban_type)

  # Join urban classification
  panel_final <- panel_final %>%
    left_join(urban_roads, by = "road_id")

  # Add weather data if available (average by month for simplicity)
  if ("Temperature(F)" %in% names(df.acc.sf) && "Precipitation(in)" %in% names(df.acc.sf)) {
    cat("Adding weather data...\n")
    weather_data <- df.acc.sf %>%
      mutate(month = floor_date(Start_Time, "month")) %>%
      group_by(month) %>%
      summarize(
        avg_temperature = mean(`Temperature(F)`, na.rm = TRUE),
        avg_precipitation = mean(`Precipitation(in)`, na.rm = TRUE),
        pct_adverse_weather = mean(Weather_Condition %in% 
                                  c("Rain", "Snow", "Fog", "Thunderstorm"), na.rm = TRUE),
        .groups = "drop"
      ) %>%
      st_set_geometry(NULL)

    panel_final <- panel_final %>%
      left_join(weather_data, by = "month")
  }

  # Print diagnostic information
  cat("\nSummary Statistics:\n")
  cat("Total unique roads:", nrow(simplified_roads), "\n")
  cat("Roads with construction:", length(unique(road_construction$road_id)), "\n")
  cat("Road-months with construction:", sum(panel_final$has_construction), "\n")
  cat("Roads with accidents:", length(unique(accident_counts$road_id)), "\n")
  cat("Total accidents matched to roads:", sum(accident_counts$accident_count), "\n")
  cat("Final panel dimensions:", nrow(panel_final), "x", ncol(panel_final), "\n")

  return(list(
    panel_data = panel_final,
    simplified_roads = simplified_roads,
    road_construction = road_construction,
    accident_counts = accident_counts
  ))
}
```

## 3. Regression and Robustness Functions

```{r regression_functions}
# Run all main regression models
run_regression_models <- function(panel_data) {
  # 1. Base DiD Model
  model1 <- feols(
    accident_count ~ has_construction | road_id + month, 
    data = panel_data
  )

  # 2. Extended Model with Urban/Rural Heterogeneity
  model2 <- feols(
    accident_count ~ has_construction + has_construction:is_urban | road_id + month, 
    data = panel_data
  )

  # 3. Full Model with additional controls
  # Check if weather data is available
  if (all(c("avg_temperature", "avg_precipitation", "pct_adverse_weather") %in% names(panel_data))) {
    model3 <- feols(
      accident_count ~ has_construction + has_construction:is_urban + 
                      has_construction:is_winter + has_construction:is_summer +
                      avg_temperature + avg_precipitation + pct_adverse_weather | 
                      road_id + month,
      data = panel_data
    )
  } else {
    model3 <- feols(
      accident_count ~ has_construction + has_construction:is_urban + 
                      has_construction:is_winter + has_construction:is_summer | 
                      road_id + month,
      data = panel_data
    )
  }

  # 4. Accident rate model (normalized by segment length)
  model4 <- feols(
    accidents_per_km ~ has_construction + has_construction:is_urban | road_id + month, 
    data = panel_data
  )

  # 5. Severity model
  model5 <- feols(
    avg_severity ~ has_construction + has_construction:is_urban | road_id + month, 
    data = panel_data
  )

  return(list(
    base_model = model1,
    urban_het_model = model2,
    full_model = model3,
    rate_model = model4,
    severity_model = model5,
    all_models = list(model1, model2, model3, model4, model5)
  ))
}

# Helper functions for robustness checks
create_event_study_data <- function(panel_data) {
  panel_data %>%
    group_by(road_id) %>%
    mutate(
      first_construction = if(any(has_construction == 1)) min(month[has_construction == 1]) else NA,
      rel_month = as.numeric(difftime(month, first_construction, units = "days")) / 30,
      rel_period = case_when(
        is.na(rel_month) ~ NA_character_,
        rel_month <= -3 ~ "t-3+",
        rel_month <= -2 ~ "t-2",
        rel_month <= -1 ~ "t-1",
        rel_month < 0 ~ "t-0",
        rel_month < 1 ~ "t+0",
        rel_month < 2 ~ "t+1",
        rel_month < 3 ~ "t+2",
        TRUE ~ "t+3+"
      )
    ) %>%
    ungroup() %>%
    mutate(
      rel_period = factor(rel_period, 
                        levels = c("t-3+", "t-2", "t-1", "t-0", "t+0", "t+1", "t+2", "t+3+"))
    )
}

run_placebo_tests <- function(panel_data, iterations = 100) {
  placebo_results <- data.frame(iteration = 1:iterations, estimate = NA, std.error = NA)
  for(i in 1:iterations) {
    set.seed(i)
    temp_data <- panel_data %>%
      mutate(placebo_construction = sample(has_construction))

    temp_model <- feols(
      accident_count ~ placebo_construction | road_id + month, 
      data = temp_data
    )

    placebo_results$estimate[i] <- coef(temp_model)[1]
    placebo_results$std.error[i] <- se(temp_model)[1]
  }
  return(placebo_results)
}

run_fe_variations <- function(panel_data) {
  list(
    "No FE" = feols(accident_count ~ has_construction, data = panel_data),
    "Road FE" = feols(accident_count ~ has_construction | road_id, data = panel_data),
    "Time FE" = feols(accident_count ~ has_construction | month, data = panel_data),
    "Both FE" = feols(accident_count ~ has_construction | road_id + month, data = panel_data)
  )
}

run_transform_variations <- function(panel_data) {
  list(
    "Linear" = feols(accident_count ~ has_construction | road_id + month, data = panel_data),
    "Log" = feols(log(accident_count + 1) ~ has_construction | road_id + month, data = panel_data),
    "Poisson" = fepois(accident_count ~ has_construction | road_id + month, data = panel_data)
  )
}

# Run all robustness checks
run_robustness_checks <- function(panel_data, base_model) {
  # Results list
  robustness_results <- list()

  # 1. Event study approach
  panel_data_event <- create_event_study_data(panel_data)
  event_model <- feols(
    accident_count ~ i(rel_period, ref = "t-1") | road_id + month,
    data = panel_data_event %>% filter(!is.na(rel_period))
  )
  robustness_results$event_model <- event_model
  robustness_results$panel_data_event <- panel_data_event

  # 2. Placebo tests
  placebo_results <- run_placebo_tests(panel_data, 100)
  robustness_results$placebo_results <- placebo_results

  # 3. Fixed effects variations
  fe_models <- run_fe_variations(panel_data)
  robustness_results$fe_models <- fe_models

  # 4. Transform outcome variables
  transform_models <- run_transform_variations(panel_data)
  robustness_results$transform_models <- transform_models

  # 5. Add matching (simplified, assumes is_urban and road_type are factors)
  matched_data <- matchit(
    has_construction ~ road_type + road_length_km + is_urban,
    data = panel_data %>% 
      filter(month == min(month)) %>%
      mutate(
        road_type = factor(road_type),
        is_urban = factor(is_urban)
      ),
    method = "nearest"
  )
  robustness_results$matched_data <- matched_data

  # Return all results
  return(robustness_results)
}
```

## 4. Visualization Functions

```{r vis_functions}
# Create static maps
create_static_maps <- function(df.acc.sf, df.const.sf, df.const.lines, ca_roads) {
  # Accident map
  ca_plot <- ggplot() +
    geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
    geom_sf(data = df.acc.sf, 
            aes(color = factor(year)), alpha = 0.7, size = 0.5) +
    scale_color_viridis_d(name = "Year") +
    theme_minimal() +
    labs(title = "California Road Accidents (2021)") +
    theme(legend.position = "bottom")

  # Construction map
  ca_const_plot <- ggplot() +
    geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
    geom_sf(data = df.const.sf, 
            aes(color = factor(year)), alpha = 0.7, size = 0.5) +
    scale_color_viridis_d(name = "Year") +
    theme_minimal() +
    labs(title = "California Construction Sites (2021)") +
    theme(legend.position = "bottom")

  # Construction line segments map
  lines_plot <- ggplot() +
    geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
    geom_sf(data = df.const.lines, 
            color = "orange", size = 1) +
    theme_minimal() +
    labs(title = "California Construction Segments (2021)")

  return(list(
    accidents_map = ca_plot,
    construction_map = ca_const_plot,
    construction_lines_map = lines_plot
  ))
}

# Create heatmaps
create_heatmaps <- function(df.acc, df.const, ca_roads) {
  # Accident heatmap
  accident_heatmap <- ggplot() +
    geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
    stat_density_2d(
      data = df.acc %>% filter(!is.na(Start_Lat) & !is.na(Start_Lng)),
      aes(x = Start_Lng, y = Start_Lat, fill = after_stat(density), alpha = after_stat(density)),
      geom = "tile", 
      contour = FALSE,
      h = 0.1,
      n = 200
    ) +
    scale_alpha_continuous(range = c(0, 0.9), guide = "none") +
    scale_fill_gradientn(
      colors = c("#0000FF", "#00FFFF", "#00FF00", "#FFFF00", "#FF0000"),
      name = "Accident\nDensity"
    ) +
    coord_sf(
      xlim = c(-124.5, -114.5),
      ylim = c(32.5, 42.5)
    ) +
    theme_minimal() +
    labs(
      title = "Accident Density Heatmap in California (2021)",
      x = NULL,
      y = NULL
    ) +
    theme(
      legend.position = "right",
      plot.title = element_text(face = "bold")
    )

  # Construction heatmap
  construction_heatmap <- ggplot() +
    geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
    stat_density_2d(
      data = df.const %>% filter(!is.na(Start_Lat) & !is.na(Start_Lng)),
      aes(x = Start_Lng, y = Start_Lat, fill = after_stat(density)),
      geom = "tile", 
      contour = FALSE,
      alpha = 0.7,
      h = 0.1,
      n = 200
    ) +
    scale_alpha_continuous(range = c(0, 0.9), guide = "none") +
    scale_fill_gradientn(
      colors = c("yellow", "orange", "red"),
      name = "Construction\nDensity"
    ) +
    coord_sf(
      xlim = c(-124.5, -114.5),
      ylim = c(32.5, 42.5)
    ) +
    theme_minimal() +
    labs(
      title = "Construction Density Heatmap in California (2021)",
      x = NULL,
      y = NULL
    ) +
    theme(
      legend.position = "right",
      plot.title = element_text(face = "bold")
    )

  return(list(
    accident_heatmap = accident_heatmap,
    construction_heatmap = construction_heatmap
  ))
}

# Create interactive maps
create_interactive_maps <- function(df.acc.sf, df.const.sf, df.const.lines) {
  # Accidents map
  accident_map <- leaflet() %>%
    addTiles() %>%
    addHeatmap(data = df.acc.sf, 
               intensity = ~1,
               radius = 8, 
               blur = 10) %>%
    setView(lng = -119.4179, lat = 36.7783, zoom = 6) %>%
    setMaxBounds(lng1 = -124.6, lat1 = 42.0,
                 lng2 = -114.1, lat2 = 32.5)

  # Construction sites map
  construction_map <- leaflet() %>%
    addTiles() %>%
    addHeatmap(data = df.const.sf, 
               intensity = ~1,
               radius = 8, 
               blur = 10,
               gradient = c("yellow", "orange", "red")) %>%
    setView(lng = -119.4179, lat = 36.7783, zoom = 6) %>%
    setMaxBounds(lng1 = -124.6, lat1 = 42.0,
                 lng2 = -114.1, lat2 = 32.5)

  # Construction lines map
  construction_lines_map <- leaflet() %>%
    addTiles() %>%
    addPolylines(data = df.const.lines, 
                 color = "orange",
                 weight = 3,
                 opacity = 0.7,
                 popup = ~paste("ID:", ID, "<br>Duration:", duration)) %>%
    setView(lng = -119.4179, lat = 36.7783, zoom = 6) %>%
    setMaxBounds(lng1 = -124.6, lat1 = 42.0,
                 lng2 = -114.1, lat2 = 32.5)

  return(list(
    accident_map = accident_map,
    construction_map = construction_map,
    construction_lines_map = construction_lines_map
  ))
}

# Create buffer analysis visualizations
create_buffer_analysis_plots <- function(accidents_per_zone, ca_roads, analysis_results, severity_comparison, temporal_analysis) {
  # Accident counts in construction zones
  accident_zones_plot <- ggplot() +
    geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
    geom_sf(data = accidents_per_zone %>% arrange(accident_count), 
            aes(color = accident_count), size = 1.5) +
    scale_color_viridis_c() +
    theme_minimal() +
    labs(title = "Accident Counts within Construction Zones (2021)")

  # Accident rate per construction segment
  if (!is.null(analysis_results$construction_segment_accident_rates)) {
    segment_rate_plot <- ggplot() +
      geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
      geom_sf(data = analysis_results$construction_segment_accident_rates, 
              aes(color = accidents_per_km), size = 1.5) +
      scale_color_viridis_c() +
      theme_minimal() +
      labs(title = "Accident Rate per Construction Segment (2021)",
           color = "Accidents/km")
  } else {
    segment_rate_plot <- NULL
  }

  # Temporal analysis
  temporal_plot <- ggplot(temporal_analysis, aes(x = month, y = accident_count)) +
    geom_line() +
    geom_point() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = "Accidents Near Construction Over Time",
         x = "Month",
         y = "Accident Count")

  # Severity comparison
  severity_plot <- ggplot(severity_comparison, aes(x = Severity, y = count, fill = location)) +
    geom_bar(stat = "identity", position = "dodge") +
    scale_fill_brewer(palette = "Set1") +
    theme_minimal() +
    labs(title = "Accident Severity: Near vs. Away from Construction",
         x = "Severity Level",
         y = "Count",
         fill = "Location")

  return(list(
    accident_zones_plot = accident_zones_plot,
    segment_rate_plot = segment_rate_plot,
    temporal_plot = temporal_plot,
    severity_plot = severity_plot
  ))
}

# Create visualizations for new panel dataset
visualize_panel_data <- function(panel_results) {
  # 1. Show roads with construction
  road_construction_map <- ggplot() +
    geom_sf(data = panel_results$simplified_roads, color = "gray90", size = 0.3) +
    geom_sf(data = panel_results$simplified_roads %>%
              filter(road_id %in% panel_results$road_construction$road_id),
            color = "red", size = 1) +
    theme_minimal() +
    labs(title = "Roads with Construction Activities (2021)")

  # 2. Show roads with accidents
  road_accident_map <- ggplot() +
    geom_sf(data = panel_results$simplified_roads, color = "gray90", size = 0.3) +
    geom_sf(data = panel_results$simplified_roads %>%
              filter(road_id %in% panel_results$accident_counts$road_id),
            aes(color = "Roads with Accidents"), size = 0.8) +
    geom_sf(data = panel_results$simplified_roads %>%
              filter(road_id %in% panel_results$road_construction$road_id),
            aes(color = "Roads with Construction"), size = 0.8) +
    scale_color_manual(values = c("Roads with Accidents" = "blue", 
                                 "Roads with Construction" = "red")) +
    theme_minimal() +
    labs(title = "Roads with Accidents vs Construction (2021)",
         color = "Road Type")

  # 3. Accident counts by month with construction indicator
  monthly_stats <- panel_results$panel_data %>%
    group_by(month) %>%
    summarize(
      total_accidents = sum(accident_count),
      construction_count = sum(has_construction),
      .groups = "drop"
    )

  temporal_plot <- ggplot(monthly_stats) +
    geom_col(aes(x = month, y = total_accidents), fill = "steelblue") +
    geom_line(aes(x = month, y = construction_count * 10), 
              color = "red", size = 1.2) + # scaled for visualization
    scale_y_continuous(
      name = "Accident Count",
      sec.axis = sec_axis(~./10, name = "Construction Count")
    ) +
    theme_minimal() +
    labs(title = "Monthly Accidents vs Construction Activities (2021)",
         x = "Month")

  # 4. Urban vs rural comparison
  urban_rural_comparison <- panel_results$panel_data %>%
    group_by(is_urban, has_construction) %>%
    summarize(
      total_accidents = sum(accident_count),
      accident_rate = mean(accidents_per_km),
      total_roads = n_distinct(road_id),
      .groups = "drop"
    ) %>%
    mutate(
      road_category = case_when(
        is_urban == TRUE & has_construction == 1 ~ "Urban with Construction",
        is_urban == TRUE & has_construction == 0 ~ "Urban without Construction",
        is_urban == FALSE & has_construction == 1 ~ "Rural with Construction",
        is_urban == FALSE & has_construction == 0 ~ "Rural without Construction"
      )
    )

  urban_rural_plot <- ggplot(urban_rural_comparison, 
                            aes(x = road_category, y = accident_rate, fill = road_category)) +
    geom_col() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = "Accident Rates by Urban/Rural and Construction Status",
         x = NULL,
         y = "Accidents per km",
         fill = "Road Category")

  return(list(
    road_construction_map = road_construction_map,
    road_accident_map = road_accident_map,
    temporal_plot = temporal_plot,
    urban_rural_plot = urban_rural_plot
  ))
}

# Create robustness check visualizations
# Create robustness check visualizations
visualize_robustness_checks <- function(robustness_results, model1) {
  # 1. Event study plot - keep as separate object, don't include in patchwork
  coef_plot <- iplot(robustness_results$event_model, 
                     main = "Event Study: Effect of Construction on Accidents",
                     xlab = "Months Relative to Construction Start", 
                     ylab = "Effect on Accident Count")

  # 2. Placebo plot
  placebo_plot <- ggplot(robustness_results$placebo_results, aes(x = estimate)) +
    geom_histogram(bins = 30, fill = "lightblue", color = "black") +
    geom_vline(xintercept = coef(model1)[1], color = "red", linewidth = 1.5) +
    theme_minimal() +
    labs(title = "Distribution of Placebo Effects vs. Actual Effect",
         x = "Estimated Effect", y = "Count",
         caption = "Red line shows actual estimated effect")

  # 3. Buffer sensitivity plot (example data - replace with actual results)
  buffer_results <- data.frame(
    buffer_dist = c(200, 300, 400, 500),
    estimate = c(0.12, 0.15, 0.11, 0.09),
    lower_ci = c(0.05, 0.08, 0.04, 0.01),
    upper_ci = c(0.19, 0.22, 0.18, 0.17)
  )

  buffer_plot <- ggplot(buffer_results, aes(x = factor(buffer_dist), y = estimate)) +
    geom_point(size = 3) +
    geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2) +
    theme_minimal() +
    labs(title = "Sensitivity to Buffer Distance",
         x = "Buffer Distance (meters)", y = "Estimated Effect on Accident Count")

  # 4. Alternative FE specifications table
  fe_comparison <- modelsummary(robustness_results$fe_models, stars = TRUE)

  # 5. Different outcome transformations
  transform_comparison <- modelsummary(robustness_results$transform_models, stars = TRUE)

  # 6. Create a simplified panel with just standard ggplot objects
  # Avoid using the iplot result in patchwork
  robustness_panel <- placebo_plot + buffer_plot + plot_layout(ncol = 2)

  return(list(
    coef_plot = coef_plot,  # Keep separate
    placebo_plot = placebo_plot,
    buffer_plot = buffer_plot,
    fe_comparison = fe_comparison,
    transform_comparison = transform_comparison,
    robustness_panel = robustness_panel
  ))
}
```

## 5. Data Loading and Processing

Let's load and process the data:

```{r load_data, message=FALSE, warning=FALSE}
# Check if processed data exists to avoid reprocessing
if (file.exists("data/processed/df_acc.rds") && 
    file.exists("data/processed/df_const.rds") && 
    file.exists("data/processed/ca_roads.rds")) {

  df.acc <- readRDS("data/processed/df_acc.rds")
  df.const <- readRDS("data/processed/df_const.rds")
  ca_roads <- readRDS("data/processed/ca_roads.rds")
  df.acc.sf <- readRDS("data/processed/df_acc_sf.rds")
  df.const.sf <- readRDS("data/processed/df_const_sf.rds")
  df.const.lines <- readRDS("data/processed/df_const_lines.rds")

  cat("Loaded preprocessed data\n")

} else {
  # Load accident data
  cat("Loading accident data...\n")
  df.acc <- fread("data/us_accidents/US_accidents_March23.csv")[
    # Filter date range of 2021
    lubridate::year(as.Date(Start_Time)) == 2021 & 
    # And California
    State == "CA"
  ][, `:=`(
    # Add year, quarter, month columns
    year = data.table::year(Start_Time),
    quarter = data.table::quarter(Start_Time),
    month = data.table::month(Start_Time),
    # Calculate duration
    duration = as.numeric(difftime(End_Time, Start_Time, units = "days"))
  )] %>% 
    as_tibble()

  # Load construction data
  cat("Loading construction data...\n")
  df.const <- fread("data/us_constructions/US_constructions_Dec21.csv")[
    # Filter date range of 2021
    lubridate::year(as.Date(Start_Time)) == 2021 & 
    # And California
    State == "CA"
  ][, `:=`(
    # Add year, quarter, month columns
    year = year(Start_Time),
    quarter = quarter(Start_Time),
    month = month(Start_Time),
    # Calculate duration
    duration = as.numeric(difftime(End_Time, Start_Time, units = "days"))
  )] %>% 
    as_tibble()

  # Convert to SF objects
  cat("Converting to SF objects...\n")
  df.acc.sf <- df.acc %>%
    filter(!is.na(Start_Lat) & !is.na(Start_Lng)) %>%
    st_as_sf(coords = c("Start_Lng", "Start_Lat"), crs = 4326)

  df.const.sf <- df.const %>%
    filter(!is.na(Start_Lat) & !is.na(Start_Lng)) %>%
    st_as_sf(coords = c("Start_Lng", "Start_Lat"), crs = 4326)

  # Create construction line objects
  cat("Creating construction linestrings...\n")
  df.const.lines <- df.const %>%
    filter(!is.na(Start_Lat) & !is.na(Start_Lng) & !is.na(End_Lat) & !is.na(End_Lng)) %>%
    mutate(geometry = pmap(list(Start_Lng, Start_Lat, End_Lng, End_Lat), 
                           function(start_lng, start_lat, end_lng, end_lat) {
                             coords <- matrix(c(start_lng, start_lat, 
                                               end_lng, end_lat), 
                                             ncol = 2, byrow = TRUE)
                             st_linestring(coords)
                           })) %>%
    st_sf(crs = 4326)

  # Load California roads
  cat("Loading California roads...\n")
  ca_roads <- primary_secondary_roads("CA", year = 2021)

  # Save processed data for future use
  saveRDS(df.acc, "data/processed/df_acc.rds")
  saveRDS(df.const, "data/processed/df_const.rds")
  saveRDS(ca_roads, "data/processed/ca_roads.rds")
  saveRDS(df.acc.sf, "data/processed/df_acc_sf.rds")
  saveRDS(df.const.sf, "data/processed/df_const_sf.rds")
  saveRDS(df.const.lines, "data/processed/df_const_lines.rds")
}
```

## 6. Spatial Analysis

Now let's perform the spatial analysis:

```{r spatial_analysis, message=FALSE, warning=FALSE}
# Check if we've already done the spatial analysis
if (file.exists("data/processed/buffer_results.rds") && 
    file.exists("data/processed/spatial_matches.rds") &&
    file.exists("data/processed/analysis_results.rds")) {

  buffer_results <- readRDS("data/processed/buffer_results.rds")
  spatial_matches <- readRDS("data/processed/spatial_matches.rds")
  analysis_results <- readRDS("data/processed/analysis_results.rds")

  cat("Loaded preprocessed spatial analysis results\n")

} else {
  # Set up parallel processing
  future::plan(future::multisession, workers = parallel::detectCores() - 1)

  # Step 1: Create construction buffers (unified approach)
  cat("Creating construction buffers...\n")
  buffer_results <- create_all_buffers(df.const.sf, df.const.lines, buffer_dist = 300)

  # Step 2: Find accidents near construction sites
  cat("Finding accidents near construction...\n")
  spatial_matches <- find_accidents_near_construction(df.acc.sf, buffer_results)

  # Step 3: Analyze accidents near construction
  cat("Analyzing accidents near construction...\n")
  analysis_results <- analyze_accidents_near_construction(
    spatial_matches, 
    df.acc.sf, 
    buffer_results$line_data
  )

  # Return to sequential processing
  future::plan(future::sequential)

  # Save results for future use
  saveRDS(buffer_results, "data/processed/buffer_results.rds")
  saveRDS(spatial_matches, "data/processed/spatial_matches.rds")
  saveRDS(analysis_results, "data/processed/analysis_results.rds")
}

# Extract key accident statistics
accidents_per_zone <- analysis_results$construction_segment_accident_rates %>%
  filter(accident_count > 0) %>%
  select(construction_id, accident_count, geometry)

cat("Total accidents in dataset:", analysis_results$total_accidents, "\n")
cat("Accidents near active construction:", analysis_results$accidents_near_construction, "\n")
cat("Percentage of accidents near active construction:", 
    analysis_results$pct_near_construction, "%\n")

# Prepare severity comparison data
accidents_near <- analysis_results$accidents_with_construction %>%
  filter(near_construction == 1) %>%
  st_set_geometry(NULL) %>%
  mutate(location = "Near construction") %>%
  select(Severity, location)

accidents_away <- analysis_results$accidents_with_construction %>%
  filter(near_construction == 0) %>%
  st_set_geometry(NULL) %>%
  mutate(location = "Away from construction") %>%
  select(Severity, location)

severity_comparison <- bind_rows(accidents_near, accidents_away) %>%
  group_by(Severity, location) %>%
  summarize(count = n(), .groups = "drop") %>%
  arrange(Severity, location)

# Prepare temporal analysis data
temporal_analysis <- spatial_matches$all_matches %>%
  st_set_geometry(NULL) %>%
  mutate(month = floor_date(accident_time, "month")) %>%
  group_by(month) %>%
  summarize(accident_count = n()) %>%
  arrange(month)
```

## 7. Data Visualization

Let's create the visualizations:

```{r visualizations, message=FALSE, warning=FALSE, fig.width=10, fig.height=8}
# Create all visualizations
static_maps <- create_static_maps(df.acc.sf, df.const.sf, df.const.lines, ca_roads)
heatmaps <- create_heatmaps(df.acc, df.const, ca_roads)
interactive_maps <- create_interactive_maps(df.acc.sf, df.const.sf, df.const.lines)
buffer_plots <- create_buffer_analysis_plots(
  accidents_per_zone, 
  ca_roads, 
  analysis_results, 
  severity_comparison, 
  temporal_analysis
)

# Display accident map
static_maps$accidents_map
ggsave("imgs/california_accidents.png", static_maps$accidents_map, width = 10, height = 8, dpi = 300)

# Display construction map
static_maps$construction_map
ggsave("imgs/california_construction.png", static_maps$construction_map, width = 10, height = 8, dpi = 300)

# Display accident heatmap
heatmaps$accident_heatmap
ggsave("imgs/accident_heatmap.png", heatmaps$accident_heatmap, width = 10, height = 8, dpi = 300)

# Display construction heatmap
heatmaps$construction_heatmap
ggsave("imgs/construction_heatmap.png", heatmaps$construction_heatmap, width = 10, height = 8, dpi = 300)

# Display interactive accident map
interactive_maps$accident_map

# Display interactive construction map
interactive_maps$construction_map

# Display accident zones plot
buffer_plots$accident_zones_plot
ggsave("imgs/accident_counts_construction_zones.png", buffer_plots$accident_zones_plot, width = 10, height = 8, dpi = 300)

# Display construction line segments
static_maps$construction_lines_map

# Display construction segment map
interactive_maps$construction_lines_map

# Display segment accident rate map
if (!is.null(buffer_plots$segment_rate_plot)) {
  buffer_plots$segment_rate_plot
}

# Display temporal analysis
buffer_plots$temporal_plot

# Display severity comparison
buffer_plots$severity_plot
```

## 8. Panel Data Creation (Improved Approach)

Now let's create the panel dataset using our improved approach:

```{r panel_data, message=FALSE, warning=FALSE}
# Check if panel data already exists
if (file.exists("data/processed/panel_results_v2.rds")) {
  panel_results <- readRDS("data/processed/panel_results_v2.rds")
  panel_data <- panel_results$panel_data

  cat("Loaded preprocessed panel data (improved version)\n")
} else {
  # Create panel dataset with improved spatial matching
  cat("Creating panel dataset with improved approach...\n")
  panel_results <- create_analysis_panel(df.const.sf, df.acc.sf, ca_roads, buffer_dist = 300)
  panel_data <- panel_results$panel_data

  # Save panel data for future use
  saveRDS(panel_results, "data/processed/panel_results_v2.rds")
}

# Generate visualizations of the new panel data
panel_viz <- visualize_panel_data(panel_results)

# Display visualizations
panel_viz$road_construction_map
ggsave("imgs/road_construction_map.png", panel_viz$road_construction_map, width = 10, height = 8, dpi = 300)

panel_viz$road_accident_map
ggsave("imgs/road_accident_map.png", panel_viz$road_accident_map, width = 10, height = 8, dpi = 300)

panel_viz$temporal_plot
ggsave("imgs/temporal_analysis.png", panel_viz$temporal_plot, width = 10, height = 6, dpi = 300)

panel_viz$urban_rural_plot
ggsave("imgs/urban_rural_comparison.png", panel_viz$urban_rural_plot, width = 10, height = 6, dpi = 300)

# Print panel data structure
cat("Panel data structure:\n")
str(panel_data)
cat("Panel data dimensions:", dim(panel_data), "\n")

# Summary statistics
cat("\nConstruction statistics:\n")
table(panel_data$has_construction)

cat("\nUrban/rural distribution:\n")
table(panel_data$is_urban, panel_data$has_construction)

cat("\nAccident summary:\n")
summary(panel_data$accident_count)
```

## 9. Regression Analysis

Let's run the regression models:

```{r regression_models, message=FALSE, warning=FALSE}
# Run regression models
cat("Running regression models...\n")
regression_models <- run_regression_models(panel_data)

# Display regression results
modelsummary(
  regression_models$all_models,
  stars = TRUE,
  title = "DiD Estimates of Construction Impacts on Accidents"
)
```

### Interpretation of Regression Models

The table shows five different regression models examining the relationship between road construction and traffic accidents in California:

### Model 1: Base DiD Model
- **has_construction**: There's a highly significant positive effect (6.291***) of construction activity on accident counts. This suggests that when construction is present on a road segment, there are approximately 6.3 more accidents compared to when there is no construction, controlling for road-specific and time-specific factors.
- The model explains a substantial portion of the variation (R² = 0.956), though much of this is due to the fixed effects rather than the construction variable itself (Within R² = 0.001).

### Model 2: Urban/Rural Heterogeneity Model
- **has_construction**: The baseline effect of construction increases to 7.436** accidents in rural areas (the reference category).
- **has_construction × is_urbanTRUE**: The interaction term (-1.664) suggests that the effect of construction on accidents is somewhat smaller in urban areas compared to rural areas, though this difference is not statistically significant.

### Model 3: Full Model with Seasonal Interactions
- **has_construction**: The baseline coefficient becomes strongly negative (-14.187***), but this must be interpreted carefully alongside the seasonal interaction terms.
- **has_construction × is_winterTRUE**: There's a very large and significant increase in accidents (57.193***) when construction occurs during winter months.
- **has_construction × is_summerTRUE**: Summer construction also shows a significant positive effect (8.024**) on accidents.
- These seasonal interactions indicate that construction's impact on accidents varies significantly by season, with winter construction being particularly associated with increased accidents.

### Model 4: Accident Rate Model (per km)
- **has_construction**: The effect on accident rates per kilometer is negative but not statistically significant (-0.047).
- This suggests that while the absolute number of accidents increases during construction (as seen in Models 1-3), when normalized by road length, the effect becomes insignificant.

### Model 5: Accident Severity Model
- **has_construction**: There is no significant effect (-0.033) of construction on accident severity.
- **has_construction × is_urbanTRUE**: There's a small positive but non-significant effect (0.045) on severity in urban areas.
- This indicates that while construction may increase the frequency of accidents, it doesn't meaningfully affect their severity.

### Key Findings:

1. **Construction generally increases accident frequency**: Models 1 and 2 show that active construction sites are associated with more accidents.

2. **Seasonal variation is critical**: Model 3 reveals that construction's impact varies substantially by season, with winter construction having a dramatically larger effect on accidents than construction in other seasons.

3. **Urban/rural differences are present but modest**: The effect of construction on accidents appears somewhat smaller in urban areas compared to rural areas, though this difference isn't statistically significant.

4. **Normalized effects are weaker**: When considering accidents per kilometer (Model 4), the effect of construction disappears, suggesting that the absolute increase in accidents may be partly related to road length or exposure.

5. **No impact on severity**: Model 5 indicates that construction doesn't significantly affect how severe accidents are.

These findings have important implications for construction planning and traffic safety management, particularly regarding seasonal timing of road projects and potentially different approaches needed for urban versus rural environments.

<hr>

## 10. Robustness Checks

Let's perform the robustness checks:

```{r robustness_checks, message=FALSE, warning=FALSE, fig.width=12, fig.height=10}
# Run robustness checks
cat("Running robustness checks...\n")
robustness_results <- run_robustness_checks(panel_data, regression_models$base_model)

# Create visualizations
robustness_plots <- visualize_robustness_checks(robustness_results, regression_models$base_model)

# Display event study plot
robustness_plots$coef_plot

# Display placebo tests plot
robustness_plots$placebo_plot

# Display buffer sensitivity plot
robustness_plots$buffer_plot

# Display fixed effects comparison
robustness_plots$fe_comparison

# Display outcome transformations comparison
robustness_plots$transform_comparison

# Display combined robustness panel
robustness_plots$robustness_panel
```

## 11. Propensity Score Matching

Finally, let's examine the propensity score matching results:

```{r matching, message=FALSE, warning=FALSE}
# Examine matching results
cat("Matching diagnostics:\n")
summary(robustness_results$matched_data)

# Get matched data
matched_data <- match.data(robustness_results$matched_data)

# Re-run main model on matched data
matched_model <- feols(
  accident_count ~ has_construction | road_id + month, 
  data = panel_data %>% 
    inner_join(
      matched_data %>% 
        select(road_id, weights), 
      by = "road_id"
    )
)

# Compare results
modelsummary(
  list("Original" = regression_models$base_model, "Matched" = matched_model),
  stars = TRUE,
  title = "Comparison of Original and Matched Models"
)
```

## 12. Conclusion

This analysis has investigated the causal impact of road construction on traffic accident rates in California during 2021. Our key findings include:

1. We found that approximately `r analysis_results$pct_near_construction`% of accidents occurred within active construction zones, suggesting a significant relationship between construction activities and traffic safety.

2. The difference-in-differences models showed consistent effects of construction on accident rates, with differential impacts between urban and rural areas.

3. The event study analysis shows that accident rates begin to change [before/after] construction starts, with the strongest effects during [early/middle/late] phases of construction.

4. Our robustness checks confirmed the stability of our findings through placebo tests, fixed effects variations, and propensity score matching.

5. The results indicate that construction activities [increase/decrease] accident rates by [X]%, with [stronger/weaker] effects in urban areas compared to rural areas.

These findings have important implications for construction planning, traffic management strategies, and safety policies, particularly regarding the differential effects between urban and rural environments.

### Future Research Directions

1. Investigating specific construction characteristics (duration, scope, traffic management) that impact safety outcomes.
2. Exploring the spatial spillover effects of construction on neighboring road segments.
3. Analyzing how different weather conditions interact with construction activities to influence accident risk.
4. Developing predictive models for construction-related accident risk to inform preventive measures.

```{r}
# Import write xlsx
library(writexl)
# Store panel data in Excel
write_xlsx(panel_data, "data/panel_data.xlsx")
```

```{r}
# Comprehensive Propensity Score Matching Analysis for Road Construction Safety

# Load required packages
library(tidyverse)
library(sf)
library(fixest)
library(MatchIt)
library(cobalt)
library(WeightIt)
library(data.table)
library(patchwork)

############################
# 1. DATA PREPARATION
############################

# Assuming panel_data is already created from your previous code
# If needed, load panel data from saved RDS file
if (!exists("panel_data")) {
  if (file.exists("data/processed/panel_results_v2.rds")) {
    panel_results <- readRDS("data/processed/panel_results_v2.rds")
    panel_data <- panel_results$panel_data
    print("Loaded panel data")
  } else {
    stop("Please run the panel data creation step first")
  }
}

# Create pre-construction accident metrics for each road segment
road_metrics <- panel_data %>%
  # Get baseline period (use the first 3 months if possible)
  filter(month %in% sort(unique(month))[1:3]) %>%
  group_by(road_id) %>%
  summarize(
    pre_accident_count = mean(accident_count, na.rm = TRUE),
    pre_severe_accidents = mean(severe_accident_count, na.rm = TRUE),
    pre_accident_rate = mean(accidents_per_km, na.rm = TRUE),
    road_length_km = first(road_length_km),
    road_type = first(road_type),
    is_urban = first(is_urban),
    .groups = "drop"
  )

# Identify which roads ever had construction
construction_roads <- panel_data %>%
  group_by(road_id) %>%
  summarize(
    ever_construction = any(has_construction == 1),
    construction_count = sum(has_construction),
    total_periods = n(),
    construction_intensity = construction_count / total_periods,
    .groups = "drop"
  )

# Combine road characteristics with construction status
matching_data <- road_metrics %>%
  left_join(construction_roads, by = "road_id") %>%
  # Convert factors as needed
  mutate(
    road_type = factor(road_type),
    is_urban = factor(is_urban),
    ever_construction = factor(ever_construction),
    # Create categorized variables for exact matching
    accident_level = cut(pre_accident_count, 
                        breaks = c(-Inf, 0, 1, 5, 10, Inf),
                        labels = c("None", "Very Low", "Low", "Medium", "High")),
    length_category = cut(road_length_km,
                         breaks = c(0, 1, 5, 10, 25, Inf),
                         labels = c("Very Short", "Short", "Medium", "Long", "Very Long"))
  )

# Print summary to verify data preparation
cat("Matching data preparation complete\n")
cat("Total roads:", nrow(matching_data), "\n")
cat("Roads with construction:", sum(matching_data$ever_construction == TRUE), "\n")
cat("Roads without construction:", sum(matching_data$ever_construction == FALSE), "\n")

############################
# 2. PROPENSITY SCORE MATCHING
############################

# Set seed for reproducibility
set.seed(123)

# Run multiple matching methods to compare
matching_methods <- list(
  # 1. Nearest neighbor matching
  nearest = matchit(
    ever_construction ~ pre_accident_count + pre_severe_accidents + 
                      road_length_km + road_type + is_urban,
    data = matching_data,
    method = "nearest",
    ratio = 3,  # Match each treated unit to 3 control units
    replace = FALSE
  ),

  # 2. Nearest neighbor with exact matching on key variables
  nearest_exact = matchit(
    ever_construction ~ pre_accident_count + pre_severe_accidents + road_length_km,
    data = matching_data,
    method = "nearest",
    exact = ~road_type + is_urban,  # Exact matching on these variables
    ratio = 2
  ),

  # 3. Optimal matching (may be slow with large datasets)
  optimal = matchit(
    ever_construction ~ pre_accident_count + pre_severe_accidents + 
                      road_length_km + road_type + is_urban,
    data = matching_data,
    method = "optimal",
    ratio = 1
  ),

  # 4. Full matching (flexible ratio)
  full = matchit(
    ever_construction ~ pre_accident_count + pre_severe_accidents + 
                      road_length_km + road_type + is_urban,
    data = matching_data,
    method = "full"
  ),

  # 5. Genetic matching (may be slow but often produces best balance)
  genetic = matchit(
    ever_construction ~ pre_accident_count + pre_severe_accidents + road_length_km,
    data = matching_data,
    method = "genetic",
    pop.size = 100  # Smaller population for faster convergence
  )
)

# Save matching results
saveRDS(matching_methods, "data/processed/matching_methods.rds")

############################
# 3. BALANCE ASSESSMENT
############################

# Function to assess balance for a matching method
assess_balance <- function(match_obj, name) {
  # Create love plot
  love_plot <- love.plot(
    match_obj, 
    binary = "std", 
    thresholds = c(m = 0.1),
    var.order = "unadjusted",
    abs = TRUE,
    title = paste("Covariate Balance -", name)
  )

  # Summarize balance
  balance_summary <- summary(match_obj)

  # Safely get match counts
  get_matched_counts <- function(bs) {
    # Try to extract from nn matrix if available
    if (is.list(bs) && "nn" %in% names(bs) && is.matrix(bs$nn)) {
      matched_treated <- sum(bs$nn[1,])

      # Check if "Unmatched" column exists
      if ("Unmatched" %in% colnames(bs$nn)) {
        matched_control <- sum(bs$nn[2,]) - bs$nn[2, "Unmatched"]
        unmatched_control <- bs$nn[2, "Unmatched"]
      } else {
        matched_control <- sum(bs$nn[2,])
        unmatched_control <- NA
      }
    } else {
      # Extract from match.data directly
      matched_data <- match.data(match_obj)
      matched_treated <- sum(matched_data$ever_construction == TRUE)
      matched_control <- sum(matched_data$ever_construction == FALSE)
      unmatched_control <- NA
    }

    return(list(
      matched_treated = matched_treated,
      matched_control = matched_control,
      unmatched_control = unmatched_control
    ))
  }

  # Safely get mean differences
  get_mean_diffs <- function(bs) {
    if (is.list(bs) && "sum.all" %in% names(bs) && is.data.frame(bs$sum.all) && "std.mean.diff" %in% names(bs$sum.all)) {
      mean_diff_unadj <- mean(abs(bs$sum.all$std.mean.diff))
      max_diff_unadj <- max(abs(bs$sum.all$std.mean.diff))
    } else if (is.list(bs) && "sum.across" %in% names(bs)) {
      mean_diff_unadj <- mean(abs(bs$sum.across[, "Std. Mean Diff."]))
      max_diff_unadj <- max(abs(bs$sum.across[, "Std. Mean Diff."]))
    } else {
      mean_diff_unadj <- NA
      max_diff_unadj <- NA
    }

    if (is.list(bs) && "sum.matched" %in% names(bs) && is.data.frame(bs$sum.matched) && "std.mean.diff" %in% names(bs$sum.matched)) {
      mean_diff_adj <- mean(abs(bs$sum.matched$std.mean.diff))
      max_diff_adj <- max(abs(bs$sum.matched$std.mean.diff))
    } else if (is.list(bs) && "sum.across" %in% names(bs) && "Std. Mean Diff. (Matched)" %in% colnames(bs$sum.across)) {
      mean_diff_adj <- mean(abs(bs$sum.across[, "Std. Mean Diff. (Matched)"]))
      max_diff_adj <- max(abs(bs$sum.across[, "Std. Mean Diff. (Matched)"]))
    } else {
      mean_diff_adj <- NA
      max_diff_adj <- NA
    }

    return(list(
      mean_diff_unadj = mean_diff_unadj,
      mean_diff_adj = mean_diff_adj,
      max_diff_unadj = max_diff_unadj,
      max_diff_adj = max_diff_adj
    ))
  }

  # Get match counts
  counts <- get_matched_counts(balance_summary)

  # Get mean differences
  diffs <- get_mean_diffs(balance_summary)

  # Create metrics data frame
  balance_metrics <- data.frame(
    Method = name,
    Mean_Diff_Unadj = diffs$mean_diff_unadj,
    Mean_Diff_Adj = diffs$mean_diff_adj,
    Max_Diff_Unadj = diffs$max_diff_unadj,
    Max_Diff_Adj = diffs$max_diff_adj,
    Matched_Treated = counts$matched_treated,
    Matched_Control = counts$matched_control,
    Unmatched_Control = counts$unmatched_control
  )

  # Return results
  return(list(
    love_plot = love_plot,
    balance_summary = balance_summary,
    balance_metrics = balance_metrics
  ))
}

# Assess balance for each method
balance_results <- lapply(names(matching_methods), function(method) {
  assess_balance(matching_methods[[method]], method)
})
names(balance_results) <- names(matching_methods)

# Combine balance metrics
balance_comparison <- do.call(rbind, lapply(balance_results, function(x) x$balance_metrics))

# Plot balance comparison
balance_plot <- ggplot(balance_comparison, aes(x = Method)) +
  geom_segment(aes(y = Mean_Diff_Unadj, yend = Mean_Diff_Adj, 
                  x = Method, xend = Method), 
               color = "gray50", size = 1.5) +
  geom_point(aes(y = Mean_Diff_Unadj), color = "red", size = 3) +
  geom_point(aes(y = Mean_Diff_Adj), color = "blue", size = 3) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Mean Absolute Standardized Difference",
       subtitle = "Before (red) vs. After (blue) Matching",
       y = "Mean Absolute Std. Difference",
       x = "")

# Save the plot
ggsave("imgs/balance_comparison.png", balance_plot, width = 10, height = 6, dpi = 300)

# Display balance metrics
print(balance_comparison)

############################
# 4. REGRESSION ANALYSIS WITH MATCHED DATA
############################

# Function to run regressions on matched data
run_matched_regressions <- function(match_obj, method_name) {
  # Get matched data
  matched_data <- match.data(match_obj)

  # Get matched road IDs
  matched_road_ids <- matched_data$road_id

  # Filter panel data to matched roads
  matched_panel <- panel_data %>%
    filter(road_id %in% matched_road_ids)

  # Add weights if available
  if ("weights" %in% names(matched_data)) {
    matched_panel <- matched_panel %>%
      left_join(select(matched_data, road_id, weights), by = "road_id")
  } else {
    matched_panel$weights <- 1
  }

  # 1. Base DiD model
  model1 <- feols(
    accident_count ~ has_construction | road_id + month, 
    data = matched_panel,
    weights = ~ weights
  )

  # 2. Urban/rural heterogeneity
  model2 <- feols(
    accident_count ~ has_construction + has_construction:is_urban | road_id + month, 
    data = matched_panel,
    weights = ~ weights
  )

  # 3. Seasonal heterogeneity
  model3 <- feols(
    accident_count ~ has_construction + has_construction:is_winter + 
                    has_construction:is_summer | road_id + month, 
    data = matched_panel,
    weights = ~ weights
  )

  # 4. Full interaction model
  model4 <- feols(
    accident_count ~ has_construction + has_construction:is_urban + 
                    has_construction:is_winter + has_construction:is_summer | 
                    road_id + month,
    data = matched_panel,
    weights = ~ weights
  )

  # 5. Road type heterogeneity
  model5 <- feols(
    accident_count ~ has_construction + has_construction:i(road_type, ref = "Other") | 
                    road_id + month,
    data = matched_panel,
    weights = ~ weights
  )

  # Return models
  return(list(
    method = method_name,
    base_model = model1,
    urban_model = model2,
    seasonal_model = model3,
    full_model = model4,
    roadtype_model = model5,
    matched_panel = matched_panel
  ))
}

# Run regressions for each matching method
regression_results <- lapply(names(matching_methods), function(method) {
  run_matched_regressions(matching_methods[[method]], method)
})
names(regression_results) <- names(matching_methods)

# Comparison table of main effects across methods
main_effects <- data.frame(
  Method = names(regression_results),
  Effect = sapply(regression_results, function(x) coef(x$base_model)["has_construction"]),
  SE = sapply(regression_results, function(x) se(x$base_model)["has_construction"]),
  CI_Lower = sapply(regression_results, function(x) 
    coef(x$base_model)["has_construction"] - 1.96 * se(x$base_model)["has_construction"]),
  CI_Upper = sapply(regression_results, function(x) 
    coef(x$base_model)["has_construction"] + 1.96 * se(x$base_model)["has_construction"]),
  Matched_Roads = sapply(regression_results, function(x) length(unique(x$matched_panel$road_id)))
)

# Save main effects
write.csv(main_effects, "data/processed/psm_main_effects.csv", row.names = FALSE)

############################
# 5. VISUALIZATIONS
############################

# Plot main effects
effects_plot <- ggplot(main_effects, aes(x = reorder(Method, Effect), y = Effect)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = CI_Lower, ymax = CI_Upper), width = 0.2) +
  theme_minimal() +
  coord_flip() +
  labs(title = "Construction Effect on Accidents Across Matching Methods",
       subtitle = "Point Estimates with 95% Confidence Intervals",
       y = "Effect on Accident Count",
       x = "")

# Save effects plot
ggsave("imgs/psm_effects_comparison.png", effects_plot, width = 10, height = 6, dpi = 300)

# Function to create trend plots for matched data
create_trend_plot <- function(matched_panel, method_name) {
  # Get treatment timing
  treatment_timing <- matched_panel %>%
    filter(has_construction == 1) %>%
    group_by(road_id) %>%
    summarize(first_treatment = min(month), .groups = "drop")

  # Add treatment timing to all roads
  matched_panel_with_timing <- matched_panel %>%
    left_join(treatment_timing, by = "road_id") %>%
    mutate(
      # For roads without construction, use median of treated roads
      first_treatment = if_else(is.na(first_treatment), 
                              median(treatment_timing$first_treatment), 
                              first_treatment),
      # Calculate relative time to treatment
      rel_month = as.numeric(difftime(month, first_treatment, units = "days")) / 30,
      # Create period groups
      rel_period = cut(rel_month, 
                     breaks = c(-Inf, -3, -2, -1, 0, 1, 2, 3, Inf),
                     labels = c("t-3+", "t-2", "t-1", "t0", "t+1", "t+2", "t+3", "t+3+"))
    )

  # Calculate average outcomes by relative period and treatment status
  trend_data <- matched_panel_with_timing %>%
    group_by(rel_period, has_construction) %>%
    summarize(
      mean_accidents = mean(accident_count, na.rm = TRUE),
      se = sd(accident_count, na.rm = TRUE) / sqrt(n()),
      .groups = "drop"
    )

  # Create trend plot
  trend_plot <- ggplot(trend_data, 
                      aes(x = rel_period, y = mean_accidents, 
                          group = factor(has_construction),
                          color = factor(has_construction))) +
    geom_line() +
    geom_point(size = 3) +
    geom_errorbar(aes(ymin = mean_accidents - 1.96*se, 
                     ymax = mean_accidents + 1.96*se), width = 0.2) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = paste("Accident Trends by Time to Construction -", method_name),
         x = "Months Relative to Construction Start",
         y = "Average Accident Count",
         color = "Construction")

  return(trend_plot)
}

# Create trend plots for each matching method
trend_plots <- lapply(names(regression_results), function(method) {
  create_trend_plot(regression_results[[method]]$matched_panel, method)
})
names(trend_plots) <- names(regression_results)

# Save trend plots
for (method in names(trend_plots)) {
  ggsave(paste0("imgs/trends_", method, ".png"), trend_plots[[method]], width = 10, height = 6, dpi = 300)
}

############################
# 6. HETEROGENEITY ANALYSIS
############################

# Function to extract heterogeneous effects
extract_heterogeneity <- function(model_list, method_name) {
  # Urban heterogeneity
  urban_effects <- data.frame(
    Method = method_name,
    Variable = c("Base Effect", "Urban Interaction"),
    Estimate = c(
      coef(model_list$urban_model)["has_construction"],
      coef(model_list$urban_model)["has_construction:is_urbanTRUE"]
    ),
    SE = c(
      se(model_list$urban_model)["has_construction"],
      se(model_list$urban_model)["has_construction:is_urbanTRUE"]
    )
  )

  # Seasonal heterogeneity
  season_effects <- data.frame(
    Method = method_name,
    Variable = c("Base Effect", "Winter Interaction", "Summer Interaction"),
    Estimate = c(
      coef(model_list$seasonal_model)["has_construction"],
      coef(model_list$seasonal_model)["has_construction:is_winterTRUE"],
      coef(model_list$seasonal_model)["has_construction:is_summerTRUE"]
    ),
    SE = c(
      se(model_list$seasonal_model)["has_construction"],
      se(model_list$seasonal_model)["has_construction:is_winterTRUE"],
      se(model_list$seasonal_model)["has_construction:is_summerTRUE"]
    )
  )

  # Combine
  return(rbind(urban_effects, season_effects))
}

# Extract heterogeneity for each method
heterogeneity_results <- do.call(rbind, lapply(names(regression_results), function(method) {
  extract_heterogeneity(regression_results[[method]], method)
}))

# Add confidence intervals
heterogeneity_results <- heterogeneity_results %>%
  mutate(
    CI_Lower = Estimate - 1.96 * SE,
    CI_Upper = Estimate + 1.96 * SE
  )

# Create heterogeneity plot
het_plot <- ggplot(heterogeneity_results, 
                  aes(x = Variable, y = Estimate, color = Method)) +
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  geom_errorbar(aes(ymin = CI_Lower, ymax = CI_Upper), 
               position = position_dodge(width = 0.5), width = 0.2) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Heterogeneous Effects Across Matching Methods",
       y = "Effect on Accident Count",
       x = "")

# Save heterogeneity plot
ggsave("imgs/psm_heterogeneity.png", het_plot, width = 12, height = 8, dpi = 300)

############################
# 7. SENSITIVITY ANALYSIS
############################

# Sensitivity to caliper width (using nearest neighbor matching)
calipers <- c(0.1, 0.2, 0.3, 0.5)

caliper_results <- lapply(calipers, function(caliper) {
  # Create matching with this caliper
  match_obj <- matchit(
    ever_construction ~ pre_accident_count + pre_severe_accidents + 
                      road_length_km + road_type + is_urban,
    data = matching_data,
    method = "nearest",
    caliper = caliper,
    ratio = 1
  )

  # Get matched data
  matched_data <- match.data(match_obj)

  # Get matched road IDs
  matched_road_ids <- matched_data$road_id

  # Filter panel data to matched roads
  matched_panel <- panel_data %>%
    filter(road_id %in% matched_road_ids)

  # Run basic model
  model <- feols(
    accident_count ~ has_construction | road_id + month, 
    data = matched_panel
  )

  # Return results
  return(list(
    caliper = caliper,
    match_obj = match_obj,
    model = model,
    estimate = coef(model)["has_construction"],
    se = se(model)["has_construction"],
    matched_treated = sum(matched_data$ever_construction == TRUE),
    matched_control = sum(matched_data$ever_construction == FALSE),
    mean_distance = mean(matched_data$distance)
  ))
})

# Extract caliper sensitivity results
caliper_comparison <- data.frame(
  Caliper = sapply(caliper_results, function(x) x$caliper),
  Estimate = sapply(caliper_results, function(x) x$estimate),
  SE = sapply(caliper_results, function(x) x$se),
  CI_Lower = sapply(caliper_results, function(x) x$estimate - 1.96 * x$se),
  CI_Upper = sapply(caliper_results, function(x) x$estimate + 1.96 * x$se),
  Matched_Treated = sapply(caliper_results, function(x) x$matched_treated),
  Matched_Control = sapply(caliper_results, function(x) x$matched_control),
  Mean_Distance = sapply(caliper_results, function(x) x$mean_distance)
)

# Caliper sensitivity plot
caliper_plot <- ggplot(caliper_comparison, aes(x = factor(Caliper), y = Estimate)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = CI_Lower, ymax = CI_Upper), width = 0.2) +
  theme_minimal() +
  labs(title = "Sensitivity to Caliper Width",
       subtitle = "Point Estimates with 95% Confidence Intervals",
       y = "Effect on Accident Count",
       x = "Caliper Width")

# Save caliper sensitivity plot
ggsave("imgs/caliper_sensitivity.png", caliper_plot, width = 10, height = 6, dpi = 300)

############################
# 8. COMPARISON WITH ORIGINAL ESTIMATE
############################

# Run the original model for comparison
original_model <- feols(
  accident_count ~ has_construction | road_id + month, 
  data = panel_data
)

# Get original and matched estimates
comparison_data <- data.frame(
  Model = c("Original (Unmatched)", names(regression_results)),
  Estimate = c(
    coef(original_model)["has_construction"],
    sapply(regression_results, function(x) coef(x$base_model)["has_construction"])
  ),
  SE = c(
    se(original_model)["has_construction"],
    sapply(regression_results, function(x) se(x$base_model)["has_construction"])
  )
)

# Add confidence intervals
comparison_data <- comparison_data %>%
  mutate(
    CI_Lower = Estimate - 1.96 * SE,
    CI_Upper = Estimate + 1.96 * SE
  )

# Create comparison plot
comparison_plot <- ggplot(comparison_data, 
                         aes(x = reorder(Model, Estimate), y = Estimate)) +
  geom_point(size = 4, aes(color = Model == "Original (Unmatched)")) +
  geom_errorbar(aes(ymin = CI_Lower, ymax = CI_Upper, 
                   color = Model == "Original (Unmatched)"), width = 0.2) +
  scale_color_manual(values = c("blue", "red")) +
  theme_minimal() +
  coord_flip() +
  theme(legend.position = "none") +
  labs(title = "Effect of Construction on Accidents: Matched vs. Unmatched",
       subtitle = "Point Estimates with 95% Confidence Intervals",
       y = "Effect on Accident Count",
       x = "")

# Save comparison plot
ggsave("imgs/original_vs_matched.png", comparison_plot, width = 10, height = 6, dpi = 300)

############################
# 9. FINAL RESULTS AND OUTPUT
############################

# Choose the best matching method based on balance metrics
best_method <- balance_comparison %>%
  arrange(Mean_Diff_Adj) %>%
  pull(Method) %>%
  .[1]

cat("Best matching method based on balance:", best_method, "\n")

# Final model results using the best method
final_results <- regression_results[[best_method]]

# Create final results table
final_models <- list(
  "Base" = final_results$base_model,
  "Urban" = final_results$urban_model,
  "Seasonal" = final_results$seasonal_model,
  "Full" = final_results$full_model
)

# Save comprehensive results
saveRDS(list(
  matching_methods = matching_methods,
  balance_results = balance_results,
  regression_results = regression_results,
  main_effects = main_effects,
  heterogeneity_results = heterogeneity_results,
  caliper_comparison = caliper_comparison,
  comparison_data = comparison_data,
  final_results = final_results
), "data/processed/comprehensive_psm_results.rds")

# Print key findings
cat("\n=== PROPENSITY SCORE MATCHING ANALYSIS COMPLETE ===\n")
cat("Original construction effect:", round(coef(original_model)["has_construction"], 2), 
    "(SE:", round(se(original_model)["has_construction"], 2), ")\n")
cat("Best matched construction effect:", 
    round(coef(final_results$base_model)["has_construction"], 2),
    "(SE:", round(se(final_results$base_model)["has_construction"], 2), ")\n")

# Print heterogeneity summary for best method
cat("\nHeterogeneity in construction effects (", best_method, "):\n", sep = "")
cat("Urban areas:", 
    round(coef(final_results$urban_model)["has_construction:is_urbanTRUE"], 2),
    "(SE:", round(se(final_results$urban_model)["has_construction:is_urbanTRUE"], 2), ")\n")
cat("Winter effect:", 
    round(coef(final_results$seasonal_model)["has_construction:is_winterTRUE"], 2),
    "(SE:", round(se(final_results$seasonal_model)["has_construction:is_winterTRUE"], 2), ")\n")
cat("Summer effect:", 
    round(coef(final_results$seasonal_model)["has_construction:is_summerTRUE"], 2),
    "(SE:", round(se(final_results$seasonal_model)["has_construction:is_summerTRUE"], 2), ")\n")

cat("\nAnalysis complete. Results saved to 'data/processed/comprehensive_psm_results.rds'\n")
cat("Visualizations saved to the 'imgs' directory\n")
```

### 2025-03-25 final/finalSIMON_MoritzPeistNoemiLucchiSimonVellin .Rmd
Duplicated .Rmd file to have a fresh start

---
title: "Geospatial Data Science - Final project"
author:
- Moritz Peist
- Noemi Lucchi
- Simon Vellin
subtitle: Analysis of US road accidents and construction sites
output:
  html_document:
    toc: true
    df_print: paged
  html_notebook:
    df_print: paged
    toc: true
---

<hr>

```{css, echo=FALSE}
h1.title {
  text-align: center;
}

h4.author, h4.date {
  text-align: center;
}

.author {
  text-align: center;
}

.date {
  text-align: center;
}

/* Add styling for subtitle */
h3.subtitle {
  text-align: center;
  font-style: italic;
  margin-top: -10px;
  color: #606060;
}
```

<div style="text-align: center;">
  <img src="../Barcelona_School_of_Economics_logo.svg" alt="BSE logo">
</div>

<hr>

Load packages

```{r Packages}
library(data.table)   # For faster data manipulation
library(tidyverse)    # For data manipulation and visualization
library(sf)           # For spatial data handling
library(leaflet)      # For interactive maps
library(leaflet.extras) # For additional leaflet features
library(mapview)      # For easier map visualization
library(tmap)         # For thematic maps
library(tigris)       # For US road networks
library(future)       # For parallel processing
library(future.apply) # For parallel processing with apply functions
library(sf)           # For spatial data handling
library(RPostgres)    # For PostgreSQL connection
library(DBI)          # For database interface

# Create directories if they don't exist
if (!dir.exists("./data/tigris")) {
  dir.create("./data/tigris", recursive = TRUE)
}

# Set custom cache directory (optional)
options(tigris_cache_dir = "./data/tigris")
# Configure tigris to use caching
options(tigris_use_cache = TRUE)
```

# Set-up convenience functions for parallel processing (as data is quite large)
```{r Parallel Buffering}
# Function to create buffers in batches with proper projection
create_buffers_in_batches <- function(sf_object, buffer_dist, batch_size = 500) {
  # First, reproject to an appropriate projected CRS for the region
  # For California, UTM Zone 10N (EPSG:26910) or 11N (EPSG:26911) works well
  # If we're unsure about your region, a web mercator projection (3857) is a reasonable default
  message("Reprojecting data to a meter-based CRS...")
  sf_object_projected <- st_transform(sf_object, 3310)  # Web Mercator

  n_features <- nrow(sf_object_projected)
  n_batches <- ceiling(n_features / batch_size)

  # Create empty list to store results
  buffer_list <- vector("list", n_batches)

  # Process in batches
  for (i in 1:n_batches) {
    start_idx <- (i-1) * batch_size + 1
    end_idx <- min(i * batch_size, n_features)

    cat(sprintf("Processing batch %d of %d (features %d to %d)\n", 
                i, n_batches, start_idx, end_idx))

    # Extract batch
    batch <- sf_object_projected[start_idx:end_idx, ]

    # Create buffer (with parallel processing within sf)
    buffer_list[[i]] <- st_buffer(batch, dist = buffer_dist)
  }

  # Combine results
  result <- do.call(rbind, buffer_list)

  # Reproject back to original CRS if needed
  message("Reprojecting results back to original CRS...")
  st_transform(result, st_crs(sf_object))
}
```

Load data (filter for California for now to limit data set size)

```{r Data Load Accidents}
# Efficient approach
df.acc <- fread("data/us_accidents/US_accidents_March23.csv")[
  # Filter date range of 2021
  lubridate::year(as.Date(Start_Time)) == 2021 & 
  # And California
  State == "CA"
][, `:=`(
  # Add year, quarter, month columns
  year = data.table::year(Start_Time),
  quarter = data.table::quarter(Start_Time),
  month = data.table::month(Start_Time),
  # Calculate duration (assuming End_Time exists in the dataset)
  duration = as.numeric(difftime(End_Time, Start_Time, units = "days"))
)] %>% 
  as_tibble()  # Convert to tibble only at the end for performance
```

```{r Data Load Constructions}
df.const <- fread("data/us_constructions/US_constructions_Dec21.csv")[
  # Filter date range of 2021
  lubridate::year(as.Date(Start_Time)) == 2021 & 
  # And California
  State == "CA"
][, `:=`(
  # Add year, quarter, month columns
  year = year(Start_Time),
  quarter = quarter(Start_Time),
  month = month(Start_Time),
  # Calculate duration (assuming End_Time exists in the dataset)
  duration = as.numeric(difftime(End_Time, Start_Time, units = "days"))
)] %>% 
  as_tibble()  # Convert to tibble only at the end for performance
```

Convert dataframes to sf objects

```{r Geospatial Data}
# Convert accident data to sf object
df.acc.sf <- df.acc %>%
  filter(!is.na(Start_Lat) & !is.na(Start_Lng)) %>%
  st_as_sf(coords = c("Start_Lng", "Start_Lat"), crs = 4326)

# Convert construction data to sf object
df.const.sf <- df.const %>%
  filter(!is.na(Start_Lat) & !is.na(Start_Lng)) %>%
  st_as_sf(coords = c("Start_Lng", "Start_Lat"), crs = 4326)
```

```{r Data Load Roads}
# Get US roads (this can be slow for the entire US, so we might want to limit by state)
# We also only get Primary and Secondary roads as they are the most important (i.e., where most constructions and accidents happen). Otherwise, the dataset would become to large (1.1GB)

ca_roads <- primary_secondary_roads("CA", year = 2021)

# Below code to get ALL roads for California (not recommended due to size)

# First get all counties in California
#ca_counties <- counties("CA", year = 2021)

# Then get roads for each county
#ca_roads_list <- lapply(ca_counties$COUNTYFP, function(county_code) {
#  roads("CA", county = county_code, year = 2021)
#})

# Optionally combine all county road data into one object
#ca_roads <- do.call(rbind, ca_roads_list)
```

```{r Static Maps}
# Static map for Accidents with year coloring
# First, store your plot in a variable
ca_plot <- ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  geom_sf(data = df.acc.sf %>% filter(State == "CA"), 
          aes(color = factor(year)), alpha = 0.7, size = 0.5) +
  scale_color_viridis_d(name = "Year") +  # Use viridis color palette for discrete values
  theme_minimal() +
  labs(title = "US Road Accidents by Year (2016-2021)") +
  theme(legend.position = "bottom")
print(ca_plot)
# Then save it using ggsave
ggsave(filename = "imgs/california_accidents.png", 
       plot = ca_plot,
       width = 10, # width in inches
       height = 8, # height in inches
       dpi = 300)  # resolution

# Construction sites map with year coloring
# First, store our plot in a variable
ca_const_plot <- ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  geom_sf(data = df.const.sf %>% filter(State == "CA"), 
          aes(color = factor(year)), alpha = 0.7, size = 0.5) +
  scale_color_viridis_d(name = "Year") +  # Use viridis color palette for discrete values
  theme_minimal() +
  labs(title = "US Construction Sites by Year (2016-2021)") +
  theme(legend.position = "bottom")
print(ca_const_plot)
# Then save it using ggsave
ggsave(filename = "imgs/california_construction.png", 
       plot = ca_const_plot,
       width = 10, # width in inches
       height = 8, # height in inches
       dpi = 300)  # resolution
```

```{r Heatmap static}
# Create heatmap of accidents using ggplot2
# Modified accident heatmap with transparency for low density areas
accident_heatmap <- ggplot() +
  # Add California roads as base layer
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  # Create density heatmap with transparency threshold
  stat_density_2d(
    data = df.acc %>% filter(!is.na(Start_Lat) & !is.na(Start_Lng)),
    aes(x = Start_Lng, y = Start_Lat, fill = after_stat(density), alpha = after_stat(density)),
    geom = "tile", 
    contour = FALSE,
    h = 0.1,
    n = 200
  ) +
  # Set alpha scaling with a minimum threshold
  scale_alpha_continuous(range = c(0, 0.9), guide = "none") +
  # Use color scheme similar to leaflet's default heatmap
  scale_fill_gradientn(
    colors = c("#0000FF", "#00FFFF", "#00FF00", "#FFFF00", "#FF0000"),
    name = "Accident\nDensity"
  ) +
  coord_sf(
    xlim = c(-124.5, -114.5),
    ylim = c(32.5, 42.5)
  ) +
  theme_minimal() +
  labs(
    title = "Accident Density Heatmap in California (2016-2021)",
    x = NULL,
    y = NULL
  ) +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold")
  )
# Display the plot
print(accident_heatmap)

ggsave(filename = "imgs/accident_heatmap.png", 
       plot = accident_heatmap,
       width = 10, # width in inches
       height = 8, # height in inches
       dpi = 300)  # resolution

```

```{r Construction Heatmap static}
# Create heatmap of construction sites using ggplot2
construction_heatmap <- ggplot() +
  # Add California roads as base layer
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  # Create density heatmap
  stat_density_2d(
    data = df.const %>% filter(!is.na(Start_Lat) & !is.na(Start_Lng)),
    aes(x = Start_Lng, y = Start_Lat, fill = after_stat(density)),
    geom = "tile", 
    contour = FALSE,
    alpha = 0.7,
    h = 0.1,  # Bandwidth
    n = 200   # Resolution
  ) +
  # Set alpha scaling with a minimum threshold
  scale_alpha_continuous(range = c(0, 0.9), guide = "none") +
  # Use color scheme matching your leaflet construction map
  scale_fill_gradientn(
    colors = c("yellow", "orange", "red"),
    name = "Construction\nDensity"
  ) +
  # Set appropriate boundaries for California
  coord_sf(
    xlim = c(-124.5, -114.5),
    ylim = c(32.5, 42.5)
  ) +
  theme_minimal() +
  labs(
    title = "Construction Density Heatmap in California (2016-2021)",
    x = NULL,
    y = NULL
  ) +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold")
  )
# Display the plot
print(construction_heatmap)
```

## Heatmaps - interactive

```{r Interactive Maps I}
# For interactive maps (often better for large datasets)
# Accidents map
accident_map <- leaflet() %>%
  addTiles() %>%
  addHeatmap(data = df.acc.sf, 
             intensity = ~1,
             radius = 8, 
             blur = 10) %>%
  setView(lng = -119.4179, lat = 36.7783, zoom = 6) %>% # Center to CA
  setMaxBounds(lng1 = -124.6, lat1 = 42.0,    # Northwest corner of CA
               lng2 = -114.1, lat2 = 32.5)    # Southeast corner of CA

# Display maps
accident_map
```

```{r Interactive Maps II}
# Construction sites map with California focus and bounds constraints
construction_map <- leaflet() %>%
  addTiles() %>%
  addHeatmap(data = df.const.sf, 
             intensity = ~1,
             radius = 8, 
             blur = 10,
             gradient = c("yellow", "orange", "red")) %>%
  setView(lng = -119.4179, lat = 36.7783, zoom = 6) %>%  # Center on California
  setMaxBounds(lng1 = -124.6, lat1 = 42.0,    # Northwest corner of CA
               lng2 = -114.1, lat2 = 32.5)    # Southeast corner of CA

# Display the map
construction_map
```

We construct construction sites now also as line objects since they have a start and end point

```{r Construction linestrings}
# Convert construction data to linestring sf object
df.const.lines <- df.const %>%
  filter(!is.na(Start_Lat) & !is.na(Start_Lng) & !is.na(End_Lat) & !is.na(End_Lng)) %>%
  mutate(geometry = pmap(list(Start_Lng, Start_Lat, End_Lng, End_Lat), 
                         function(start_lng, start_lat, end_lng, end_lat) {
                           coords <- matrix(c(start_lng, start_lat, 
                                             end_lng, end_lat), 
                                           ncol = 2, byrow = TRUE)
                           st_linestring(coords)
                         })) %>%
  st_sf(crs = 4326)
```

## Spatial operations in PostgreSQL

Since file sizes are so large and Spatial operations in R incredibly slow, we retreat to a PostgreSQL database for the following operations (PostGIS also uses the SF framework, thus R are almost identical to the following SQL functions used).

```{r DB ops}
# Create a connection to your PostgreSQL database
con <- dbConnect(
  Postgres(),
  dbname = "geospatial_final",
  host = "localhost",
  port = 5432,
  # Uncomment and add your credentials if needed
  user = "postgres",
  password = Sys.getenv("POSTGRES_PWD")
)

# Set connection parameters for performance optimization
dbExecute(con, "SET work_mem = '128MB'")           # Increase working memory
dbExecute(con, "SET maintenance_work_mem = '256MB'") # For index creation
dbExecute(con, "SET random_page_cost = 1.1")       # For SSD storage
dbExecute(con, "SET effective_cache_size = '2GB'") # Estimate of memory available
dbExecute(con, "SET max_parallel_workers_per_gather = 4") # Use multiple cores

# Check if PostGIS extension is enabled
postgis_check <- dbGetQuery(con, "SELECT PostGIS_version()")
cat("PostGIS version:", postgis_check[[1]], "\n")

# Check if SRID 3310 is available and add if not
dbExecute(con, "
  DO $$
  BEGIN
    IF NOT EXISTS (SELECT 1 FROM spatial_ref_sys WHERE srid = 3310) THEN
      INSERT INTO spatial_ref_sys (srid, auth_name, auth_srid, proj4text, srtext)
      VALUES (3310, 'EPSG', 3310, 
      '+proj=aea +lat_1=34 +lat_2=40.5 +lat_0=0 +lon_0=-120 +x_0=0 +y_0=-4000000 +datum=NAD83 +units=m +no_defs',
      'PROJCS[\"NAD83 / California Albers\",GEOGCS[\"NAD83\",DATUM[\"North_American_Datum_1983\",SPHEROID[\"GRS 1980\",6378137,298.257222101,AUTHORITY[\"EPSG\",\"7019\"]],AUTHORITY[\"EPSG\",\"6269\"]],PRIMEM[\"Greenwich\",0,AUTHORITY[\"EPSG\",\"8901\"]],UNIT[\"degree\",0.01745329251994328,AUTHORITY[\"EPSG\",\"9122\"]],AUTHORITY[\"EPSG\",\"4269\"]],PROJECTION[\"Albers_Conic_Equal_Area\"],PARAMETER[\"standard_parallel_1\",34],PARAMETER[\"standard_parallel_2\",40.5],PARAMETER[\"latitude_of_center\",0],PARAMETER[\"longitude_of_center\",-120],PARAMETER[\"false_easting\",0],PARAMETER[\"false_northing\",-4000000],UNIT[\"metre\",1,AUTHORITY[\"EPSG\",\"9001\"]],AUTHORITY[\"EPSG\",\"3310\"]]');
    END IF;
  END
  $$;
")

# Check if tables exist and only create them if they don't
# Begin transaction for data processing
dbExecute(con, "BEGIN")

# Check if accidents table exists, if not create it
has_accidents <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'accidents')")$exists
if (!has_accidents) {
  # Write accidents data to database
  sf::st_write(df.acc.sf, con, "accidents", 
               layer_options = c("GEOM_TYPE=GEOMETRY", 
                                 "GEOMETRY_NAME=geometry",
                                 "SRID=4326"))
  cat("Written accidents data to database\n")

  # Create spatial index for accidents
  dbExecute(con, "CREATE INDEX idx_accidents_geometry ON accidents USING GIST(geometry)")
  dbExecute(con, 'CREATE INDEX idx_accidents_start_time ON accidents("Start_Time");')
  dbExecute(con, "ANALYZE accidents")
  cat("Created spatial index for accidents table\n")
}

# Check if construction table exists, if not create it
has_construction <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'construction')")$exists

if (!has_construction) {
  # Write construction data to database
  sf::st_write(df.const.sf, con, "construction", 
               layer_options = c("GEOM_TYPE=GEOMETRY", 
                                 "GEOMETRY_NAME=geometry",
                                 "SRID=4326"))
  cat("Written construction data to database\n")

  # Create spatial index for construction
  dbExecute(con, "CREATE INDEX idx_construction_geometry ON construction USING GIST(geometry)")
  dbExecute(con, 'CREATE INDEX idx_construction_start_time ON construction("Start_Time");')
  dbExecute(con, 'CREATE INDEX idx_construction_end_time ON construction("End_Time");')
  dbExecute(con, "ANALYZE construction")
  cat("Created spatial index for construction table\n")
}

# Check if construction table with lines exists, if not create it
has_construction_lines <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'construction_lines')")$exists

if (!has_construction_lines) {
  # Write construction data to database
  sf::st_write(df.const.lines, con, "construction_lines", 
               layer_options = c("GEOM_TYPE=GEOMETRY", 
                                 "GEOMETRY_NAME=geometry",
                                 "SRID=4326"))
  cat("Written construction_lines data to database\n")

  # Create spatial index for construction
  dbExecute(con, "CREATE INDEX idx_construction_lines_geometry ON construction_lines USING GIST(geometry)")
  dbExecute(con, 'CREATE INDEX idx_construction_lines_start_time ON construction_lines("Start_Time");')
  dbExecute(con, 'CREATE INDEX idx_construction_lines_end_time ON construction_lines("End_Time");')
  dbExecute(con, "ANALYZE construction_lines")
  cat("Created spatial index for construction_lines table\n")
}

# Commit transaction for base tables
dbExecute(con, "COMMIT")

#---------------------------------------------------------------------------
# STEP 1A: Create enhanced construction buffers table with time ranges
#---------------------------------------------------------------------------
has_construction_buffers <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'construction_buffers')")$exists
if (!has_construction_buffers) {
  cat("Step 1A: Creating enhanced construction_buffers table with time ranges...\n")

  dbExecute(con, "BEGIN")

  # Create construction buffers table with appropriate projection and time ranges
  dbExecute(con, "
    CREATE TABLE construction_buffers AS
    SELECT 
      \"ID\" as construction_id,
      \"Start_Time\" as start_time,
      \"End_Time\" as end_time,
      ST_Transform(
        ST_Buffer(
          ST_Transform(geometry, 3310), -- Transform to California Albers for accurate distance
          300                           -- 300 meters buffer
        ), 
        4326                            -- Transform back to WGS 84 for joining
      ) AS geometry
    FROM construction
  ")

  # Add primary key and spatial index
  dbExecute(con, "ALTER TABLE construction_buffers ADD PRIMARY KEY (construction_id)")
  dbExecute(con, "CREATE INDEX idx_construction_buffers_geometry ON construction_buffers USING GIST(geometry)")
  dbExecute(con, "CREATE INDEX idx_construction_buffers_times ON construction_buffers(start_time, end_time)")
  dbExecute(con, "ANALYZE construction_buffers")

  dbExecute(con, "COMMIT")
  cat("Created enhanced construction_buffers table with spatial index and time ranges\n")
} else {
  cat("Construction buffers table already exists, skipping creation\n")
}

#---------------------------------------------------------------------------
# STEP 1B: Create spatial matches table (using enhanced pre-computed buffers)
#---------------------------------------------------------------------------
has_spatial_matches <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'spatial_matches')")$exists
if (!has_spatial_matches) {
  cat("Step 1B: Creating spatial_matches table using enhanced pre-computed buffers...\n")

  dbExecute(con, "BEGIN")

  # Simplified join using pre-computed buffers with time information
  dbExecute(con, "
    CREATE TABLE spatial_matches AS
    SELECT
      a.\"ID\" AS accident_id,
      a.\"Severity\" AS accident_severity,
      a.\"Start_Time\" AS accident_time,
      b.construction_id
    FROM 
      accidents a
    JOIN 
      construction_buffers b ON 
        ST_Within(a.geometry, b.geometry) AND
        a.\"Start_Time\" BETWEEN b.start_time AND b.end_time
  ")

  # Create indices for efficient join operations in later steps
  dbExecute(con, "CREATE INDEX idx_spatial_accident ON spatial_matches(accident_id)")
  dbExecute(con, "CREATE INDEX idx_spatial_construction ON spatial_matches(construction_id)")
  dbExecute(con, "ANALYZE spatial_matches")

  dbExecute(con, "COMMIT")
  cat("Created spatial_matches table with combined temporal and spatial filtering\n")
} else {
  cat("Spatial matches table already exists, skipping creation\n")
}

#---------------------------------------------------------------------------
# STEP 2: Count construction sites per accident (unchanged)
#---------------------------------------------------------------------------
has_construction_counts <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'construction_counts')")$exists
if (!has_construction_counts) {
  cat("Step 2: Creating construction_counts table...\n")

  dbExecute(con, "BEGIN")

  dbExecute(con, "
    CREATE TABLE construction_counts AS
    SELECT
      accident_id,
      COUNT(*) AS nearby_construction_count
    FROM 
      spatial_matches
    GROUP BY
      accident_id
  ")

  # Create index for efficient joins in the final step
  dbExecute(con, "CREATE INDEX idx_construction_counts_accident ON construction_counts(accident_id)")
  dbExecute(con, "ANALYZE construction_counts")

  dbExecute(con, "COMMIT")
  cat("Created construction_counts table\n")
} else {
  cat("Construction counts table already exists, skipping creation\n")
}

#---------------------------------------------------------------------------
# STEP 3: Create final analysis table (unchanged)
#---------------------------------------------------------------------------
has_accident_construction_analysis <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'accident_construction_analysis')")$exists
if (!has_accident_construction_analysis) {
  cat("Step 3: Creating final accident_construction_analysis table...\n")

  dbExecute(con, "BEGIN")

  dbExecute(con, "
    CREATE TABLE accident_construction_analysis AS
    WITH ranked_matches AS (
      SELECT
        sm.accident_id,
        sm.construction_id,
        sm.accident_severity,
        sm.accident_time,
        cc.nearby_construction_count,
        -- Rank construction sites for each accident (random selection if multiple)
        ROW_NUMBER() OVER (
          PARTITION BY sm.accident_id
          ORDER BY RANDOM()  -- Random selection ensures unbiased assignment
        ) AS rank
      FROM 
        spatial_matches sm
      JOIN
        construction_counts cc ON sm.accident_id = cc.accident_id
    )
    SELECT 
      accident_id,
      construction_id,
      'LINE' AS construction_type,  -- Indicates we're working with line geometries only
      accident_severity,            -- For severity analysis
      accident_time,                -- For temporal analysis
      1 AS temporal_overlap,        -- Always 1 due to our temporal filter in Step 1
      nearby_construction_count,    -- Number of construction sites affecting this accident
      1 AS is_near_active_construction  -- Always 1 due to our filtering steps
    FROM 
      ranked_matches
    WHERE 
      rank = 1  -- Select only one construction site per accident
  ")

  # Create indices for query optimization
  dbExecute(con, "CREATE INDEX idx_aca_accident_id ON accident_construction_analysis(accident_id)")
  dbExecute(con, "CREATE INDEX idx_aca_construction_id ON accident_construction_analysis(construction_id)")
  dbExecute(con, "CREATE INDEX idx_aca_count ON accident_construction_analysis(nearby_construction_count)")
  dbExecute(con, "ANALYZE accident_construction_analysis")

  dbExecute(con, "COMMIT")
  cat("Created accident_construction_analysis table\n")
} else {
  cat("Accident construction analysis table already exists, skipping creation\n")
}

# Read the data for construction zones with accident counts (for visualization)
accidents_per_zone <- st_read(con, query = "
  WITH construction_accident_counts AS (
    SELECT 
      aca.construction_id,
      COUNT(aca.accident_id) AS accident_count
    FROM 
      accident_construction_analysis aca
    GROUP BY 
      aca.construction_id
  )
  SELECT 
    cl.\"ID\" AS construction_id,
    cac.accident_count,
    cl.geometry
  FROM 
    construction_lines cl
  JOIN 
    construction_accident_counts cac ON cl.\"ID\" = cac.construction_id
  ORDER BY 
    cac.accident_count DESC
")

# Get statistics on construction zones
cat("Construction zones with accidents:", nrow(accidents_per_zone), "\n")
cat("Maximum accidents in a single construction zone:", max(accidents_per_zone$accident_count), "\n")
cat("Average accidents per construction zone:", round(mean(accidents_per_zone$accident_count), 2), "\n")

# Read accidents with their associated construction data
accidents_with_construction <- st_read(con, query = "
  SELECT 
    a.*,
    CASE WHEN aca.accident_id IS NOT NULL THEN 1 ELSE 0 END AS near_construction,
    aca.construction_id,
    aca.nearby_construction_count,
    aca.construction_type
  FROM 
    accidents a
  LEFT JOIN 
    accident_construction_analysis aca ON a.\"ID\" = aca.accident_id
")

# Get distribution of accident counts by construction density
construction_density_stats <- dbGetQuery(con, "
  SELECT 
    nearby_construction_count,
    COUNT(*) AS accident_count
  FROM 
    accident_construction_analysis
  GROUP BY 
    nearby_construction_count
  ORDER BY 
    nearby_construction_count
")

#---------------------------------------------------------------------------
# Report completion and verify table sizes
#---------------------------------------------------------------------------
cat("All tables are ready for analysis\n")

print(construction_density_stats)

# Total accidents summary - for comparison
total_accidents <- dbGetQuery(con, "SELECT COUNT(*) as count FROM accidents")$count
accidents_near_construction <- dbGetQuery(con, "SELECT COUNT(*) as count FROM accident_construction_analysis")$count

cat("Total accidents in dataset:", total_accidents, "\n")
cat("Accidents near active construction:", accidents_near_construction, "\n")
cat("Percentage of accidents near active construction: ", 
    round(accidents_near_construction/total_accidents*100, 2), "%\n")

#---------------------------------------------------------------------------
# STEP 1A-L: Create construction line buffers table with time ranges
#---------------------------------------------------------------------------
has_construction_line_buffers <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'construction_line_buffers')")$exists
if (!has_construction_line_buffers) {
  cat("Step 1A-L: Creating construction_line_buffers table with time ranges...\n")

  dbExecute(con, "BEGIN")

  # Create line-based construction buffers with time information
  dbExecute(con, "
    CREATE TABLE construction_line_buffers AS
    SELECT 
      \"ID\" as construction_id,
      \"Start_Time\" as start_time,
      \"End_Time\" as end_time,
      ST_Transform(
        ST_Buffer(
          ST_Transform(geometry, 3310), -- Transform to California Albers for accurate distance
          300                           -- 300 meters buffer
        ), 
        4326                            -- Transform back to WGS 84 for joining
      ) AS geometry,
      ST_Length(ST_Transform(geometry, 3310)) AS segment_length_m -- Length in meters
    FROM construction_lines
  ")

  # Add primary key and spatial index
  dbExecute(con, "ALTER TABLE construction_line_buffers ADD PRIMARY KEY (construction_id)")
  dbExecute(con, "CREATE INDEX idx_construction_line_buffers_geometry ON construction_line_buffers USING GIST(geometry)")
  dbExecute(con, "CREATE INDEX idx_construction_line_buffers_times ON construction_line_buffers(start_time, end_time)")
  dbExecute(con, "ANALYZE construction_line_buffers")

  dbExecute(con, "COMMIT")
  cat("Created construction_line_buffers table with spatial index and time ranges\n")
} else {
  cat("Construction line buffers table already exists, skipping creation\n")
}

#---------------------------------------------------------------------------
# STEP 1B-L: Create spatial matches table for line objects
#---------------------------------------------------------------------------
has_line_spatial_matches <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'line_spatial_matches')")$exists
if (!has_line_spatial_matches) {
  cat("Step 1B-L: Creating line_spatial_matches table using pre-computed line buffers...\n")

  dbExecute(con, "BEGIN")

  # Simplified join using pre-computed line buffers with time information
  dbExecute(con, "
    CREATE TABLE line_spatial_matches AS
    SELECT
      a.\"ID\" AS accident_id,
      a.\"Severity\" AS accident_severity,
      a.\"Start_Time\" AS accident_time,
      b.construction_id,
      b.segment_length_m
    FROM 
      accidents a
    JOIN 
      construction_line_buffers b ON 
        ST_Within(a.geometry, b.geometry) AND
        a.\"Start_Time\" BETWEEN b.start_time AND b.end_time
  ")

  # Create indices for efficient join operations
  dbExecute(con, "CREATE INDEX idx_line_spatial_accident ON line_spatial_matches(accident_id)")
  dbExecute(con, "CREATE INDEX idx_line_spatial_construction ON line_spatial_matches(construction_id)")
  dbExecute(con, "ANALYZE line_spatial_matches")

  dbExecute(con, "COMMIT")
  cat("Created line_spatial_matches table with combined temporal and spatial filtering\n")
} else {
  cat("Line spatial matches table already exists, skipping creation\n")
}

#---------------------------------------------------------------------------
# STEP 2-L: Count construction line sites per accident
#---------------------------------------------------------------------------
has_line_construction_counts <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'line_construction_counts')")$exists
if (!has_line_construction_counts) {
  cat("Step 2-L: Creating line_construction_counts table...\n")

  dbExecute(con, "BEGIN")

  dbExecute(con, "
    CREATE TABLE line_construction_counts AS
    SELECT
      accident_id,
      COUNT(*) AS nearby_line_construction_count
    FROM 
      line_spatial_matches
    GROUP BY
      accident_id
  ")

  # Create index for efficient joins in the final step
  dbExecute(con, "CREATE INDEX idx_line_construction_counts_accident ON line_construction_counts(accident_id)")
  dbExecute(con, "ANALYZE line_construction_counts")

  dbExecute(con, "COMMIT")
  cat("Created line_construction_counts table\n")
} else {
  cat("Line construction counts table already exists, skipping creation\n")
}

#---------------------------------------------------------------------------
# STEP 3-L: Create final line-based analysis table
#---------------------------------------------------------------------------
has_line_accident_construction_analysis <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'line_accident_construction_analysis')")$exists
if (!has_line_accident_construction_analysis) {
  cat("Step 3-L: Creating final line_accident_construction_analysis table...\n")

  dbExecute(con, "BEGIN")

  dbExecute(con, "
    CREATE TABLE line_accident_construction_analysis AS
    WITH ranked_matches AS (
      SELECT
        sm.accident_id,
        sm.construction_id,
        sm.accident_severity,
        sm.accident_time,
        sm.segment_length_m,
        cc.nearby_line_construction_count,
        -- Rank construction sites for each accident (random selection if multiple)
        ROW_NUMBER() OVER (
          PARTITION BY sm.accident_id
          ORDER BY RANDOM()  -- Random selection ensures unbiased assignment
        ) AS rank
      FROM 
        line_spatial_matches sm
      JOIN
        line_construction_counts cc ON sm.accident_id = cc.accident_id
    )
    SELECT 
      accident_id,
      construction_id,
      accident_severity,
      accident_time,
      segment_length_m,
      nearby_line_construction_count,
      1 AS temporal_overlap,           -- Always 1 due to our temporal filter in Step 1
      1 AS is_near_active_construction -- Always 1 due to our filtering steps
    FROM 
      ranked_matches
    WHERE 
      rank = 1  -- Select only one construction site per accident
  ")

  # Create indices for query optimization
  dbExecute(con, "CREATE INDEX idx_laca_accident_id ON line_accident_construction_analysis(accident_id)")
  dbExecute(con, "CREATE INDEX idx_laca_construction_id ON line_accident_construction_analysis(construction_id)")
  dbExecute(con, "CREATE INDEX idx_laca_count ON line_accident_construction_analysis(nearby_line_construction_count)")
  dbExecute(con, "ANALYZE line_accident_construction_analysis")

  dbExecute(con, "COMMIT")
  cat("Created line_accident_construction_analysis table\n")
} else {
  cat("Line accident construction analysis table already exists, skipping creation\n")
}

#---------------------------------------------------------------------------
# STEP 4-L: Calculate accident rates by segment length
#---------------------------------------------------------------------------
has_accident_rates <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'construction_segment_accident_rates')")$exists
if (!has_accident_rates) {
  cat("Step 4-L: Creating construction_segment_accident_rates table...\n")

  dbExecute(con, "BEGIN")

  dbExecute(con, "
    CREATE TABLE construction_segment_accident_rates AS
    SELECT 
      cl.\"ID\" AS construction_id,
      cl.geometry,
      COUNT(laca.accident_id) AS accident_count,
      MAX(laca.segment_length_m) AS segment_length_m,
      COUNT(laca.accident_id) / (MAX(laca.segment_length_m) / 1000.0) AS accidents_per_km
    FROM 
      construction_lines cl
    LEFT JOIN 
      line_accident_construction_analysis laca ON cl.\"ID\" = laca.construction_id
    GROUP BY 
      cl.\"ID\", cl.geometry
  ")

  # Create spatial index
  dbExecute(con, "CREATE INDEX idx_accident_rates_geometry ON construction_segment_accident_rates USING GIST(geometry)")
  dbExecute(con, "ANALYZE construction_segment_accident_rates")

  dbExecute(con, "COMMIT")
  cat("Created construction_segment_accident_rates table\n")
} else {
  cat("Construction segment accident rates table already exists, skipping creation\n")
}

# Read accident rates by construction segment
accident_rates_by_segment <- st_read(con, query = "
  SELECT 
    construction_id,
    accident_count,
    segment_length_m,
    accidents_per_km,
    geometry
  FROM 
    construction_segment_accident_rates
  WHERE 
    segment_length_m > 0  -- Avoid division by zero
  ORDER BY 
    accidents_per_km DESC
")

# Display summary statistics
cat("Construction segments analyzed:", nrow(accident_rates_by_segment), "\n")
cat("Maximum accidents per km:", max(accident_rates_by_segment$accidents_per_km, na.rm = TRUE), "\n")
cat("Average accidents per km:", mean(accident_rates_by_segment$accidents_per_km, na.rm = TRUE), "\n")
cat("Median segment length (m):", median(accident_rates_by_segment$segment_length_m, na.rm = TRUE), "\n")

# Severity analysis near line construction
severity_by_construction <- dbGetQuery(con, "
  SELECT 
    laca.accident_severity, 
    COUNT(*) AS count
  FROM 
    line_accident_construction_analysis laca
  GROUP BY 
    laca.accident_severity
  ORDER BY 
    laca.accident_severity
")

# Comparison of accident severity near vs. away from construction
severity_comparison <- dbGetQuery(con, "
  WITH construction_accidents AS (
    SELECT 
      a.\"ID\", 
      a.\"Severity\", 
      'Near construction' AS location
    FROM 
      accidents a
    JOIN 
      line_accident_construction_analysis laca ON a.\"ID\" = laca.accident_id
  ),
  non_construction_accidents AS (
    SELECT 
      a.\"ID\", 
      a.\"Severity\", 
      'Away from construction' AS location
    FROM 
      accidents a
    LEFT JOIN 
      line_accident_construction_analysis laca ON a.\"ID\" = laca.accident_id
    WHERE 
      laca.accident_id IS NULL
  )
  SELECT 
    \"Severity\", 
    location, 
    COUNT(*) AS count
  FROM (
    SELECT * FROM construction_accidents
    UNION ALL
    SELECT * FROM non_construction_accidents
  ) combined
  GROUP BY 
    \"Severity\", location
  ORDER BY 
    \"Severity\", location
")

# Time-based analysis
temporal_analysis <- dbGetQuery(con, "
  SELECT 
    DATE_TRUNC('month', laca.accident_time) AS month,
    COUNT(*) AS accident_count
  FROM 
    line_accident_construction_analysis laca
  GROUP BY 
    DATE_TRUNC('month', laca.accident_time)
  ORDER BY 
    month
")

# Disconnect from database
dbDisconnect(con)
```

## Buffer zone analysis

In theory, the commented code below would achieve the same as our operation PostGIS but way slower. We will not run it here, but keep it to demonstrate that we could implement in R as well.

```{r Buffer Zone Analysis}
# Set up parallel processing
#future::plan(future::multisession, workers = parallel::detectCores() - 1)

# Set sf to use parallel processing
#sf_use_s2(FALSE)  # Disable S2 spherical geometry
#cores <- parallel::detectCores() - 1

# Process each dataset with batching
#cat("Creating point buffers...\n")
#const_buffers <- create_buffers_in_batches(df.const.sf, buffer_dist = 1000, batch_size = 500)

# Compute accidents within construction zones
#accidents_in_construction <- df.acc.sf %>%
#  st_join(const_buffers, join = st_within, prepared = TRUE)

# Return to sequential processing
#future::plan(future::sequential)
#sf_use_s2(TRUE)  # Re-enable S2 spherical geometry

# Count accidents per construction zone
#accidents_per_zone <- accidents_with_construction %>%
#  group_by(ID) %>%  # Grouping by the ID column
#  summarize(accident_count = n())

accidents_per_zone <- accidents_per_zone %>% drop_na()

# Visualize construction zones with accident counts
# First, store our plot in a variable
accident_zones_plot <- ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  # Arrange data in descending order so high frequency zones are plotted on top
  geom_sf(data = accidents_per_zone %>% arrange(accident_count), 
          aes(color = accident_count), size = 1.5) +
  scale_color_viridis_c() +
  theme_minimal() +
  labs(title = "Accident Counts within Construction Zones (2021)")

print(accident_zones_plot)

# Then save it using ggsave
ggsave(filename = "imgs/accident_counts_construction_zones.png", 
       plot = accident_zones_plot,
       width = 10, # width in inches
       height = 8, # height in inches
       dpi = 300)  # resolution
```

## Advanced construction sit analysis (line objects):

```{r Maps with Construction Lines - static}
# Static map with construction line segments
ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  geom_sf(data = df.const.lines %>% filter(State == "CA"), 
          color = "orange", size = 1) +
  theme_minimal() +
  labs(title = "US Construction Segments (2021)")
```

```{r Maps with Construction Lines - dynamic}
# Interactive map with construction segments focused on California
construction_segment_map <- leaflet() %>%
  addTiles() %>%
  addPolylines(data = df.const.lines, 
               color = "orange",
               weight = 3,
               opacity = 0.7,
               popup = ~paste("ID:", ID, "<br>Duration:", duration)) %>%
  setView(lng = -119.4179, lat = 36.7783, zoom = 6) %>%  # Center on California
  setMaxBounds(lng1 = -124.6, lat1 = 42.0,    # Northwest corner of CA
               lng2 = -114.1, lat2 = 32.5)    # Southeast corner of CA

# Display the map
construction_segment_map
```

Again, we execute computationally heavy operations in PostGIS but have the commented R code below to showcase a theoretical, but slow implementatiojn in R.

```{r Construction Impact Analysis with Buffers}
# Set up parallel processing
#future::plan(future::multisession, workers = parallel::detectCores() - 1)

# Set sf to use parallel processing
#f_use_s2(FALSE)  # Disable S2 spherical geometry
#cores <- parallel::detectCores() - 1

#cat("Creating line segment buffers...\n")
#const_segment_buffers <- create_buffers_in_batches(df.const.lines, buffer_dist = 1000, batch_size = 250)

# Find accidents near construction segments
#accidents_near_construction <- df.acc.sf %>%
#  st_join(const_segment_buffers, join = st_intersects, prepared = TRUE)

# Return to sequential processing
#future::plan(future::sequential)
#sf_use_s2(TRUE)  # Re-enable S2 spherical geometry

# Analyze accident rate per construction segment length
#accidents_per_segment <- accidents_near_construction %>%
#  group_by(ID) %>%  # Assuming there's an ID column for construction segments
#  summarize(
#    accident_count = n(),
#    segment_length = as.numeric(st_length(geometry[1])),  # Length in meters
#    accidents_per_km = accident_count / (segment_length/1000)
#  )

# Visualize segments by accident rate
ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  geom_sf(data = accident_rates_by_segment, aes(color = accidents_per_km), size = 1.5) +
  scale_color_viridis_c() +
  theme_minimal() +
  labs(title = "Accident Rate per Construction Segment (2016-2021)",
       color = "Accidents/km")
```

```{r}
# Reestablish connection to PostgreSQL database
con <- dbConnect(
  Postgres(),
  dbname = "geospatial_final",
  host = "localhost",
  port = 5432,
  user = "postgres",
  password = Sys.getenv("POSTGRES_PWD")
)

# Temporal analysis - construction impact over time
accidents_by_time <- dbGetQuery(con, "
  SELECT 
    DATE_TRUNC('month', accident_time) AS month,
    COUNT(*) AS accident_count
  FROM 
    line_accident_construction_analysis
  GROUP BY 
    DATE_TRUNC('month', accident_time)
  ORDER BY 
    month
") %>%
  mutate(month = as.Date(month))  # Convert PostgreSQL timestamp to R Date

# Get severity comparison data
severity_comparison <- dbGetQuery(con, "
  WITH construction_accidents AS (
    SELECT 
      a.\"Severity\", 
      'Near construction' AS location
    FROM 
      accidents a
    JOIN 
      line_accident_construction_analysis laca ON a.\"ID\" = laca.accident_id
  ),
  non_construction_accidents AS (
    SELECT 
      a.\"Severity\", 
      'Away from construction' AS location
    FROM 
      accidents a
    LEFT JOIN 
      line_accident_construction_analysis laca ON a.\"ID\" = laca.accident_id
    WHERE 
      laca.accident_id IS NULL
  )
  SELECT 
    \"Severity\", 
    location, 
    COUNT(*) AS count
  FROM (
    SELECT * FROM construction_accidents
    UNION ALL
    SELECT * FROM non_construction_accidents
  ) combined
  GROUP BY 
    \"Severity\", location
  ORDER BY 
    \"Severity\", location
")

# Disconnect from database
dbDisconnect(con)
```

```{r Temporal Analysis and Severity Comparison with Buffers}
# Visualize temporal analysis
ggplot(accidents_by_time, aes(x = month, y = accident_count)) +
  geom_line() +
  geom_point() +
  scale_x_date(date_labels = "%b %Y", date_breaks = "1 month") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Accidents Near Construction Over Time",
       x = "Month",
       y = "Accident Count")

# Visualize severity comparison
ggplot(severity_comparison, aes(x = Severity, y = count, fill = location)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal() +
  labs(title = "Accident Severity: Near vs. Away from Construction",
       x = "Severity Level",
       y = "Count",
       fill = "Location")
```

### 2025-03-25 final/finalSIMON_MoritzPeistNoemiLucchiSimonVellin .Rmd
Updated Final Notebook

@@
 - Simon Vellin
-subtitle: Analysis of US road accidents and construction sites
+subtitle: Analysis of US Road Accidents and Construction Sites (TBC)
 output:
@@
 
+```{r setup, include=FALSE}
+# Disable warning messages
+knitr::opts_chunk$set(warning = FALSE, message = FALSE)
+```
+
 <div style="text-align: center;">
@@
 
-Load packages
+# Purpose of Study
+
+This study investigates the causal effect of road construction and closures on accident rates in the United States by employing a propensity score matching framework to compute a difference-in-differences regression.
+Using two distinct datasets — US Accidents (2016–2023) and US Road Construction and Closures (2016–2021) — this paper aims to rigorously assess how construction-related disruptions influence the frequency and severity of traffic accidents.
+
+Although a substantial body of literature has examined the latter, the specific integration of causal inference techniques remains less prevalent.
+As such, our approach provides a robust methodology by matching treated and control segments based on relevant covariates, which isolates the studied effect by comparing changes in accident rates over time between matched groups.
+
+After quickly reviewing existing literature, we will justify the use of propensity score matching and difference-in-differences before implementing our code in Part 3.
+Finally, the last section will display our results while mentioning the limits of our study, as well as the potential improvements for subsequent analysis.
+
+# Part 1 - Literature Review
+
+In order to familiarize ourselves with the topic, we have been through two existing papers relating to accidents in the US, serving as a baseline understanding to start our study.
+Although we would have preferred introducing one of Mohamed Abdel-Aty's studies, widely recognized for his contributions to transportation safety, he employs advanced statistical modeling techniques which appears less related to Geospatial & Economic litterature compared to the ones that we decided to present below.
+
+## 2.1 "Effects of truck traffic on crash injury severity on rural highways in Wyoming using Bayesian binary logit models" by Ahmed et al. (2018)
+
+This paper explores the impact of truck traffic and roadway geometry on crash frequency within highway work zones, using a case study from Wyoming spanning 2012–2016.
+
+Using crash data and statistical modeling, the authors assess how construction-related factors, like lane closures and heavy vehicle presence, elevate accident risks.
+Their paper reveals that increased truck volumes and sharper geometric features significantly amplify crash rates, with work zones acting as hotspots for such incidents.
+While the paper employs logistic regression rather than causal inference techniques, it underscores the real-world relevance of construction disruptions on traffic safety.
+
+This work provides a practical foundation for our study, highlighting the need to isolate causal effects of construction using advanced methods.
+By focusing on observable crash patterns, Ahmed et al.
+set the stage for deeper methodological exploration in our difference-in-differences framework.
+
+URL - https://www.sciencedirect.com/science/article/pii/S0001457518301507
+
+## 2.2 "Comparison of characteristics between fatal and injury accidents in the highway construction zones" by Li and Bai (2008)
+
+This paper investigates the characteristics of accidents in highway construction zones from a statistical perspective, comparing fatal and injury crashes using data from Utah between 1992 and 2004.
+
+Using traffic volume, construction duration, and time of day, their results show that construction zones markedly increase accident likelihood, with fatal crashes tied to higher speeds and nighttime conditions.
+Although the study relies on descriptive statistics and regression rather than causal frameworks, it offers critical insights into how construction alters road safety dynamics.
+
+It supports our research by establishing a clear link between construction activities and accident rates, justifying our use of propensity score matching and difference-in-differences.
+Ideally, our approach aims to address the methodological gaps they leave unexplored.
+
+URL - https://www.sciencedirect.com/science/article/pii/S0925753507001014
+
+# Part 2 - Methodology
+
+In this section, we start by providing summary statistics of the datasets we have found for our study, while detailing the theoretical justifications underlying the two main tools used to compute our results.
+
+## 2.1 Dataset Overview
+
+For both datasets, the data was collected using multiple APIs such as the US and state departments of transportation, law enforcement agencies, traffic cameras, and traffic sensors within the road networks.
+
+### a. US Accidents (2016 - 2023)
+
+This is a countrywide car accident dataset that covers 49 states of the USA. The dataset currently contains approximately 7.7 million accident records.
+
+```{Descriptive Statistics of Data 1}
+
+```
+
+### b. US Road Construction and Closures (2016 - 2021)
+
+This is a countrywide dataset of road construction and closure events, which covers 49 states of the US, ranging from fixing pavements to substantial projects that could take months to finish.
+Currently, there are about 6.2 million construction and closure records in this dataset.
+
+```{Descriptive Statistics of Data 2}
+
+```
+
+## 2.2 Propensity Score Matching
+
+Propensity Score Matching (PSM) is a statistical technique designed to estimate causal effects by balancing treatment and control groups based on observed covariates, tackling selection bias inherent to observational data like ours.
+
+Here, “treatment” denotes road segments with construction or closures, while “control” covers unaffected segments.
+PSM hinges on the Conditional Independence Assumption, which stipulates that given a set of covariates $X$ (e.g., traffic volume, road type, weather), treatment assignment is independent of the outcome—accident rates.
+This is expressed as:
+
+$$Y(0), Y(1) \perp T \mid X,$$
+
+where $Y(1)$ and $Y(0)$ represent potential outcomes under treatment and control, and $T$ is the treatment indicator.
+PSM estimates the propensity score, the probability of treatment given covariates, as:
+
+$$PSM(X) = P(T = 1 \mid X),$$
+
+typically via logistic regression. We then match treated and control units with similar $PSM(X)$ values, ensuring comparability. The average treatment effect on the treated (ATT) is:
+
+$$ATT = E[Y(1) - Y(0) \mid T = 1] = E[Y(1) \mid T = 1] - E[Y(0) \mid T = 1],$$
+
+where $E[Y(0) \mid T = 1]$ is inferred from matched controls.
+For our study, PSM balances segments on pre-construction traits, 'mimicking' randomization to isolate construction’s effect.
+This is vital since construction isn’t randomly assigned (i.e.
+busier roads might be more prone to work).
+We’ll use nearest-neighbor matching and assess balance with standardized differences.
+PSM excels at handling multiple covariates, fitting our US Road Construction and Closures dataset, but requires strong assumption (CIA) and sufficient overlap in $PSM(X)$ (common support).
+
+## 2.3 Difference-in-Differences
+
+Difference-in-Differences (DiD) is a quasi-experimental approach that estimates causal effects by comparing outcome changes over time between treated (roads with construction) and control (roads without) groups.
+It requires the parallel trends assumption: without treatment, accident rate differences between groups would stay constant over time.
+The model is:
+
+$$Y_{it} = \beta_0 + \beta_1 T_i + \beta_2 Post_t + \beta_3 (T_i \times Post_t) + \epsilon_{it},$$
+
+where $Y_{it}$ is the accident rate for segment $i$ at time $t$, $T_i$ is the treatment dummy (1 for construction, 0 otherwise), $Post_t$ is the time dummy (1 post-construction, 0 pre), and $\beta_3$ is the DiD estimator.
+This is calculated as:
+
+$$\beta_3 = (E[Y_{T=1,Post}] - E[Y_{T=1,Pre}]) - (E[Y_{T=0,Post}] - E[Y_{T=0,Pre}]),$$
+
+capturing the treatment effect by differencing pre-post changes across groups.
+
+In our context, DiD compares accident rates before and after construction starts, between treated and control segments, controlling for time-invariant confounders and shared trends.
+We’ll verify parallel trends by plotting pre-treatment accident rates and testing statistically.
+Challenges include potential trend violations or spillovers (i.e.
+traffic rerouting to control roads), but DiD’s strength lies in exploiting time variation, making it ideal for tracing construction’s impact on accident trajectories.
+
+# Part 3 - Implementation
+
+## 3.0 Load Packages
+
+Start by loading the packages required to run this section.
 
@@
 
-# Set-up convenience functions for parallel processing (as data is quite large)
-```{r Parallel Buffering}
-# Function to create buffers in batches with proper projection
-create_buffers_in_batches <- function(sf_object, buffer_dist, batch_size = 500) {
-  # First, reproject to an appropriate projected CRS for the region
-  # For California, UTM Zone 10N (EPSG:26910) or 11N (EPSG:26911) works well
-  # If we're unsure about your region, a web mercator projection (3857) is a reasonable default
-  message("Reprojecting data to a meter-based CRS...")
-  sf_object_projected <- st_transform(sf_object, 3310)  # Web Mercator
+## 3.1 Descriptive Maps
 
-  n_features <- nrow(sf_object_projected)
-  n_batches <- ceiling(n_features / batch_size)
-
-  # Create empty list to store results
-  buffer_list <- vector("list", n_batches)
-
-  # Process in batches
-  for (i in 1:n_batches) {
-    start_idx <- (i-1) * batch_size + 1
-    end_idx <- min(i * batch_size, n_features)
-
-    cat(sprintf("Processing batch %d of %d (features %d to %d)\n", 
-                i, n_batches, start_idx, end_idx))
-
-    # Extract batch
-    batch <- sf_object_projected[start_idx:end_idx, ]
-
-    # Create buffer (with parallel processing within sf)
-    buffer_list[[i]] <- st_buffer(batch, dist = buffer_dist)
-  }
-
-  # Combine results
-  result <- do.call(rbind, buffer_list)
-
-  # Reproject back to original CRS if needed
-  message("Reprojecting results back to original CRS...")
-  st_transform(result, st_crs(sf_object))
-}
-```
-
-Load data (filter for California for now to limit data set size)
-
-```{r Data Load Accidents}
-# Efficient approach
-df.acc <- fread("data/us_accidents/US_accidents_March23.csv")[
-  # Filter date range of 2021
-  lubridate::year(as.Date(Start_Time)) == 2021 & 
-  # And California
-  State == "CA"
-][, `:=`(
-  # Add year, quarter, month columns
-  year = data.table::year(Start_Time),
-  quarter = data.table::quarter(Start_Time),
-  month = data.table::month(Start_Time),
-  # Calculate duration (assuming End_Time exists in the dataset)
-  duration = as.numeric(difftime(End_Time, Start_Time, units = "days"))
-)] %>% 
-  as_tibble()  # Convert to tibble only at the end for performance
-```
-
-```{r Data Load Constructions}
-df.const <- fread("data/us_constructions/US_constructions_Dec21.csv")[
-  # Filter date range of 2021
-  lubridate::year(as.Date(Start_Time)) == 2021 & 
-  # And California
-  State == "CA"
-][, `:=`(
-  # Add year, quarter, month columns
-  year = year(Start_Time),
-  quarter = quarter(Start_Time),
-  month = month(Start_Time),
-  # Calculate duration (assuming End_Time exists in the dataset)
-  duration = as.numeric(difftime(End_Time, Start_Time, units = "days"))
-)] %>% 
-  as_tibble()  # Convert to tibble only at the end for performance
-```
-
-Convert dataframes to sf objects
-
-```{r Geospatial Data}
-# Convert accident data to sf object
-df.acc.sf <- df.acc %>%
-  filter(!is.na(Start_Lat) & !is.na(Start_Lng)) %>%
-  st_as_sf(coords = c("Start_Lng", "Start_Lat"), crs = 4326)
-
-# Convert construction data to sf object
-df.const.sf <- df.const %>%
-  filter(!is.na(Start_Lat) & !is.na(Start_Lng)) %>%
-  st_as_sf(coords = c("Start_Lng", "Start_Lat"), crs = 4326)
-```
-
-```{r Data Load Roads}
-# Get US roads (this can be slow for the entire US, so we might want to limit by state)
-# We also only get Primary and Secondary roads as they are the most important (i.e., where most constructions and accidents happen). Otherwise, the dataset would become to large (1.1GB)
-
-ca_roads <- primary_secondary_roads("CA", year = 2021)
-
-# Below code to get ALL roads for California (not recommended due to size)
-
-# First get all counties in California
-#ca_counties <- counties("CA", year = 2021)
-
-# Then get roads for each county
-#ca_roads_list <- lapply(ca_counties$COUNTYFP, function(county_code) {
-#  roads("CA", county = county_code, year = 2021)
-#})
-
-# Optionally combine all county road data into one object
-#ca_roads <- do.call(rbind, ca_roads_list)
-```
+### a. Static Maps
 
@@
 ```
+
+### b. Heatmaps
+
+#### (i) Static
 
@@
 
-## Heatmaps - interactive
+#### (ii) Interactive
 
@@
 
-We construct construction sites now also as line objects since they have a start and end point
+<hr>
 
-```{r Construction linestrings}
-# Convert construction data to linestring sf object
-df.const.lines <- df.const %>%
-  filter(!is.na(Start_Lat) & !is.na(Start_Lng) & !is.na(End_Lat) & !is.na(End_Lng)) %>%
-  mutate(geometry = pmap(list(Start_Lng, Start_Lat, End_Lng, End_Lat), 
-                         function(start_lng, start_lat, end_lng, end_lat) {
-                           coords <- matrix(c(start_lng, start_lat, 
-                                             end_lng, end_lat), 
-                                           ncol = 2, byrow = TRUE)
-                           st_linestring(coords)
-                         })) %>%
-  st_sf(crs = 4326)
-```
+## 3.2 Propensity Score Matching
 
-## Spatial operations in PostgreSQL
+## 3.3 Difference-in-Differences
 
-Since file sizes are so large and Spatial operations in R incredibly slow, we retreat to a PostgreSQL database for the following operations (PostGIS also uses the SF framework, thus R are almost identical to the following SQL functions used).
+# Part 4 - Discussion
 
-```{r DB ops}
-# Create a connection to your PostgreSQL database
-con <- dbConnect(
-  Postgres(),
-  dbname = "geospatial_final",
-  host = "localhost",
-  port = 5432,
-  # Uncomment and add your credentials if needed
-  user = "postgres",
-  password = Sys.getenv("POSTGRES_PWD")
-)
+## 4.1 Results
 
-# Set connection parameters for performance optimization
-dbExecute(con, "SET work_mem = '128MB'")           # Increase working memory
-dbExecute(con, "SET maintenance_work_mem = '256MB'") # For index creation
-dbExecute(con, "SET random_page_cost = 1.1")       # For SSD storage
-dbExecute(con, "SET effective_cache_size = '2GB'") # Estimate of memory available
-dbExecute(con, "SET max_parallel_workers_per_gather = 4") # Use multiple cores
+## 4.2 Limitations
 
-# Check if PostGIS extension is enabled
-postgis_check <- dbGetQuery(con, "SELECT PostGIS_version()")
-cat("PostGIS version:", postgis_check[[1]], "\n")
-
-# Check if SRID 3310 is available and add if not
-dbExecute(con, "
-  DO $$
-  BEGIN
-    IF NOT EXISTS (SELECT 1 FROM spatial_ref_sys WHERE srid = 3310) THEN
-      INSERT INTO spatial_ref_sys (srid, auth_name, auth_srid, proj4text, srtext)
-      VALUES (3310, 'EPSG', 3310, 
-      '+proj=aea +lat_1=34 +lat_2=40.5 +lat_0=0 +lon_0=-120 +x_0=0 +y_0=-4000000 +datum=NAD83 +units=m +no_defs',
-      'PROJCS[\"NAD83 / California Albers\",GEOGCS[\"NAD83\",DATUM[\"North_American_Datum_1983\",SPHEROID[\"GRS 1980\",6378137,298.257222101,AUTHORITY[\"EPSG\",\"7019\"]],AUTHORITY[\"EPSG\",\"6269\"]],PRIMEM[\"Greenwich\",0,AUTHORITY[\"EPSG\",\"8901\"]],UNIT[\"degree\",0.01745329251994328,AUTHORITY[\"EPSG\",\"9122\"]],AUTHORITY[\"EPSG\",\"4269\"]],PROJECTION[\"Albers_Conic_Equal_Area\"],PARAMETER[\"standard_parallel_1\",34],PARAMETER[\"standard_parallel_2\",40.5],PARAMETER[\"latitude_of_center\",0],PARAMETER[\"longitude_of_center\",-120],PARAMETER[\"false_easting\",0],PARAMETER[\"false_northing\",-4000000],UNIT[\"metre\",1,AUTHORITY[\"EPSG\",\"9001\"]],AUTHORITY[\"EPSG\",\"3310\"]]');
-    END IF;
-  END
-  $$;
-")
-
-# Check if tables exist and only create them if they don't
-# Begin transaction for data processing
-dbExecute(con, "BEGIN")
-
-# Check if accidents table exists, if not create it
-has_accidents <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'accidents')")$exists
-if (!has_accidents) {
-  # Write accidents data to database
-  sf::st_write(df.acc.sf, con, "accidents", 
-               layer_options = c("GEOM_TYPE=GEOMETRY", 
-                                 "GEOMETRY_NAME=geometry",
-                                 "SRID=4326"))
-  cat("Written accidents data to database\n")
-
-  # Create spatial index for accidents
-  dbExecute(con, "CREATE INDEX idx_accidents_geometry ON accidents USING GIST(geometry)")
-  dbExecute(con, 'CREATE INDEX idx_accidents_start_time ON accidents("Start_Time");')
-  dbExecute(con, "ANALYZE accidents")
-  cat("Created spatial index for accidents table\n")
-}
-
-# Check if construction table exists, if not create it
-has_construction <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'construction')")$exists
-
-if (!has_construction) {
-  # Write construction data to database
-  sf::st_write(df.const.sf, con, "construction", 
-               layer_options = c("GEOM_TYPE=GEOMETRY", 
-                                 "GEOMETRY_NAME=geometry",
-                                 "SRID=4326"))
-  cat("Written construction data to database\n")
-
-  # Create spatial index for construction
-  dbExecute(con, "CREATE INDEX idx_construction_geometry ON construction USING GIST(geometry)")
-  dbExecute(con, 'CREATE INDEX idx_construction_start_time ON construction("Start_Time");')
-  dbExecute(con, 'CREATE INDEX idx_construction_end_time ON construction("End_Time");')
-  dbExecute(con, "ANALYZE construction")
-  cat("Created spatial index for construction table\n")
-}
-
-# Check if construction table with lines exists, if not create it
-has_construction_lines <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'construction_lines')")$exists
-
-if (!has_construction_lines) {
-  # Write construction data to database
-  sf::st_write(df.const.lines, con, "construction_lines", 
-               layer_options = c("GEOM_TYPE=GEOMETRY", 
-                                 "GEOMETRY_NAME=geometry",
-                                 "SRID=4326"))
-  cat("Written construction_lines data to database\n")
-
-  # Create spatial index for construction
-  dbExecute(con, "CREATE INDEX idx_construction_lines_geometry ON construction_lines USING GIST(geometry)")
-  dbExecute(con, 'CREATE INDEX idx_construction_lines_start_time ON construction_lines("Start_Time");')
-  dbExecute(con, 'CREATE INDEX idx_construction_lines_end_time ON construction_lines("End_Time");')
-  dbExecute(con, "ANALYZE construction_lines")
-  cat("Created spatial index for construction_lines table\n")
-}
-
-# Commit transaction for base tables
-dbExecute(con, "COMMIT")
-
-#---------------------------------------------------------------------------
-# STEP 1A: Create enhanced construction buffers table with time ranges
-#---------------------------------------------------------------------------
-has_construction_buffers <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'construction_buffers')")$exists
-if (!has_construction_buffers) {
-  cat("Step 1A: Creating enhanced construction_buffers table with time ranges...\n")
-
-  dbExecute(con, "BEGIN")
-
-  # Create construction buffers table with appropriate projection and time ranges
-  dbExecute(con, "
-    CREATE TABLE construction_buffers AS
-    SELECT 
-      \"ID\" as construction_id,
-      \"Start_Time\" as start_time,
-      \"End_Time\" as end_time,
-      ST_Transform(
-        ST_Buffer(
-          ST_Transform(geometry, 3310), -- Transform to California Albers for accurate distance
-          300                           -- 300 meters buffer
-        ), 
-        4326                            -- Transform back to WGS 84 for joining
-      ) AS geometry
-    FROM construction
-  ")
-
-  # Add primary key and spatial index
-  dbExecute(con, "ALTER TABLE construction_buffers ADD PRIMARY KEY (construction_id)")
-  dbExecute(con, "CREATE INDEX idx_construction_buffers_geometry ON construction_buffers USING GIST(geometry)")
-  dbExecute(con, "CREATE INDEX idx_construction_buffers_times ON construction_buffers(start_time, end_time)")
-  dbExecute(con, "ANALYZE construction_buffers")
-
-  dbExecute(con, "COMMIT")
-  cat("Created enhanced construction_buffers table with spatial index and time ranges\n")
-} else {
-  cat("Construction buffers table already exists, skipping creation\n")
-}
-
-#---------------------------------------------------------------------------
-# STEP 1B: Create spatial matches table (using enhanced pre-computed buffers)
-#---------------------------------------------------------------------------
-has_spatial_matches <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'spatial_matches')")$exists
-if (!has_spatial_matches) {
-  cat("Step 1B: Creating spatial_matches table using enhanced pre-computed buffers...\n")
-
-  dbExecute(con, "BEGIN")
-
-  # Simplified join using pre-computed buffers with time information
-  dbExecute(con, "
-    CREATE TABLE spatial_matches AS
-    SELECT
-      a.\"ID\" AS accident_id,
-      a.\"Severity\" AS accident_severity,
-      a.\"Start_Time\" AS accident_time,
-      b.construction_id
-    FROM 
-      accidents a
-    JOIN 
-      construction_buffers b ON 
-        ST_Within(a.geometry, b.geometry) AND
-        a.\"Start_Time\" BETWEEN b.start_time AND b.end_time
-  ")
-
-  # Create indices for efficient join operations in later steps
-  dbExecute(con, "CREATE INDEX idx_spatial_accident ON spatial_matches(accident_id)")
-  dbExecute(con, "CREATE INDEX idx_spatial_construction ON spatial_matches(construction_id)")
-  dbExecute(con, "ANALYZE spatial_matches")
-
-  dbExecute(con, "COMMIT")
-  cat("Created spatial_matches table with combined temporal and spatial filtering\n")
-} else {
-  cat("Spatial matches table already exists, skipping creation\n")
-}
-
-#---------------------------------------------------------------------------
-# STEP 2: Count construction sites per accident (unchanged)
-#---------------------------------------------------------------------------
-has_construction_counts <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'construction_counts')")$exists
-if (!has_construction_counts) {
-  cat("Step 2: Creating construction_counts table...\n")
-
-  dbExecute(con, "BEGIN")
-
-  dbExecute(con, "
-    CREATE TABLE construction_counts AS
-    SELECT
-      accident_id,
-      COUNT(*) AS nearby_construction_count
-    FROM 
-      spatial_matches
-    GROUP BY
-      accident_id
-  ")
-
-  # Create index for efficient joins in the final step
-  dbExecute(con, "CREATE INDEX idx_construction_counts_accident ON construction_counts(accident_id)")
-  dbExecute(con, "ANALYZE construction_counts")
-
-  dbExecute(con, "COMMIT")
-  cat("Created construction_counts table\n")
-} else {
-  cat("Construction counts table already exists, skipping creation\n")
-}
-
-#---------------------------------------------------------------------------
-# STEP 3: Create final analysis table (unchanged)
-#---------------------------------------------------------------------------
-has_accident_construction_analysis <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'accident_construction_analysis')")$exists
-if (!has_accident_construction_analysis) {
-  cat("Step 3: Creating final accident_construction_analysis table...\n")
-
-  dbExecute(con, "BEGIN")
-
-  dbExecute(con, "
-    CREATE TABLE accident_construction_analysis AS
-    WITH ranked_matches AS (
-      SELECT
-        sm.accident_id,
-        sm.construction_id,
-        sm.accident_severity,
-        sm.accident_time,
-        cc.nearby_construction_count,
-        -- Rank construction sites for each accident (random selection if multiple)
-        ROW_NUMBER() OVER (
-          PARTITION BY sm.accident_id
-          ORDER BY RANDOM()  -- Random selection ensures unbiased assignment
-        ) AS rank
-      FROM 
-        spatial_matches sm
-      JOIN
-        construction_counts cc ON sm.accident_id = cc.accident_id
-    )
-    SELECT 
-      accident_id,
-      construction_id,
-      'LINE' AS construction_type,  -- Indicates we're working with line geometries only
-      accident_severity,            -- For severity analysis
-      accident_time,                -- For temporal analysis
-      1 AS temporal_overlap,        -- Always 1 due to our temporal filter in Step 1
-      nearby_construction_count,    -- Number of construction sites affecting this accident
-      1 AS is_near_active_construction  -- Always 1 due to our filtering steps
-    FROM 
-      ranked_matches
-    WHERE 
-      rank = 1  -- Select only one construction site per accident
-  ")
-
-  # Create indices for query optimization
-  dbExecute(con, "CREATE INDEX idx_aca_accident_id ON accident_construction_analysis(accident_id)")
-  dbExecute(con, "CREATE INDEX idx_aca_construction_id ON accident_construction_analysis(construction_id)")
-  dbExecute(con, "CREATE INDEX idx_aca_count ON accident_construction_analysis(nearby_construction_count)")
-  dbExecute(con, "ANALYZE accident_construction_analysis")
-
-  dbExecute(con, "COMMIT")
-  cat("Created accident_construction_analysis table\n")
-} else {
-  cat("Accident construction analysis table already exists, skipping creation\n")
-}
-
-# Read the data for construction zones with accident counts (for visualization)
-accidents_per_zone <- st_read(con, query = "
-  WITH construction_accident_counts AS (
-    SELECT 
-      aca.construction_id,
-      COUNT(aca.accident_id) AS accident_count
-    FROM 
-      accident_construction_analysis aca
-    GROUP BY 
-      aca.construction_id
-  )
-  SELECT 
-    cl.\"ID\" AS construction_id,
-    cac.accident_count,
-    cl.geometry
-  FROM 
-    construction_lines cl
-  JOIN 
-    construction_accident_counts cac ON cl.\"ID\" = cac.construction_id
-  ORDER BY 
-    cac.accident_count DESC
-")
-
-# Get statistics on construction zones
-cat("Construction zones with accidents:", nrow(accidents_per_zone), "\n")
-cat("Maximum accidents in a single construction zone:", max(accidents_per_zone$accident_count), "\n")
-cat("Average accidents per construction zone:", round(mean(accidents_per_zone$accident_count), 2), "\n")
-
-# Read accidents with their associated construction data
-accidents_with_construction <- st_read(con, query = "
-  SELECT 
-    a.*,
-    CASE WHEN aca.accident_id IS NOT NULL THEN 1 ELSE 0 END AS near_construction,
-    aca.construction_id,
-    aca.nearby_construction_count,
-    aca.construction_type
-  FROM 
-    accidents a
-  LEFT JOIN 
-    accident_construction_analysis aca ON a.\"ID\" = aca.accident_id
-")
-
-# Get distribution of accident counts by construction density
-construction_density_stats <- dbGetQuery(con, "
-  SELECT 
-    nearby_construction_count,
-    COUNT(*) AS accident_count
-  FROM 
-    accident_construction_analysis
-  GROUP BY 
-    nearby_construction_count
-  ORDER BY 
-    nearby_construction_count
-")
-
-#---------------------------------------------------------------------------
-# Report completion and verify table sizes
-#---------------------------------------------------------------------------
-cat("All tables are ready for analysis\n")
-
-print(construction_density_stats)
-
-# Total accidents summary - for comparison
-total_accidents <- dbGetQuery(con, "SELECT COUNT(*) as count FROM accidents")$count
-accidents_near_construction <- dbGetQuery(con, "SELECT COUNT(*) as count FROM accident_construction_analysis")$count
-
-cat("Total accidents in dataset:", total_accidents, "\n")
-cat("Accidents near active construction:", accidents_near_construction, "\n")
-cat("Percentage of accidents near active construction: ", 
-    round(accidents_near_construction/total_accidents*100, 2), "%\n")
-
-#---------------------------------------------------------------------------
-# STEP 1A-L: Create construction line buffers table with time ranges
-#---------------------------------------------------------------------------
-has_construction_line_buffers <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'construction_line_buffers')")$exists
-if (!has_construction_line_buffers) {
-  cat("Step 1A-L: Creating construction_line_buffers table with time ranges...\n")
-
-  dbExecute(con, "BEGIN")
-
-  # Create line-based construction buffers with time information
-  dbExecute(con, "
-    CREATE TABLE construction_line_buffers AS
-    SELECT 
-      \"ID\" as construction_id,
-      \"Start_Time\" as start_time,
-      \"End_Time\" as end_time,
-      ST_Transform(
-        ST_Buffer(
-          ST_Transform(geometry, 3310), -- Transform to California Albers for accurate distance
-          300                           -- 300 meters buffer
-        ), 
-        4326                            -- Transform back to WGS 84 for joining
-      ) AS geometry,
-      ST_Length(ST_Transform(geometry, 3310)) AS segment_length_m -- Length in meters
-    FROM construction_lines
-  ")
-
-  # Add primary key and spatial index
-  dbExecute(con, "ALTER TABLE construction_line_buffers ADD PRIMARY KEY (construction_id)")
-  dbExecute(con, "CREATE INDEX idx_construction_line_buffers_geometry ON construction_line_buffers USING GIST(geometry)")
-  dbExecute(con, "CREATE INDEX idx_construction_line_buffers_times ON construction_line_buffers(start_time, end_time)")
-  dbExecute(con, "ANALYZE construction_line_buffers")
-
-  dbExecute(con, "COMMIT")
-  cat("Created construction_line_buffers table with spatial index and time ranges\n")
-} else {
-  cat("Construction line buffers table already exists, skipping creation\n")
-}
-
-#---------------------------------------------------------------------------
-# STEP 1B-L: Create spatial matches table for line objects
-#---------------------------------------------------------------------------
-has_line_spatial_matches <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'line_spatial_matches')")$exists
-if (!has_line_spatial_matches) {
-  cat("Step 1B-L: Creating line_spatial_matches table using pre-computed line buffers...\n")
-
-  dbExecute(con, "BEGIN")
-
-  # Simplified join using pre-computed line buffers with time information
-  dbExecute(con, "
-    CREATE TABLE line_spatial_matches AS
-    SELECT
-      a.\"ID\" AS accident_id,
-      a.\"Severity\" AS accident_severity,
-      a.\"Start_Time\" AS accident_time,
-      b.construction_id,
-      b.segment_length_m
-    FROM 
-      accidents a
-    JOIN 
-      construction_line_buffers b ON 
-        ST_Within(a.geometry, b.geometry) AND
-        a.\"Start_Time\" BETWEEN b.start_time AND b.end_time
-  ")
-
-  # Create indices for efficient join operations
-  dbExecute(con, "CREATE INDEX idx_line_spatial_accident ON line_spatial_matches(accident_id)")
-  dbExecute(con, "CREATE INDEX idx_line_spatial_construction ON line_spatial_matches(construction_id)")
-  dbExecute(con, "ANALYZE line_spatial_matches")
-
-  dbExecute(con, "COMMIT")
-  cat("Created line_spatial_matches table with combined temporal and spatial filtering\n")
-} else {
-  cat("Line spatial matches table already exists, skipping creation\n")
-}
-
-#---------------------------------------------------------------------------
-# STEP 2-L: Count construction line sites per accident
-#---------------------------------------------------------------------------
-has_line_construction_counts <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'line_construction_counts')")$exists
-if (!has_line_construction_counts) {
-  cat("Step 2-L: Creating line_construction_counts table...\n")
-
-  dbExecute(con, "BEGIN")
-
-  dbExecute(con, "
-    CREATE TABLE line_construction_counts AS
-    SELECT
-      accident_id,
-      COUNT(*) AS nearby_line_construction_count
-    FROM 
-      line_spatial_matches
-    GROUP BY
-      accident_id
-  ")
-
-  # Create index for efficient joins in the final step
-  dbExecute(con, "CREATE INDEX idx_line_construction_counts_accident ON line_construction_counts(accident_id)")
-  dbExecute(con, "ANALYZE line_construction_counts")
-
-  dbExecute(con, "COMMIT")
-  cat("Created line_construction_counts table\n")
-} else {
-  cat("Line construction counts table already exists, skipping creation\n")
-}
-
-#---------------------------------------------------------------------------
-# STEP 3-L: Create final line-based analysis table
-#---------------------------------------------------------------------------
-has_line_accident_construction_analysis <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'line_accident_construction_analysis')")$exists
-if (!has_line_accident_construction_analysis) {
-  cat("Step 3-L: Creating final line_accident_construction_analysis table...\n")
-
-  dbExecute(con, "BEGIN")
-
-  dbExecute(con, "
-    CREATE TABLE line_accident_construction_analysis AS
-    WITH ranked_matches AS (
-      SELECT
-        sm.accident_id,
-        sm.construction_id,
-        sm.accident_severity,
-        sm.accident_time,
-        sm.segment_length_m,
-        cc.nearby_line_construction_count,
-        -- Rank construction sites for each accident (random selection if multiple)
-        ROW_NUMBER() OVER (
-          PARTITION BY sm.accident_id
-          ORDER BY RANDOM()  -- Random selection ensures unbiased assignment
-        ) AS rank
-      FROM 
-        line_spatial_matches sm
-      JOIN
-        line_construction_counts cc ON sm.accident_id = cc.accident_id
-    )
-    SELECT 
-      accident_id,
-      construction_id,
-      accident_severity,
-      accident_time,
-      segment_length_m,
-      nearby_line_construction_count,
-      1 AS temporal_overlap,           -- Always 1 due to our temporal filter in Step 1
-      1 AS is_near_active_construction -- Always 1 due to our filtering steps
-    FROM 
-      ranked_matches
-    WHERE 
-      rank = 1  -- Select only one construction site per accident
-  ")
-
-  # Create indices for query optimization
-  dbExecute(con, "CREATE INDEX idx_laca_accident_id ON line_accident_construction_analysis(accident_id)")
-  dbExecute(con, "CREATE INDEX idx_laca_construction_id ON line_accident_construction_analysis(construction_id)")
-  dbExecute(con, "CREATE INDEX idx_laca_count ON line_accident_construction_analysis(nearby_line_construction_count)")
-  dbExecute(con, "ANALYZE line_accident_construction_analysis")
-
-  dbExecute(con, "COMMIT")
-  cat("Created line_accident_construction_analysis table\n")
-} else {
-  cat("Line accident construction analysis table already exists, skipping creation\n")
-}
-
-#---------------------------------------------------------------------------
-# STEP 4-L: Calculate accident rates by segment length
-#---------------------------------------------------------------------------
-has_accident_rates <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'construction_segment_accident_rates')")$exists
-if (!has_accident_rates) {
-  cat("Step 4-L: Creating construction_segment_accident_rates table...\n")
-
-  dbExecute(con, "BEGIN")
-
-  dbExecute(con, "
-    CREATE TABLE construction_segment_accident_rates AS
-    SELECT 
-      cl.\"ID\" AS construction_id,
-      cl.geometry,
-      COUNT(laca.accident_id) AS accident_count,
-      MAX(laca.segment_length_m) AS segment_length_m,
-      COUNT(laca.accident_id) / (MAX(laca.segment_length_m) / 1000.0) AS accidents_per_km
-    FROM 
-      construction_lines cl
-    LEFT JOIN 
-      line_accident_construction_analysis laca ON cl.\"ID\" = laca.construction_id
-    GROUP BY 
-      cl.\"ID\", cl.geometry
-  ")
-
-  # Create spatial index
-  dbExecute(con, "CREATE INDEX idx_accident_rates_geometry ON construction_segment_accident_rates USING GIST(geometry)")
-  dbExecute(con, "ANALYZE construction_segment_accident_rates")
-
-  dbExecute(con, "COMMIT")
-  cat("Created construction_segment_accident_rates table\n")
-} else {
-  cat("Construction segment accident rates table already exists, skipping creation\n")
-}
-
-# Read accident rates by construction segment
-accident_rates_by_segment <- st_read(con, query = "
-  SELECT 
-    construction_id,
-    accident_count,
-    segment_length_m,
-    accidents_per_km,
-    geometry
-  FROM 
-    construction_segment_accident_rates
-  WHERE 
-    segment_length_m > 0  -- Avoid division by zero
-  ORDER BY 
-    accidents_per_km DESC
-")
-
-# Display summary statistics
-cat("Construction segments analyzed:", nrow(accident_rates_by_segment), "\n")
-cat("Maximum accidents per km:", max(accident_rates_by_segment$accidents_per_km, na.rm = TRUE), "\n")
-cat("Average accidents per km:", mean(accident_rates_by_segment$accidents_per_km, na.rm = TRUE), "\n")
-cat("Median segment length (m):", median(accident_rates_by_segment$segment_length_m, na.rm = TRUE), "\n")
-
-# Severity analysis near line construction
-severity_by_construction <- dbGetQuery(con, "
-  SELECT 
-    laca.accident_severity, 
-    COUNT(*) AS count
-  FROM 
-    line_accident_construction_analysis laca
-  GROUP BY 
-    laca.accident_severity
-  ORDER BY 
-    laca.accident_severity
-")
-
-# Comparison of accident severity near vs. away from construction
-severity_comparison <- dbGetQuery(con, "
-  WITH construction_accidents AS (
-    SELECT 
-      a.\"ID\", 
-      a.\"Severity\", 
-      'Near construction' AS location
-    FROM 
-      accidents a
-    JOIN 
-      line_accident_construction_analysis laca ON a.\"ID\" = laca.accident_id
-  ),
-  non_construction_accidents AS (
-    SELECT 
-      a.\"ID\", 
-      a.\"Severity\", 
-      'Away from construction' AS location
-    FROM 
-      accidents a
-    LEFT JOIN 
-      line_accident_construction_analysis laca ON a.\"ID\" = laca.accident_id
-    WHERE 
-      laca.accident_id IS NULL
-  )
-  SELECT 
-    \"Severity\", 
-    location, 
-    COUNT(*) AS count
-  FROM (
-    SELECT * FROM construction_accidents
-    UNION ALL
-    SELECT * FROM non_construction_accidents
-  ) combined
-  GROUP BY 
-    \"Severity\", location
-  ORDER BY 
-    \"Severity\", location
-")
-
-# Time-based analysis
-temporal_analysis <- dbGetQuery(con, "
-  SELECT 
-    DATE_TRUNC('month', laca.accident_time) AS month,
-    COUNT(*) AS accident_count
-  FROM 
-    line_accident_construction_analysis laca
-  GROUP BY 
-    DATE_TRUNC('month', laca.accident_time)
-  ORDER BY 
-    month
-")
-
-# Disconnect from database
-dbDisconnect(con)
-```
-
-## Buffer zone analysis
-
-In theory, the commented code below would achieve the same as our operation PostGIS but way slower. We will not run it here, but keep it to demonstrate that we could implement in R as well.
-
-```{r Buffer Zone Analysis}
-# Set up parallel processing
-#future::plan(future::multisession, workers = parallel::detectCores() - 1)
-
-# Set sf to use parallel processing
-#sf_use_s2(FALSE)  # Disable S2 spherical geometry
-#cores <- parallel::detectCores() - 1
-
-# Process each dataset with batching
-#cat("Creating point buffers...\n")
-#const_buffers <- create_buffers_in_batches(df.const.sf, buffer_dist = 1000, batch_size = 500)
-
-# Compute accidents within construction zones
-#accidents_in_construction <- df.acc.sf %>%
-#  st_join(const_buffers, join = st_within, prepared = TRUE)
-
-# Return to sequential processing
-#future::plan(future::sequential)
-#sf_use_s2(TRUE)  # Re-enable S2 spherical geometry
-
-# Count accidents per construction zone
-#accidents_per_zone <- accidents_with_construction %>%
-#  group_by(ID) %>%  # Grouping by the ID column
-#  summarize(accident_count = n())
-
-accidents_per_zone <- accidents_per_zone %>% drop_na()
-
-# Visualize construction zones with accident counts
-# First, store our plot in a variable
-accident_zones_plot <- ggplot() +
-  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
-  # Arrange data in descending order so high frequency zones are plotted on top
-  geom_sf(data = accidents_per_zone %>% arrange(accident_count), 
-          aes(color = accident_count), size = 1.5) +
-  scale_color_viridis_c() +
-  theme_minimal() +
-  labs(title = "Accident Counts within Construction Zones (2021)")
-
-print(accident_zones_plot)
-
-# Then save it using ggsave
-ggsave(filename = "imgs/accident_counts_construction_zones.png", 
-       plot = accident_zones_plot,
-       width = 10, # width in inches
-       height = 8, # height in inches
-       dpi = 300)  # resolution
-```
-
-## Advanced construction sit analysis (line objects):
-
-```{r Maps with Construction Lines - static}
-# Static map with construction line segments
-ggplot() +
-  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
-  geom_sf(data = df.const.lines %>% filter(State == "CA"), 
-          color = "orange", size = 1) +
-  theme_minimal() +
-  labs(title = "US Construction Segments (2021)")
-```
-
-```{r Maps with Construction Lines - dynamic}
-# Interactive map with construction segments focused on California
-construction_segment_map <- leaflet() %>%
-  addTiles() %>%
-  addPolylines(data = df.const.lines, 
-               color = "orange",
-               weight = 3,
-               opacity = 0.7,
-               popup = ~paste("ID:", ID, "<br>Duration:", duration)) %>%
-  setView(lng = -119.4179, lat = 36.7783, zoom = 6) %>%  # Center on California
-  setMaxBounds(lng1 = -124.6, lat1 = 42.0,    # Northwest corner of CA
-               lng2 = -114.1, lat2 = 32.5)    # Southeast corner of CA
-
-# Display the map
-construction_segment_map
-```
-
-Again, we execute computationally heavy operations in PostGIS but have the commented R code below to showcase a theoretical, but slow implementatiojn in R.
-
-```{r Construction Impact Analysis with Buffers}
-# Set up parallel processing
-#future::plan(future::multisession, workers = parallel::detectCores() - 1)
-
-# Set sf to use parallel processing
-#f_use_s2(FALSE)  # Disable S2 spherical geometry
-#cores <- parallel::detectCores() - 1
-
-#cat("Creating line segment buffers...\n")
-#const_segment_buffers <- create_buffers_in_batches(df.const.lines, buffer_dist = 1000, batch_size = 250)
-
-# Find accidents near construction segments
-#accidents_near_construction <- df.acc.sf %>%
-#  st_join(const_segment_buffers, join = st_intersects, prepared = TRUE)
-
-# Return to sequential processing
-#future::plan(future::sequential)
-#sf_use_s2(TRUE)  # Re-enable S2 spherical geometry
-
-# Analyze accident rate per construction segment length
-#accidents_per_segment <- accidents_near_construction %>%
-#  group_by(ID) %>%  # Assuming there's an ID column for construction segments
-#  summarize(
-#    accident_count = n(),
-#    segment_length = as.numeric(st_length(geometry[1])),  # Length in meters
-#    accidents_per_km = accident_count / (segment_length/1000)
-#  )
-
-# Visualize segments by accident rate
-ggplot() +
-  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
-  geom_sf(data = accident_rates_by_segment, aes(color = accidents_per_km), size = 1.5) +
-  scale_color_viridis_c() +
-  theme_minimal() +
-  labs(title = "Accident Rate per Construction Segment (2016-2021)",
-       color = "Accidents/km")
-```
-
-```{r}
-# Reestablish connection to PostgreSQL database
-con <- dbConnect(
-  Postgres(),
-  dbname = "geospatial_final",
-  host = "localhost",
-  port = 5432,
-  user = "postgres",
-  password = Sys.getenv("POSTGRES_PWD")
-)
-
-# Temporal analysis - construction impact over time
-accidents_by_time <- dbGetQuery(con, "
-  SELECT 
-    DATE_TRUNC('month', accident_time) AS month,
-    COUNT(*) AS accident_count
-  FROM 
-    line_accident_construction_analysis
-  GROUP BY 
-    DATE_TRUNC('month', accident_time)
-  ORDER BY 
-    month
-") %>%
-  mutate(month = as.Date(month))  # Convert PostgreSQL timestamp to R Date
-
-# Get severity comparison data
-severity_comparison <- dbGetQuery(con, "
-  WITH construction_accidents AS (
-    SELECT 
-      a.\"Severity\", 
-      'Near construction' AS location
-    FROM 
-      accidents a
-    JOIN 
-      line_accident_construction_analysis laca ON a.\"ID\" = laca.accident_id
-  ),
-  non_construction_accidents AS (
-    SELECT 
-      a.\"Severity\", 
-      'Away from construction' AS location
-    FROM 
-      accidents a
-    LEFT JOIN 
-      line_accident_construction_analysis laca ON a.\"ID\" = laca.accident_id
-    WHERE 
-      laca.accident_id IS NULL
-  )
-  SELECT 
-    \"Severity\", 
-    location, 
-    COUNT(*) AS count
-  FROM (
-    SELECT * FROM construction_accidents
-    UNION ALL
-    SELECT * FROM non_construction_accidents
-  ) combined
-  GROUP BY 
-    \"Severity\", location
-  ORDER BY 
-    \"Severity\", location
-")
-
-# Disconnect from database
-dbDisconnect(con)
-```
-
-```{r Temporal Analysis and Severity Comparison with Buffers}
-# Visualize temporal analysis
-ggplot(accidents_by_time, aes(x = month, y = accident_count)) +
-  geom_line() +
-  geom_point() +
-  scale_x_date(date_labels = "%b %Y", date_breaks = "1 month") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  labs(title = "Accidents Near Construction Over Time",
-       x = "Month",
-       y = "Accident Count")
-
-# Visualize severity comparison
-ggplot(severity_comparison, aes(x = Severity, y = count, fill = location)) +
-  geom_bar(stat = "identity", position = "dodge") +
-  scale_fill_brewer(palette = "Set1") +
-  theme_minimal() +
-  labs(title = "Accident Severity: Near vs. Away from Construction",
-       x = "Severity Level",
-       y = "Count",
-       fill = "Location")
-```
+## 4.3 Improvements

### 2025-03-26 final/archive/streamlined_nb - Kopie.Rmd
Restructured, cleaned, added final code

---
title: "Geospatial Data Science - Final project"
author:
- Moritz Peist
- Noemi Lucchi
- Simon Vellin
subtitle: Analysis of US road accidents and construction sites
output:
  html_document:
    toc: true
    df_print: paged
  html_notebook:
    df_print: paged
    toc: true
---

<hr>

```{css, echo=FALSE}
h1.title {
  text-align: center;
}

h4.author, h4.date {
  text-align: center;
}

.author {
  text-align: center;
}

.date {
  text-align: center;
}

/* Add styling for subtitle */
h3.subtitle {
  text-align: center;
  font-style: italic;
  margin-top: -10px;
  color: #606060;
}
```

```{r}
library(data.table)   # For faster data manipulation
library(tidyverse)    # For data manipulation and visualization
library(sf)           # For spatial data handling
library(leaflet)      # For interactive maps
library(leaflet.extras) # For additional leaflet features
library(mapview)      # For easier map visualization
library(tmap)         # For thematic maps
library(tigris)       # For US road networks
library(future)       # For parallel processing
library(future.apply) # For parallel processing with apply functions
library(sf)           # For spatial data handling
library(RPostgres)    # For PostgreSQL connection
library(DBI)          # For database interface

# Create directories if they don't exist
if (!dir.exists("./data/tigris")) {
  dir.create("./data/tigris", recursive = TRUE)
}

# Set custom cache directory (optional)
options(tigris_cache_dir = "./data/tigris")
# Configure tigris to use caching
options(tigris_use_cache = TRUE)

# Load Caltrans State Highway Network
aadt <- st_read("data/Traffic_Volumes_AADT/Traffic_Volumes_AADT.shp", crs = 3310)
```

```{r}
# Load accidents

# Efficient approach
df.acc <- fread("data/us_accidents/US_accidents_March23.csv")[
  # Filter date range of 2021
  lubridate::year(as.Date(Start_Time)) == 2021 & 
  # And California
  State == "CA"
][, `:=`(
  # Add year, quarter, month columns
  year = data.table::year(Start_Time),
  quarter = data.table::quarter(Start_Time),
  month = data.table::month(Start_Time),
  # Calculate duration (assuming End_Time exists in the dataset)
  duration = as.numeric(difftime(End_Time, Start_Time, units = "days"))
)] %>% 
  as_tibble()  # Convert to tibble only at the end for performance
```

```{r}
df.const <- fread("data/us_constructions/US_constructions_Dec21.csv")[
  # Filter date range of 2021
  lubridate::year(as.Date(Start_Time)) == 2021 & 
  # And California
  State == "CA"
][, `:=`(
  # Add year, quarter, month columns
  year = year(Start_Time),
  quarter = quarter(Start_Time),
  month = month(Start_Time),
  # Calculate duration (assuming End_Time exists in the dataset)
  duration = as.numeric(difftime(End_Time, Start_Time, units = "days"))
)] %>% 
  as_tibble()  # Convert to tibble only at the end for performance
```

```{r Geospatial Data}
# Convert accident data to sf object
df.acc.sf <- df.acc %>%
  filter(!is.na(Start_Lat) & !is.na(Start_Lng)) %>%
  st_as_sf(coords = c("Start_Lng", "Start_Lat"), crs = 3310)

# Convert construction data to sf object
df.const.sf <- df.const %>%
  filter(!is.na(Start_Lat) & !is.na(Start_Lng)) %>%
  st_as_sf(coords = c("Start_Lng", "Start_Lat"), crs = 3310)
```

### 2025-03-26 final/archive/streamlined_nb.Rmd
Restructured, cleaned, added final code

--- 2025-03-18 final/streamlined_nb.Rmd
@@
 output:
+  html_document:
+    toc: true
+    df_print: paged
   html_notebook:
@@
     toc: true
-  html_document:
-    toc: true
-    df_print: paged
 ---

### 2025-03-26 final/archive/test.tex
Restructured, cleaned, added final code

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{booktabs}
\usepackage{float}
\floatplacement{figure}{H}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={California Road Construction Safety Analysis: A Spatial Econometric Approach},
  pdfauthor={Moritz Peist; Noemi Lucchi; Simon Vellin},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{California Road Construction Safety Analysis: A Spatial
Econometric Approach}
\author{Moritz Peist \and Noemi Lucchi \and Simon Vellin}
\date{März 26, 2025}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{3}
\tableofcontents
}
\subsection{Abstract}\label{abstract}

This study examines the causal relationship between road construction
activities and traffic accident rates in California using
high-resolution spatial data from 2021. Implementing a
difference-in-differences framework with spatial controls, we analyze
how construction zones affect accident frequency and severity. Our
findings indicate a significant positive relationship between active
construction and accident occurrence, with heterogeneous effects across
urban and rural environments. Specifically, we observe that the effect
of construction on accident rates is substantially larger in urban
areas. Spatial spillover analysis reveals that accident impacts extend
beyond immediate construction zones, with particularly strong effects in
areas 500-1000m from construction sites. This research provides
evidence-based insights for transportation safety planning and targeted
mitigation strategies.

\subsection{1. Introduction}\label{introduction}

Road construction is a necessary yet potentially disruptive component of
transportation infrastructure maintenance and development. While
essential for long-term safety improvements, construction activities
create temporary changes in traffic patterns, road conditions, and
driver behavior that may increase accident risk. Understanding the
causal relationship between construction activities and traffic safety
has important implications for construction scheduling, resource
allocation, traffic management strategies, and policy development.

Despite the importance of this relationship, rigorous causal evidence on
how road construction affects traffic accidents remains limited,
particularly with respect to potential heterogeneity between urban and
rural environments. This study addresses this gap by leveraging
comprehensive spatial data on road construction activities and traffic
accidents in California during 2021 to answer the following research
questions:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  How does road construction causally impact traffic accident rates?
\item
  Do these effects systematically differ between urban and rural
  environments?
\item
  Are there spatial spillover effects beyond the immediate construction
  zones?
\end{enumerate}

The findings of this research have direct implications for
transportation safety planning, particularly regarding how resources
should be allocated for traffic management during construction
activities across different environments.

\subsection{2. Data and Methods}\label{data-and-methods}

\subsubsection{2.1 Data Sources}\label{data-sources}

This study integrates multiple large-scale datasets to create a
comprehensive spatial-temporal framework for analysis:

\textbf{Primary Datasets:} 1. \textbf{US Road Construction and Closures
Dataset (2021)} - Contains detailed records of construction activities
in California with coordinates, start/end times, and affected distances
2. \textbf{US Traffic Accidents Dataset (2021)} - Comprehensive accident
records with precise locations, timestamps, and severity measures 3.
\textbf{Caltrans State Highway Network} - Annual Average Daily Traffic
(AADT) data for normalizing accident rates by traffic volume 4.
\textbf{US Census Bureau Urban Areas} - Official urban/rural
classifications for spatial heterogeneity analysis

\subsubsection{2.2 Data Processing and
Preparation}\label{data-processing-and-preparation}

We focused our analysis on California road construction and accidents
from 2021, creating a spatial-temporal framework that allows us to
examine the causal relationship between construction activities and
accidents.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Load Caltrans State Highway Network}
\NormalTok{aadt }\OtherTok{\textless{}{-}} \FunctionTok{st\_read}\NormalTok{(}\StringTok{"data/Traffic\_Volumes\_AADT/Traffic\_Volumes\_AADT.shp"}\NormalTok{)}

\CommentTok{\# Load and filter accidents data}
\NormalTok{df.acc }\OtherTok{\textless{}{-}} \FunctionTok{fread}\NormalTok{(}\StringTok{"data/us\_accidents/US\_accidents\_March23.csv"}\NormalTok{)[}
  \CommentTok{\# Filter date range of 2021}
\NormalTok{  lubridate}\SpecialCharTok{::}\FunctionTok{year}\NormalTok{(}\FunctionTok{as.Date}\NormalTok{(Start\_Time)) }\SpecialCharTok{==} \DecValTok{2021} \SpecialCharTok{\&} 
  \CommentTok{\# And California}
\NormalTok{  State }\SpecialCharTok{==} \StringTok{"CA"}
\NormalTok{][, }\StringTok{\textasciigrave{}}\AttributeTok{:=}\StringTok{\textasciigrave{}}\NormalTok{(}
  \CommentTok{\# Add year, quarter, month columns}
  \AttributeTok{year =}\NormalTok{ data.table}\SpecialCharTok{::}\FunctionTok{year}\NormalTok{(Start\_Time),}
  \AttributeTok{quarter =}\NormalTok{ data.table}\SpecialCharTok{::}\FunctionTok{quarter}\NormalTok{(Start\_Time),}
  \AttributeTok{month =}\NormalTok{ data.table}\SpecialCharTok{::}\FunctionTok{month}\NormalTok{(Start\_Time),}
  \CommentTok{\# Calculate duration}
  \AttributeTok{duration =} \FunctionTok{as.numeric}\NormalTok{(}\FunctionTok{difftime}\NormalTok{(End\_Time, Start\_Time, }\AttributeTok{units =} \StringTok{"days"}\NormalTok{))}
\NormalTok{)] }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{as\_tibble}\NormalTok{()}

\CommentTok{\# Load and filter construction data}
\NormalTok{df.const }\OtherTok{\textless{}{-}} \FunctionTok{fread}\NormalTok{(}\StringTok{"data/us\_constructions/US\_constructions\_Dec21.csv"}\NormalTok{)[}
  \CommentTok{\# Filter date range of 2021}
\NormalTok{  lubridate}\SpecialCharTok{::}\FunctionTok{year}\NormalTok{(}\FunctionTok{as.Date}\NormalTok{(Start\_Time)) }\SpecialCharTok{==} \DecValTok{2021} \SpecialCharTok{\&} 
  \CommentTok{\# And California}
\NormalTok{  State }\SpecialCharTok{==} \StringTok{"CA"}
\NormalTok{][, }\StringTok{\textasciigrave{}}\AttributeTok{:=}\StringTok{\textasciigrave{}}\NormalTok{(}
  \CommentTok{\# Add year, quarter, month columns}
  \AttributeTok{year =} \FunctionTok{year}\NormalTok{(Start\_Time),}
  \AttributeTok{quarter =} \FunctionTok{quarter}\NormalTok{(Start\_Time),}
  \AttributeTok{month =} \FunctionTok{month}\NormalTok{(Start\_Time),}
  \CommentTok{\# Calculate duration}
  \AttributeTok{duration =} \FunctionTok{as.numeric}\NormalTok{(}\FunctionTok{difftime}\NormalTok{(End\_Time, Start\_Time, }\AttributeTok{units =} \StringTok{"days"}\NormalTok{))}
\NormalTok{)][}
\NormalTok{  duration }\SpecialCharTok{\textgreater{}} \DecValTok{1}  \CommentTok{\# Filter out constructions lasting less than a day}
\NormalTok{] }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{as\_tibble}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Display data summary (using pre{-}computed values for reproducibility)}
\NormalTok{accidents\_count }\OtherTok{\textless{}{-}} \DecValTok{341876}
\NormalTok{construction\_count }\OtherTok{\textless{}{-}} \DecValTok{33513}
\NormalTok{aadt\_count }\OtherTok{\textless{}{-}} \DecValTok{13874}

\FunctionTok{cat}\NormalTok{(}\StringTok{"Summary of datasets:}\SpecialCharTok{\textbackslash{}n}\StringTok{"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Summary of datasets:
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{cat}\NormalTok{(}\StringTok{"{-} Accidents dataset:"}\NormalTok{, accidents\_count, }\StringTok{"records}\SpecialCharTok{\textbackslash{}n}\StringTok{"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## - Accidents dataset: 341876 records
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{cat}\NormalTok{(}\StringTok{"{-} Construction dataset:"}\NormalTok{, construction\_count, }\StringTok{"records}\SpecialCharTok{\textbackslash{}n}\StringTok{"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## - Construction dataset: 33513 records
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{cat}\NormalTok{(}\StringTok{"{-} AADT dataset:"}\NormalTok{, aadt\_count, }\StringTok{"records}\SpecialCharTok{\textbackslash{}n}\StringTok{"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## - AADT dataset: 13874 records
\end{verbatim}

The data preparation revealed 341,876 accident records and 33,513
construction events across California in 2021. The temporal distribution
of both accidents and construction projects showed seasonal patterns,
with construction activity peaking in the spring months and accidents
showing higher frequency in winter and fall.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Temporal patterns visualization}
\NormalTok{acc\_monthly }\OtherTok{\textless{}{-}}\NormalTok{ df.acc }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{month =} \FunctionTok{floor\_date}\NormalTok{(}\FunctionTok{as.Date}\NormalTok{(Start\_Time), }\StringTok{"month"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{count}\NormalTok{(month)}

\NormalTok{const\_monthly }\OtherTok{\textless{}{-}}\NormalTok{ df.const }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{month =} \FunctionTok{floor\_date}\NormalTok{(}\FunctionTok{as.Date}\NormalTok{(Start\_Time), }\StringTok{"month"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{count}\NormalTok{(month)}

\CommentTok{\# Plot monthly patterns}
\NormalTok{p1 }\OtherTok{\textless{}{-}} \FunctionTok{ggplot}\NormalTok{(acc\_monthly, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ month, }\AttributeTok{y =}\NormalTok{ n)) }\SpecialCharTok{+}
  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{color =} \StringTok{"blue"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{geom\_point}\NormalTok{(}\AttributeTok{color =} \StringTok{"blue"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Monthly Accidents in California (2021)"}\NormalTok{,}
       \AttributeTok{x =} \StringTok{"Month"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Count"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_minimal}\NormalTok{()}

\NormalTok{p2 }\OtherTok{\textless{}{-}} \FunctionTok{ggplot}\NormalTok{(const\_monthly, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ month, }\AttributeTok{y =}\NormalTok{ n)) }\SpecialCharTok{+}
  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{color =} \StringTok{"red"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{geom\_point}\NormalTok{(}\AttributeTok{color =} \StringTok{"red"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"New Construction Projects in California (2021)"}\NormalTok{,}
       \AttributeTok{x =} \StringTok{"Month"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Count"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_minimal}\NormalTok{()}

\NormalTok{p1 }\SpecialCharTok{+}\NormalTok{ p2}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Check construction durations}
\NormalTok{df.const }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ duration)) }\SpecialCharTok{+}
  \FunctionTok{geom\_histogram}\NormalTok{(}\AttributeTok{bins =} \DecValTok{30}\NormalTok{, }\AttributeTok{fill =} \StringTok{"orange"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Distribution of Construction Duration"}\NormalTok{,}
       \AttributeTok{x =} \StringTok{"Duration (days)"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Count"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_minimal}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\subsubsection{2.3 Methodology}\label{methodology}

Our analysis employs a difference-in-differences (DiD) framework
augmented with spatial controls to identify the causal impact of
construction activities on traffic accidents. This approach allows us to
address potential endogeneity concerns by controlling for time-invariant
location characteristics and time-specific effects.

\paragraph{2.3.1 Spatial Processing}\label{spatial-processing}

To establish the spatial relationship between accidents and construction
activities, we converted both accident and construction datasets to
spatial objects and created 500-meter buffer zones around construction
locations. These buffer zones served as our treatment areas.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Create spatial objects}
\ControlFlowTok{if}\NormalTok{ (}\SpecialCharTok{!}\FunctionTok{inherits}\NormalTok{(df.acc, }\StringTok{"sf"}\NormalTok{)) \{}
\NormalTok{  df.acc.sf }\OtherTok{\textless{}{-}}\NormalTok{ df.acc }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(Start\_Lat) }\SpecialCharTok{\&} \SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(Start\_Lng)) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{st\_as\_sf}\NormalTok{(}\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(}\StringTok{"Start\_Lng"}\NormalTok{, }\StringTok{"Start\_Lat"}\NormalTok{), }\AttributeTok{crs =} \DecValTok{4326}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{st\_transform}\NormalTok{(}\AttributeTok{crs =} \DecValTok{3310}\NormalTok{)  }\CommentTok{\# CA Albers for accurate distance}
\NormalTok{\} }\ControlFlowTok{else}\NormalTok{ \{}
\NormalTok{  df.acc.sf }\OtherTok{\textless{}{-}}\NormalTok{ df.acc}
\NormalTok{\}}

\ControlFlowTok{if}\NormalTok{ (}\SpecialCharTok{!}\FunctionTok{inherits}\NormalTok{(df.const, }\StringTok{"sf"}\NormalTok{)) \{}
\NormalTok{  df.const.sf }\OtherTok{\textless{}{-}}\NormalTok{ df.const }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(Start\_Lat) }\SpecialCharTok{\&} \SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(Start\_Lng)) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{st\_as\_sf}\NormalTok{(}\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(}\StringTok{"Start\_Lng"}\NormalTok{, }\StringTok{"Start\_Lat"}\NormalTok{), }\AttributeTok{crs =} \DecValTok{4326}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{st\_transform}\NormalTok{(}\AttributeTok{crs =} \DecValTok{3310}\NormalTok{)}
\NormalTok{\} }\ControlFlowTok{else}\NormalTok{ \{}
\NormalTok{  df.const.sf }\OtherTok{\textless{}{-}}\NormalTok{ df.const}
\NormalTok{\}}

\CommentTok{\# Create construction buffers (500m)}
\NormalTok{const\_buffers }\OtherTok{\textless{}{-}}\NormalTok{ df.const.sf }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{st\_buffer}\NormalTok{(}\AttributeTok{dist =} \DecValTok{500}\NormalTok{)  }\CommentTok{\# 500m buffer}

\CommentTok{\# Add buffer metadata}
\NormalTok{const\_buffers }\OtherTok{\textless{}{-}}\NormalTok{ const\_buffers }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{buffer\_id =} \FunctionTok{paste0}\NormalTok{(}\StringTok{"c"}\NormalTok{, }\FunctionTok{row\_number}\NormalTok{()),}
    \AttributeTok{start\_date =} \FunctionTok{as.Date}\NormalTok{(Start\_Time),}
    \AttributeTok{end\_date =} \FunctionTok{as.Date}\NormalTok{(End\_Time),}
    \AttributeTok{construction\_duration =} \FunctionTok{as.numeric}\NormalTok{(}\FunctionTok{difftime}\NormalTok{(end\_date, start\_date, }\AttributeTok{units =} \StringTok{"days"}\NormalTok{))}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

\paragraph{2.3.2 Urban-Rural
Classification}\label{urban-rural-classification}

We classified construction sites as urban or rural based on their
intersection with US Census Bureau urban areas. This classification is
crucial for our heterogeneity analysis.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Download urban areas for California}
\NormalTok{ca\_urban\_areas }\OtherTok{\textless{}{-}}\NormalTok{ tigris}\SpecialCharTok{::}\FunctionTok{urban\_areas}\NormalTok{(}\AttributeTok{cb =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{year =} \DecValTok{2019}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{st\_transform}\NormalTok{(}\FunctionTok{st\_crs}\NormalTok{(const\_buffers)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}\FunctionTok{grepl}\NormalTok{(}\StringTok{", CA$"}\NormalTok{, NAME10))}

\CommentTok{\# Classify construction sites as urban or rural}
\NormalTok{const\_urban }\OtherTok{\textless{}{-}} \FunctionTok{st\_join}\NormalTok{(}
\NormalTok{  const\_buffers,}
\NormalTok{  ca\_urban\_areas }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(}\AttributeTok{urban\_name =}\NormalTok{ NAME10),}
  \AttributeTok{join =}\NormalTok{ st\_intersects}
\NormalTok{)}

\NormalTok{const\_urban }\OtherTok{\textless{}{-}}\NormalTok{ const\_urban }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{is\_urban =} \SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(urban\_name))}

\CommentTok{\# Visualize urban/rural distribution}
\NormalTok{urban\_rural\_counts }\OtherTok{\textless{}{-}}\NormalTok{ const\_urban }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{st\_drop\_geometry}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{count}\NormalTok{(is\_urban) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{percentage =}\NormalTok{ n }\SpecialCharTok{/} \FunctionTok{sum}\NormalTok{(n) }\SpecialCharTok{*} \DecValTok{100}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Urban{-}rural distribution (using pre{-}computed values)}
\NormalTok{urban\_rural\_df }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}
  \AttributeTok{is\_urban =} \FunctionTok{c}\NormalTok{(}\ConstantTok{FALSE}\NormalTok{, }\ConstantTok{TRUE}\NormalTok{),}
  \AttributeTok{n =} \FunctionTok{c}\NormalTok{(}\DecValTok{9803}\NormalTok{, }\DecValTok{23795}\NormalTok{),}
  \AttributeTok{percentage =} \FunctionTok{c}\NormalTok{(}\FloatTok{29.2}\NormalTok{, }\FloatTok{70.8}\NormalTok{)}
\NormalTok{)}

\NormalTok{kableExtra}\SpecialCharTok{::}\FunctionTok{kable}\NormalTok{(urban\_rural\_df, }
                 \AttributeTok{col.names =} \FunctionTok{c}\NormalTok{(}\StringTok{"Urban Area"}\NormalTok{, }\StringTok{"Count"}\NormalTok{, }\StringTok{"Percentage (\%)"}\NormalTok{),}
                 \AttributeTok{caption =} \StringTok{"Distribution of Construction Sites by Urban/Rural Classification"}\NormalTok{,}
                 \AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{,}
                 \AttributeTok{digits =} \DecValTok{1}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  kableExtra}\SpecialCharTok{::}\FunctionTok{kable\_styling}\NormalTok{(}\AttributeTok{latex\_options =} \FunctionTok{c}\NormalTok{(}\StringTok{"striped"}\NormalTok{, }\StringTok{"hold\_position"}\NormalTok{),}
                           \AttributeTok{full\_width =} \ConstantTok{FALSE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  kableExtra}\SpecialCharTok{::}\FunctionTok{row\_spec}\NormalTok{(}\DecValTok{0}\NormalTok{, }\AttributeTok{bold =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  kableExtra}\SpecialCharTok{::}\FunctionTok{row\_spec}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\DecValTok{2}\NormalTok{, }\AttributeTok{background =} \FunctionTok{c}\NormalTok{(}\StringTok{"\#f5f5f5"}\NormalTok{, }\StringTok{"\#e8e8e8"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[t]{lrr}
\caption{\label{tab:urban_rural_table}Distribution of Construction Sites by Urban/Rural Classification}\\
\toprule
\textbf{Urban Area} & \textbf{Count} & \textbf{Percentage (\%)}\\
\midrule
\cellcolor[HTML]{f5f5f5}{\cellcolor{gray!10}{FALSE}} & \cellcolor[HTML]{e8e8e8}{\cellcolor{gray!10}{9803}} & \cellcolor[HTML]{f5f5f5}{\cellcolor{gray!10}{29.2}}\\
\cellcolor[HTML]{f5f5f5}{TRUE} & \cellcolor[HTML]{e8e8e8}{23795} & \cellcolor[HTML]{f5f5f5}{70.8}\\
\bottomrule
\end{longtable}

Our analysis revealed that approximately 70.8\% of construction zones
were located in urban areas, with the remaining 29.2\% in rural
environments.

\paragraph{2.3.3 Panel Dataset
Construction}\label{panel-dataset-construction}

We created a panel dataset that tracks accidents in each buffer zone
over time, with indicators for pre-construction, during-construction,
and post-construction periods. This approach allows us to implement the
difference-in-differences design.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Join accidents to construction buffers}
\NormalTok{accidents\_in\_buffers }\OtherTok{\textless{}{-}} \FunctionTok{st\_join}\NormalTok{(}
\NormalTok{  df.acc.sf }\SpecialCharTok{\%\textgreater{}\%} 
    \FunctionTok{select}\NormalTok{(}\AttributeTok{accident\_id =}\NormalTok{ ID, }\AttributeTok{accident\_date =}\NormalTok{ Start\_Time, Severity),}
\NormalTok{  const\_urban }\SpecialCharTok{\%\textgreater{}\%} 
    \FunctionTok{select}\NormalTok{(buffer\_id, }\AttributeTok{construction\_id =}\NormalTok{ ID, start\_date, end\_date, is\_urban),}
  \AttributeTok{join =}\NormalTok{ st\_intersects}
\NormalTok{)}

\CommentTok{\# Classify accidents by timing relative to construction}
\NormalTok{accidents\_in\_buffers }\OtherTok{\textless{}{-}}\NormalTok{ accidents\_in\_buffers }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{accident\_date =} \FunctionTok{as.Date}\NormalTok{(accident\_date),}
    \AttributeTok{time\_period =} \FunctionTok{case\_when}\NormalTok{(}
      \FunctionTok{is.na}\NormalTok{(start\_date) }\SpecialCharTok{\textasciitilde{}} \ConstantTok{NA\_character\_}\NormalTok{,}
\NormalTok{      accident\_date }\SpecialCharTok{\textless{}}\NormalTok{ start\_date }\SpecialCharTok{\textasciitilde{}} \StringTok{"before"}\NormalTok{,}
\NormalTok{      accident\_date }\SpecialCharTok{\textgreater{}}\NormalTok{ end\_date }\SpecialCharTok{\textasciitilde{}} \StringTok{"after"}\NormalTok{,}
      \ConstantTok{TRUE} \SpecialCharTok{\textasciitilde{}} \StringTok{"during"}
\NormalTok{    ),}
    \AttributeTok{month\_year =} \FunctionTok{floor\_date}\NormalTok{(accident\_date, }\StringTok{"month"}\NormalTok{)}
\NormalTok{  )}

\CommentTok{\# Count accidents by construction period}
\NormalTok{period\_counts }\OtherTok{\textless{}{-}}\NormalTok{ accidents\_in\_buffers }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{st\_drop\_geometry}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(time\_period)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{count}\NormalTok{(time\_period)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Period counts (using pre{-}computed values)}
\NormalTok{period\_df }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}
  \AttributeTok{time\_period =} \FunctionTok{c}\NormalTok{(}\StringTok{"after"}\NormalTok{, }\StringTok{"before"}\NormalTok{, }\StringTok{"during"}\NormalTok{),}
  \AttributeTok{n =} \FunctionTok{c}\NormalTok{(}\DecValTok{381112}\NormalTok{, }\DecValTok{590600}\NormalTok{, }\DecValTok{308601}\NormalTok{)}
\NormalTok{)}

\NormalTok{period\_df }\OtherTok{\textless{}{-}}\NormalTok{ period\_df }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{percentage =}\NormalTok{ n }\SpecialCharTok{/} \FunctionTok{sum}\NormalTok{(n) }\SpecialCharTok{*} \DecValTok{100}\NormalTok{)}

\NormalTok{kableExtra}\SpecialCharTok{::}\FunctionTok{kable}\NormalTok{(period\_df, }
                 \AttributeTok{col.names =} \FunctionTok{c}\NormalTok{(}\StringTok{"Construction Period"}\NormalTok{, }\StringTok{"Count"}\NormalTok{, }\StringTok{"Percentage (\%)"}\NormalTok{),}
                 \AttributeTok{caption =} \StringTok{"Distribution of Accidents by Construction Period"}\NormalTok{,}
                 \AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{,}
                 \AttributeTok{digits =} \DecValTok{1}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  kableExtra}\SpecialCharTok{::}\FunctionTok{kable\_styling}\NormalTok{(}\AttributeTok{latex\_options =} \FunctionTok{c}\NormalTok{(}\StringTok{"striped"}\NormalTok{, }\StringTok{"hold\_position"}\NormalTok{),}
                           \AttributeTok{full\_width =} \ConstantTok{FALSE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  kableExtra}\SpecialCharTok{::}\FunctionTok{row\_spec}\NormalTok{(}\DecValTok{0}\NormalTok{, }\AttributeTok{bold =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  kableExtra}\SpecialCharTok{::}\FunctionTok{row\_spec}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\DecValTok{3}\NormalTok{, }\AttributeTok{background =} \FunctionTok{c}\NormalTok{(}\StringTok{"\#f5f5f5"}\NormalTok{, }\StringTok{"\#e8e8e8"}\NormalTok{, }\StringTok{"\#f5f5f5"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[t]{lrr}
\caption{\label{tab:period_counts_table}Distribution of Accidents by Construction Period}\\
\toprule
\textbf{Construction Period} & \textbf{Count} & \textbf{Percentage (\%)}\\
\midrule
\cellcolor[HTML]{f5f5f5}{\cellcolor{gray!10}{after}} & \cellcolor[HTML]{e8e8e8}{\cellcolor{gray!10}{381112}} & \cellcolor[HTML]{f5f5f5}{\cellcolor{gray!10}{29.8}}\\
\cellcolor[HTML]{f5f5f5}{before} & \cellcolor[HTML]{e8e8e8}{590600} & \cellcolor[HTML]{f5f5f5}{46.1}\\
\cellcolor[HTML]{f5f5f5}{\cellcolor{gray!10}{during}} & \cellcolor[HTML]{e8e8e8}{\cellcolor{gray!10}{308601}} & \cellcolor[HTML]{f5f5f5}{\cellcolor{gray!10}{24.1}}\\
\bottomrule
\end{longtable}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Generate all buffer{-}month combinations}
\NormalTok{all\_months }\OtherTok{\textless{}{-}} \FunctionTok{seq}\NormalTok{(}\FunctionTok{as.Date}\NormalTok{(}\StringTok{"2021{-}01{-}01"}\NormalTok{), }\FunctionTok{as.Date}\NormalTok{(}\StringTok{"2021{-}12{-}01"}\NormalTok{), }\AttributeTok{by =} \StringTok{"month"}\NormalTok{)}
\NormalTok{panel\_grid }\OtherTok{\textless{}{-}} \FunctionTok{expand\_grid}\NormalTok{(}
  \AttributeTok{buffer\_id =} \FunctionTok{unique}\NormalTok{(const\_urban}\SpecialCharTok{$}\NormalTok{buffer\_id),}
  \AttributeTok{month =}\NormalTok{ all\_months}
\NormalTok{)}

\CommentTok{\# Join buffer characteristics}
\NormalTok{panel\_grid }\OtherTok{\textless{}{-}}\NormalTok{ panel\_grid }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{left\_join}\NormalTok{(}
\NormalTok{    const\_urban }\SpecialCharTok{\%\textgreater{}\%}
      \FunctionTok{st\_drop\_geometry}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
      \FunctionTok{select}\NormalTok{(buffer\_id, }\AttributeTok{construction\_id =}\NormalTok{ ID, start\_date, end\_date, is\_urban),}
    \AttributeTok{by =} \StringTok{"buffer\_id"}
\NormalTok{  )}

\CommentTok{\# Create treatment indicator}
\NormalTok{panel\_grid }\OtherTok{\textless{}{-}}\NormalTok{ panel\_grid }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{treatment =}\NormalTok{ month }\SpecialCharTok{\textgreater{}=}\NormalTok{ start\_date }\SpecialCharTok{\&}\NormalTok{ month }\SpecialCharTok{\textless{}=}\NormalTok{ end\_date,}
    \AttributeTok{post\_treatment =}\NormalTok{ month }\SpecialCharTok{\textgreater{}}\NormalTok{ end\_date}
\NormalTok{  )}

\CommentTok{\# Count accidents by buffer and month}
\NormalTok{monthly\_accidents }\OtherTok{\textless{}{-}}\NormalTok{ accidents\_in\_buffers }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{st\_drop\_geometry}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{month =} \FunctionTok{floor\_date}\NormalTok{(accident\_date, }\StringTok{"month"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{count}\NormalTok{(buffer\_id, month) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{rename}\NormalTok{(}\AttributeTok{accident\_count =}\NormalTok{ n)}

\CommentTok{\# Join accident counts to panel}
\NormalTok{panel\_data }\OtherTok{\textless{}{-}}\NormalTok{ panel\_grid }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{left\_join}\NormalTok{(monthly\_accidents, }\AttributeTok{by =} \FunctionTok{c}\NormalTok{(}\StringTok{"buffer\_id"}\NormalTok{, }\StringTok{"month"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{replace\_na}\NormalTok{(}\FunctionTok{list}\NormalTok{(}\AttributeTok{accident\_count =} \DecValTok{0}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\paragraph{2.3.4 Traffic Volume
Integration}\label{traffic-volume-integration}

To control for exposure to traffic, we integrated Annual Average Daily
Traffic (AADT) data from Caltrans, allowing us to normalize accident
counts by traffic volume.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Transform AADT data to ensure matching CRS}
\NormalTok{aadt\_sf }\OtherTok{\textless{}{-}} \FunctionTok{st\_transform}\NormalTok{(aadt, }\FunctionTok{st\_crs}\NormalTok{(const\_buffers))}

\CommentTok{\# Convert AADT columns to numeric}
\NormalTok{aadt\_sf }\OtherTok{\textless{}{-}}\NormalTok{ aadt\_sf }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{BACK\_AADT\_num =} \FunctionTok{as.numeric}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(BACK\_AADT)),}
    \AttributeTok{AHEAD\_AADT\_num =} \FunctionTok{as.numeric}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(AHEAD\_AADT))}
\NormalTok{  )}

\CommentTok{\# Directly find the nearest AADT point for each construction buffer}
\NormalTok{nearest\_indices }\OtherTok{\textless{}{-}} \FunctionTok{st\_nearest\_feature}\NormalTok{(const\_buffers, aadt\_sf)}
\NormalTok{nearest\_distances }\OtherTok{\textless{}{-}} \FunctionTok{st\_distance}\NormalTok{(const\_buffers, aadt\_sf[nearest\_indices,], }\AttributeTok{by\_element =} \ConstantTok{TRUE}\NormalTok{)}

\CommentTok{\# Create a dataframe with buffer IDs and nearest AADT values}
\NormalTok{buffer\_aadt }\OtherTok{\textless{}{-}}\NormalTok{ const\_buffers }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{select}\NormalTok{(buffer\_id) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{bind\_cols}\NormalTok{(}
\NormalTok{    aadt\_sf[nearest\_indices, ] }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{select}\NormalTok{(BACK\_AADT\_num, AHEAD\_AADT\_num)}
\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{distance\_to\_nearest =} \FunctionTok{as.numeric}\NormalTok{(nearest\_distances),}
    \AttributeTok{traffic\_volume =} \FunctionTok{rowMeans}\NormalTok{(}\FunctionTok{cbind}\NormalTok{(BACK\_AADT\_num, AHEAD\_AADT\_num), }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{),}
    \AttributeTok{n\_points =} \DecValTok{1}
\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{st\_drop\_geometry}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{select}\NormalTok{(buffer\_id, traffic\_volume, n\_points, distance\_to\_nearest)}

\CommentTok{\# Use the buffer approach for points within a certain distance}
\NormalTok{buffer\_radius }\OtherTok{\textless{}{-}} \DecValTok{1000}  \CommentTok{\# meters}
\NormalTok{buffer\_aadt }\OtherTok{\textless{}{-}}\NormalTok{ buffer\_aadt }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{method =} \FunctionTok{if\_else}\NormalTok{(distance\_to\_nearest }\SpecialCharTok{\textless{}=}\NormalTok{ buffer\_radius, }\StringTok{"within\_buffer"}\NormalTok{, }\StringTok{"nearest\_feature"}\NormalTok{)}
\NormalTok{  )}

\CommentTok{\# Join to panel data}
\NormalTok{panel\_data }\OtherTok{\textless{}{-}}\NormalTok{ panel\_data }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{left\_join}\NormalTok{(buffer\_aadt, }\AttributeTok{by =} \StringTok{"buffer\_id"}\NormalTok{)}

\CommentTok{\# Create normalized accident rate}
\NormalTok{panel\_data }\OtherTok{\textless{}{-}}\NormalTok{ panel\_data }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{accident\_rate =} \FunctionTok{ifelse}\NormalTok{(traffic\_volume }\SpecialCharTok{\textgreater{}} \DecValTok{0}\NormalTok{, }
\NormalTok{                          accident\_count }\SpecialCharTok{/}\NormalTok{ traffic\_volume }\SpecialCharTok{*} \DecValTok{10000}\NormalTok{,  }\CommentTok{\# Per 10,000 vehicles}
                          \ConstantTok{NA}\NormalTok{)}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

\subsubsection{2.4 Econometric Models}\label{econometric-models}

We implemented several model specifications to ensure robustness and to
explore heterogeneity in the effects:

\paragraph{2.4.1 Base Difference-in-Differences
Model}\label{base-difference-in-differences-model}

Our baseline model takes the form:

\[accident\_count_{it} = \alpha + \beta_1 treatment_{it} + \gamma_t + \delta_i + \varepsilon_{it}\]

Where: - \(accident\_count_{it}\) is the number of accidents in buffer
\(i\) at time \(t\) - \(treatment_{it}\) is a binary indicator for
active construction - \(\gamma_t\) represents time fixed effects -
\(\delta_i\) represents location fixed effects - \(\varepsilon_{it}\) is
the error term

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Basic DiD model}
\NormalTok{did\_model }\OtherTok{\textless{}{-}} \FunctionTok{feols}\NormalTok{(}
\NormalTok{  accident\_count }\SpecialCharTok{\textasciitilde{}}\NormalTok{ treatment }\SpecialCharTok{|}\NormalTok{ buffer\_id }\SpecialCharTok{+}\NormalTok{ month,}
  \AttributeTok{data =}\NormalTok{ panel\_data}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\paragraph{2.4.2 Heterogeneity Analysis}\label{heterogeneity-analysis}

To examine differential effects between urban and rural areas, we
extended the base model with an interaction term:

\[accident\_count_{it} = \alpha + \beta_1 treatment_{it} + \beta_2 (treatment_{it} \times urban_i) + \gamma_t + \delta_i + \varepsilon_{it}\]

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# DiD with urban/rural interaction}
\NormalTok{did\_urban\_model }\OtherTok{\textless{}{-}} \FunctionTok{feols}\NormalTok{(}
\NormalTok{  accident\_count }\SpecialCharTok{\textasciitilde{}}\NormalTok{ treatment }\SpecialCharTok{*}\NormalTok{ is\_urban }\SpecialCharTok{|}\NormalTok{ buffer\_id }\SpecialCharTok{+}\NormalTok{ month,}
  \AttributeTok{data =}\NormalTok{ panel\_data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(is\_urban))}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\paragraph{2.4.3 Normalized Accident Rate
Model}\label{normalized-accident-rate-model}

To account for traffic volume differences, we also estimated models
using normalized accident rates:

\[accident\_rate_{it} = \alpha + \beta_1 treatment_{it} + \gamma_t + \delta_i + \varepsilon_{it}\]

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# DiD with normalized accident rate}
\NormalTok{did\_rate\_model }\OtherTok{\textless{}{-}} \FunctionTok{feols}\NormalTok{(}
\NormalTok{  accident\_rate }\SpecialCharTok{\textasciitilde{}}\NormalTok{ treatment }\SpecialCharTok{|}\NormalTok{ buffer\_id }\SpecialCharTok{+}\NormalTok{ month,}
  \AttributeTok{data =}\NormalTok{ panel\_data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(accident\_rate))}
\NormalTok{)}

\CommentTok{\# DiD with urban/rural heterogeneity and normalized rate}
\NormalTok{did\_rate\_urban\_model }\OtherTok{\textless{}{-}} \FunctionTok{feols}\NormalTok{(}
\NormalTok{  accident\_rate }\SpecialCharTok{\textasciitilde{}}\NormalTok{ treatment }\SpecialCharTok{*}\NormalTok{ is\_urban }\SpecialCharTok{|}\NormalTok{ buffer\_id }\SpecialCharTok{+}\NormalTok{ month,}
  \AttributeTok{data =}\NormalTok{ panel\_data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(accident\_rate), }\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(is\_urban))}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\paragraph{2.4.4 Spatial Econometric
Models}\label{spatial-econometric-models}

To account for spatial dependence, we implemented spatial lag models:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Prepare spatial weights}
\NormalTok{coords }\OtherTok{\textless{}{-}} \FunctionTok{st\_coordinates}\NormalTok{(}\FunctionTok{st\_centroid}\NormalTok{(const\_buffers))}
\NormalTok{buffer\_nb }\OtherTok{\textless{}{-}}\NormalTok{ spdep}\SpecialCharTok{::}\FunctionTok{dnearneigh}\NormalTok{(coords, }\AttributeTok{d1 =} \DecValTok{0}\NormalTok{, }\AttributeTok{d2 =} \DecValTok{5000}\NormalTok{)}
\NormalTok{buffer\_listw }\OtherTok{\textless{}{-}}\NormalTok{ spdep}\SpecialCharTok{::}\FunctionTok{nb2listw}\NormalTok{(buffer\_nb, }\AttributeTok{style =} \StringTok{"W"}\NormalTok{, }\AttributeTok{zero.policy =} \ConstantTok{TRUE}\NormalTok{)}

\CommentTok{\# Create a sample data for spatial analysis (one month snapshot)}
\NormalTok{test\_month }\OtherTok{\textless{}{-}} \FunctionTok{as.Date}\NormalTok{(}\StringTok{"2021{-}06{-}01"}\NormalTok{)}
\NormalTok{spdata }\OtherTok{\textless{}{-}}\NormalTok{ panel\_data }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(county\_name), month }\SpecialCharTok{==}\NormalTok{ test\_month)}

\CommentTok{\# Create spatial weights}
\NormalTok{spw\_coords }\OtherTok{\textless{}{-}}\NormalTok{ spdata }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{select}\NormalTok{(X, Y) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{as.matrix}\NormalTok{()}
\NormalTok{sp\_nb }\OtherTok{\textless{}{-}}\NormalTok{ spdep}\SpecialCharTok{::}\FunctionTok{dnearneigh}\NormalTok{(spw\_coords, }\AttributeTok{d1 =} \DecValTok{0}\NormalTok{, }\AttributeTok{d2 =} \DecValTok{5000}\NormalTok{)}
\NormalTok{spw }\OtherTok{\textless{}{-}}\NormalTok{ spdep}\SpecialCharTok{::}\FunctionTok{nb2listw}\NormalTok{(sp\_nb, }\AttributeTok{style =} \StringTok{"W"}\NormalTok{, }\AttributeTok{zero.policy =} \ConstantTok{TRUE}\NormalTok{)}

\CommentTok{\# Spatial lag model}
\NormalTok{spatial\_model }\OtherTok{\textless{}{-}}\NormalTok{ spatialreg}\SpecialCharTok{::}\FunctionTok{lagsarlm}\NormalTok{(}
\NormalTok{  accident\_count }\SpecialCharTok{\textasciitilde{}}\NormalTok{ treatment }\SpecialCharTok{+}\NormalTok{ is\_urban }\SpecialCharTok{+} \FunctionTok{log}\NormalTok{(traffic\_volume }\SpecialCharTok{+} \DecValTok{1}\NormalTok{),}
  \AttributeTok{data =}\NormalTok{ spdata,}
  \AttributeTok{listw =}\NormalTok{ spw,}
  \AttributeTok{method =} \StringTok{"Matrix"}\NormalTok{,}
  \AttributeTok{zero.policy =} \ConstantTok{TRUE}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\paragraph{2.4.5 Spatial Spillover
Analysis}\label{spatial-spillover-analysis}

To examine how construction impacts extend beyond the immediate
construction zone, we implemented distance band models:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Create distance bands from construction sites}
\NormalTok{panel\_data }\OtherTok{\textless{}{-}}\NormalTok{ panel\_data }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{distance\_band =} \FunctionTok{cut}\NormalTok{(}
\NormalTok{      distance\_to\_nearest,}
      \AttributeTok{breaks =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{500}\NormalTok{, }\DecValTok{1000}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{5000}\NormalTok{, }\ConstantTok{Inf}\NormalTok{),}
      \AttributeTok{labels =} \FunctionTok{c}\NormalTok{(}\StringTok{"0{-}500m"}\NormalTok{, }\StringTok{"500{-}1000m"}\NormalTok{, }\StringTok{"1{-}2km"}\NormalTok{, }\StringTok{"2{-}5km"}\NormalTok{, }\StringTok{"\textgreater{}5km"}\NormalTok{)}
\NormalTok{    )}
\NormalTok{  )}

\CommentTok{\# Test for spillover effects across distance bands}
\NormalTok{spillover\_model }\OtherTok{\textless{}{-}} \FunctionTok{feols}\NormalTok{(}
\NormalTok{  accident\_count }\SpecialCharTok{\textasciitilde{}} \FunctionTok{i}\NormalTok{(distance\_band, treatment, }\AttributeTok{ref =} \StringTok{"\textgreater{}5km"}\NormalTok{) }\SpecialCharTok{|}\NormalTok{ buffer\_id }\SpecialCharTok{+}\NormalTok{ month,}
  \AttributeTok{data =}\NormalTok{ panel\_data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(distance\_band))}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\paragraph{2.4.6 Event Study Analysis}\label{event-study-analysis}

To test for parallel trends and examine dynamic effects, we implemented
an event study approach:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Create relative time variable}
\NormalTok{panel\_data }\OtherTok{\textless{}{-}}\NormalTok{ panel\_data }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{rel\_time =} \FunctionTok{as.numeric}\NormalTok{(}\FunctionTok{difftime}\NormalTok{(month, start\_date, }\AttributeTok{units =} \StringTok{"days"}\NormalTok{)) }\SpecialCharTok{/} \DecValTok{30}\NormalTok{,}
    \AttributeTok{rel\_time\_cat =} \FunctionTok{cut}\NormalTok{(rel\_time, }
                       \AttributeTok{breaks =} \FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\ConstantTok{Inf}\NormalTok{, }\SpecialCharTok{{-}}\DecValTok{12}\NormalTok{, }\SpecialCharTok{{-}}\DecValTok{6}\NormalTok{, }\SpecialCharTok{{-}}\DecValTok{3}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{3}\NormalTok{, }\DecValTok{6}\NormalTok{, }\DecValTok{12}\NormalTok{, }\ConstantTok{Inf}\NormalTok{),}
                       \AttributeTok{labels =} \FunctionTok{c}\NormalTok{(}\StringTok{"\textless{}{-}12m"}\NormalTok{, }\StringTok{"{-}12to{-}6m"}\NormalTok{, }\StringTok{"{-}6to{-}3m"}\NormalTok{, }\StringTok{"{-}3to0m"}\NormalTok{, }
                                  \StringTok{"0to3m"}\NormalTok{, }\StringTok{"3to6m"}\NormalTok{, }\StringTok{"6to12m"}\NormalTok{, }\StringTok{"\textgreater{}12m"}\NormalTok{))}
\NormalTok{  )}

\CommentTok{\# Event study model}
\NormalTok{event\_study\_model }\OtherTok{\textless{}{-}} \FunctionTok{feols}\NormalTok{(}
\NormalTok{  accident\_count }\SpecialCharTok{\textasciitilde{}} \FunctionTok{i}\NormalTok{(rel\_time\_cat, }\AttributeTok{ref =} \StringTok{"{-}3to0m"}\NormalTok{) }\SpecialCharTok{|}\NormalTok{ buffer\_id }\SpecialCharTok{+}\NormalTok{ county\_name}\SpecialCharTok{\^{}}\NormalTok{month,}
  \AttributeTok{data =}\NormalTok{ panel\_data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(county\_name), }\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(rel\_time\_cat))}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\subsection{3. Results}\label{results}

\subsubsection{3.1 Descriptive Statistics}\label{descriptive-statistics}

Our analysis covered 341,876 accidents and 33,513 construction projects
across California in 2021. Examining the temporal distribution reveals
that accident rates peaked in winter months (January and December) and
fall (October), while construction activity was highest in spring
(March-May).

The spatial distribution analysis showed that 70.8\% of construction
activities occurred in urban areas, with the remaining 29.2\% in rural
environments. This urban-rural breakdown is critical for understanding
the heterogeneous effects of construction on accident rates.

When examining accidents by their timing relative to construction, we
observed the following distribution:

\begin{itemize}
\tightlist
\item
  Before construction: 590,600 accidents (46.2\%)
\item
  During construction: 308,601 accidents (24.1\%)
\item
  After construction: 381,112 accidents (29.7\%)
\end{itemize}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Summary statistics table}
\NormalTok{summary\_stats }\OtherTok{\textless{}{-}}\NormalTok{ panel\_data }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{group\_by}\NormalTok{(treatment, is\_urban) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{summarize}\NormalTok{(}
    \AttributeTok{num\_observations =} \FunctionTok{n}\NormalTok{(),}
    \AttributeTok{mean\_accidents =} \FunctionTok{mean}\NormalTok{(accident\_count),}
    \AttributeTok{median\_accidents =} \FunctionTok{median}\NormalTok{(accident\_count),}
    \AttributeTok{max\_accidents =} \FunctionTok{max}\NormalTok{(accident\_count),}
    \AttributeTok{mean\_accident\_rate =} \FunctionTok{mean}\NormalTok{(accident\_rate, }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{)}
\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{ungroup}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Summary statistics (using pre{-}computed values)}
\NormalTok{summary\_stats\_df }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}
  \AttributeTok{treatment =} \FunctionTok{c}\NormalTok{(}\ConstantTok{FALSE}\NormalTok{, }\ConstantTok{FALSE}\NormalTok{, }\ConstantTok{TRUE}\NormalTok{, }\ConstantTok{TRUE}\NormalTok{),}
  \AttributeTok{is\_urban =} \FunctionTok{c}\NormalTok{(}\ConstantTok{FALSE}\NormalTok{, }\ConstantTok{TRUE}\NormalTok{, }\ConstantTok{FALSE}\NormalTok{, }\ConstantTok{TRUE}\NormalTok{),}
  \AttributeTok{num\_observations =} \FunctionTok{c}\NormalTok{(}\DecValTok{91383}\NormalTok{, }\DecValTok{220203}\NormalTok{, }\DecValTok{26253}\NormalTok{, }\DecValTok{65337}\NormalTok{),}
  \AttributeTok{mean\_accidents =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.526}\NormalTok{, }\FloatTok{4.230}\NormalTok{, }\FloatTok{0.505}\NormalTok{, }\FloatTok{4.550}\NormalTok{),}
  \AttributeTok{median\_accidents =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{),}
  \AttributeTok{max\_accidents =} \FunctionTok{c}\NormalTok{(}\DecValTok{61}\NormalTok{, }\DecValTok{82}\NormalTok{, }\DecValTok{44}\NormalTok{, }\DecValTok{77}\NormalTok{),}
  \AttributeTok{mean\_accident\_rate =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.037}\NormalTok{, }\FloatTok{0.086}\NormalTok{, }\FloatTok{0.036}\NormalTok{, }\FloatTok{0.092}\NormalTok{)}
\NormalTok{)}

\NormalTok{kableExtra}\SpecialCharTok{::}\FunctionTok{kable}\NormalTok{(summary\_stats\_df, }
                 \AttributeTok{col.names =} \FunctionTok{c}\NormalTok{(}\StringTok{"Construction"}\NormalTok{, }\StringTok{"Urban"}\NormalTok{, }\StringTok{"Observations"}\NormalTok{, }\StringTok{"Mean Accidents"}\NormalTok{, }
                               \StringTok{"Median Accidents"}\NormalTok{, }\StringTok{"Max Accidents"}\NormalTok{, }\StringTok{"Mean Accident Rate"}\NormalTok{),}
                 \AttributeTok{caption =} \StringTok{"Summary Statistics by Treatment Status and Urban/Rural Classification"}\NormalTok{,}
                 \AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{,}
                 \AttributeTok{digits =} \DecValTok{3}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  kableExtra}\SpecialCharTok{::}\FunctionTok{kable\_styling}\NormalTok{(}\AttributeTok{latex\_options =} \FunctionTok{c}\NormalTok{(}\StringTok{"striped"}\NormalTok{, }\StringTok{"hold\_position"}\NormalTok{, }\StringTok{"scale\_down"}\NormalTok{),}
                           \AttributeTok{full\_width =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  kableExtra}\SpecialCharTok{::}\FunctionTok{row\_spec}\NormalTok{(}\DecValTok{0}\NormalTok{, }\AttributeTok{bold =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  kableExtra}\SpecialCharTok{::}\FunctionTok{row\_spec}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\DecValTok{4}\NormalTok{, }\AttributeTok{background =} \FunctionTok{c}\NormalTok{(}\StringTok{"\#f5f5f5"}\NormalTok{, }\StringTok{"\#e8e8e8"}\NormalTok{, }\StringTok{"\#f5f5f5"}\NormalTok{, }\StringTok{"\#e8e8e8"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{longtabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\caption{\label{tab:summary_stats_table}Summary Statistics by Treatment Status and Urban/Rural Classification}\\
\toprule
\textbf{Construction} & \textbf{Urban} & \textbf{Observations} & \textbf{Mean Accidents} & \textbf{Median Accidents} & \textbf{Max Accidents} & \textbf{Mean Accident Rate}\\
\midrule
\cellcolor[HTML]{f5f5f5}{\cellcolor{gray!10}{FALSE}} & \cellcolor[HTML]{e8e8e8}{\cellcolor{gray!10}{FALSE}} & \cellcolor[HTML]{f5f5f5}{\cellcolor{gray!10}{91383}} & \cellcolor[HTML]{e8e8e8}{\cellcolor{gray!10}{0.526}} & \cellcolor[HTML]{f5f5f5}{\cellcolor{gray!10}{0}} & \cellcolor[HTML]{e8e8e8}{\cellcolor{gray!10}{61}} & \cellcolor[HTML]{f5f5f5}{\cellcolor{gray!10}{0.037}}\\
\cellcolor[HTML]{f5f5f5}{FALSE} & \cellcolor[HTML]{e8e8e8}{TRUE} & \cellcolor[HTML]{f5f5f5}{220203} & \cellcolor[HTML]{e8e8e8}{4.230} & \cellcolor[HTML]{f5f5f5}{1} & \cellcolor[HTML]{e8e8e8}{82} & \cellcolor[HTML]{f5f5f5}{0.086}\\
\cellcolor[HTML]{f5f5f5}{\cellcolor{gray!10}{TRUE}} & \cellcolor[HTML]{e8e8e8}{\cellcolor{gray!10}{FALSE}} & \cellcolor[HTML]{f5f5f5}{\cellcolor{gray!10}{26253}} & \cellcolor[HTML]{e8e8e8}{\cellcolor{gray!10}{0.505}} & \cellcolor[HTML]{f5f5f5}{\cellcolor{gray!10}{0}} & \cellcolor[HTML]{e8e8e8}{\cellcolor{gray!10}{44}} & \cellcolor[HTML]{f5f5f5}{\cellcolor{gray!10}{0.036}}\\
\cellcolor[HTML]{f5f5f5}{TRUE} & \cellcolor[HTML]{e8e8e8}{TRUE} & \cellcolor[HTML]{f5f5f5}{65337} & \cellcolor[HTML]{e8e8e8}{4.550} & \cellcolor[HTML]{f5f5f5}{1} & \cellcolor[HTML]{e8e8e8}{77} & \cellcolor[HTML]{f5f5f5}{0.092}\\
\bottomrule
\end{longtabu}

\subsubsection{3.2 Difference-in-Differences
Results}\label{difference-in-differences-results}

Table 3 presents the results from our main difference-in-differences
models.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# DiD model results (using pre{-}computed values)}
\NormalTok{did\_models }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}
  \AttributeTok{Variable =} \FunctionTok{c}\NormalTok{(}\StringTok{"Treatment"}\NormalTok{, }\StringTok{"Treatment × Urban"}\NormalTok{, }\StringTok{"Fixed Effects"}\NormalTok{, }\StringTok{"Observations"}\NormalTok{, }\StringTok{"R²"}\NormalTok{, }\StringTok{"Within R²"}\NormalTok{),}
  \AttributeTok{Base =} \FunctionTok{c}\NormalTok{(}\StringTok{"0.096*** (0.021)"}\NormalTok{, }\StringTok{"{-}"}\NormalTok{, }\StringTok{"Buffer, Month"}\NormalTok{, }\StringTok{"403,176"}\NormalTok{, }\StringTok{"0.749"}\NormalTok{, }\StringTok{"7.93e{-}5"}\NormalTok{),}
  \AttributeTok{Urban =} \FunctionTok{c}\NormalTok{(}\StringTok{"{-}0.056*** (0.013)"}\NormalTok{, }\StringTok{"0.216*** (0.029)"}\NormalTok{, }\StringTok{"Buffer, Month"}\NormalTok{, }\StringTok{"403,176"}\NormalTok{, }\StringTok{"0.749"}\NormalTok{, }\StringTok{"1.83e{-}4"}\NormalTok{),}
  \AttributeTok{Rate =} \FunctionTok{c}\NormalTok{(}\StringTok{"0.026 (0.017)"}\NormalTok{, }\StringTok{"{-}"}\NormalTok{, }\StringTok{"Buffer, Month"}\NormalTok{, }\StringTok{"402,084"}\NormalTok{, }\StringTok{"0.178"}\NormalTok{, }\StringTok{"5.96e{-}6"}\NormalTok{),}
  \AttributeTok{Rate\_Urban =} \FunctionTok{c}\NormalTok{(}\StringTok{"{-}0.032 (0.048)"}\NormalTok{, }\StringTok{"0.082* (0.048)"}\NormalTok{, }\StringTok{"Buffer, Month"}\NormalTok{, }\StringTok{"402,084"}\NormalTok{, }\StringTok{"0.178"}\NormalTok{, }\StringTok{"2.13e{-}5"}\NormalTok{)}
\NormalTok{)}

\NormalTok{kableExtra}\SpecialCharTok{::}\FunctionTok{kable}\NormalTok{(did\_models, }
                 \AttributeTok{col.names =} \FunctionTok{c}\NormalTok{(}\StringTok{"Variable"}\NormalTok{, }\StringTok{"Base Model"}\NormalTok{, }\StringTok{"Urban Heterogeneity"}\NormalTok{, }\StringTok{"Rate Model"}\NormalTok{, }\StringTok{"Rate Urban Model"}\NormalTok{),}
                 \AttributeTok{caption =} \StringTok{"Difference{-}in{-}Differences Estimates of Construction Effects on Accidents"}\NormalTok{,}
                 \AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  kableExtra}\SpecialCharTok{::}\FunctionTok{kable\_styling}\NormalTok{(}\AttributeTok{latex\_options =} \FunctionTok{c}\NormalTok{(}\StringTok{"striped"}\NormalTok{, }\StringTok{"hold\_position"}\NormalTok{, }\StringTok{"scale\_down"}\NormalTok{),}
                           \AttributeTok{full\_width =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  kableExtra}\SpecialCharTok{::}\FunctionTok{row\_spec}\NormalTok{(}\DecValTok{0}\NormalTok{, }\AttributeTok{bold =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  kableExtra}\SpecialCharTok{::}\FunctionTok{add\_footnote}\NormalTok{(}\StringTok{"Standard errors clustered by buffer\_id in parentheses. *** p\textless{}0.001, ** p\textless{}0.01, * p\textless{}0.05"}\NormalTok{, }
                          \AttributeTok{notation =} \StringTok{"none"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{longtabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\caption{\label{tab:did_models_table}Difference-in-Differences Estimates of Construction Effects on Accidents}\\
\toprule
\textbf{Variable} & \textbf{Base Model} & \textbf{Urban Heterogeneity} & \textbf{Rate Model} & \textbf{Rate Urban Model}\\
\midrule
\cellcolor{gray!10}{Treatment} & \cellcolor{gray!10}{0.096*** (0.021)} & \cellcolor{gray!10}{-0.056*** (0.013)} & \cellcolor{gray!10}{0.026 (0.017)} & \cellcolor{gray!10}{-0.032 (0.048)}\\
Treatment × Urban & - & 0.216*** (0.029) & - & 0.082* (0.048)\\
\cellcolor{gray!10}{Fixed Effects} & \cellcolor{gray!10}{Buffer, Month} & \cellcolor{gray!10}{Buffer, Month} & \cellcolor{gray!10}{Buffer, Month} & \cellcolor{gray!10}{Buffer, Month}\\
Observations & 403,176 & 403,176 & 402,084 & 402,084\\
\cellcolor{gray!10}{R²} & \cellcolor{gray!10}{0.749} & \cellcolor{gray!10}{0.749} & \cellcolor{gray!10}{0.178} & \cellcolor{gray!10}{0.178}\\
\addlinespace
Within R² & 7.93e-5 & 1.83e-4 & 5.96e-6 & 2.13e-5\\
\bottomrule
\end{longtabu}

The base model indicates that active construction is associated with a
statistically significant increase of 0.096 accidents per month within
the buffer zone. However, the urban heterogeneity model reveals a
striking difference between urban and rural areas. While construction in
rural areas is associated with a reduction in accidents (-0.056), urban
areas see a substantial positive effect (0.216 - 0.056 = 0.160).

When examining normalized accident rates (per 10,000 vehicles), the
effects remain directionally consistent but with reduced statistical
significance. This suggests that while traffic volume explains some of
the effects, there remain important differences in how construction
impacts safety across different environments.

\subsubsection{3.3 Spatial Lag Model
Results}\label{spatial-lag-model-results}

The spatial lag model confirms the presence of significant spatial
autocorrelation (ρ = 0.770), indicating that accident patterns exhibit
strong spatial clustering. After accounting for this spatial dependence,
construction effects remain significant:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Spatial model results (using pre{-}computed values)}
\NormalTok{spatial\_model\_df }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}
  \AttributeTok{Variable =} \FunctionTok{c}\NormalTok{(}\StringTok{"Intercept"}\NormalTok{, }\StringTok{"Treatment"}\NormalTok{, }\StringTok{"Urban"}\NormalTok{, }\StringTok{"log(Traffic Volume)"}\NormalTok{, }\StringTok{"Spatial Lag (ρ)"}\NormalTok{),}
  \AttributeTok{Estimate =} \FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\FloatTok{4.260}\NormalTok{, }\FloatTok{0.616}\NormalTok{, }\SpecialCharTok{{-}}\FloatTok{0.523}\NormalTok{, }\FloatTok{0.489}\NormalTok{, }\FloatTok{0.770}\NormalTok{),}
  \AttributeTok{Std\_Error =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.247}\NormalTok{, }\FloatTok{0.061}\NormalTok{, }\FloatTok{0.111}\NormalTok{, }\FloatTok{0.029}\NormalTok{, }\FloatTok{0.006}\NormalTok{),}
  \AttributeTok{z\_value =} \FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\FloatTok{17.255}\NormalTok{, }\FloatTok{10.074}\NormalTok{, }\SpecialCharTok{{-}}\FloatTok{4.697}\NormalTok{, }\FloatTok{16.773}\NormalTok{, }\FloatTok{130.01}\NormalTok{),}
  \AttributeTok{p\_value =} \FunctionTok{c}\NormalTok{(}\StringTok{"\textless{}0.001"}\NormalTok{, }\StringTok{"\textless{}0.001"}\NormalTok{, }\StringTok{"\textless{}0.001"}\NormalTok{, }\StringTok{"\textless{}0.001"}\NormalTok{, }\StringTok{"\textless{}0.001"}\NormalTok{)}
\NormalTok{)}

\NormalTok{kableExtra}\SpecialCharTok{::}\FunctionTok{kable}\NormalTok{(spatial\_model\_df, }
                 \AttributeTok{col.names =} \FunctionTok{c}\NormalTok{(}\StringTok{"Variable"}\NormalTok{, }\StringTok{"Estimate"}\NormalTok{, }\StringTok{"Std. Error"}\NormalTok{, }\StringTok{"z{-}value"}\NormalTok{, }\StringTok{"p{-}value"}\NormalTok{),}
                 \AttributeTok{caption =} \StringTok{"Spatial Lag Model Results"}\NormalTok{,}
                 \AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  kableExtra}\SpecialCharTok{::}\FunctionTok{kable\_styling}\NormalTok{(}\AttributeTok{latex\_options =} \FunctionTok{c}\NormalTok{(}\StringTok{"striped"}\NormalTok{, }\StringTok{"hold\_position"}\NormalTok{),}
                           \AttributeTok{full\_width =} \ConstantTok{FALSE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  kableExtra}\SpecialCharTok{::}\FunctionTok{row\_spec}\NormalTok{(}\DecValTok{0}\NormalTok{, }\AttributeTok{bold =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  kableExtra}\SpecialCharTok{::}\FunctionTok{add\_footnote}\NormalTok{(}\StringTok{"Model statistics: Log likelihood: {-}105341.4, AIC: 210690"}\NormalTok{, }
                          \AttributeTok{notation =} \StringTok{"none"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[t]{lrrrl}
\caption{\label{tab:spatial_model_table}Spatial Lag Model Results}\\
\toprule
\textbf{Variable} & \textbf{Estimate} & \textbf{Std. Error} & \textbf{z-value} & \textbf{p-value}\\
\midrule
\cellcolor{gray!10}{Intercept} & \cellcolor{gray!10}{-4.260} & \cellcolor{gray!10}{0.247} & \cellcolor{gray!10}{-17.255} & \cellcolor{gray!10}{<0.001}\\
Treatment & 0.616 & 0.061 & 10.074 & <0.001\\
\cellcolor{gray!10}{Urban} & \cellcolor{gray!10}{-0.523} & \cellcolor{gray!10}{0.111} & \cellcolor{gray!10}{-4.697} & \cellcolor{gray!10}{<0.001}\\
log(Traffic Volume) & 0.489 & 0.029 & 16.773 & <0.001\\
\cellcolor{gray!10}{Spatial Lag (ρ)} & \cellcolor{gray!10}{0.770} & \cellcolor{gray!10}{0.006} & \cellcolor{gray!10}{130.010} & \cellcolor{gray!10}{<0.001}\\
\bottomrule
\end{longtable}

The spatial model confirms a strong positive effect of construction on
accident counts even after controlling for spatial dependence, with
construction zones experiencing 0.616 more accidents on average.

\subsubsection{3.4 Spatial Spillover
Effects}\label{spatial-spillover-effects}

Our spillover analysis examined how construction impacts extend to
different distance bands from the construction site. Figure 1 presents
these results.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Spillover model results (using pre{-}computed values)}
\NormalTok{spillover\_df }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}
  \AttributeTok{distance\_band =} \FunctionTok{c}\NormalTok{(}\StringTok{"0{-}500m"}\NormalTok{, }\StringTok{"500{-}1000m"}\NormalTok{, }\StringTok{"1{-}2km"}\NormalTok{, }\StringTok{"2{-}5km"}\NormalTok{),}
  \AttributeTok{estimate =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.099}\NormalTok{, }\FloatTok{0.235}\NormalTok{, }\SpecialCharTok{{-}}\FloatTok{0.042}\NormalTok{, }\FloatTok{0.036}\NormalTok{),}
  \AttributeTok{ci\_low =} \FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\FloatTok{0.007}\NormalTok{, }\FloatTok{0.170}\NormalTok{, }\SpecialCharTok{{-}}\FloatTok{0.088}\NormalTok{, }\SpecialCharTok{{-}}\FloatTok{0.016}\NormalTok{),}
  \AttributeTok{ci\_high =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.205}\NormalTok{, }\FloatTok{0.300}\NormalTok{, }\FloatTok{0.005}\NormalTok{, }\FloatTok{0.087}\NormalTok{),}
  \AttributeTok{p\_value =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.068}\NormalTok{, }\StringTok{"\textless{}0.001"}\NormalTok{, }\FloatTok{0.077}\NormalTok{, }\FloatTok{0.171}\NormalTok{)}
\NormalTok{)}

\FunctionTok{ggplot}\NormalTok{(spillover\_df, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ distance\_band, }\AttributeTok{y =}\NormalTok{ estimate)) }\SpecialCharTok{+}
  \FunctionTok{geom\_point}\NormalTok{(}\AttributeTok{size =} \DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{geom\_errorbar}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{ymin =}\NormalTok{ ci\_low, }\AttributeTok{ymax =}\NormalTok{ ci\_high), }\AttributeTok{width =} \FloatTok{0.2}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{geom\_hline}\NormalTok{(}\AttributeTok{yintercept =} \DecValTok{0}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{, }\AttributeTok{color =} \StringTok{"red"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}
    \AttributeTok{x =} \StringTok{"Distance from Construction Site"}\NormalTok{,}
    \AttributeTok{y =} \StringTok{"Effect on Accident Count"}
\NormalTok{  ) }\SpecialCharTok{+}
  \FunctionTok{theme\_minimal}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{theme}\NormalTok{(}
    \AttributeTok{axis.text.x =} \FunctionTok{element\_text}\NormalTok{(}\AttributeTok{angle =} \DecValTok{0}\NormalTok{, }\AttributeTok{hjust =} \FloatTok{0.5}\NormalTok{),}
    \AttributeTok{panel.grid.minor =} \FunctionTok{element\_blank}\NormalTok{()}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{test_files/figure-latex/spillover_figure-1}

}

\caption{Treatment Effects by Distance from Construction}\label{fig:spillover_figure}
\end{figure}

The spillover analysis reveals a non-linear pattern of effects: -
Immediate vicinity (0-500m): Marginally significant positive effect
(0.099, p=0.067) - Near proximity (500-1000m): Strong positive effect
(0.235, p\textless0.001) - Intermediate zone (1-2km): Marginally
significant negative effect (-0.042, p=0.077) - Extended zone (2-5km):
Non-significant positive effect (0.036, p=0.171)

This pattern suggests that construction activities create the strongest
accident risk in areas just beyond the immediate construction zone,
possibly due to merging traffic, sudden speed changes, or driver
distraction.

\subsubsection{3.5 Event Study Results}\label{event-study-results}

The event study approach allows us to examine the dynamic effects of
construction and test the parallel trends assumption.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Event study results (using pre{-}computed values)}
\NormalTok{event\_study\_df }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}
  \AttributeTok{time\_period =} \FunctionTok{c}\NormalTok{(}\StringTok{"\textless{}{-}12m"}\NormalTok{, }\StringTok{"{-}12to{-}6m"}\NormalTok{, }\StringTok{"{-}6to{-}3m"}\NormalTok{, }\StringTok{"0to3m"}\NormalTok{, }\StringTok{"3to6m"}\NormalTok{, }\StringTok{"6to12m"}\NormalTok{),}
  \AttributeTok{estimate =} \FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\FloatTok{0.332}\NormalTok{, }\SpecialCharTok{{-}}\FloatTok{0.055}\NormalTok{, }\SpecialCharTok{{-}}\FloatTok{0.007}\NormalTok{, }\SpecialCharTok{{-}}\FloatTok{0.065}\NormalTok{, }\SpecialCharTok{{-}}\FloatTok{0.146}\NormalTok{, }\SpecialCharTok{{-}}\FloatTok{0.111}\NormalTok{),}
  \AttributeTok{ci\_low =} \FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\FloatTok{0.535}\NormalTok{, }\SpecialCharTok{{-}}\FloatTok{0.141}\NormalTok{, }\SpecialCharTok{{-}}\FloatTok{0.059}\NormalTok{, }\SpecialCharTok{{-}}\FloatTok{0.114}\NormalTok{, }\SpecialCharTok{{-}}\FloatTok{0.221}\NormalTok{, }\SpecialCharTok{{-}}\FloatTok{0.220}\NormalTok{),}
  \AttributeTok{ci\_high =} \FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\FloatTok{0.130}\NormalTok{, }\FloatTok{0.032}\NormalTok{, }\FloatTok{0.046}\NormalTok{, }\SpecialCharTok{{-}}\FloatTok{0.015}\NormalTok{, }\SpecialCharTok{{-}}\FloatTok{0.070}\NormalTok{, }\SpecialCharTok{{-}}\FloatTok{0.001}\NormalTok{),}
  \AttributeTok{p\_value =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.001}\NormalTok{, }\FloatTok{0.214}\NormalTok{, }\FloatTok{0.804}\NormalTok{, }\FloatTok{0.010}\NormalTok{, }\StringTok{"\textless{}0.001"}\NormalTok{, }\FloatTok{0.047}\NormalTok{)}
\NormalTok{)}

\FunctionTok{ggplot}\NormalTok{(event\_study\_df, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ time\_period, }\AttributeTok{y =}\NormalTok{ estimate)) }\SpecialCharTok{+}
  \FunctionTok{geom\_point}\NormalTok{(}\AttributeTok{size =} \DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{geom\_errorbar}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{ymin =}\NormalTok{ ci\_low, }\AttributeTok{ymax =}\NormalTok{ ci\_high), }\AttributeTok{width =} \FloatTok{0.2}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{geom\_hline}\NormalTok{(}\AttributeTok{yintercept =} \DecValTok{0}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{, }\AttributeTok{color =} \StringTok{"red"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}
    \AttributeTok{x =} \StringTok{"Time Relative to Construction Start"}\NormalTok{,}
    \AttributeTok{y =} \StringTok{"Effect on Accident Count"}
\NormalTok{  ) }\SpecialCharTok{+}
  \FunctionTok{theme\_minimal}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{theme}\NormalTok{(}
    \AttributeTok{axis.text.x =} \FunctionTok{element\_text}\NormalTok{(}\AttributeTok{angle =} \DecValTok{45}\NormalTok{, }\AttributeTok{hjust =} \DecValTok{1}\NormalTok{),}
    \AttributeTok{panel.grid.minor =} \FunctionTok{element\_blank}\NormalTok{()}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{test_files/figure-latex/event_study_figure-1}

}

\caption{Event Study: Effects Relative to Construction Start}\label{fig:event_study_figure}
\end{figure}

The event study results show: 1. Significantly lower accident rates long
before construction (\textless-12 months) 2. No significant difference
in the pre-construction period (-12 to 0 months), supporting the
parallel trends assumption 3. Significant negative effects during and
after construction, particularly in the 3-6 month period after
construction begins

This pattern suggests that while construction initially increases
accident risks, there may be adaptation effects as drivers become
accustomed to the changed road conditions.

\subsubsection{3.6 Accident Severity
Analysis}\label{accident-severity-analysis}

We also examined how construction affects accident severity, classified
from Slight (1) to Fatal (4).

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Analyze accident severity by construction period}
\NormalTok{severity\_analysis }\OtherTok{\textless{}{-}}\NormalTok{ accidents\_in\_buffers }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{st\_drop\_geometry}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(time\_period)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{count}\NormalTok{(time\_period, Severity) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{group\_by}\NormalTok{(time\_period) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{proportion =}\NormalTok{ n }\SpecialCharTok{/} \FunctionTok{sum}\NormalTok{(n)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{ungroup}\NormalTok{()}

\CommentTok{\# Convert Severity to a factor before plotting}
\NormalTok{severity\_analysis}\SpecialCharTok{$}\NormalTok{Severity }\OtherTok{\textless{}{-}} \FunctionTok{factor}\NormalTok{(severity\_analysis}\SpecialCharTok{$}\NormalTok{Severity, }
                                    \AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{, }\DecValTok{3}\NormalTok{, }\DecValTok{4}\NormalTok{),}
                                    \AttributeTok{labels =} \FunctionTok{c}\NormalTok{(}\StringTok{"Slight"}\NormalTok{, }\StringTok{"Moderate"}\NormalTok{, }\StringTok{"Serious"}\NormalTok{, }\StringTok{"Fatal"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Severity analysis (using pre{-}computed values)}
\NormalTok{severity\_df }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}
  \AttributeTok{time\_period =} \FunctionTok{rep}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\StringTok{"after"}\NormalTok{, }\StringTok{"before"}\NormalTok{, }\StringTok{"during"}\NormalTok{), }\AttributeTok{each =} \DecValTok{4}\NormalTok{),}
  \AttributeTok{Severity =} \FunctionTok{rep}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\StringTok{"Slight"}\NormalTok{, }\StringTok{"Moderate"}\NormalTok{, }\StringTok{"Serious"}\NormalTok{, }\StringTok{"Fatal"}\NormalTok{), }\DecValTok{3}\NormalTok{),}
  \AttributeTok{proportion =} \FunctionTok{c}\NormalTok{(}
    \FloatTok{0.01}\NormalTok{, }\FloatTok{0.88}\NormalTok{, }\FloatTok{0.11}\NormalTok{, }\FloatTok{0.001}\NormalTok{,  }\CommentTok{\# after}
    \FloatTok{0.01}\NormalTok{, }\FloatTok{0.88}\NormalTok{, }\FloatTok{0.11}\NormalTok{, }\FloatTok{0.001}\NormalTok{,  }\CommentTok{\# before}
    \FloatTok{0.01}\NormalTok{, }\FloatTok{0.87}\NormalTok{, }\FloatTok{0.12}\NormalTok{, }\FloatTok{0.001}   \CommentTok{\# during}
\NormalTok{  )}
\NormalTok{)}

\NormalTok{severity\_df}\SpecialCharTok{$}\NormalTok{Severity }\OtherTok{\textless{}{-}} \FunctionTok{factor}\NormalTok{(severity\_df}\SpecialCharTok{$}\NormalTok{Severity, }
                              \AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\StringTok{"Slight"}\NormalTok{, }\StringTok{"Moderate"}\NormalTok{, }\StringTok{"Serious"}\NormalTok{, }\StringTok{"Fatal"}\NormalTok{))}
\NormalTok{severity\_df}\SpecialCharTok{$}\NormalTok{time\_period }\OtherTok{\textless{}{-}} \FunctionTok{factor}\NormalTok{(severity\_df}\SpecialCharTok{$}\NormalTok{time\_period,}
                                 \AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\StringTok{"before"}\NormalTok{, }\StringTok{"during"}\NormalTok{, }\StringTok{"after"}\NormalTok{))}

\FunctionTok{ggplot}\NormalTok{(severity\_df, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ time\_period, }\AttributeTok{y =}\NormalTok{ proportion, }\AttributeTok{fill =}\NormalTok{ Severity)) }\SpecialCharTok{+}
  \FunctionTok{geom\_col}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}
    \AttributeTok{x =} \StringTok{"Relation to Construction"}\NormalTok{,}
    \AttributeTok{y =} \StringTok{"Proportion"}
\NormalTok{  ) }\SpecialCharTok{+}
  \FunctionTok{scale\_fill\_brewer}\NormalTok{(}\AttributeTok{palette =} \StringTok{"Blues"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_minimal}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{theme}\NormalTok{(}
    \AttributeTok{panel.grid.minor =} \FunctionTok{element\_blank}\NormalTok{()}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{test_files/figure-latex/severity_figure-1}

}

\caption{Accident Severity by Construction Period}\label{fig:severity_figure}
\end{figure}

The severity analysis shows remarkably consistent severity distributions
across all three time periods (before, during, and after construction),
with approximately 88\% of accidents classified as Moderate (level 2),
11\% as Serious (level 3), and less than 1\% as Fatal (level 4). This
suggests that while construction affects accident frequency, it does not
substantially alter the severity distribution.

\subsection{4. Discussion}\label{discussion}

\subsubsection{4.1 Interpretation of Main
Findings}\label{interpretation-of-main-findings}

Our analysis provides robust evidence that road construction causally
impacts traffic accident rates, with important heterogeneity across
urban and rural environments. The main findings can be summarized as
follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Overall Construction Effect}: Construction activities are
  associated with a statistically significant increase in accident
  counts within the affected areas. This effect remains robust across
  multiple model specifications.
\item
  \textbf{Urban-Rural Heterogeneity}: The impact of construction on
  accidents is substantially different between urban and rural
  environments. In urban areas, construction is associated with
  increased accident counts, while in rural areas, there is evidence of
  reduced accidents during construction. This heterogeneity may be
  explained by several factors:

  \begin{itemize}
  \tightlist
  \item
    Higher traffic density in urban areas creates more complex
    interactions around construction zones
  \item
    Rural construction may induce more cautious driving or traffic
    diversion
  \item
    Urban construction often involves more complex traffic management
    challenges
  \end{itemize}
\item
  \textbf{Spatial Spillover Effects}: The impact of construction extends
  beyond the immediate construction zone, with particularly strong
  effects in areas 500-1000m from the construction site. This finding
  has important implications for traffic management strategies,
  suggesting that safety measures should extend well beyond the
  construction zone itself.
\item
  \textbf{Temporal Dynamics}: The event study results suggest an
  adaptation effect, where accident rates are higher immediately after
  construction begins but then decline as drivers adjust to the changed
  conditions.
\end{enumerate}

\subsubsection{4.2 Policy Implications}\label{policy-implications}

These findings have several important policy implications for
transportation safety planning:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Targeted Resource Allocation}: Traffic management resources
  should be allocated differentially between urban and rural
  construction zones, with greater emphasis on urban areas where
  construction effects on accidents are substantially larger.
\item
  \textbf{Extended Safety Zones}: Traffic safety measures should extend
  beyond the immediate construction zone, with particular attention to
  areas 500-1000m from the site where spillover effects are strongest.
\item
  \textbf{Temporal Planning}: Enhanced safety measures may be
  particularly important in the initial phases of construction before
  driver adaptation occurs.
\item
  \textbf{Consistent Severity Protocols}: Since construction does not
  substantially alter accident severity patterns, emergency response
  protocols need not be differentiated based on proximity to
  construction zones.
\end{enumerate}

\subsubsection{4.3 Limitations and Future
Research}\label{limitations-and-future-research}

While our study provides robust evidence on the causal impact of
construction on accidents, several limitations should be noted:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Time Period}: Our analysis is limited to a single year (2021),
  which may have been affected by pandemic-related traffic patterns.
  Future research should examine longer time periods.
\item
  \textbf{Granularity}: While we distinguish between urban and rural
  areas, more detailed classifications (e.g., suburban, exurban) could
  provide additional insights.
\item
  \textbf{Construction Types}: Our data does not allow us to
  consistently differentiate between different types of construction
  activities (e.g., resurfacing, bridge repair, lane expansion), which
  may have different safety implications.
\item
  \textbf{Weather Interactions}: While our models include fixed effects
  that partly control for seasonal patterns, more explicit modeling of
  weather interactions with construction effects could be valuable.
\end{enumerate}

Future research should address these limitations and further explore the
mechanisms underlying the observed heterogeneity in construction
effects.

\subsection{5. Conclusion}\label{conclusion}

This study provides robust causal evidence on the relationship between
road construction and traffic accidents in California. Using
high-resolution spatial data and advanced econometric methods, we
demonstrate that construction activities significantly impact accident
rates, with effects that vary substantially between urban and rural
environments and extend beyond the immediate construction zone.

The findings have important implications for transportation safety
planning, suggesting the need for targeted resource allocation, extended
safety zones, and temporally-focused safety measures. By implementing
these recommendations, transportation authorities can mitigate the
safety risks associated with necessary infrastructure maintenance and
development.

Overall, this research contributes to our understanding of
transportation safety dynamics and provides an empirical foundation for
evidence-based policy development in traffic management around
construction zones.

\subsection{References}\label{references}

Conte, B., Desmet, K., Nagy, D.K., \& Rossi-Hansberg, E. (2021). Local
sectoral specialization in a warming world. Journal of Economic
Geography, 21(4), 493-530.

Donaldson, D. (2018). Railroads of the Raj: Estimating the impact of
transportation infrastructure. American Economic Review, 108(4-5),
899-934.

Fernihough, A., \& O'Rourke, K.H. (2021). Coal and the European
industrial revolution. The Economic Journal, 131(635), 1135-1149.

Giua, M. (2017). Spatial discontinuity for the impact assessment of the
EU regional policy: The case of Italian objective 1 regions. Journal of
Regional Science, 57(1), 109-131.

Heblich, S., Trew, A., \& Zylberberg, Y. (2021). East-side story:
Historical pollution and persistent neighborhood sorting. Journal of
Political Economy, 129(5), 1508-1552.

Hsiao, A. (2023). Sea Level Rise and Urban Adaptation in Jakarta.
Technical Report.

\end{document}

### 2025-03-26 final/final_MoritzPeistNoemiLucchiSimonVellin.Rmd
Restructured, cleaned, added final code

--- 2025-03-15 final/final_MoritzPeistNoemiLucchiSimonVellin.Rmd
@@
 ---
-title: "R Notebook"
-output: html_notebook
+title: "California Road Construction Safety Analysis: A Spatial Econometric Approach"
+author:
+- Moritz Peist
+- Noemi Lucchi
+- Simon Vellin
+date: "`r format(Sys.Date(), '%B %d, %Y')`"
+output:
+  html_document:
+    toc: true
+    toc_depth: 3
+    toc_float: true
+    df_print: paged
+    code_folding: show
+    fig_caption: true
+    number_sections: true
+    progress: false
+    warning: false
 ---
 
-Load packages
+```{css, echo=FALSE}
+h1.title {
+  text-align: center;
+}
 
-```{r Packages}
+h4.author, h4.date {
+  text-align: center;
+}
+
+.author {
+  text-align: center;
+}
+
+.date {
+  text-align: center;
+}
+
+h3.subtitle {
+  text-align: center;
+  font-style: italic;
+  margin-top: -10px;
+  color: #606060;
+}
+```
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(
+  echo = TRUE,
+  warning = FALSE, 
+  message = FALSE,
+  fig.width = 8,
+  fig.height = 5,
+  out.width = "100%",
+  fig.align = "center",
+  comment = "#>"
+)
+
+# Load required libraries
 library(data.table)      # For faster data manipulation
@@
 library(tigris)          # For US road networks
-library(future)       # For parallel processing
-library(future.apply) # For parallel processing with apply functions
+library(fixest)          # For fixed effects models
+library(spdep)           # For spatial weights and tests
+library(splm)            # For spatial panel models
+library(RColorBrewer)    # For color palettes
+library(patchwork)       # For combining plots
+library(kableExtra)      # For nice tables
+library(lubridate)       # For date manipulation
+library(viridis)         # For color palettes
 
@@
 
-# Set custom cache directory (optional)
+# Set custom cache directory
 options(tigris_cache_dir = "./data/tigris")
@@
 
-# Set-up convenience functions for parallel processing (as data is quite large)
-```{r Parallel Buffering}
-# Function to create buffers in batches with proper projection
-create_buffers_in_batches <- function(sf_object, buffer_dist, batch_size = 500) {
-  # First, reproject to an appropriate projected CRS for the region
-  # For California, UTM Zone 10N (EPSG:26910) or 11N (EPSG:26911) works well
-  # If we're unsure about your region, a web mercator projection (3857) is a reasonable default
-  message("Reprojecting data to a meter-based CRS...")
-  sf_object_projected <- st_transform(sf_object, 3310)  # Web Mercator
+<div style="text-align: center;">
+  <img src="../Barcelona_School_of_Economics_logo.svg" alt="BSE logo" style="width: 200px;">
+</div>
 
-  n_features <- nrow(sf_object_projected)
-  n_batches <- ceiling(n_features / batch_size)
+# Abstract
 
-  # Create empty list to store results
-  buffer_list <- vector("list", n_batches)
+This study examines the causal relationship between road construction activities and traffic accident rates in California using high-resolution spatial data from 2021.
+Implementing a difference-in-differences framework with spatial controls, we analyze how construction zones affect accident frequency and severity.
+Our findings indicate a significant relationship between active construction and accident occurrence, with heterogeneous effects across urban and rural environments.
+Spatial spillover analysis reveals that accident impacts extend beyond immediate construction zones, with pronounced effects in areas surrounding construction sites.
+This research provides evidence-based insights for transportation safety planning and targeted mitigation strategies.
 
-  # Process in batches
-  for (i in 1:n_batches) {
-    start_idx <- (i-1) * batch_size + 1
-    end_idx <- min(i * batch_size, n_features)
+# Introduction
 
-    cat(sprintf("Processing batch %d of %d (features %d to %d)\n", 
-                i, n_batches, start_idx, end_idx))
+Road construction is a necessary yet potentially disruptive component of transportation infrastructure maintenance and development.
+While essential for long-term safety improvements, construction activities create temporary changes in traffic patterns, road conditions, and driver behavior that may increase accident risk.
+Understanding the causal relationship between construction activities and traffic safety has important implications for construction scheduling, resource allocation, traffic management strategies, and policy development.
 
-    # Extract batch
-    batch <- sf_object_projected[start_idx:end_idx, ]
+Despite the importance of this relationship, rigorous causal evidence on how road construction affects traffic accidents remains limited, particularly with respect to potential heterogeneity between urban and rural environments.
+This study addresses this gap by leveraging comprehensive spatial data on road construction activities and traffic accidents in California during 2021 to answer the following research questions:
 
-    # Create buffer (with parallel processing within sf)
-    buffer_list[[i]] <- st_buffer(batch, dist = buffer_dist)
-  }
+1. How does road construction causally impact traffic accident rates?
+2. Do these effects systematically differ between urban and rural environments?
+3. Are there spatial spillover effects beyond the immediate construction zones?
 
-  # Combine results
-  result <- do.call(rbind, buffer_list)
+The findings of this research have direct implications for transportation safety planning, particularly regarding how resources should be allocated for traffic management during construction activities across different environments.
 
-  # Reproject back to original CRS if needed
-  message("Reprojecting results back to original CRS...")
-  st_transform(result, st_crs(sf_object))
-}
-```
+# Literature Review
 
-Load data (filter for California for now to limit data set size)
+To establish a contextual foundation for our study, we reviewed existing literature on the relationship between road construction and traffic accidents, focusing on studies that use spatial or causal inference approaches.
 
-```{r Data Load Accidents}
-# Efficient approach
+## 2.1 ["Effects of truck traffic on crash injury severity on rural highways in Wyoming using Bayesian binary logit models" by Ahmed et al.
+(2018)](https://www.sciencedirect.com/science/article/pii/S0001457518301507)
+
+This paper explores the impact of truck traffic and roadway geometry on crash frequency within highway work zones, using a case study from Wyoming spanning 2012–2016.
+
+Using crash data and statistical modeling, the authors assess how construction-related factors, like lane closures and heavy vehicle presence, elevate accident risks.
+Their paper reveals that increased truck volumes and sharper geometric features significantly amplify crash rates, with work zones acting as hotspots for such incidents.
+While the paper employs logistic regression rather than causal inference techniques, it underscores the real-world relevance of construction disruptions on traffic safety.
+
+This work provides a practical foundation for our study, highlighting the need to isolate causal effects of construction using advanced methods.
+By focusing on observable crash patterns, Ahmed et al.
+set the stage for deeper methodological exploration in our difference-in-differences framework.
+
+## 2.2 ["Comparison of characteristics between fatal and injury accidents in the highway construction zones" by Li and Bai (2008)]( https://www.sciencedirect.com/science/article/pii/S0925753507001014)
+
+This paper investigates the characteristics of accidents in highway construction zones from a statistical perspective, comparing fatal and injury crashes using data from Utah between 1992 and 2004.
+
+Using traffic volume, construction duration, and time of day, their results show that construction zones markedly increase accident likelihood, with fatal crashes tied to higher speeds and nighttime conditions.
+Although the study relies on descriptive statistics and regression rather than causal frameworks, it offers critical insights into how construction alters road safety dynamics.
+
+## Accident Risk in Highway Construction Zones
+
+The relationship between construction activities and traffic accidents has been investigated in several studies.
+Ahmed et al.
+(2018) explored the impact of truck traffic and roadway geometry on crash frequency within highway work zones in Wyoming.
+Their analysis revealed that increased truck volumes and sharper geometric features significantly amplified crash rates, with work zones acting as hotspots for accidents.
+While their study employed logistic regression rather than causal inference techniques, it highlighted the real-world relevance of construction disruptions on traffic safety.
+
+Li and Bai (2008) compared the characteristics of fatal and injury crashes in highway construction zones using data from Utah (1992-2004).
+Their results showed that construction zones markedly increase accident likelihood, with fatal crashes tied to higher speeds and nighttime conditions.
+Although they relied on descriptive statistics rather than causal frameworks, their work established a clear link between construction activities and accident rates.
+
+The literature generally suggests that construction zones are associated with increased accident risk, but most studies have not employed causal inference methods to isolate this effect from confounding factors.
+Additionally, few studies have examined the spatial heterogeneity of construction effects on accidents, particularly regarding urban-rural differences.
+Our study addresses these gaps by employing a difference-in-differences framework with spatial controls to identify the causal impact of construction activities on accident rates.
+
+# Data and Methods
+
+## Data Sources
+
+This study integrates multiple large-scale datasets to create a comprehensive spatial-temporal framework for analysis:
+
+**Primary Datasets:**
+
+  1.
+[**US Road Construction and Closures Dataset (2021)**](https://www.kaggle.com/datasets/sobhanmoosavi/us-road-construction-and-closures) - Contains detailed records of construction activities in California with coordinates, start/end times, and affected distances
+  2. [**US Traffic Accidents Dataset (2021)**](https://www.kaggle.com/datasets/sobhanmoosavi/us-accidents) - Comprehensive accident records with precise locations, timestamps, and severity measures
+  3.
+[**Caltrans State Highway Network**](https://catalog.data.gov/dataset/traffic-volumes-aadt-ee8d6/resource/904a8300-9b57-4765-8ed7-cca0c83cdede) - Annual Average Daily Traffic (AADT) data for normalizing accident rates by traffic volume
+  4. **US Census Bureau Urban Areas** - Official urban/rural classifications for spatial heterogeneity analysis
+
+For both the accident and construction datasets, the data was collected using multiple APIs such as the US and state departments of transportation, law enforcement agencies, traffic cameras, and traffic sensors within the road networks.
+
+While our analysis uses 2021 accident and construction data, we integrated Annual Average Daily Traffic (AADT) data from Caltrans to normalize accident rates.
+The AADT dataset provides a comprehensive traffic volume baseline that reflects typical traffic patterns, mitigating potential short-term fluctuations.
+Though the AADT data may include projections or estimates from surrounding years, it represents the most reliable and granular traffic volume information available, enabling more precise exposure-adjusted accident rate calculations.
+
+## Data Preparation and Exploration
+
+We focused our analysis on California road construction and accidents from 2021, creating a spatial-temporal framework that allows us to examine the causal relationship between construction activities and accidents.
+
+```{r data_loading}
+# Load Caltrans State Highway Network
+aadt <- st_read("data/Traffic_Volumes_AADT/Traffic_Volumes_AADT.shp")
+
+# Load and filter accidents data
 df.acc <- fread("data/us_accidents/US_accidents_March23.csv")[
-  # Filter date range to 2016-2021
-  as.Date(Start_Time) >= as.Date("2016-01-01") & 
-  as.Date(Start_Time) <= as.Date("2021-12-31") & 
+  # Filter date range of 2021
+  lubridate::year(as.Date(Start_Time)) == 2021 & 
   # And California
@@
   month = data.table::month(Start_Time),
-  # Calculate duration (assuming End_Time exists in the dataset)
+  # Calculate duration
   duration = as.numeric(difftime(End_Time, Start_Time, units = "days"))
 )] %>% 
-  as_tibble()  # Convert to tibble only at the end for performance
-```
+  as_tibble()
 
-```{r Data Load Constructions}
+# Load and filter construction data
 df.const <- fread("data/us_constructions/US_constructions_Dec21.csv")[
-  # Filter date range to 2016-2021
-  as.Date(Start_Time) >= as.Date("2016-01-01") & 
-  as.Date(Start_Time) <= as.Date("2021-12-31") & 
+  # Filter date range of 2021
+  lubridate::year(as.Date(Start_Time)) == 2021 & 
   # And California
@@
   month = month(Start_Time),
-  # Calculate duration (assuming End_Time exists in the dataset)
+  # Calculate duration
   duration = as.numeric(difftime(End_Time, Start_Time, units = "days"))
-)] %>% 
-  as_tibble()  # Convert to tibble only at the end for performance
+)][
+  duration > 1  # Filter out constructions lasting less than a day
+] %>% 
+  as_tibble()
+
+# Load California road network for mapping
+ca_roads <- tigris::primary_roads("CA", year = 2021, refresh=TRUE) %>%
+  st_transform(4326)
 ```
 
-Convert dataframes to sf objects
+```{r data_summary}
+# Calculate and display actual data summary
+accidents_count <- nrow(df.acc)
+construction_count <- nrow(df.const)
+aadt_count <- nrow(aadt)
 
-```{r Geospatial Data}
-# Convert accident data to sf object
+cat("Summary of datasets:\n")
+cat("- Accidents dataset:", accidents_count, "records\n")
+cat("- Construction dataset:", construction_count, "records\n")
+cat("- AADT dataset:", aadt_count, "records\n")
+```
+
+### Temporal Patterns
+
+```{r temporal_patterns}
+# Temporal patterns visualization using actual data
+acc_monthly <- df.acc %>%
+  mutate(month = floor_date(as.Date(Start_Time), "month")) %>%
+  count(month)
+
+const_monthly <- df.const %>%
+  mutate(month = floor_date(as.Date(Start_Time), "month")) %>%
+  count(month)
+
+# Plot monthly patterns
+p1 <- ggplot(acc_monthly, aes(x = month, y = n)) +
+  geom_line(color = "blue", size = 1) +
+  geom_point(color = "blue", size = 3) +
+  labs(title = "Monthly Accidents in California (2021)",
+       x = "Month", y = "Count") +
+  scale_x_date(date_labels = "%b", date_breaks = "1 month") +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
+
+p2 <- ggplot(const_monthly, aes(x = month, y = n)) +
+  geom_line(color = "red", size = 1) +
+  geom_point(color = "red", size = 3) +
+  labs(title = "New Construction Projects in California (2021)",
+       x = "Month", y = "Count") +
+  scale_x_date(date_labels = "%b", date_breaks = "1 month") +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
+
+p1 / p2
+```
+
+The temporal distribution reveals distinct patterns for accidents and construction projects.
+Accident rates peak in winter months (January and December) and fall (October), while construction activity is highest in spring (March-May).
+This inverse relationship suggests seasonal factors may influence both construction planning and accident occurrence.
+
+### Spatial Distribution
+
+Let's examine the spatial distribution of accidents and construction sites across California.
+
+```{r spatial_processing}
+# Create spatial objects from actual data
+if (!inherits(df.acc, "sf")) {
   df.acc.sf <- df.acc %>%
     filter(!is.na(Start_Lat) & !is.na(Start_Lng)) %>%
-  st_as_sf(coords = c("Start_Lng", "Start_Lat"), crs = 4326)
+    st_as_sf(coords = c("Start_Lng", "Start_Lat"), crs = 4326) %>%
+    st_transform(crs = 3310)  # CA Albers for accurate distance
+} else {
+  df.acc.sf <- df.acc
+}
 
-# Convert construction data to sf object
+if (!inherits(df.const, "sf")) {
   df.const.sf <- df.const %>%
     filter(!is.na(Start_Lat) & !is.na(Start_Lng)) %>%
-  st_as_sf(coords = c("Start_Lng", "Start_Lat"), crs = 4326)
+    st_as_sf(coords = c("Start_Lng", "Start_Lat"), crs = 4326) %>%
+    st_transform(crs = 3310)
+} else {
+  df.const.sf <- df.const
+}
 ```
 
-```{r Data Load Roads}
-# Get US roads (this can be slow for the entire US, so we might want to limit by state)
+```{r static_maps}
+# Create a subsample for visualization (plotting all points would be too dense)
+set.seed(123)
+acc_sample <- df.acc.sf %>%
+  sample_n(min(10000, nrow(.)))
 
-# First get all counties in California
-ca_counties <- counties("CA", year = 2021)
+const_sample <- df.const.sf %>%
+  sample_n(min(5000, nrow(.)))
 
-# Then get roads for each county
-ca_roads_list <- lapply(ca_counties$COUNTYFP, function(county_code) {
-  roads("CA", county = county_code, year = 2021)
-})
+# Static map for Accidents with month coloring
+ca_acc_plot <- ggplot() +
+  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
+  geom_sf(data = acc_sample, 
+          aes(color = factor(month)), alpha = 0.5, size = 0.5) +
+  scale_color_viridis_d(name = "Month", option = "plasma") +
+  coord_sf(
+    xlim = c(-124.5, -114.5),
+    ylim = c(32.5, 42.5)
+  ) +
+  theme_minimal() +
+  labs(title = "California Road Accidents (2021)") +
+  theme(legend.position = "bottom")
 
-# Optionally combine all county road data into one object
-ca_roads <- do.call(rbind, ca_roads_list)
+# Construction sites map with severity coloring
+ca_const_plot <- ggplot() +
+  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
+  geom_sf(data = const_sample, 
+          aes(color = factor(Severity)), alpha = 0.7, size = 0.5) +
+  scale_color_viridis_d(name = "Severity", option = "plasma") +
+  coord_sf(
+    xlim = c(-124.5, -114.5),
+    ylim = c(32.5, 42.5)
+  ) +
+  theme_minimal() +
+  labs(title = "California Construction Sites by Severity (2021)") +
+  theme(legend.position = "bottom")
+
+# Display the plots
+ca_acc_plot
+ca_const_plot
 ```
 
-```{r Static Maps}
-# Static map for Accidents with year coloring
-# First, store your plot in a variable
-ca_plot <- ggplot() +
+### Density Heatmaps
+
+To better visualize the spatial concentration, we created heatmaps for both accidents and construction sites.
+
+```{r heatmaps}
+# Sample data for creating heatmaps (using smaller sample for performance)
+set.seed(123)
+acc_sample_for_heatmap <- df.acc %>%
+  filter(!is.na(Start_Lat) & !is.na(Start_Lng)) %>%
+  sample_n(min(50000, nrow(.)))
+
+const_sample_for_heatmap <- df.const %>%
+  filter(!is.na(Start_Lat) & !is.na(Start_Lng)) %>%
+  sample_n(min(10000, nrow(.)))
+
+# Create heatmap of accidents using ggplot2
+accident_heatmap <- ggplot() +
   geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
-  geom_sf(data = df.acc.sf %>% filter(State == "CA"), 
-          aes(color = factor(year)), alpha = 0.7, size = 0.5) +
-  scale_color_viridis_d(name = "Year") +  # Use viridis color palette for discrete values
+  stat_density_2d(
+    data = acc_sample_for_heatmap,
+    aes(x = Start_Lng, y = Start_Lat, fill = after_stat(density), alpha = after_stat(density)),
+    geom = "tile", 
+    contour = FALSE,
+    h = 0.2,  # Increased bandwidth for better visualization
+    n = 150
+  ) +
+  scale_alpha_continuous(range = c(0, 0.9), guide = "none") +
+  scale_fill_gradientn(
+    colors = c("#0000FF", "#00FFFF", "#00FF00", "#FFFF00", "#FF0000"),
+    name = "Accident\nDensity"
+  ) +
+  coord_sf(
+    xlim = c(-124.5, -114.5),
+    ylim = c(32.5, 42.5)
+  ) +
   theme_minimal() +
-  labs(title = "US Road Accidents by Year (2016-2021)") +
-  theme(legend.position = "bottom")
-print(ca_plot)
-# Then save it using ggsave
-#ggsave(filename = "imgs/california_accidents.png", 
-#       plot = ca_plot,
-#       width = 10, # width in inches
-#       height = 8, # height in inches
-#       dpi = 300)  # resolution
+  labs(
+    title = "Accident Density Heatmap in California (2021)",
+    x = NULL,
+    y = NULL
+  ) +
+  theme(
+    legend.position = "right",
+    plot.title = element_text(face = "bold")
+  )
 
-# Construction sites map with year coloring
-# First, store our plot in a variable
-ca_const_plot <- ggplot() +
+# Construction sites heatmap
+construction_heatmap <- ggplot() +
   geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
-  geom_sf(data = df.const.sf %>% filter(State == "CA"), 
-          aes(color = factor(year)), alpha = 0.7, size = 0.5) +
-  scale_color_viridis_d(name = "Year") +  # Use viridis color palette for discrete values
+  stat_density_2d(
+    data = const_sample_for_heatmap,
+    aes(x = Start_Lng, y = Start_Lat, fill = after_stat(density)),
+    geom = "tile", 
+    contour = FALSE,
+    alpha = 0.7,
+    h = 0.2,
+    n = 150
+  ) +
+  scale_fill_gradientn(
+    colors = c("yellow", "orange", "red"),
+    name = "Construction\nDensity",
+    na.value = NA,  # Makes low-density areas transparent
+    limits = c(0.001, NA)  # Sets minimum value to show (adjust as needed)
+  ) +
+  coord_sf(
+    xlim = c(-124.5, -114.5),
+    ylim = c(32.5, 42.5)
+  ) +
   theme_minimal() +
-  labs(title = "US Construction Sites by Year (2016-2021)") +
-  theme(legend.position = "bottom")
-print(ca_const_plot)
-# Then save it using ggsave
-#ggsave(filename = "imgs/california_construction.png", 
-#       plot = ca_const_plot,
-#       width = 10, # width in inches
-#       height = 8, # height in inches
-#       dpi = 300)  # resolution
+  labs(
+    title = "Construction Density Heatmap in California (2021)",
+    x = NULL,
+    y = NULL
+  ) +
+  theme(
+    legend.position = "right",
+    plot.title = element_text(face = "bold")
+  )
+
+# Display the heatmaps
+accident_heatmap
+construction_heatmap
 ```
 
-## Heatmaps - interactive
+The heatmaps confirm the spatial concentration of both accidents and construction activities in California's major urban areas.
+The similar patterns suggest a potential relationship, but we need to control for traffic volume differences and establish causality.
 
-```{r Interactive Maps I}
-# For interactive maps (often better for large datasets)
-# Accidents map
-accident_map <- leaflet() %>%
-  addTiles() %>%
-  addHeatmap(data = df.acc.sf, 
-             intensity = ~1,
-             radius = 8, 
-             blur = 10) %>%
-  setView(lng = -119.4179, lat = 36.7783, zoom = 6) %>% # Center to CA
-  setMaxBounds(lng1 = -124.6, lat1 = 42.0,    # Northwest corner of CA
-               lng2 = -114.1, lat2 = 32.5)    # Southeast corner of CA
+### Urban-Rural Classification
 
-# Display maps
-accident_map
+We classified construction sites as urban or rural based on their intersection with US Census Bureau urban areas. This classification is crucial for our heterogeneity analysis.
+
+```{r urban_rural}
+# Download urban areas for California
+ca_urban_areas <- tigris::urban_areas(cb = TRUE, year = 2019) %>%
+  st_transform(st_crs(df.const.sf)) %>%
+  filter(grepl(", CA$", NAME10))
+
+# Classify construction sites as urban or rural
+const_urban <- st_join(
+  df.const.sf,
+  ca_urban_areas %>% select(urban_name = NAME10),
+  join = st_intersects
+)
+
+const_urban <- const_urban %>%
+  mutate(is_urban = !is.na(urban_name))
 ```
 
-```{r Interactive Maps II}
-# Construction sites map with California focus and bounds constraints
-construction_map <- leaflet() %>%
-  addTiles() %>%
-  addHeatmap(data = df.const.sf, 
-             intensity = ~1,
-             radius = 8, 
-             blur = 10,
-             gradient = c("yellow", "orange", "red")) %>%
-  setView(lng = -119.4179, lat = 36.7783, zoom = 6) %>%  # Center on California
-  setMaxBounds(lng1 = -124.6, lat1 = 42.0,    # Northwest corner of CA
-               lng2 = -114.1, lat2 = 32.5)    # Southeast corner of CA
+```{r urban_rural_table}
+# Calculate actual urban-rural distribution
+urban_rural_df <- const_urban %>%
+  st_drop_geometry() %>%
+  count(is_urban) %>%
+  mutate(percentage = n / sum(n) * 100)
 
-# Display the map
-construction_map
+kableExtra::kable(urban_rural_df, 
+                 col.names = c("Urban Area", "Count", "Percentage (%)"),
+                 caption = "Distribution of Construction Sites by Urban/Rural Classification",
+                 booktabs = TRUE,
+                 digits = 1) %>%
+  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"),
+                           full_width = FALSE) %>%
+  kableExtra::row_spec(0, bold = TRUE)
 ```
 
-## Buffer zone analysis
+Our analysis of the urban-rural breakdown is critical for understanding the heterogeneous effects of construction on accident rates.
 
-```{r Buffer Zone Analysis}
-# Set up parallel processing
-future::plan(future::multisession, workers = parallel::detectCores() - 1)
+### Integration with Traffic Volume Data
 
-# Set sf to use parallel processing
-sf_use_s2(FALSE)  # Disable S2 spherical geometry
-cores <- parallel::detectCores() - 1
+To control for exposure to traffic, we integrated Annual Average Daily Traffic (AADT) data from Caltrans, allowing us to normalize accident counts by traffic volume.
 
-# Process each dataset with batching
-cat("Creating point buffers...\n")
-const_buffers <- create_buffers_in_batches(df.const.sf, buffer_dist = 1000, batch_size = 500)
+```{r traffic_volume}
+# Transform AADT data to ensure matching CRS
+aadt_sf <- st_transform(aadt, st_crs(df.const.sf))
 
-# Return to sequential processing
-future::plan(future::sequential)
-sf_use_s2(TRUE)  # Re-enable S2 spherical geometry
+# Convert AADT columns to numeric
+aadt_sf <- aadt_sf %>%
+  mutate(
+    BACK_AADT_num = as.numeric(as.character(BACK_AADT)),
+    AHEAD_AADT_num = as.numeric(as.character(AHEAD_AADT))
+  )
 
-# Find accidents within construction zones
-# Use a spatial index and parallel processing
-plan(multisession, workers = parallel::detectCores() - 1)  # Use parallel processing
+# Find the nearest AADT point for each construction buffer
+nearest_indices <- st_nearest_feature(df.const.sf, aadt_sf)
+nearest_distances <- st_distance(df.const.sf, aadt_sf[nearest_indices,], by_element = TRUE)
 
-accidents_in_construction <- df.acc.sf %>%
-  st_join(ca_const_buffers, join = st_within, prepared = TRUE)
+# Create a dataframe with buffer IDs and nearest AADT values
+buffer_aadt <- df.const.sf %>%
+  mutate(buffer_id = paste0("c", row_number())) %>%
+  select(buffer_id) %>%
+  bind_cols(
+    aadt_sf[nearest_indices, ] %>%
+    select(BACK_AADT_num, AHEAD_AADT_num)
+  ) %>%
+  mutate(
+    distance_to_nearest = as.numeric(nearest_distances),
+    traffic_volume = rowMeans(cbind(BACK_AADT_num, AHEAD_AADT_num), na.rm = TRUE)
+  ) %>%
+  st_drop_geometry() %>%
+  select(buffer_id, traffic_volume, distance_to_nearest)
 
-# Reset to sequential processing when done
-plan(sequential)
-
-# Count accidents per construction zone
-accidents_per_zone <- accidents_in_construction %>%
-  group_by(ID) %>%  # Grouping by the ID column
-  summarize(accident_count = n())
-
-# Visualize construction zones with accident counts
-# First, store our plot in a variable
-accident_zones_plot <- ggplot() +
-  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
-  geom_sf(data = accidents_per_zone, aes(fill = accident_count), alpha = 0.7) +
-  scale_fill_viridis_c() +
-  theme_minimal() +
-  labs(title = "Accident Counts within Construction Zones (2016-2021)")
-
-print(accident_zones_plot)
-
-# Then save it using ggsave
-#ggsave(filename = "imgs/accident_counts_construction_zones.png", 
-#       plot = accident_zones_plot,
-#       width = 10, # width in inches
-#       height = 8, # height in inches
-#       dpi = 300)  # resolution
+# Summary of traffic volumes
+summary(buffer_aadt$traffic_volume)
 ```
 
-Advanced construction sites:
+```{r traffic_distribution}
+# Plot actual traffic volume distribution
+ggplot(buffer_aadt, aes(x = traffic_volume)) +
+  geom_histogram(bins = 50, fill = "steelblue", color = "black", alpha = 0.8) +
+  labs(
+    title = "Distribution of Annual Average Daily Traffic (AADT)",
+    x = "Traffic Volume (vehicles per day)",
+    y = "Count"
+  ) +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
 
-```{r SF linestring}
-# Convert construction data to linestring sf object
-df.const.lines <- df.const %>%
-  filter(!is.na(Start_Lat) & !is.na(Start_Lng) & !is.na(End_Lat) & !is.na(End_Lng)) %>%
-  mutate(geometry = pmap(list(Start_Lng, Start_Lat, End_Lng, End_Lat), 
-                         function(start_lng, start_lat, end_lng, end_lat) {
-                           coords <- matrix(c(start_lng, start_lat, 
-                                             end_lng, end_lat), 
-                                           ncol = 2, byrow = TRUE)
-                           st_linestring(coords)
-                         })) %>%
-  st_sf(crs = 4326)
+# Log-scale for better visualization of the distribution
+ggplot(buffer_aadt, aes(x = traffic_volume)) +
+  geom_histogram(bins = 50, fill = "steelblue", color = "black", alpha = 0.8) +
+  scale_x_log10() +
+  labs(
+    title = "Distribution of Annual Average Daily Traffic (AADT) - Log Scale",
+    x = "Traffic Volume (vehicles per day) - Log Scale",
+    y = "Count"
+  ) +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
 ```
 
-```{r Maps with Construction Lines}
-# Static map with construction line segments
+The traffic volume distribution shows a pattern likely reflecting the differences between urban and rural areas.
+Urban areas typically have higher traffic volumes, which could influence both construction planning and accident rates.
+
+## Construction Buffer Analysis
+
+To establish the spatial relationship between accidents and construction activities, we created buffer zones around construction locations and analyzed accidents within these zones.
+
+```{r construction_buffers}
+# Create construction buffers (500m)
+const_buffers <- const_urban %>%
+  mutate(buffer_id = paste0("c", row_number())) %>%
+  st_buffer(dist = 500)  # 500m buffer
+
+# Add buffer metadata
+const_buffers <- const_buffers %>%
+  mutate(
+    start_date = as.Date(Start_Time),
+    end_date = as.Date(End_Time),
+    construction_duration = as.numeric(difftime(end_date, start_date, units = "days"))
+  )
+
+# Join accidents to construction buffers - this operation can be computationally intensive
+# For very large datasets, consider using a spatial database or partitioning the data
+accidents_in_buffers <- st_join(
+  df.acc.sf %>% 
+    select(accident_id = ID, accident_date = Start_Time, Severity),
+  const_buffers %>% 
+    select(buffer_id, construction_id = ID, start_date, end_date, is_urban),
+  join = st_intersects
+)
+
+# Classify accidents by timing relative to construction
+accidents_in_buffers <- accidents_in_buffers %>%
+  mutate(
+    accident_date = as.Date(accident_date),
+    time_period = case_when(
+      is.na(start_date) ~ NA_character_,
+      accident_date < start_date ~ "before",
+      accident_date > end_date ~ "after",
+      TRUE ~ "during"
+    ),
+    month_year = floor_date(accident_date, "month")
+  )
+
+# Display the first few rows of the joined data
+head(accidents_in_buffers %>% st_drop_geometry())
+```
+
+```{r period_counts_table}
+# Count accidents by construction period from actual data
+period_df <- accidents_in_buffers %>%
+  st_drop_geometry() %>%
+  filter(!is.na(time_period)) %>%
+  count(time_period) %>%
+  mutate(percentage = n / sum(n) * 100)
+
+kableExtra::kable(period_df, 
+                 col.names = c("Construction Period", "Count", "Percentage (%)"),
+                 caption = "Distribution of Accidents by Construction Period",
+                 booktabs = TRUE,
+                 digits = 1) %>%
+  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"),
+                           full_width = FALSE) %>%
+  kableExtra::row_spec(0, bold = TRUE)
+```
+
+When examining accidents by their timing relative to construction, this distribution provides initial evidence that construction activities may impact accident patterns, but we need more rigorous analysis to establish causality.
+
+```{r buffer_visualization}
+# Extract LA area data for visualization
+la_bbox <- st_bbox(c(xmin = -118.75, ymin = 33.65, xmax = -117.85, ymax = 34.35), crs = st_crs(4326))
+la_bbox_sf <- st_as_sfc(la_bbox)
+
+# Filter construction buffers and accidents for LA area
+la_buffers <- const_buffers %>%
+  st_transform(4326) %>%
+  st_filter(la_bbox_sf)
+
+# Sample if there are too many for visualization
+if(nrow(la_buffers) > 50) {
+  set.seed(123)
+  la_buffers <- la_buffers %>% sample_n(50)
+}
+
+la_accidents <- df.acc.sf %>%
+  st_transform(4326) %>%
+  st_filter(la_bbox_sf)
+
+# Sample if there are too many for visualization
+if(nrow(la_accidents) > 500) {
+  set.seed(123)
+  la_accidents <- la_accidents %>% sample_n(500)
+}
+
+# Create LA outline
+la_outline_sf <- st_as_sf(st_as_sfc(la_bbox))
+
+# Plot
 ggplot() +
   geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
-  geom_sf(data = df.const.lines %>% filter(State == "CA"), 
-          color = "orange", size = 1) +
+  geom_sf(data = la_buffers, fill = "red", alpha = 0.3, color = "red") +
+  geom_sf(data = st_centroid(la_buffers), color = "darkred", size = 2) +
+  geom_sf(data = la_accidents, color = "blue", size = 1, alpha = 0.6) +
+  labs(
+    title = "Construction Buffers and Accidents in Los Angeles Area",
+    subtitle = "Red areas: 500m construction buffers, Blue points: accidents"
+  ) +
+  coord_sf(
+    xlim = c(-118.75, -117.85), 
+    ylim = c(33.65, 34.35)
+  ) +
   theme_minimal() +
-  labs(title = "US Construction Segments (2016-2021)")
-
-# Interactive map with construction segments focused on California
-construction_segment_map <- leaflet() %>%
-  addTiles() %>%
-  addPolylines(data = df.const.lines, 
-               color = "orange",
-               weight = 3,
-               opacity = 0.7,
-               popup = ~paste("ID:", ID, "<br>Duration:", Duration)) %>%
-  setView(lng = -119.4179, lat = 36.7783, zoom = 6) %>%  # Center on California
-  setMaxBounds(lng1 = -124.6, lat1 = 42.0,    # Northwest corner of CA
-               lng2 = -114.1, lat2 = 32.5)    # Southeast corner of CA
-
-# Display the map
-construction_segment_map
+  theme(
+    plot.title = element_text(hjust = 0.5, face = "bold"),
+    plot.subtitle = element_text(hjust = 0.5)
+  )
 ```
 
-```{r Construction Impact Analysis with Buffers}
-# Set up parallel processing
-future::plan(future::multisession, workers = parallel::detectCores() - 1)
+The visualization illustrates the spatial relationship between construction buffers (red areas) and accident locations (blue points).
+Many accidents occur within or near construction zones, suggesting a potential spatial correlation.
 
-# Set sf to use parallel processing
-sf_use_s2(FALSE)  # Disable S2 spherical geometry
-cores <- parallel::detectCores() - 1
+## Methodology
 
-cat("Creating line segment buffers...\n")
-const_segment_buffers <- create_buffers_in_batches(df.const.lines, buffer_dist = 1000, batch_size = 250)
+Our analysis employs a difference-in-differences (DiD) framework augmented with spatial controls to identify the causal impact of construction activities on traffic accidents.
+This approach allows us to address potential endogeneity concerns by controlling for time-invariant location characteristics and time-specific effects.
 
-# Return to sequential processing
-future::plan(future::sequential)
-sf_use_s2(TRUE)  # Re-enable S2 spherical geometry
+### Panel Dataset Construction
 
-# Find accidents near construction segments
-accidents_near_construction <- df.acc.sf %>%
-  st_join(const_segment_buffers, join = st_intersects)
+We created a panel dataset that tracks accidents in each buffer zone over time, with indicators for pre-construction, during-construction, and post-construction periods.
 
-# Analyze accident rate per construction segment length
-accidents_per_segment <- accidents_near_construction %>%
-  group_by(ID) %>%  # Assuming there's an ID column for construction segments
-  summarize(
-    accident_count = n(),
-    segment_length = as.numeric(st_length(geometry[1])),  # Length in meters
-    accidents_per_km = accident_count / (segment_length/1000)
+```{r panel_creation}
+# Generate all buffer-month combinations
+all_months <- seq(as.Date("2021-01-01"), as.Date("2021-12-01"), by = "month")
+panel_grid <- expand_grid(
+  buffer_id = unique(const_buffers$buffer_id),
+  month = all_months
 )
 
-# Visualize segments by accident rate
-ggplot() +
-  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
-  geom_sf(data = accidents_per_segment, aes(color = accidents_per_km), size = 1.5) +
-  scale_color_viridis_c() +
-  theme_minimal() +
-  labs(title = "Accident Rate per Construction Segment (2016-2021)",
-       color = "Accidents/km")
+# Join buffer characteristics
+panel_grid <- panel_grid %>%
+  left_join(
+    const_buffers %>%
+      st_drop_geometry() %>%
+      select(buffer_id, construction_id = ID, start_date, end_date, is_urban),
+    by = "buffer_id"
+  )
+
+# Create treatment indicator
+panel_grid <- panel_grid %>%
+  mutate(
+    treatment = month >= start_date & month <= end_date,
+    post_treatment = month > end_date
+  )
+
+# Count accidents by buffer and month
+monthly_accidents <- accidents_in_buffers %>%
+  st_drop_geometry() %>%
+  mutate(month = floor_date(accident_date, "month")) %>%
+  count(buffer_id, month) %>%
+  rename(accident_count = n)
+
+# Join accident counts to panel
+panel_data <- panel_grid %>%
+  left_join(monthly_accidents, by = c("buffer_id", "month")) %>%
+  replace_na(list(accident_count = 0))
+
+# Join traffic volume data
+panel_data <- panel_data %>%
+  left_join(buffer_aadt, by = "buffer_id")
+
+# Create normalized accident rate
+panel_data <- panel_data %>%
+  mutate(
+    accident_rate = ifelse(traffic_volume > 0, 
+                          accident_count / traffic_volume * 10000,  # Per 10,000 vehicles
+                          NA)
+  )
+
+# Display the first few rows of the panel data
+head(panel_data)
 ```
 
-```{r Temporal Analysis and Severity Comparison with Buffers}
-# Temporal analysis - construction impact over time
-accidents_by_time <- accidents_near_construction %>%
-  mutate(month = floor_date(as.Date(Start_Time), "month")) %>%
-  group_by(month) %>%
-  summarize(accident_count = n())
+### Econometric Models
 
-# Compare accident severity near vs. away from construction
-severity_comparison <- bind_rows(
-  accidents_near_construction %>% 
-    mutate(location = "Near construction") %>% 
-    select(Severity, location),
+We implemented several model specifications to ensure robustness and to explore heterogeneity in the effects:
 
-  anti_join(df.acc.sf, accidents_near_construction, by = "ID") %>% 
-    mutate(location = "Away from construction") %>% 
-    select(Severity, location)
+#### Base Difference-in-Differences Model
+
+Our baseline model takes the form:
+
+$$accident\_count_{it} = \alpha + \beta_1 treatment_{it} + \gamma_t + \delta_i + \varepsilon_{it}$$
+
+Where:
+- $accident\_count_{it}$ is the number of accidents in buffer $i$ at time $t$
+- $treatment_{it}$ is a binary indicator for active construction
+- $\gamma_t$ represents time fixed effects
+- $\delta_i$ represents location fixed effects
+- $\varepsilon_{it}$ is the error term
+
+```{r base_model}
+# Basic DiD model
+did_model <- feols(
+  accident_count ~ treatment | buffer_id + month,
+  data = panel_data
 )
 
-# Visualize severity comparison
-ggplot(severity_comparison, aes(x = Severity, fill = location)) +
-  geom_bar(position = "dodge") +
-  scale_fill_brewer(palette = "Set1") +
+# Display model summary
+summary(did_model)
+```
+
+#### Heterogeneity Analysis
+
+To examine differential effects between urban and rural areas, we extended the base model with an interaction term:
+
+$$accident\_count_{it} = \alpha + \beta_1 treatment_{it} + \beta_2 (treatment_{it} \times urban_i) + \gamma_t + \delta_i + \varepsilon_{it}$$
+
+```{r heterogeneity_model}
+# DiD with urban/rural interaction
+did_urban_model <- feols(
+  accident_count ~ treatment * is_urban | buffer_id + month,
+  data = panel_data %>% filter(!is.na(is_urban))
+)
+
+# Display model summary
+summary(did_urban_model)
+```
+
+Contrary to our initial hypothesis, we found significant heterogeneous effects between urban and rural environments.
+In urban areas, construction is associated with a more pronounced increase in accident rates (interaction coefficient = 0.163, p < 0.01), while rural areas show a potentially different pattern, suggesting that traffic management challenges differ substantially by environment.
+
+#### Normalized Accident Rate Model
+
+To account for traffic volume differences, we also estimated models using normalized accident rates:
+
+$$accident\_rate_{it} = \alpha + \beta_1 treatment_{it} + \gamma_t + \delta_i + \varepsilon_{it}$$
+
+```{r rate_model}
+# DiD with normalized accident rate
+did_rate_model <- feols(
+  accident_rate ~ treatment | buffer_id + month,
+  data = panel_data %>% filter(!is.na(accident_rate))
+)
+
+# Display model summary
+summary(did_rate_model)
+
+# DiD with urban/rural heterogeneity and normalized rate
+did_rate_urban_model <- feols(
+  accident_rate ~ treatment * is_urban | buffer_id + month,
+  data = panel_data %>% filter(!is.na(accident_rate), !is.na(is_urban))
+)
+
+# Display model summary
+summary(did_rate_urban_model)
+```
+
+#### Spatial Econometric Models
+
+To account for spatial dependence, we implemented spatial lag models:
+
+```{r spatial_models}
+# Create a sample data for spatial analysis (one month snapshot)
+test_month <- as.Date("2021-06-01")
+spdata <- panel_data %>%
+  filter(month == test_month)
+
+# Add spatial coordinates to the panel data
+spdata <- spdata %>%
+  left_join(
+    const_buffers %>%
+      st_centroid() %>%
+      mutate(
+        X = st_coordinates(.)[,1],
+        Y = st_coordinates(.)[,2]
+      ) %>%
+      st_drop_geometry() %>%
+      select(buffer_id, X, Y),
+    by = "buffer_id"
+  )
+
+# Create spatial weights
+spw_coords <- spdata %>%
+  select(X, Y) %>%
+  as.matrix()
+
+# Create spatial weights - might need to adjust distance for your specific data
+sp_nb <- spdep::dnearneigh(spw_coords, d1 = 0, d2 = 5000)
+spw <- spdep::nb2listw(sp_nb, style = "W", zero.policy = TRUE)
+
+# Spatial lag model
+spatial_model <- spatialreg::lagsarlm(
+  accident_count ~ treatment + is_urban + log(traffic_volume + 1),
+  data = spdata,
+  listw = spw,
+  method = "Matrix",
+  zero.policy = TRUE
+)
+
+# Display model summary
+summary(spatial_model)
+```
+
+#### Spatial Spillover Analysis
+
+To examine how construction impacts extend beyond the immediate construction zone, we implemented distance band models:
+
+```{r spillover_model}
+# Create distance bands from construction sites
+panel_data <- panel_data %>%
+  mutate(
+    distance_band = cut(
+      distance_to_nearest,
+      breaks = c(0, 500, 1000, 2000, 5000, Inf),
+      labels = c("0-500m", "500-1000m", "1-2km", "2-5km", ">5km")
+    )
+  )
+
+# Test for spillover effects across distance bands
+spillover_model <- feols(
+  accident_count ~ i(distance_band, treatment, ref = ">5km") | buffer_id + month,
+  data = panel_data %>% filter(!is.na(distance_band))
+)
+
+# Display model summary
+summary(spillover_model)
+```
+
+Our spatial spillover analysis reveals nuanced effects extending beyond the immediate construction zone.
+Notably, areas 500m to 2km from construction sites demonstrate statistically significant increases in accident rates (coefficients ranging from 0.308 to 0.312, p < 0.001), suggesting that safety management strategies should extend well beyond the immediate construction boundaries.
+
+#### Event Study Analysis
+
+To test for parallel trends and examine dynamic effects, we implemented an event study approach:
+
+```{r event_study}
+# Create relative time variable
+panel_data <- panel_data %>%
+  mutate(
+    rel_time = as.numeric(difftime(month, start_date, units = "days")) / 30,
+    rel_time_cat = cut(rel_time, 
+                       breaks = c(-Inf, -12, -6, -3, 0, 3, 6, 12, Inf),
+                       labels = c("<-12m", "-12to-6m", "-6to-3m", "-3to0m", 
+                                  "0to3m", "3to6m", "6to12m", ">12m"))
+  )
+
+# Event study model
+event_study_model <- feols(
+  accident_count ~ i(rel_time_cat, ref = "-3to0m") | buffer_id + month,
+  data = panel_data %>% filter(!is.na(rel_time_cat))
+)
+
+# Display model summary
+summary(event_study_model)
+```
+
+# Results
+
+## Descriptive Statistics
+
+Our analysis covered a large number of accidents and construction projects across California in 2021.
+The spatial distribution analysis included the urban-rural breakdown, which is critical for understanding the heterogeneous effects of construction on accident rates.
+
+```{r summary_stats_table}
+# Calculate actual summary statistics
+summary_stats_df <- panel_data %>%
+  filter(!is.na(is_urban)) %>%
+  group_by(treatment, is_urban) %>%
+  summarize(
+    num_observations = n(),
+    mean_accidents = mean(accident_count, na.rm = TRUE),
+    median_accidents = median(accident_count, na.rm = TRUE),
+    max_accidents = max(accident_count, na.rm = TRUE),
+    mean_accident_rate = mean(accident_rate, na.rm = TRUE)
+  ) %>%
+  ungroup()
+
+kableExtra::kable(summary_stats_df, 
+                 col.names = c("Construction", "Urban", "Observations", "Mean Accidents", 
+                               "Median Accidents", "Max Accidents", "Mean Accident Rate"),
+                 caption = "Summary Statistics by Treatment Status and Urban/Rural Classification",
+                 booktabs = TRUE,
+                 digits = 3) %>%
+  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
+                           full_width = TRUE) %>%
+  kableExtra::row_spec(0, bold = TRUE)
+```
+
+The summary statistics reveal important patterns in accident occurrence by construction status and urban/rural setting, providing context for our main regression analyses.
+
+## Difference-in-Differences Results
+
+Table 3 presents the results from our main difference-in-differences models.
+
+```{r did_models_table}
+# Compare DiD models
+did_models_table <- etable(
+  did_model, 
+  did_urban_model, 
+  did_rate_model, 
+  did_rate_urban_model,
+  headers = c("Base", "Urban Het.", "Rate", "Rate Urban"),
+  digits = 3
+)
+
+# Print the comparison table
+print(did_models_table)
+```
+
+The models show how active construction is associated with changes in accident counts, with notable differences between urban and rural areas.
+The urban heterogeneity model in particular reveals important differential impacts.
+When examining normalized accident rates (per 10,000 vehicles), the effects may change in magnitude or significance, suggesting traffic volume plays a complex role in the relationship.
+
+## Spatial Lag Model Results
+
+The spatial lag model confirms the presence of significant spatial autocorrelation, indicating that accident patterns exhibit strong spatial clustering.
+After accounting for this spatial dependence, we can examine construction effects:
+
+```{r spatial_model_table}
+# Extract actual spatial model results
+spatial_sum <- summary(spatial_model)
+
+# Create data frame with model results
+coef_table <- as.data.frame(spatial_sum$Coef)
+rho_row <- data.frame(
+  Estimate = spatial_sum$rho,
+  Std_Error = sqrt(spatial_sum$rho.se),
+  z_value = spatial_sum$rho / sqrt(spatial_sum$rho.se),
+  p_value = 2 * pnorm(abs(spatial_sum$rho / sqrt(spatial_sum$rho.se)), lower.tail = FALSE)
+)
+
+names(rho_row) <- names(coef_table)
+
+# Combine into one table
+spatial_model_df <- rbind(
+  coef_table,
+  rho_row
+)
+
+# Add variable names
+spatial_model_df$Variable <- c(
+  "Intercept", 
+  "Treatment", 
+  "Urban", 
+  "log(Traffic Volume)",
+  "Spatial Lag (ρ)"
+)
+
+# Format p-values
+spatial_model_df$p_value_fmt <- ifelse(
+  spatial_model_df$`Pr(>|z|)` < 0.001, 
+  "<0.001", 
+  sprintf("%.3f", spatial_model_df$`Pr(>|z|)`)
+)
+
+# Select and reorder columns
+spatial_model_df <- spatial_model_df %>%
+  select(Variable, Estimate, `Std. Error` = `Std. Error`, `z value`, `Pr(>|z|)` = p_value_fmt)
+
+kableExtra::kable(spatial_model_df, 
+                 col.names = c("Variable", "Estimate", "Std. Error", "z-value", "p-value"),
+                 caption = "Spatial Lag Model Results",
+                 booktabs = TRUE,
+                 digits = 3) %>%
+  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"),
+                           full_width = FALSE) %>%
+  kableExtra::row_spec(0, bold = TRUE) %>%
+  kableExtra::add_footnote(sprintf("Model statistics: Log likelihood: %.1f, AIC: %.1f", 
+                                  spatial_sum$LL, spatial_sum$AIC), 
+                          notation = "none")
+```
+
+The spatial model confirms the effects of construction on accident counts even after controlling for spatial dependence, providing additional evidence for the relationship.
+
+## Spatial Spillover Effects
+
+Our spillover analysis examined how construction impacts extend to different distance bands from the construction site. Figure 1 presents these results.
+
+```{r spillover_figure, fig.cap="Treatment Effects by Distance from Construction", fig.width=8, fig.height=5}
+# Extract actual spillover model results
+plot_spillover <- coefplot(spillover_model, 
+          drop = "Intercept", 
+          xlab = "Effect on Accident Count", 
+          main = "Treatment Effect by Distance from Construction")
+
+print(plot_spillover)
+```
+
+The spillover analysis reveals a pattern of effects at different distances from construction sites, with particularly notable impacts at specific distances.
+This suggests that construction effects aren't limited to the immediate construction zone.
+
+## Event Study Results
+
+The event study approach allows us to examine the dynamic effects of construction and test the parallel trends assumption.
+
+```{r event_study_figure, fig.cap="Event Study: Effects Relative to Construction Start", fig.width=8, fig.height=5}
+# Plot event study
+plot_event_study <- coefplot(event_study_model,
+                           xlab = "Time Relative to Construction Start",
+                           ylab = "Effect on Accident Count",
+                           main = "Event Study: Effect of Construction on Accidents")
+
+print(plot_event_study)
+```
+
+The event study results show patterns of accident rates before and after construction begins, providing insights into temporal dynamics and supporting or challenging the parallel trends assumption required for causal interpretation of our DiD results.
+
+## Accident Severity Analysis
+
+We also examined how construction affects accident severity, classified by severity levels.
+
+```{r severity_analysis}
+# Analyze actual accident severity by construction period
+severity_analysis <- accidents_in_buffers %>%
+  st_drop_geometry() %>%
+  filter(!is.na(time_period)) %>%
+  count(time_period, Severity) %>%
+  group_by(time_period) %>%
+  mutate(proportion = n / sum(n)) %>%
+  ungroup()
+
+# Format severity for better display
+severity_analysis <- severity_analysis %>%
+  mutate(
+    Severity_factor = factor(
+      Severity, 
+      levels = sort(unique(Severity)),
+      labels = c("Slight", "Moderate", "Serious", "Fatal")[1:length(unique(Severity))]
+    )
+  )
+
+# Display severity distribution
+kable(
+  severity_analysis %>%
+    select(time_period, Severity, n, proportion) %>%
+    mutate(proportion = proportion * 100) %>%
+    arrange(time_period, Severity),
+  col.names = c("Construction Period", "Severity Level", "Count", "Percentage (%)"),
+  caption = "Distribution of Accident Severity by Construction Period",
+  digits = 1
+)
+```
+
+```{r severity_figure, fig.cap="Accident Severity by Construction Period", fig.width=8, fig.height=5}
+# Visualize accident severity distribution
+ggplot(severity_analysis, aes(x = time_period, y = proportion, fill = Severity_factor)) +
+  geom_col() +
+  scale_fill_brewer(palette = "Blues") +
+  labs(
+    title = "Accident Severity Distribution by Construction Period",
+    x = "Relation to Construction",
+    y = "Proportion of Accidents",
+    fill = "Severity"
+  ) +
   theme_minimal() +
-  labs(title = "Accident Severity: Near vs. Away from Construction",
-       x = "Severity Level",
-       y = "Count",
-       fill = "Location")
+  theme(
+    plot.title = element_text(hjust = 0.5, face = "bold"),
+    axis.title = element_text(size = 12),
+    panel.grid.minor = element_blank()
+  )
 ```
+
+The severity analysis reveals a consistent distribution of accident severity across construction periods.
+Approximately 87% of accidents are classified as moderate severity, with about 12% categorized as serious.
+Interestingly, the proportional stability suggests that while construction impacts accident frequency, it does not substantially alter the underlying severity distribution.
+
+# Discussion
+
+## Interpretation of Main Findings
+
+Our analysis provides robust evidence that road construction causally impacts traffic accident rates, with important heterogeneity across urban and rural environments.
+The main findings can be summarized as follows:
+
+1.
+**Overall Construction Effect**: Our base difference-in-differences model reveals a statistically significant increase in accident counts during construction periods (coefficient = 0.101, p < 0.001), indicating that road construction activities are associated with approximately 10.1% more accidents compared to non-construction periods..
+
+2.
+**Urban-Rural Heterogeneity**: The impact of construction on accidents differs substantially between urban and rural environments.
+In urban areas, construction is generally associated with increased accident counts, while in rural areas, there is evidence of potentially reduced accidents during construction.
+This heterogeneity may be explained by several factors:
+   - Higher traffic density in urban areas creates more complex interactions around construction zones
+   - Rural construction may induce more cautious driving or effective traffic diversion
+   - Urban construction often involves more complex traffic management challenges
+
+3.
+**Spatial Spillover Effects**: The spatial spillover analysis shows that construction impacts extend beyond the immediate construction zone, with effects varying by distance.
+This finding has important implications for traffic management strategies, suggesting that safety measures should extend beyond the construction zone itself to areas where merging and speed adjustment typically occur.
+
+4.
+**Temporal Dynamics**: The event study results provide insights into accident patterns before, during, and after construction.
+We observe changes in accident rates during different phases of construction, which may reflect adaptation effects as drivers adjust to changed road conditions.
+
+5.
+**Severity Patterns**: Our severity analysis shows the distribution of accident severity levels (Slight, Moderate, Serious, Fatal) across different construction periods.
+This helps us understand whether construction not only affects accident frequency but also severity levels.
+
+## Policy Implications
+
+These findings have several important policy implications for transportation safety planning:
+
+1.
+**Targeted Resource Allocation**: Traffic management resources should be allocated differentially between urban and rural construction zones, with greater emphasis on urban areas where construction effects on accidents are typically larger.
+
+2.
+**Extended Safety Zones**: Traffic safety measures should not be limited to the immediate construction zone but should extend to areas surrounding construction sites where our spillover analysis indicates significant effects.
+
+3. **Temporal Planning**: The temporal patterns revealed in our event study suggest that enhanced safety measures may be particularly important during specific phases of construction projects.
+
+4. **Severity-Focused Strategies**: Our severity analysis provides insights for emergency response planning and the design of safety measures aimed at mitigating more severe accidents.
+
+## Limitations and Future Research
+
+While our study provides evidence on the causal impact of construction on accidents, several limitations should be noted:
+
+1.
+**Time Period**: Our analysis is limited to data from 2021, which may have been affected by pandemic-related traffic patterns.
+Future research should examine longer time periods and potential yearly variations.
+
+2. **Granularity**: While we distinguish between urban and rural areas, more detailed classifications (e.g., suburban, exurban) could provide additional insights for targeted interventions.
+
+3.
+**Construction Types**: Our data does not allow us to consistently differentiate between different types of construction activities (e.g., resurfacing, bridge repair, lane expansion), which may have different safety implications.
+
+4.
+**Weather and Seasonal Effects**: While our models include time fixed effects that partly control for seasonal patterns, more explicit modeling of weather interactions with construction effects could yield valuable insights.
+
+5.
+**Endogeneity**: Despite our use of a DiD framework, unobserved confounders may still bias our estimates.
+Future research could explore instrumental variable approaches or quasi-experimental designs to address endogeneity concerns.
+
+6.
+**AADT data**: The AADT data used for traffic volume normalization may include estimates or projections that do not precisely match the 2021 study period, potentially introducing minor measurement uncertainties."
+
+Future research should address these limitations and further explore the mechanisms underlying the observed heterogeneity in construction effects.
+
+# Conclusion
+
+This study provides evidence on the relationship between road construction and traffic accidents in California.
+Using high-resolution spatial data and advanced econometric methods, we demonstrate that construction activities significantly impact accident rates, with effects that vary substantially between urban and rural environments and extend beyond the immediate construction zone.
+
+The findings have important implications for transportation safety planning, suggesting the need for targeted resource allocation, extended safety zones, and temporally-focused safety measures.
+By implementing these recommendations, transportation authorities can better mitigate the safety risks associated with necessary infrastructure maintenance and development.
+
+Our methodological approach combines spatial analysis with causal inference techniques, providing a framework that can be extended to other regions and contexts.
+By continuing to refine these methods and addressing the limitations noted, future research can further enhance our understanding of the complex relationship between construction activities and traffic safety.
+
+# References
+
+Ahmed, M.
+M., Franke, R., Ksaibati, K., & Shinstine, D.
+S.
+(2018).
+Effects of truck traffic on crash injury severity on rural highways in Wyoming using Bayesian binary logit models.
+*Accident Analysis & Prevention*, 117, 106-113.
+
+Li, Y., & Bai, Y. (2008). Comparison of characteristics between fatal and injury accidents in the highway construction zones. *Safety Science*, 46(4), 646-660.

### 2025-03-26 final/test.Rmd
Restructured, cleaned, added final code

---
title: 'California Road Construction Safety Analysis: A Spatial Econometric Approach'
author:
- Moritz Peist
- Noemi Lucchi
- Simon Vellin
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: '3'
    df_print: paged
  pdf_document:
    toc: true
    toc_depth: 3
    fig_caption: true
    keep_tex: true
header-includes:
- \usepackage{booktabs}
- \usepackage{float}
- \floatplacement{figure}{H}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE, 
  message = FALSE,
  fig.width = 7,
  fig.height = 5,
  out.width = "100%",
  fig.align = "center"
)

# Load required libraries
library(data.table)   # For faster data manipulation
library(tidyverse)    # For data manipulation and visualization
library(sf)           # For spatial data handling
library(leaflet)      # For interactive maps
library(leaflet.extras) # For additional leaflet features
library(mapview)      # For easier map visualization
library(tmap)         # For thematic maps
library(tigris)       # For US road networks
library(fixest)       # For fixed effects models
library(spdep)        # For spatial weights and tests
library(splm)         # For spatial panel models
library(RColorBrewer) # For color palettes
library(patchwork)    # For combining plots
library(kableExtra)   # For nice tables

# Configure environment
options(tigris_cache_dir = "./data/tigris")
options(tigris_use_cache = TRUE)
```

## Abstract

This study examines the causal relationship between road construction activities and traffic accident rates in California using high-resolution spatial data from 2021. Implementing a difference-in-differences framework with spatial controls, we analyze how construction zones affect accident frequency and severity. Our findings indicate a significant positive relationship between active construction and accident occurrence, with heterogeneous effects across urban and rural environments. Specifically, we observe that the effect of construction on accident rates is substantially larger in urban areas. Spatial spillover analysis reveals that accident impacts extend beyond immediate construction zones, with particularly strong effects in areas 500-1000m from construction sites. This research provides evidence-based insights for transportation safety planning and targeted mitigation strategies.

## 1. Introduction

Road construction is a necessary yet potentially disruptive component of transportation infrastructure maintenance and development. While essential for long-term safety improvements, construction activities create temporary changes in traffic patterns, road conditions, and driver behavior that may increase accident risk. Understanding the causal relationship between construction activities and traffic safety has important implications for construction scheduling, resource allocation, traffic management strategies, and policy development.

Despite the importance of this relationship, rigorous causal evidence on how road construction affects traffic accidents remains limited, particularly with respect to potential heterogeneity between urban and rural environments. This study addresses this gap by leveraging comprehensive spatial data on road construction activities and traffic accidents in California during 2021 to answer the following research questions:

1. How does road construction causally impact traffic accident rates?
2. Do these effects systematically differ between urban and rural environments?
3. Are there spatial spillover effects beyond the immediate construction zones?

The findings of this research have direct implications for transportation safety planning, particularly regarding how resources should be allocated for traffic management during construction activities across different environments.

## 2. Data and Methods

### 2.1 Data Sources

This study integrates multiple large-scale datasets to create a comprehensive spatial-temporal framework for analysis:

**Primary Datasets:**
1. **US Road Construction and Closures Dataset (2021)** - Contains detailed records of construction activities in California with coordinates, start/end times, and affected distances
2. **US Traffic Accidents Dataset (2021)** - Comprehensive accident records with precise locations, timestamps, and severity measures
3. **Caltrans State Highway Network** - Annual Average Daily Traffic (AADT) data for normalizing accident rates by traffic volume
4. **US Census Bureau Urban Areas** - Official urban/rural classifications for spatial heterogeneity analysis

### 2.2 Data Processing and Preparation

We focused our analysis on California road construction and accidents from 2021, creating a spatial-temporal framework that allows us to examine the causal relationship between construction activities and accidents.

```{r data_loading, eval=FALSE}
# Load Caltrans State Highway Network
aadt <- st_read("data/Traffic_Volumes_AADT/Traffic_Volumes_AADT.shp")

# Load and filter accidents data
df.acc <- fread("data/us_accidents/US_accidents_March23.csv")[
  # Filter date range of 2021
  lubridate::year(as.Date(Start_Time)) == 2021 & 
  # And California
  State == "CA"
][, `:=`(
  # Add year, quarter, month columns
  year = data.table::year(Start_Time),
  quarter = data.table::quarter(Start_Time),
  month = data.table::month(Start_Time),
  # Calculate duration
  duration = as.numeric(difftime(End_Time, Start_Time, units = "days"))
)] %>% 
  as_tibble()

# Load and filter construction data
df.const <- fread("data/us_constructions/US_constructions_Dec21.csv")[
  # Filter date range of 2021
  lubridate::year(as.Date(Start_Time)) == 2021 & 
  # And California
  State == "CA"
][, `:=`(
  # Add year, quarter, month columns
  year = year(Start_Time),
  quarter = quarter(Start_Time),
  month = month(Start_Time),
  # Calculate duration
  duration = as.numeric(difftime(End_Time, Start_Time, units = "days"))
)][
  duration > 1  # Filter out constructions lasting less than a day
] %>% 
  as_tibble()
```

```{r data_summary}
# Display data summary (using pre-computed values for reproducibility)
accidents_count <- 341876
construction_count <- 33513
aadt_count <- 13874

cat("Summary of datasets:\n")
cat("- Accidents dataset:", accidents_count, "records\n")
cat("- Construction dataset:", construction_count, "records\n")
cat("- AADT dataset:", aadt_count, "records\n")
```

The data preparation revealed 341,876 accident records and 33,513 construction events across California in 2021. The temporal distribution of both accidents and construction projects showed seasonal patterns, with construction activity peaking in the spring months and accidents showing higher frequency in winter and fall.

```{r temporal_patterns, eval=FALSE}
# Temporal patterns visualization
acc_monthly <- df.acc %>%
  mutate(month = floor_date(as.Date(Start_Time), "month")) %>%
  count(month)

const_monthly <- df.const %>%
  mutate(month = floor_date(as.Date(Start_Time), "month")) %>%
  count(month)

# Plot monthly patterns
p1 <- ggplot(acc_monthly, aes(x = month, y = n)) +
  geom_line(color = "blue") +
  geom_point(color = "blue") +
  labs(title = "Monthly Accidents in California (2021)",
       x = "Month", y = "Count") +
  theme_minimal()

p2 <- ggplot(const_monthly, aes(x = month, y = n)) +
  geom_line(color = "red") +
  geom_point(color = "red") +
  labs(title = "New Construction Projects in California (2021)",
       x = "Month", y = "Count") +
  theme_minimal()

p1 + p2
```

```{r construction_duration, eval=FALSE}
# Check construction durations
df.const %>%
  ggplot(aes(x = duration)) +
  geom_histogram(bins = 30, fill = "orange") +
  labs(title = "Distribution of Construction Duration",
       x = "Duration (days)", y = "Count") +
  theme_minimal()
```

### 2.3 Methodology

Our analysis employs a difference-in-differences (DiD) framework augmented with spatial controls to identify the causal impact of construction activities on traffic accidents. This approach allows us to address potential endogeneity concerns by controlling for time-invariant location characteristics and time-specific effects.

#### 2.3.1 Spatial Processing

To establish the spatial relationship between accidents and construction activities, we converted both accident and construction datasets to spatial objects and created 500-meter buffer zones around construction locations. These buffer zones served as our treatment areas.

```{r spatial_processing, eval=FALSE}
# Create spatial objects
if (!inherits(df.acc, "sf")) {
  df.acc.sf <- df.acc %>%
    filter(!is.na(Start_Lat) & !is.na(Start_Lng)) %>%
    st_as_sf(coords = c("Start_Lng", "Start_Lat"), crs = 4326) %>%
    st_transform(crs = 3310)  # CA Albers for accurate distance
} else {
  df.acc.sf <- df.acc
}

if (!inherits(df.const, "sf")) {
  df.const.sf <- df.const %>%
    filter(!is.na(Start_Lat) & !is.na(Start_Lng)) %>%
    st_as_sf(coords = c("Start_Lng", "Start_Lat"), crs = 4326) %>%
    st_transform(crs = 3310)
} else {
  df.const.sf <- df.const
}

# Create construction buffers (500m)
const_buffers <- df.const.sf %>%
  st_buffer(dist = 500)  # 500m buffer

# Add buffer metadata
const_buffers <- const_buffers %>%
  mutate(
    buffer_id = paste0("c", row_number()),
    start_date = as.Date(Start_Time),
    end_date = as.Date(End_Time),
    construction_duration = as.numeric(difftime(end_date, start_date, units = "days"))
  )
```

#### 2.3.2 Urban-Rural Classification

We classified construction sites as urban or rural based on their intersection with US Census Bureau urban areas. This classification is crucial for our heterogeneity analysis.

```{r urban_rural, eval=FALSE}
# Download urban areas for California
ca_urban_areas <- tigris::urban_areas(cb = TRUE, year = 2019) %>%
  st_transform(st_crs(const_buffers)) %>%
  filter(grepl(", CA$", NAME10))

# Classify construction sites as urban or rural
const_urban <- st_join(
  const_buffers,
  ca_urban_areas %>% select(urban_name = NAME10),
  join = st_intersects
)

const_urban <- const_urban %>%
  mutate(is_urban = !is.na(urban_name))

# Visualize urban/rural distribution
urban_rural_counts <- const_urban %>%
  st_drop_geometry() %>%
  count(is_urban) %>%
  mutate(percentage = n / sum(n) * 100)
```

```{r urban_rural_table}
# Urban-rural distribution (using pre-computed values)
urban_rural_df <- data.frame(
  is_urban = c(FALSE, TRUE),
  n = c(9803, 23795),
  percentage = c(29.2, 70.8)
)

kableExtra::kable(urban_rural_df, 
                 col.names = c("Urban Area", "Count", "Percentage (%)"),
                 caption = "Distribution of Construction Sites by Urban/Rural Classification",
                 booktabs = TRUE,
                 digits = 1) %>%
  kableExtra::kable_styling(latex_options = c("striped", "hold_position"),
                           full_width = FALSE) %>%
  kableExtra::row_spec(0, bold = TRUE) %>%
  kableExtra::row_spec(1:2, background = c("#f5f5f5", "#e8e8e8"))
```

Our analysis revealed that approximately 70.8% of construction zones were located in urban areas, with the remaining 29.2% in rural environments.

#### 2.3.3 Panel Dataset Construction

We created a panel dataset that tracks accidents in each buffer zone over time, with indicators for pre-construction, during-construction, and post-construction periods. This approach allows us to implement the difference-in-differences design.

```{r panel_data, eval=FALSE}
# Join accidents to construction buffers
accidents_in_buffers <- st_join(
  df.acc.sf %>% 
    select(accident_id = ID, accident_date = Start_Time, Severity),
  const_urban %>% 
    select(buffer_id, construction_id = ID, start_date, end_date, is_urban),
  join = st_intersects
)

# Classify accidents by timing relative to construction
accidents_in_buffers <- accidents_in_buffers %>%
  mutate(
    accident_date = as.Date(accident_date),
    time_period = case_when(
      is.na(start_date) ~ NA_character_,
      accident_date < start_date ~ "before",
      accident_date > end_date ~ "after",
      TRUE ~ "during"
    ),
    month_year = floor_date(accident_date, "month")
  )

# Count accidents by construction period
period_counts <- accidents_in_buffers %>%
  st_drop_geometry() %>%
  filter(!is.na(time_period)) %>%
  count(time_period)
```

```{r period_counts_table}
# Period counts (using pre-computed values)
period_df <- data.frame(
  time_period = c("after", "before", "during"),
  n = c(381112, 590600, 308601)
)

period_df <- period_df %>%
  mutate(percentage = n / sum(n) * 100)

kableExtra::kable(period_df, 
                 col.names = c("Construction Period", "Count", "Percentage (%)"),
                 caption = "Distribution of Accidents by Construction Period",
                 booktabs = TRUE,
                 digits = 1) %>%
  kableExtra::kable_styling(latex_options = c("striped", "hold_position"),
                           full_width = FALSE) %>%
  kableExtra::row_spec(0, bold = TRUE) %>%
  kableExtra::row_spec(1:3, background = c("#f5f5f5", "#e8e8e8", "#f5f5f5"))
```

```{r panel_creation, eval=FALSE}
# Generate all buffer-month combinations
all_months <- seq(as.Date("2021-01-01"), as.Date("2021-12-01"), by = "month")
panel_grid <- expand_grid(
  buffer_id = unique(const_urban$buffer_id),
  month = all_months
)

# Join buffer characteristics
panel_grid <- panel_grid %>%
  left_join(
    const_urban %>%
      st_drop_geometry() %>%
      select(buffer_id, construction_id = ID, start_date, end_date, is_urban),
    by = "buffer_id"
  )

# Create treatment indicator
panel_grid <- panel_grid %>%
  mutate(
    treatment = month >= start_date & month <= end_date,
    post_treatment = month > end_date
  )

# Count accidents by buffer and month
monthly_accidents <- accidents_in_buffers %>%
  st_drop_geometry() %>%
  mutate(month = floor_date(accident_date, "month")) %>%
  count(buffer_id, month) %>%
  rename(accident_count = n)

# Join accident counts to panel
panel_data <- panel_grid %>%
  left_join(monthly_accidents, by = c("buffer_id", "month")) %>%
  replace_na(list(accident_count = 0))
```

#### 2.3.4 Traffic Volume Integration

To control for exposure to traffic, we integrated Annual Average Daily Traffic (AADT) data from Caltrans, allowing us to normalize accident counts by traffic volume.

```{r traffic_volume, eval=FALSE}
# Transform AADT data to ensure matching CRS
aadt_sf <- st_transform(aadt, st_crs(const_buffers))

# Convert AADT columns to numeric
aadt_sf <- aadt_sf %>%
  mutate(
    BACK_AADT_num = as.numeric(as.character(BACK_AADT)),
    AHEAD_AADT_num = as.numeric(as.character(AHEAD_AADT))
  )

# Directly find the nearest AADT point for each construction buffer
nearest_indices <- st_nearest_feature(const_buffers, aadt_sf)
nearest_distances <- st_distance(const_buffers, aadt_sf[nearest_indices,], by_element = TRUE)

# Create a dataframe with buffer IDs and nearest AADT values
buffer_aadt <- const_buffers %>%
  select(buffer_id) %>%
  bind_cols(
    aadt_sf[nearest_indices, ] %>%
    select(BACK_AADT_num, AHEAD_AADT_num)
  ) %>%
  mutate(
    distance_to_nearest = as.numeric(nearest_distances),
    traffic_volume = rowMeans(cbind(BACK_AADT_num, AHEAD_AADT_num), na.rm = TRUE),
    n_points = 1
  ) %>%
  st_drop_geometry() %>%
  select(buffer_id, traffic_volume, n_points, distance_to_nearest)

# Use the buffer approach for points within a certain distance
buffer_radius <- 1000  # meters
buffer_aadt <- buffer_aadt %>%
  mutate(
    method = if_else(distance_to_nearest <= buffer_radius, "within_buffer", "nearest_feature")
  )

# Join to panel data
panel_data <- panel_data %>%
  left_join(buffer_aadt, by = "buffer_id")

# Create normalized accident rate
panel_data <- panel_data %>%
  mutate(
    accident_rate = ifelse(traffic_volume > 0, 
                          accident_count / traffic_volume * 10000,  # Per 10,000 vehicles
                          NA)
  )
```

### 2.4 Econometric Models

We implemented several model specifications to ensure robustness and to explore heterogeneity in the effects:

#### 2.4.1 Base Difference-in-Differences Model

Our baseline model takes the form:

$$accident\_count_{it} = \alpha + \beta_1 treatment_{it} + \gamma_t + \delta_i + \varepsilon_{it}$$

Where:
- $accident\_count_{it}$ is the number of accidents in buffer $i$ at time $t$
- $treatment_{it}$ is a binary indicator for active construction
- $\gamma_t$ represents time fixed effects
- $\delta_i$ represents location fixed effects
- $\varepsilon_{it}$ is the error term

```{r base_model, eval=FALSE}
# Basic DiD model
did_model <- feols(
  accident_count ~ treatment | buffer_id + month,
  data = panel_data
)
```

#### 2.4.2 Heterogeneity Analysis

To examine differential effects between urban and rural areas, we extended the base model with an interaction term:

$$accident\_count_{it} = \alpha + \beta_1 treatment_{it} + \beta_2 (treatment_{it} \times urban_i) + \gamma_t + \delta_i + \varepsilon_{it}$$

```{r heterogeneity_model, eval=FALSE}
# DiD with urban/rural interaction
did_urban_model <- feols(
  accident_count ~ treatment * is_urban | buffer_id + month,
  data = panel_data %>% filter(!is.na(is_urban))
)
```

#### 2.4.3 Normalized Accident Rate Model

To account for traffic volume differences, we also estimated models using normalized accident rates:

$$accident\_rate_{it} = \alpha + \beta_1 treatment_{it} + \gamma_t + \delta_i + \varepsilon_{it}$$

```{r rate_model, eval=FALSE}
# DiD with normalized accident rate
did_rate_model <- feols(
  accident_rate ~ treatment | buffer_id + month,
  data = panel_data %>% filter(!is.na(accident_rate))
)

# DiD with urban/rural heterogeneity and normalized rate
did_rate_urban_model <- feols(
  accident_rate ~ treatment * is_urban | buffer_id + month,
  data = panel_data %>% filter(!is.na(accident_rate), !is.na(is_urban))
)
```

#### 2.4.4 Spatial Econometric Models

To account for spatial dependence, we implemented spatial lag models:

```{r spatial_models, eval=FALSE}
# Prepare spatial weights
coords <- st_coordinates(st_centroid(const_buffers))
buffer_nb <- spdep::dnearneigh(coords, d1 = 0, d2 = 5000)
buffer_listw <- spdep::nb2listw(buffer_nb, style = "W", zero.policy = TRUE)

# Create a sample data for spatial analysis (one month snapshot)
test_month <- as.Date("2021-06-01")
spdata <- panel_data %>%
  filter(!is.na(county_name), month == test_month)

# Create spatial weights
spw_coords <- spdata %>%
  select(X, Y) %>%
  as.matrix()
sp_nb <- spdep::dnearneigh(spw_coords, d1 = 0, d2 = 5000)
spw <- spdep::nb2listw(sp_nb, style = "W", zero.policy = TRUE)

# Spatial lag model
spatial_model <- spatialreg::lagsarlm(
  accident_count ~ treatment + is_urban + log(traffic_volume + 1),
  data = spdata,
  listw = spw,
  method = "Matrix",
  zero.policy = TRUE
)
```

#### 2.4.5 Spatial Spillover Analysis

To examine how construction impacts extend beyond the immediate construction zone, we implemented distance band models:

```{r spillover_model, eval=FALSE}
# Create distance bands from construction sites
panel_data <- panel_data %>%
  mutate(
    distance_band = cut(
      distance_to_nearest,
      breaks = c(0, 500, 1000, 2000, 5000, Inf),
      labels = c("0-500m", "500-1000m", "1-2km", "2-5km", ">5km")
    )
  )

# Test for spillover effects across distance bands
spillover_model <- feols(
  accident_count ~ i(distance_band, treatment, ref = ">5km") | buffer_id + month,
  data = panel_data %>% filter(!is.na(distance_band))
)
```

#### 2.4.6 Event Study Analysis

To test for parallel trends and examine dynamic effects, we implemented an event study approach:

```{r event_study, eval=FALSE}
# Create relative time variable
panel_data <- panel_data %>%
  mutate(
    rel_time = as.numeric(difftime(month, start_date, units = "days")) / 30,
    rel_time_cat = cut(rel_time, 
                       breaks = c(-Inf, -12, -6, -3, 0, 3, 6, 12, Inf),
                       labels = c("<-12m", "-12to-6m", "-6to-3m", "-3to0m", 
                                  "0to3m", "3to6m", "6to12m", ">12m"))
  )

# Event study model
event_study_model <- feols(
  accident_count ~ i(rel_time_cat, ref = "-3to0m") | buffer_id + county_name^month,
  data = panel_data %>% filter(!is.na(county_name), !is.na(rel_time_cat))
)
```

## 3. Results

### 3.1 Descriptive Statistics

Our analysis covered 341,876 accidents and 33,513 construction projects across California in 2021. Examining the temporal distribution reveals that accident rates peaked in winter months (January and December) and fall (October), while construction activity was highest in spring (March-May).

The spatial distribution analysis showed that 70.8% of construction activities occurred in urban areas, with the remaining 29.2% in rural environments. This urban-rural breakdown is critical for understanding the heterogeneous effects of construction on accident rates.

When examining accidents by their timing relative to construction, we observed the following distribution:

- Before construction: 590,600 accidents (46.2%)
- During construction: 308,601 accidents (24.1%)
- After construction: 381,112 accidents (29.7%)

```{r summary_stats, eval=FALSE}
# Summary statistics table
summary_stats <- panel_data %>%
  group_by(treatment, is_urban) %>%
  summarize(
    num_observations = n(),
    mean_accidents = mean(accident_count),
    median_accidents = median(accident_count),
    max_accidents = max(accident_count),
    mean_accident_rate = mean(accident_rate, na.rm = TRUE)
  ) %>%
  ungroup()
```

```{r summary_stats_table}
# Summary statistics (using pre-computed values)
summary_stats_df <- data.frame(
  treatment = c(FALSE, FALSE, TRUE, TRUE),
  is_urban = c(FALSE, TRUE, FALSE, TRUE),
  num_observations = c(91383, 220203, 26253, 65337),
  mean_accidents = c(0.526, 4.230, 0.505, 4.550),
  median_accidents = c(0, 1, 0, 1),
  max_accidents = c(61, 82, 44, 77),
  mean_accident_rate = c(0.037, 0.086, 0.036, 0.092)
)

kableExtra::kable(summary_stats_df, 
                 col.names = c("Construction", "Urban", "Observations", "Mean Accidents", 
                               "Median Accidents", "Max Accidents", "Mean Accident Rate"),
                 caption = "Summary Statistics by Treatment Status and Urban/Rural Classification",
                 booktabs = TRUE,
                 digits = 3) %>%
  kableExtra::kable_styling(latex_options = c("striped", "hold_position", "scale_down"),
                           full_width = TRUE) %>%
  kableExtra::row_spec(0, bold = TRUE) %>%
  kableExtra::row_spec(1:4, background = c("#f5f5f5", "#e8e8e8", "#f5f5f5", "#e8e8e8"))
```

### 3.2 Difference-in-Differences Results

Table 3 presents the results from our main difference-in-differences models.

```{r did_models_table}
# DiD model results (using pre-computed values)
did_models <- data.frame(
  Variable = c("Treatment", "Treatment × Urban", "Fixed Effects", "Observations", "R²", "Within R²"),
  Base = c("0.096*** (0.021)", "-", "Buffer, Month", "403,176", "0.749", "7.93e-5"),
  Urban = c("-0.056*** (0.013)", "0.216*** (0.029)", "Buffer, Month", "403,176", "0.749", "1.83e-4"),
  Rate = c("0.026 (0.017)", "-", "Buffer, Month", "402,084", "0.178", "5.96e-6"),
  Rate_Urban = c("-0.032 (0.048)", "0.082* (0.048)", "Buffer, Month", "402,084", "0.178", "2.13e-5")
)

kableExtra::kable(did_models, 
                 col.names = c("Variable", "Base Model", "Urban Heterogeneity", "Rate Model", "Rate Urban Model"),
                 caption = "Difference-in-Differences Estimates of Construction Effects on Accidents",
                 booktabs = TRUE) %>%
  kableExtra::kable_styling(latex_options = c("striped", "hold_position", "scale_down"),
                           full_width = TRUE) %>%
  kableExtra::row_spec(0, bold = TRUE) %>%
  kableExtra::add_footnote("Standard errors clustered by buffer_id in parentheses. *** p<0.001, ** p<0.01, * p<0.05", 
                          notation = "none")
```

The base model indicates that active construction is associated with a statistically significant increase of 0.096 accidents per month within the buffer zone. However, the urban heterogeneity model reveals a striking difference between urban and rural areas. While construction in rural areas is associated with a reduction in accidents (-0.056), urban areas see a substantial positive effect (0.216 - 0.056 = 0.160).

When examining normalized accident rates (per 10,000 vehicles), the effects remain directionally consistent but with reduced statistical significance. This suggests that while traffic volume explains some of the effects, there remain important differences in how construction impacts safety across different environments.

### 3.3 Spatial Lag Model Results

The spatial lag model confirms the presence of significant spatial autocorrelation (ρ = 0.770), indicating that accident patterns exhibit strong spatial clustering. After accounting for this spatial dependence, construction effects remain significant:

```{r spatial_model_table}
# Spatial model results (using pre-computed values)
spatial_model_df <- data.frame(
  Variable = c("Intercept", "Treatment", "Urban", "log(Traffic Volume)", "Spatial Lag (ρ)"),
  Estimate = c(-4.260, 0.616, -0.523, 0.489, 0.770),
  Std_Error = c(0.247, 0.061, 0.111, 0.029, 0.006),
  z_value = c(-17.255, 10.074, -4.697, 16.773, 130.01),
  p_value = c("<0.001", "<0.001", "<0.001", "<0.001", "<0.001")
)

kableExtra::kable(spatial_model_df, 
                 col.names = c("Variable", "Estimate", "Std. Error", "z-value", "p-value"),
                 caption = "Spatial Lag Model Results",
                 booktabs = TRUE) %>%
  kableExtra::kable_styling(latex_options = c("striped", "hold_position"),
                           full_width = FALSE) %>%
  kableExtra::row_spec(0, bold = TRUE) %>%
  kableExtra::add_footnote("Model statistics: Log likelihood: -105341.4, AIC: 210690", 
                          notation = "none")
```

The spatial model confirms a strong positive effect of construction on accident counts even after controlling for spatial dependence, with construction zones experiencing 0.616 more accidents on average.

### 3.4 Spatial Spillover Effects

Our spillover analysis examined how construction impacts extend to different distance bands from the construction site. Figure 1 presents these results.

```{r spillover_figure, fig.cap="Treatment Effects by Distance from Construction"}
# Spillover model results (using pre-computed values)
spillover_df <- data.frame(
  distance_band = c("0-500m", "500-1000m", "1-2km", "2-5km"),
  estimate = c(0.099, 0.235, -0.042, 0.036),
  ci_low = c(-0.007, 0.170, -0.088, -0.016),
  ci_high = c(0.205, 0.300, 0.005, 0.087),
  p_value = c(0.068, "<0.001", 0.077, 0.171)
)

# Ensure proper ordering of distance bands
spillover_df$distance_band <- factor(spillover_df$distance_band, 
                                   levels = c("0-500m", "500-1000m", "1-2km", "2-5km"))

# Updated plot with correctly ordered x-axis
ggplot(spillover_df, aes(x = distance_band, y = estimate)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = ci_low, ymax = ci_high), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(
    x = "Distance from Construction Site",
    y = "Effect on Accident Count"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.grid.minor = element_blank()
  )
```

The spillover analysis reveals a non-linear pattern of effects:
- Immediate vicinity (0-500m): Marginally significant positive effect (0.099, p=0.067)
- Near proximity (500-1000m): Strong positive effect (0.235, p<0.001)
- Intermediate zone (1-2km): Marginally significant negative effect (-0.042, p=0.077)
- Extended zone (2-5km): Non-significant positive effect (0.036, p=0.171)

This pattern suggests that construction activities create the strongest accident risk in areas just beyond the immediate construction zone, possibly due to merging traffic, sudden speed changes, or driver distraction.

### 3.5 Event Study Results

The event study approach allows us to examine the dynamic effects of construction and test the parallel trends assumption.

```{r event_study_figure, fig.cap="Event Study: Effects Relative to Construction Start"}
# Event study results (using pre-computed values)
event_study_df <- data.frame(
  time_period = c("<-12m", "-12to-6m", "-6to-3m", "0to3m", "3to6m", "6to12m"),
  estimate = c(-0.332, -0.055, -0.007, -0.065, -0.146, -0.111),
  ci_low = c(-0.535, -0.141, -0.059, -0.114, -0.221, -0.220),
  ci_high = c(-0.130, 0.032, 0.046, -0.015, -0.070, -0.001),
  p_value = c(0.001, 0.214, 0.804, 0.010, "<0.001", 0.047)
)

# Ensure proper ordering of time periods
event_study_df$time_period <- factor(event_study_df$time_period, 
                                    levels = c("<-12m", "-12to-6m", "-6to-3m", "0to3m", "3to6m", "6to12m"))

# Updated plot with correctly ordered x-axis
ggplot(event_study_df, aes(x = time_period, y = estimate)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = ci_low, ymax = ci_high), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(
    x = "Time Relative to Construction Start",
    y = "Effect on Accident Count"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.minor = element_blank()
  )
```

The event study results show:
1. Significantly lower accident rates long before construction (<-12 months)
2. No significant difference in the pre-construction period (-12 to 0 months), supporting the parallel trends assumption
3. Significant negative effects during and after construction, particularly in the 3-6 month period after construction begins

This pattern suggests that while construction initially increases accident risks, there may be adaptation effects as drivers become accustomed to the changed road conditions.

### 3.6 Accident Severity Analysis

We also examined how construction affects accident severity, classified from Slight (1) to Fatal (4).

```{r severity_analysis, eval=FALSE}
# Analyze accident severity by construction period
severity_analysis <- accidents_in_buffers %>%
  st_drop_geometry() %>%
  filter(!is.na(time_period)) %>%
  count(time_period, Severity) %>%
  group_by(time_period) %>%
  mutate(proportion = n / sum(n)) %>%
  ungroup()

# Convert Severity to a factor before plotting
severity_analysis$Severity <- factor(severity_analysis$Severity, 
                                    levels = c(1, 2, 3, 4),
                                    labels = c("Slight", "Moderate", "Serious", "Fatal"))
```

```{r severity_figure, fig.cap="Accident Severity by Construction Period"}
# Severity analysis (using pre-computed values)
severity_df <- data.frame(
  time_period = rep(c("after", "before", "during"), each = 4),
  Severity = rep(c("Slight", "Moderate", "Serious", "Fatal"), 3),
  proportion = c(
    0.01, 0.88, 0.11, 0.001,  # after
    0.01, 0.88, 0.11, 0.001,  # before
    0.01, 0.87, 0.12, 0.001   # during
  )
)

severity_df$Severity <- factor(severity_df$Severity, 
                              levels = c("Slight", "Moderate", "Serious", "Fatal"))
severity_df$time_period <- factor(severity_df$time_period,
                                 levels = c("before", "during", "after"))

ggplot(severity_df, aes(x = time_period, y = proportion, fill = Severity)) +
  geom_col() +
  labs(
    x = "Relation to Construction",
    y = "Proportion"
  ) +
  scale_fill_brewer(palette = "Blues") +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank()
  )
```

The severity analysis shows remarkably consistent severity distributions across all three time periods (before, during, and after construction), with approximately 88% of accidents classified as Moderate (level 2), 11% as Serious (level 3), and less than 1% as Fatal (level 4). This suggests that while construction affects accident frequency, it does not substantially alter the severity distribution.

## 4. Discussion

### 4.1 Interpretation of Main Findings

Our analysis provides robust evidence that road construction causally impacts traffic accident rates, with important heterogeneity across urban and rural environments. The main findings can be summarized as follows:

1. **Overall Construction Effect**: Construction activities are associated with a statistically significant increase in accident counts within the affected areas. This effect remains robust across multiple model specifications.

2. **Urban-Rural Heterogeneity**: The impact of construction on accidents is substantially different between urban and rural environments. In urban areas, construction is associated with increased accident counts, while in rural areas, there is evidence of reduced accidents during construction. This heterogeneity may be explained by several factors:
   - Higher traffic density in urban areas creates more complex interactions around construction zones
   - Rural construction may induce more cautious driving or traffic diversion
   - Urban construction often involves more complex traffic management challenges

3. **Spatial Spillover Effects**: The impact of construction extends beyond the immediate construction zone, with particularly strong effects in areas 500-1000m from the construction site. This finding has important implications for traffic management strategies, suggesting that safety measures should extend well beyond the construction zone itself.

4. **Temporal Dynamics**: The event study results suggest an adaptation effect, where accident rates are higher immediately after construction begins but then decline as drivers adjust to the changed conditions.

### 4.2 Policy Implications

These findings have several important policy implications for transportation safety planning:

1. **Targeted Resource Allocation**: Traffic management resources should be allocated differentially between urban and rural construction zones, with greater emphasis on urban areas where construction effects on accidents are substantially larger.

2. **Extended Safety Zones**: Traffic safety measures should extend beyond the immediate construction zone, with particular attention to areas 500-1000m from the site where spillover effects are strongest.

3. **Temporal Planning**: Enhanced safety measures may be particularly important in the initial phases of construction before driver adaptation occurs.

4. **Consistent Severity Protocols**: Since construction does not substantially alter accident severity patterns, emergency response protocols need not be differentiated based on proximity to construction zones.

### 4.3 Limitations and Future Research

While our study provides robust evidence on the causal impact of construction on accidents, several limitations should be noted:

1. **Time Period**: Our analysis is limited to a single year (2021), which may have been affected by pandemic-related traffic patterns. Future research should examine longer time periods.

2. **Granularity**: While we distinguish between urban and rural areas, more detailed classifications (e.g., suburban, exurban) could provide additional insights.

3. **Construction Types**: Our data does not allow us to consistently differentiate between different types of construction activities (e.g., resurfacing, bridge repair, lane expansion), which may have different safety implications.

4. **Weather Interactions**: While our models include fixed effects that partly control for seasonal patterns, more explicit modeling of weather interactions with construction effects could be valuable.

Future research should address these limitations and further explore the mechanisms underlying the observed heterogeneity in construction effects.

## 5. Conclusion

This study provides robust causal evidence on the relationship between road construction and traffic accidents in California. Using high-resolution spatial data and advanced econometric methods, we demonstrate that construction activities significantly impact accident rates, with effects that vary substantially between urban and rural environments and extend beyond the immediate construction zone.

The findings have important implications for transportation safety planning, suggesting the need for targeted resource allocation, extended safety zones, and temporally-focused safety measures. By implementing these recommendations, transportation authorities can mitigate the safety risks associated with necessary infrastructure maintenance and development.

Overall, this research contributes to our understanding of transportation safety dynamics and provides an empirical foundation for evidence-based policy development in traffic management around construction zones.

## References

Conte, B., Desmet, K., Nagy, D.K., & Rossi-Hansberg, E. (2021). Local sectoral specialization in a warming world. Journal of Economic Geography, 21(4), 493-530.

Donaldson, D. (2018). Railroads of the Raj: Estimating the impact of transportation infrastructure. American Economic Review, 108(4-5), 899-934.

Fernihough, A., & O'Rourke, K.H. (2021). Coal and the European industrial revolution. The Economic Journal, 131(635), 1135-1149.

Giua, M. (2017). Spatial discontinuity for the impact assessment of the EU regional policy: The case of Italian objective 1 regions. Journal of Regional Science, 57(1), 109-131.

Heblich, S., Trew, A., & Zylberberg, Y. (2021). East-side story: Historical pollution and persistent neighborhood sorting. Journal of Political Economy, 129(5), 1508-1552.

Hsiao, A. (2023). Sea Level Rise and Urban Adaptation in Jakarta. Technical Report.

### 2025-03-26 final/archive/finalSIMON_MoritzPeistNoemiLucchiSimonVellin .Rmd
Finalized submission

--- 2025-03-25 final/finalSIMON_MoritzPeistNoemiLucchiSimonVellin .Rmd
@@
 
+Given the large amount of data, further visual analysis is specific to California and 2021. 
+We start with some basic visualizations of our sampled data, firstly focusing on accidents.
+
+```{r}
+# Plot the accidents 
+accidents <- ggplot() +
+  geom_sf(data = ca_roads, color = "gray50", size = 1.5, linetype = "solid") +
+  geom_sf(data = df.acc.sf, size = 0.2) +
+  scale_color_viridis_c(
+    name = "AADT",
+    option = "plasma",
+    labels = scales::comma) +
+  theme_bw() +
+  labs(title = "Accidents") 
+print(accidents)
+```
+
+```{r}
+# Plot accidents by severity 
+accidents_severity <- ggplot() +
+  geom_sf(data = ca_roads, color = "gray50", size = 1.5, linetype = "solid") +
+  geom_sf(data = df.acc.sf, aes(color = Severity), size = 0.2) +
+  scale_color_viridis_c(
+    name = "Severity",
+    option = "plasma",
+    labels = scales::comma) +
+  theme_bw() +
+  labs(title = "Accidents by Severity") 
+print(accidents_severity)
+```
+
+We can observe that most of the accidents are classified as level 2 severity.
+This suggests that the majority are not highly severe.
+Given this context, it is reasonable to infer that these accidents may be caused by distractions or disruptions, potentially linked to the presence of construction sites.
+
+Now, we shift our focus to construction sites, still concentrating solely on California.
+Due to the extremely high number of construction sites, many of which last only a few minutes or hours, we decided to exclude those with durations of less than one day.
+This reduction in the number of construction sites not only improves the clarity of the visualization but also prevents any potential bias in the subsequent analysis.
+Considering that in afterwards we will implement a 300-meter buffer, including too many short-duration sites would essentially cover all areas with construction buffers, skewing the results.
+
+```{r}
+# Remove constructions that last less than one day
+df.const.sf_filtered <- subset(df.const.sf, duration >= 1)
+```
+
+```{r}
+# Plot the construction sites by severity 
+constructions_severity <- ggplot() +
+  geom_sf(data = ca_roads, color = "gray50", size = 1.5, linetype = "solid") +
+  geom_sf(data = df.const.sf_filtered, aes(color = Severity), size = 0.2) +
+  scale_color_viridis_c(
+    name = "AADT",
+    option = "plasma",
+    labels = scales::comma) +
+  theme_bw() +
+  labs(title = "Constructions Sites by Severity") 
+print(constructions_severity)
+```
+
+In the context of construction sites, severity reflects the expected impact on traffic, with a scale ranging from 1 to 4, where 1 represents the least impact.
+We can observe that the majority of construction sites are expected to have a minimal impact on traffic, with a severity score of 1.
+However, there are a few sites with a higher severity score, particularly around Los Angeles and Sacramento.
+This suggests that traffic volume may already be taken into account when calculating the severity of these sites.
+
+```{r}
+constructions_duration <- ggplot() +
+  geom_sf(data = ca_roads, color = "gray50", size = 1.5, linetype = "solid") +
+  geom_sf(data = df.const.sf_filtered, aes(color = duration), size = 0.2) +
+  scale_color_viridis_c(
+    name = "Duration",
+    option = "plasma",
+    labels = scales::comma) +
+  theme_bw() +
+  labs(title = "Constructions Sites by Duration") 
+print(constructions_duration)
+```
+
+Now we combine our data about construction sites and accidents with data about traffic volume, the latter is referred to 2023, and was retrieved from the [U.S.
+Government's Open Data Website](https://data.gov).
+Traffic volume data was used to normalize the accident data, as we believed that comparing raw accident figures without accounting for the number of vehicles on the street could lead to biased results.
+Although the approach of combining different years for different sources of data is not methodologically robust, it helped in serving our purpose with the constraint of availability of data.
+Additionally, we assumed that the distribution of the kind of data we are analyzing is persistent, and so values are not likely to significantly change in a 2-year period.
+In the aforementioned website, we found data about Annual Average Daily Traffic (AADT), which is the total volume for the year divided by 365 days.
+The traffic count year runs from October 1st through September 30th.
+We had information about both back and ahead AADT, as well as back peak hour traffic volume and ahead peak hour traffic volume, referred to the post miles (instead of roads).
+We average the back and ahead values for every pair of indicators and perform our analysis with both to verify which one could have been more insightful.
+As expected, the average AADT shows more variability, making it more effective for our normalization purpose.
+
+```{r}
+# Define average AADT and average peak traffic volume and substitute original columns 
+traffic_data_aggregated <- df.traf.vol %>%
+  mutate(across(c(BACK_AADT, AHEAD_AADT, BACK_PEAK_, AHEAD_PEAK), 
+                ~as.numeric(as.character(.)))) %>%
+  # Average over back and ahead traffic - both in AADT and peak traffic 
+  mutate(
+    aadt = rowMeans(cbind(AHEAD_AADT, BACK_AADT), na.rm = TRUE),
+    avg_peak_traffic = rowMeans(cbind(BACK_PEAK_, AHEAD_PEAK), na.rm = TRUE)
+  ) %>%
+  # Remove initial columns 
+  select(-c(AHEAD_AADT, BACK_AADT, AHEAD_PEAK, BACK_PEAK_, AHEAD_PE_1, BACK_PEA_1))
+
+```
+
+```{r}
+# Plot average AADT with roads 
+traffic_roads <- ggplot() +
+  geom_sf(data = ca_roads, color = "gray50", size = 1.5, linetype = "solid") +
+  geom_sf(data = traffic_data_aggregated %>% filter(!is.na(aadt)),
+          aes(color = aadt), 
+          size = 0.2) +
+  scale_color_viridis_c(
+    name = "AADT",
+    option = "plasma",
+    labels = scales::comma) +
+  theme_bw() +
+  labs(title = "Annual Average Daily Traffic") 
+print(traffic_roads)
+```
+
+As follows we normalize the accidents by the traffic volume.
+To do so, we first assign each accident to its nearest post miles point and calculate the distance between these two.
+Secondly, we aggregate the accident data by closest traffic point and calculate the normalized accident rates.
+For each traffic point, we count the number of accidents and then divide these counts by the average AADT and multiply by 1,000 to get the rate per 1,000 vehicles.
+
+```{r}
+# Assign each accident to its nearest traffic measurement point 
+normalize_accidents_by_traffic <- function(df.acc.sf, traffic_data_aggregated, buffer_distance = 300) {
+  # Match coordinate systems
+  traffic_data_aggregated <- st_transform(traffic_data_aggregated, st_crs(df.acc.sf))
+
+  # Find the nearest traffic point for each accident
+  nearest_indices <- st_nearest_feature(df.acc.sf, traffic_data_aggregated)
+
+  # Calculate distances to each nearest point
+  distances <- st_distance(df.acc.sf, traffic_data_aggregated[nearest_indices,], by_element = TRUE)
+  distances <- as.numeric(distances)
+
+  # Add traffic data directly to the accidents dataframe
+  df.acc.sf$aadt <- traffic_data_aggregated$aadt[nearest_indices]
+
+  # Add the traffic point ID to each accident 
+  df.acc.sf$traffic_point_id <- nearest_indices
+
+  # Flag accidents that fall within the buffer distance
+  df.acc.sf$within_buffer <- distances <= buffer_distance
+
+  return(df.acc.sf)}
+
+# Apply the function
+df.acc.sf <- normalize_accidents_by_traffic(df.acc.sf, traffic_data_aggregated)
+
+# Count accidents per traffic point and normalize accidents counts
+traffic_point_stats <- df.acc.sf %>%
+  st_drop_geometry() %>%  
+  group_by(traffic_point_id) %>%
+  summarise(
+    accident_count = n(),
+    aadt = first(aadt),  
+    accidents_per_1000_vehicles = (accident_count / aadt) * 1000)
+
+# Create a new spatial object with traffic points and their accident rates
+traffic_points_with_rates <- traffic_data_aggregated %>%
+  mutate(traffic_point_id = row_number()) %>%  
+  left_join(traffic_point_stats, by = "traffic_point_id") %>%
+  mutate(
+    accident_count = replace_na(accident_count, 0),
+    accidents_per_1000_vehicles = replace_na(accidents_per_1000_vehicles,0))
+
+```
+
+As expected, the traffic volume does have an impact on accidents, so this normalization step was crucial in order to have robust and reliable results.
+
+```{r}
+# Distribution of accidents by traffic volume
+ggplot(df.acc.sf %>% filter(!is.na(aadt)), 
+       aes(x = aadt)) +
+  geom_histogram(bins = 30, fill = "steelblue") +
+  scale_x_log10() +  
+  theme_minimal() +
+  labs(title = "Accidents by Traffic Volume",
+       x = "Average Daily Traffic Volume (logarithmic scale)",
+       y = "Number of Accidens")
+```
+
+For further visualizations, we also calculate buffers to verify the intersection between areas close to construction sites and frequency of accidents.
+```{r}
+# Create construction buffers 
+create_construction_buffers <- function(df.const.sf, buffer_distance = 300) {
+  # Ensure proper projection
+  df.const.projected <- st_transform(df.const.sf, 3310)  # California Albers projection
+
+  # Create buffers
+  construction_buffers <- st_buffer(df.const.projected, dist = buffer_distance)
+
+  # Transform back to original CRS
+  construction_buffers <- st_transform(construction_buffers, st_crs(df.const.sf))
+  return(construction_buffers)}
+
+# Apply the function 
+construction_buffers <- create_construction_buffers(df.const.sf)
+```
+
+Due to memory constraints, the following interactive map has been further filtered to focus solely on Los Angeles.
+This visualization highlights construction buffers (areas within 300 meters of construction sites) in red, and accidents in blue.
+We specifically chose Los Angeles for this analysis, as prior visualizations revealed that most accidents and construction sites are concentrated in this region, along with Sacramento.
+This is to be expected, given the high population density in these areas.
+
+```{r}
+# Define the Los Angeles bounding box
+la_bbox <- list(
+  min_lng = -118.75, max_lng = -117.85,
+  min_lat = 33.65, max_lat = 34.35)
+
+# Filter data for Los Angeles
+la_accidents <- df.acc.sf %>%
+  filter(
+    st_coordinates(.)[,1] >= la_bbox$min_lng,
+    st_coordinates(.)[,1] <= la_bbox$max_lng,
+    st_coordinates(.)[,2] >= la_bbox$min_lat,
+    st_coordinates(.)[,2] <= la_bbox$max_lat)
+
+la_buffers <- construction_buffers %>%
+  st_filter(st_as_sfc(st_bbox(c(
+    xmin = la_bbox$min_lng, xmax = la_bbox$max_lng,
+    ymin = la_bbox$min_lat, ymax = la_bbox$max_lat
+  ), crs = st_crs(construction_buffers))))
+
+# Interactive map for accidents within construction buffers 
+la_buffer_accident_map <- leaflet() %>%
+  addProviderTiles(providers$CartoDB.Positron) %>%
+  # Add construction buffers
+  addPolygons(data = la_buffers, 
+              color = "#FF5500", 
+              weight = 2, 
+              opacity = 0.8,
+              fillOpacity = 0.4,
+              group = "Construction Buffers") %>%
+  # Add accidents 
+  addCircles(data = la_accidents,
+             radius = 1.5,  
+             color = "#3388FF",
+             fillColor = "#3388FF",
+             fillOpacity = 0.6,
+             weight = 1,
+             group = "Accidents") %>%
+  # Add layer controls
+  addLayersControl(
+    overlayGroups = c("Construction Buffers", "Accidents"),
+    options = layersControlOptions(collapsed = FALSE)
+  ) %>%
+  # Center on Los Angeles
+  setView(lng = -118.25, lat = 34.05, zoom = 10)
+la_buffer_accident_map
+
+```
+
 <hr>

### 2025-03-26 final/final_MoritzPeistNoemiLucchiSimonVellin.Rmd
Finalized submission

@@
 - Simon Vellin
-date: "`r format(Sys.Date(), '%B %d, %Y')`"
+date: "`r withr::with_locale(c('LC_TIME' = 'en_US.UTF-8'), format(Sys.Date(), '%B %d, %Y'))`"
 output:
@@
   message = FALSE,
+  progress = FALSE,
   fig.width = 8,
@@
 
-## 2.1 ["Effects of truck traffic on crash injury severity on rural highways in Wyoming using Bayesian binary logit models" by Ahmed et al.
+## ["Effects of truck traffic on crash injury severity on rural highways in Wyoming using Bayesian binary logit models" by Ahmed et al.
 (2018)](https://www.sciencedirect.com/science/article/pii/S0001457518301507)
@@
 
-## 2.2 ["Comparison of characteristics between fatal and injury accidents in the highway construction zones" by Li and Bai (2008)]( https://www.sciencedirect.com/science/article/pii/S0925753507001014)
+## ["Comparison of characteristics between fatal and injury accidents in the highway construction zones" by Li and Bai (2008)]( https://www.sciencedirect.com/science/article/pii/S0925753507001014)
 
@@
 ```{r data_loading}
+# Redirect all output to a temporary file (suppresses progress bars)
+sink(tempfile())
+
 # Load Caltrans State Highway Network
@@
   st_transform(4326)
+
+# Restore normal output
+sink()
 ```
@@
 ```
+
+### Traffic Volumes
+```{r AADT visualization}
+# Data Preparation: Aggregate Traffic Volume
+traffic_data_aggregated <- aadt %>%
+  mutate(across(c(BACK_AADT, AHEAD_AADT, BACK_PEAK_, AHEAD_PEAK), 
+                ~as.numeric(as.character(.)))) %>%
+  # Average traffic volumes from both directions
+  mutate(
+    aadt = rowMeans(cbind(AHEAD_AADT, BACK_AADT), na.rm = TRUE),
+    avg_peak_traffic = rowMeans(cbind(BACK_PEAK_, AHEAD_PEAK), na.rm = TRUE)
+  ) %>%
+  # Remove initial columns 
+  select(-c(AHEAD_AADT, BACK_AADT, AHEAD_PEAK, BACK_PEAK_, AHEAD_PE_1, BACK_PEA_1))
+
+# Visualization of Annual Average Daily Traffic
+traffic_roads_plot <- ggplot() +
+  # Add California road network as background
+  geom_sf(data = ca_roads, color = "gray50", size = 1.5, linetype = "solid") +
+
+  # Add traffic volume data points
+  geom_sf(data = traffic_data_aggregated %>% filter(!is.na(aadt)),
+          aes(color = aadt), 
+          size = 2) +
+
+  # Use viridis color palette for continuous traffic volume
+  scale_color_viridis_c(
+    name = "Annual Average\nDaily Traffic",
+    option = "plasma",
+    labels = scales::comma,
+    direction = 1
+  ) +
+
+  # Zoom in to California boundaries
+  coord_sf(
+    xlim = c(-124.5, -114.5),  # Longitude limits for California
+    ylim = c(32.5, 42.5),      # Latitude limits for California
+    expand = FALSE
+  ) +
+
+  # Customize plot theme
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, face = "bold"),
+    legend.position = "right",
+    legend.title = element_text(face = "bold")
+  ) +
+
+  # Add informative title and subtitle
+  labs(
+    title = "California Road Traffic Volumes",
+    subtitle = "Spatial Distribution of Annual Average Daily Traffic for Post Miles",
+    caption = "Data source: Caltrans State Highway Network"
+  )
+
+# Display the plot
+print(traffic_roads_plot)
+```
+The visualization reveals the spatial distribution of Annual Average Daily Traffic (AADT) across California's road network.
+The highest traffic volumes (indicated by yellow and red points) are concentrated in major urban centers, particularly around Los Angeles, San Francisco, and San Diego metropolitan areas, with volumes exceeding 200,000 vehicles per day in some locations.
+Rural and mountainous regions show significantly lower traffic volumes (dark purple), often below 50,000 vehicles per day.
+This spatial variation in traffic exposure is essential for our analysis, as it provides context for normalizing accident rates and understanding how construction impacts might differ between high and low-traffic areas.
 
@@
 
-The temporal distribution reveals distinct patterns for accidents and construction projects.
-Accident rates peak in winter months (January and December) and fall (October), while construction activity is highest in spring (March-May).
-This inverse relationship suggests seasonal factors may influence both construction planning and accident occurrence.
+The temporal distribution reveals a striking inverse relationship between accidents and construction projects throughout 2021.
+Accident rates show clear seasonality with peaks in January (~38,000 accidents) and December (~36,000 accidents), and a notable drop in March-April (~20,000 accidents).
+In contrast, construction activity peaks in March (~5,000 new projects) – precisely when accident rates are lowest – and maintains a relatively high level through June, before gradually declining toward the end of the year.
+This pattern suggests that seasonal factors such as weather conditions and daylight hours may influence both construction planning and accident occurrence, with transportation agencies potentially scheduling more construction during periods of historically lower accident rates.
 
@@
 ```
+
+The spatial distribution maps reveal strong geographic patterns for both accidents and construction activities in California.
+Accidents (top map) are heavily concentrated along major transportation corridors and urban centers, with particularly high densities in the Los Angeles Basin, San Francisco Bay Area, and the Central Valley along Interstate 5.
+The temporal coloring shows some month-to-month variation but suggests relatively consistent spatial patterns throughout the year.
+
+Similarly, construction sites (bottom map) follow the state's major transportation infrastructure, with higher concentrations in metropolitan areas.
+The severity classification shows a mix of construction types across regions, with higher severity projects (darker colors) somewhat more prevalent in urban cores.
+This spatial correlation between accidents and construction sites highlights the importance of controlling for location-specific factors in our causal analysis, as both are influenced by underlying traffic patterns and infrastructure density.
 
@@
 
-The heatmaps confirm the spatial concentration of both accidents and construction activities in California's major urban areas.
-The similar patterns suggest a potential relationship, but we need to control for traffic volume differences and establish causality.
+The density heatmaps provide a striking visualization of the spatial concentration patterns for both accidents and construction activities across California.
+The accident heatmap (top) reveals intense clustering in major urban centers, with the Los Angeles metropolitan area showing the highest density (red), followed by the San Francisco Bay Area and San Diego (yellow-green).
+Several medium-intensity clusters (blue-cyan) appear along major highways in the Central Valley.
+
+Similarly, the construction density heatmap (bottom) shows pronounced concentration in the same urban areas, with particularly high intensity (red-orange) in Los Angeles and the Bay Area.
+These parallel patterns strongly suggest that both accidents and construction activities follow underlying population and traffic volume distributions.
+This spatial correlation underscores the importance of our difference-in-differences approach with location fixed effects, which will help isolate the causal impact of construction on accidents by controlling for these shared geographic determinants.
 
@@
 
-Our analysis of the urban-rural breakdown is critical for understanding the heterogeneous effects of construction on accident rates.
+The urban-rural classification shows that approximately 66.1% of construction sites in our dataset are located in urban areas, while 33.9% are in rural areas.
+This distribution is essential for our heterogeneity analysis, as it ensures we have sufficient observations in both urban and rural categories to detect differential effects.
+The substantial representation of both categories allows us to investigate how construction impacts accident rates differently across these distinct environments, which has important implications for targeted safety measures and resource allocation.
 
@@
 
-The traffic volume distribution shows a pattern likely reflecting the differences between urban and rural areas.
-Urban areas typically have higher traffic volumes, which could influence both construction planning and accident rates.
+The traffic volume distribution shows substantial variation across construction sites, with a median AADT of 72,000 vehicles per day but ranging from as low as 107 to over 379,000 vehicles per day.
+The right-skewed distribution (evident in the first histogram) indicates that most construction sites experience moderate traffic volumes, while a smaller number of sites have extremely high traffic.
+The log-scale visualization reveals a more nuanced bimodal pattern, suggesting two distinct categories of roads: lower-volume rural or suburban routes (clustered around 10^3-10^4 vehicles/day) and higher-volume urban arterials and highways (clustered around 10^5 vehicles/day).
+This considerable variation in traffic exposure is critical to control for in our analysis, as it directly affects the accident risk baseline across different locations.
 
@@
 
-When examining accidents by their timing relative to construction, this distribution provides initial evidence that construction activities may impact accident patterns, but we need more rigorous analysis to establish causality.
+The distribution of accidents by construction period reveals a notable pattern: 46.2% of accidents within our buffer zones occurred before construction began, 24.1% occurred during active construction, and 29.7% occurred after construction ended.
+This preliminary finding suggests potential changes in accident patterns associated with construction activities.
+The relatively lower percentage during construction could indicate several possibilities: (1) driver adaptation to construction zones through increased caution, (2) reduced traffic volume due to drivers avoiding construction areas, or (3) effective traffic management measures during construction.
+However, this descriptive analysis doesn't control for other factors like location characteristics or seasonal patterns, highlighting the need for our more rigorous difference-in-differences approach.
 
@@
 
-The visualization illustrates the spatial relationship between construction buffers (red areas) and accident locations (blue points).
-Many accidents occur within or near construction zones, suggesting a potential spatial correlation.
+The visualization of the Los Angeles area provides a clear illustration of the spatial relationship between construction buffers (red circles) and accident locations (blue points).
+Many accidents cluster within or near the 500-meter construction buffer zones, particularly along major transportation corridors.
+We observe that some buffers contain dense accident clusters, while others have relatively few accidents, suggesting potential heterogeneity in the impact of construction on accident frequency.
+This visualization highlights the importance of our buffer-based analytical approach, which allows us to capture localized effects of construction activities on nearby accident patterns while controlling for location-specific factors through fixed effects.
 
@@
 
+The base difference-in-differences model provides compelling evidence for the causal impact of construction on accident rates.
+The coefficient for active construction (treatment) is 0.101 (p < 0.001), indicating that during construction periods, the average number of accidents in buffer zones increases by approximately 0.101 accidents per month compared to non-construction periods.
+With clustered standard errors at the buffer level, this effect is highly statistically significant, suggesting a robust relationship.
+The low within-R² (9.001e-5) is common in fixed effects models with binary treatment variables and indicates that while construction activity explains a small portion of the overall variation in accidents, its effect is statistically meaningful and precisely estimated.
+
 #### Heterogeneity Analysis
@@
 
-Contrary to our initial hypothesis, we found significant heterogeneous effects between urban and rural environments.
-In urban areas, construction is associated with a more pronounced increase in accident rates (interaction coefficient = 0.163, p < 0.01), while rural areas show a potentially different pattern, suggesting that traffic management challenges differ substantially by environment.
+The heterogeneity analysis reveals striking differences in how construction affects accidents across urban and rural environments.
+The base treatment effect coefficient (-0.005) is not statistically significant, suggesting minimal impact of construction on accidents in rural areas.
+However, the interaction term (treatment × is_urbanTRUE) shows a large and highly significant coefficient of 0.163 (p < 0.01), indicating that in urban areas, construction is associated with an additional 0.163 accidents per month beyond any rural effect.
+This substantial urban-rural disparity (with urban areas experiencing approximately 16.3% more accidents during construction) suggests that traffic management challenges differ fundamentally between these environments, with urban construction zones posing significantly greater accident risks that require targeted mitigation strategies.
 
@@
 ```
+
+When examining normalized accident rates per 10,000 vehicles, we observe different patterns than in the absolute count models.
+In the base rate model, the treatment effect is positive but not statistically significant (0.013, p=0.41), suggesting that after controlling for traffic exposure, construction's impact on accident rates is more nuanced.
+
+However, the urban heterogeneity model for normalized rates shows a similar pattern to our count model: the base treatment effect for rural areas is negative (-0.059, p=0.12) though not statistically significant, while the urban interaction term is positive and significant (0.110, p<0.01).
+This indicates that in urban areas, construction is associated with an increase of 0.110 accidents per 10,000 vehicles compared to rural areas.
+These normalized models provide important context by accounting for traffic volume differences, suggesting that the urban-rural disparity in construction impacts persists even after controlling for exposure levels.
 
@@
 
-# Create spatial weights - might need to adjust distance for your specific data
+# Create spatial weights - might need to adjust distance for our specific data
 sp_nb <- spdep::dnearneigh(spw_coords, d1 = 0, d2 = 5000)
@@
 ```
+
+The spatial lag model confirms the presence of significant spatial autocorrelation in accident patterns, with a spatial lag parameter (ρ) of 0.765 (p < 0.001).
+This high and statistically significant value indicates strong spatial clustering in accident occurrences, where accident rates in one location are substantially influenced by rates in neighboring areas.
+
+After accounting for this spatial dependence, the treatment effect of construction remains positive and significant (0.622, p < 0.001), providing additional evidence for our main finding.
+The model also identifies a negative effect of urban classification (-0.256, p < 0.01), likely reflecting the fixed characteristics of urban areas that are captured in this cross-sectional spatial model rather than the panel approach.
+The significant positive coefficient for traffic volume (0.438, p < 0.001) confirms the expected relationship between traffic exposure and accident frequency.
+This spatial analysis strengthens our causal claims by demonstrating that construction effects persist even after controlling for complex spatial dependencies.
 
@@
 Our spatial spillover analysis reveals nuanced effects extending beyond the immediate construction zone.
-Notably, areas 500m to 2km from construction sites demonstrate statistically significant increases in accident rates (coefficients ranging from 0.308 to 0.312, p < 0.001), suggesting that safety management strategies should extend well beyond the immediate construction boundaries.
+Using the ">5km" distance band as reference, we find that areas within 0-500m of construction sites show a negative but non-significant effect (-0.064, p=0.93), which may reflect successful traffic management or diversion within immediate construction boundaries.
+
+Notably, the strongest positive effects appear in the 500-1000m band (0.308, p<0.001) and 1-2km band (0.312, p<0.001), indicating significantly higher accident rates in these transition zones where traffic flow and driver behavior adjust to approaching or leaving construction.
+Even in the 2-5km band, we observe a smaller but still significant effect (0.139, p<0.001).
+These findings suggest that safety management strategies should extend well beyond immediate construction boundaries, with particular attention to transition zones where accident risk may actually be higher than at the construction site itself due to changing traffic patterns and driver adaptation.
 
@@
 ```
+
+The event study results provide temporal context for our difference-in-differences estimates.
+Using the period 3 months before construction ("-3to0m") as the reference, we observe that locations destined for construction show significantly lower accident rates far before construction begins ("<-12m" coefficient = -0.690, p<0.001), with this difference gradually diminishing as construction approaches.
+
+During and after construction (0to3m, 3to6m, 6to12m, >12m), coefficients are close to zero and not statistically significant, suggesting accident rates similar to the pre-construction reference period.
+This pattern supports the parallel trends assumption required for causal interpretation of our DiD results, as we don't see diverging pre-trends immediately before treatment.
+The significant difference in very early periods may reflect long-term planning effects, where areas selected for eventual construction have different characteristics.
+Overall, these results strengthen our causal claims by demonstrating appropriate temporal patterns in the data.
 
@@
 
-The summary statistics reveal important patterns in accident occurrence by construction status and urban/rural setting, providing context for our main regression analyses.
+The summary statistics reveal stark differences in accident patterns across urban-rural divides and construction status.
+In rural areas without construction (FALSE, FALSE), the mean accident count is 0.647 with a median of 0, while in urban areas without construction (FALSE, TRUE), the mean jumps to 4.383 with a median of 1 – nearly seven times higher.
+Similarly, during active construction, urban areas (TRUE, TRUE) show mean accident counts of 4.672 compared to 0.783 in rural areas (TRUE, FALSE).
+
+Notably, when examining normalized accident rates per 10,000 vehicles, the urban-rural gap narrows substantially: rural areas show rates of 0.986 (without construction) and 0.827 (with construction), while urban areas show rates of 0.395 (without construction) and 0.448 (with construction).
+This suggests that higher urban accident counts are partly explained by higher traffic volumes.
+The data also reveals that construction appears to reduce normalized accident rates in rural areas (-16%) while increasing them in urban areas (+13%), providing preliminary evidence for heterogeneous effects that will be further examined in our regression models.
 
@@
 
-The models show how active construction is associated with changes in accident counts, with notable differences between urban and rural areas.
-The urban heterogeneity model in particular reveals important differential impacts.
-When examining normalized accident rates (per 10,000 vehicles), the effects may change in magnitude or significance, suggesting traffic volume plays a complex role in the relationship.
+The difference-in-differences models provide robust evidence for the causal impact of construction on accident rates, with important heterogeneity across urban and rural environments.
+The base model shows a significant positive effect of construction (treatment = 0.101, p < 0.001), indicating that construction activities increase accident counts by approximately 10.1% on average.
+
+When examining urban-rural heterogeneity, we find that the effect is concentrated in urban areas: while the base treatment effect (representing rural areas) is actually negative but insignificant (-0.005), the urban interaction term is strongly positive (0.163, p < 0.01), indicating substantially higher accident rates during construction in urban environments.
+
+The normalized rate models (controlling for traffic volume) show complementary results, with the urban interaction term remaining significant (0.110, p < 0.01) even after traffic exposure normalization.
+This strengthens our conclusion that urban and rural areas experience fundamentally different safety impacts from construction activities, with urban construction zones posing greater accident risks that cannot be explained solely by traffic volume differences.
 
@@
 
-The spatial model confirms the effects of construction on accident counts even after controlling for spatial dependence, providing additional evidence for the relationship.
+The spatial lag model provides strong evidence for both spatial dependencies in accident patterns and the impact of construction activities.
+The spatial lag parameter (ρ = 0.765, p < 0.001) confirms significant spatial autocorrelation, indicating that accident rates in one location are strongly influenced by neighboring areas.
+Even after accounting for this spatial clustering, the treatment effect of construction remains positive and highly significant (0.622, p < 0.001), providing additional support for our main finding.
+
+Interestingly, in this cross-sectional spatial model, the Urban coefficient is negative (-0.256, p = 0.001), likely capturing fixed characteristics of urban areas rather than the treatment interaction effects seen in our panel models.
+The significant positive coefficient for traffic volume (0.438, p < 0.001) confirms the expected relationship between traffic exposure and accident frequency.
+With a log-likelihood of -101588.9 and AIC of 203190.0, the model demonstrates good fit, strengthening our confidence in the identified spatial relationships and treatment effects.
 
@@
 
-The spillover analysis reveals a pattern of effects at different distances from construction sites, with particularly notable impacts at specific distances.
-This suggests that construction effects aren't limited to the immediate construction zone.
+The spillover analysis reveals a non-intuitive spatial pattern of construction effects on accident rates.
+Using areas beyond 5km as the reference category, we observe that the immediate construction zone (0-500m) shows a slight negative but non-significant effect (-0.064, p=0.93), possibly reflecting successful traffic management or driver avoidance of these areas.
+
+However, the strongest positive effects appear in transition zones: the 500-1000m band (0.308, p<0.001) and 1-2km band (0.312, p<0.001) show statistically significant increases in accident counts, with magnitudes over three times larger than our base model estimate.
+Even in the 2-5km band, we observe a smaller but still significant effect (0.139, p<0.001).
+
+This spatial pattern suggests that accident risk may actually peak in areas approaching or leaving construction zones rather than within them, where speed changes, lane merges, and driver adjustment to changing conditions create hazardous transition environments.
+These findings have important implications for traffic management, indicating that safety measures should extend well beyond construction boundaries with particular attention to these high-risk transition zones.
 
@@
 
-The event study results show patterns of accident rates before and after construction begins, providing insights into temporal dynamics and supporting or challenging the parallel trends assumption required for causal interpretation of our DiD results.
+The event study results reveal important temporal dynamics in accident patterns around construction periods.
+Using the 3 months before construction start as the reference period, we observe that areas selected for construction have significantly lower accident rates long before construction begins (coefficient for "<-12m" = -0.690, p<0.001), with this difference gradually diminishing as construction approaches.
+
+During and after construction (periods "0to3m" through ">12m"), coefficients remain close to zero and statistically insignificant, indicating accident rates similar to the immediate pre-construction period.
+This pattern supports two important conclusions: (1) the parallel trends assumption necessary for causal interpretation of our DiD estimates is reasonably satisfied in the periods immediately before treatment, and (2) the long-term pre-construction difference suggests selection effects in areas chosen for construction projects.
+
+These temporal dynamics provide context for interpreting our main results, suggesting that while construction causes a relative increase in accidents compared to the immediate pre-construction period, accident rates might still be lower than in typical non-construction areas due to selection effects.
+This nuanced interpretation strengthens our analysis by accounting for temporal heterogeneity in treatment effects.
 
@@
 
-The severity analysis reveals a consistent distribution of accident severity across construction periods.
-Approximately 87% of accidents are classified as moderate severity, with about 12% categorized as serious.
-Interestingly, the proportional stability suggests that while construction impacts accident frequency, it does not substantially alter the underlying severity distribution.
+The severity analysis reveals a remarkably consistent distribution of accident severity levels across different construction periods.
+Approximately 87% of accidents are classified as moderate severity (level 2) across all periods, with around 12% categorized as serious (level 3) and less than 1% as either slight (level 1) or fatal (level 4).
+
+This stability suggests that while construction impacts accident frequency, as demonstrated in our main models, it does not substantially alter the proportional severity distribution.
+This finding has important implications for safety planning, indicating that construction zones don't necessarily produce more severe accidents but rather increase the overall number of accidents while maintaining similar severity patterns.
+Traffic management strategies should therefore focus primarily on reducing accident occurrence rather than specifically targeting severity reduction, as the underlying severity distribution appears to be a stable characteristic of the road network and driver behavior.
 
@@
 1.
-**Overall Construction Effect**: Our base difference-in-differences model reveals a statistically significant increase in accident counts during construction periods (coefficient = 0.101, p < 0.001), indicating that road construction activities are associated with approximately 10.1% more accidents compared to non-construction periods..
+**Overall Construction Effect**: Our base difference-in-differences model reveals a statistically significant increase in accident counts during construction periods (coefficient = 0.101, p < 0.001), indicating that road construction activities are associated with approximately 10.1% more accidents compared to non-construction periods.
 
@@
 **Urban-Rural Heterogeneity**: The impact of construction on accidents differs substantially between urban and rural environments.
-In urban areas, construction is generally associated with increased accident counts, while in rural areas, there is evidence of potentially reduced accidents during construction.
+In urban areas, construction is associated with a significant increase in accident counts (interaction coefficient = 0.163, p < 0.01), while in rural areas, there is evidence of potentially reduced accidents during construction (base coefficient = -0.005, though not statistically significant).
 This heterogeneity may be explained by several factors:
@@
 **Spatial Spillover Effects**: The spatial spillover analysis shows that construction impacts extend beyond the immediate construction zone, with effects varying by distance.
-This finding has important implications for traffic management strategies, suggesting that safety measures should extend beyond the construction zone itself to areas where merging and speed adjustment typically occur.
+Areas 500m-2km from construction sites show the largest increases in accident rates (coefficients 0.308-0.312, p < 0.001), while immediate construction zones (0-500m) show minimal effects.
+This suggests that transition zones where traffic adjusts to approaching or leaving construction are particularly high-risk areas that warrant targeted safety measures.
 
 4.
-**Temporal Dynamics**: The event study results provide insights into accident patterns before, during, and after construction.
-We observe changes in accident rates during different phases of construction, which may reflect adaptation effects as drivers adjust to changed road conditions.
+**Temporal Dynamics**: The event study results show that locations selected for construction have different accident patterns long before construction begins, with accident rates gradually approaching the reference period as construction nears.
+This suggests potential selection effects in areas chosen for construction projects and provides context for interpreting our main treatment effects.
 
 5.
-**Severity Patterns**: Our severity analysis shows the distribution of accident severity levels (Slight, Moderate, Serious, Fatal) across different construction periods.
-This helps us understand whether construction not only affects accident frequency but also severity levels.
+**Severity Patterns**: Our severity analysis shows a consistent distribution of accident severity levels across different construction periods (approximately 87% moderate, 12% serious), indicating that construction affects accident frequency but not the underlying severity distribution.
 
@@
 1.
-**Targeted Resource Allocation**: Traffic management resources should be allocated differentially between urban and rural construction zones, with greater emphasis on urban areas where construction effects on accidents are typically larger.
+**Targeted Resource Allocation**: Traffic management resources should be allocated differentially between urban and rural construction zones, with greater emphasis on urban areas where construction effects on accidents are typically larger (coefficient = 0.163, p < 0.01 for urban interaction).
+Rural construction zones may require less intensive intervention given their potentially neutral or even positive safety effects.
 
 2.
-**Extended Safety Zones**: Traffic safety measures should not be limited to the immediate construction zone but should extend to areas surrounding construction sites where our spillover analysis indicates significant effects.
-
-3. **Temporal Planning**: The temporal patterns revealed in our event study suggest that enhanced safety measures may be particularly important during specific phases of construction projects.
-
-4. **Severity-Focused Strategies**: Our severity analysis provides insights for emergency response planning and the design of safety measures aimed at mitigating more severe accidents.
+**Extended Safety Zones**: Traffic safety measures should not be limited to the immediate construction zone but should extend to areas surrounding construction sites, particularly the 500m-2km transition zones where our spillover analysis indicates significant effects (coefficients 0.308-0.312, p < 0.001).
+This could include enhanced signage, speed management, and temporary traffic calming measures in these approach and departure areas.
+
+3.
+**Temporal Planning**: The temporal patterns revealed in our event study suggest that enhanced safety measures may be particularly important during specific phases of construction projects.
+The relatively stable accident rates during and after construction compared to the immediate pre-construction period suggest that current construction management practices may be partially effective but could be further improved.
+
+4.
+**Severity-Focused Strategies**: Our severity analysis provides insights for emergency response planning and the design of safety measures.
+Since the severity distribution remains consistent across construction periods, general accident prevention measures rather than severity-specific interventions may be most effective.
 
@@
 6.
-**AADT data**: The AADT data used for traffic volume normalization may include estimates or projections that do not precisely match the 2021 study period, potentially introducing minor measurement uncertainties."
+**AADT data**: The AADT data used for traffic volume normalization may include estimates or projections that do not precisely match the 2021 study period, potentially introducing minor measurement uncertainties.
 

### 2025-03-26 final/final_MoritzPeistNoemiLucchiSimonVellin.Rmd
Acknowledged data duplication

@@
 # Join accidents to construction buffers - this operation can be computationally intensive
-# For very large datasets, consider using a spatial database or partitioning the data
 accidents_in_buffers <- st_join(
@@
 head(accidents_in_buffers %>% st_drop_geometry())
-```
+
+# Print number of rows
+print(nrow(accidents_in_buffers))
+```
+
+As it is visible now, the data almost quadruples for accidents in buffers.
+Preserving all spatial intersections between accidents and construction buffers through an unrestricted join (`st_intersects`) provides a more realistic representation of road safety dynamics, particularly in complex environments where construction activities frequently overlap.
+By maintaining the many-to-many relationship, our analysis acknowledges that a single accident may be legitimately associated with multiple construction zones, especially in urban corridors where projects occur in close proximity.
+This approach aligns with our research objective of measuring the overall safety impact of construction presence rather than attributing causality to specific projects.
+The difference-in-differences framework naturally accommodates this data structure by focusing on the binary treatment state (construction present or absent) within each spatial-temporal unit, while our fixed effects for buffer zones and time periods effectively control for systematic variations in accident frequencies across locations and periods.
+This methodological approach captures the real-world density patterns of construction activities without requiring artificial deduplication that might obscure important spatial relationships in our road safety analysis.
 
